### 2022-12-23 paper/main.tex
setting up paper

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{geometry}
\usepackage[sort&compress, numbers, super]{natbib}
\usepackage{pxfonts}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{lineno}

\newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits}

%\newcommand{\demo}{S1}

\doublespacing
\linenumbers

\title{Carryover clustering effects in free recall reveal how prior experiences influence memory for new experiences}
\author{Jeremy R. Manning\textsuperscript{1, *}, Andrew C. Heusser\textsuperscript{1, 2}, Kirsten Ziman\textsuperscript{1, 3},\\Emily Whitaker\textsuperscript{1}, and Paxton C. Fitzpatrick\textsuperscript{1}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili Interactive\\\textsuperscript{3}Princeton University\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}

\date{}

\begin{document}
\maketitle

\begin{abstract}
Insert abstract here
\end{abstract}

\section*{Main text}
Insert text here (Figs.~\ref{fig:demo}, \demo).

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/recall_dynamics}
\caption{\textbf{Recall dynamics in free recall.}}
\label{fig:recall-dynamics}
\end{figure}

\bibliographystyle{apa}
\bibliography{cdl}
\end{document}

### 2022-12-23 paper/supplement.tex
setting up paper

--- 2022-12-23 paper/main.tex
@@
-\documentclass[10pt]{article}
+\documentclass{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{geometry}
-\usepackage[sort&compress, numbers, super]{natbib}
+\usepackage{natbib}
 \usepackage{pxfonts}
 \usepackage{graphicx}
+\usepackage{newfloat}
 \usepackage{setspace}
 \usepackage{hyperref}
-\usepackage{lineno}
+\usepackage{placeins}
 
 \newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits}
+\newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}
 
-%\newcommand{\demo}{S1}
+\newcommand{\demo}{1}
 
-\doublespacing
-\linenumbers
-
-\title{Carryover clustering effects in free recall reveal how prior experiences influence memory for new experiences}
+\title{\textit{Supplementary materials for}: Carryover clustering effects in free recall reveal how prior experiences influence memory for new experiences}
 \author{Jeremy R.
@@
 
-\date{}
+\bibliographystyle{apa}
 
 \begin{document}
+
+\renewcommand{\figurename}{Supplementary Figure}
+
+%\begin{titlepage}
+%  \maketitle
+%  \thispagestyle{empty}
+%\end{titlepage}
+
+\setcounter{equation}{0}
+\setcounter{figure}{0}
+\setcounter{table}{0}
+\setcounter{page}{1}
+\setcounter{section}{0}
+\makeatletter
+\renewcommand{\theequation}{S\arabic{equation}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\bibnumfmt}[1]{[S#1]}
+\renewcommand{\citenumfont}[1]{S#1}
+
+\newcommand{\demo}{1}
+
 \maketitle
 
-\begin{abstract}
-Insert abstract here
-\end{abstract}
-
-\section*{Main text}
-Insert text here (Figs.~\ref{fig:demo}, \demo).
-
-\begin{figure}[tp]
-\centering
-\includegraphics[width=\textwidth]{figures/recall_dynamics}
-\caption{\textbf{Recall dynamics in free recall.}}
-\label{fig:recall-dynamics}
-\end{figure}
-
+\newpage
+\renewcommand{\refname}{Supplementary references}
 \bibliographystyle{apa}
 \bibliography{cdl}
+
 \end{document}

### 2022-12-23 paper/main.tex

@@
 
-\title{Carryover clustering effects in free recall reveal how prior experiences influence memory for new experiences}
+\title{Carryover effects in free recall reveal how prior experiences influence memory for new experiences}
 \author{Jeremy R.
@@
 \begin{abstract}
-Insert abstract here
+We perceive, interpet, and remember ongoing experiences through the lens of our prior experiences.
+Inferring that we are one type of situation versus another can lead us to interpret the same physical
+experience differently.  In turn, this can affect how we focus our attention, form expectations of what will happen next, 
+remember what is happening now, draw on our prior related experiences, and so on.  To study these phenomena,
+we asked participants to perform simple word list learning tasks.  Across different experimental conditions, we held the set of
+to-be-learned words constant, but we manipulated the orders in which the words were studied.  We found that 
+these order manipulations affected not only how the participants recalled the ordered lists, but also how they recalled later randomly
+ordered lists.  Our work shows how structure in our ongoing experiences can exert influence on how we remember unrelated
+subsequent experiences.
 \end{abstract}
 
-\section*{Main text}
-Insert text here (Figs.~\ref{fig:demo}, \demo).
+\section*{Introduction}
+
+% the role of context and prior experience in memory
+
+% situation models: forming expectations, predicting ambiguous future experiences
+
+% gap between "classic" free recall tasks and naturalistic (or real-world) memory tasks
+
+% feature-rich free recall, basic manipulation conditions, preview of findings
+
+\section*{Results}
 
@@
 
+% figure: clustering effects
+
+% figure: recall initiation
+
+% figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
+
+% figure: fingerprint trajectories
+
+% figure: carryover effects: clustering early vs. clustering late
+
+% figure: clustering carryover vs. accuracy difference
+
+% figure: clustering carryover vs. temporal clustering differences
+
+\section*{Discussion}
+
+% recap
+
+% connections to prior work: context effects, priming, situation models
+
+% implications for adaptive learning, education and training
+
+\section*{Materials and methods}
+\subsection*{Participants}
+
+\subsection*{Experimental design}
+
+\subsection*{Analysis}
+
 \bibliographystyle{apa}

### 2022-12-23 paper/main.tex
backing up initial attempts at introduction

@@
 % the role of context and prior experience in memory
+Experience is subjective: different people who encounter identical physical experiences
+can take away very different meanings and memories.  One reason is that our subjective
+experiences in the moment are shaped in part the idiosyncratic prior experiences, memories,
+goals, thoughts, expectations, and emotions that we bring with us into the present moment.
+These factors collectively define a \textit{context} for our experiences~\citep{Mann20}.
 
 % situation models: forming expectations, predicting ambiguous future experiences
+The contexts we encounter help us to construct \textit{situation models}~\citep{RangRitc12, MannEtal15}
+or \textit{schemas}~\citep{MasiEtal22, BaldEtal18} that describe how experiences are likely to unfold
+based on our prior experiences with similar contextual cues.  For example, when we enter a sit-down restaurant,
+we might expect to be seated at a table, given a menu, and served food.  Priming someone
+to expect a particular situation or context can also influence how they resolve potentail ambiguities
+in their ongoing experiences, including ambiguous movies and narratives~\citep{YeshEtal17}.
 
@@
 \bibliographystyle{apa}
-\bibliography{cdl}
+\bibliography{CDL-bibliography/cdl}
 \end{document}

### 2022-12-23 paper/boneyard.tex
added some potentially useful text from a former grant application

% text copied from early NSF grant draft (unused in final grant)

Memory researchers, like cognitive neuropsychologists in many domains, have historically taken a ``less is more'' approach.  They follow the logic that, given a memory topic of interest (e.g.\ memory for real-world autobiographical events), we can ask: what is the simplest experimental paradigm that we might reasonably expect to involve a shared or analogous set of behavioral tendencies and brain structures?  There are many benefits to taking a simplified approach to studying complex cognitive phenomena.  By composing a set of experimental circumstances where every aspect of the stimuli and timing is carefully controlled by the experimentalist, and where the behaviors a participant can exhibit are highly constrained, it becomes possible to systematically link specific experimental observations to their corresponding neural and behavioral mechanisms~\citep{OReiMuna00}.  However, there is also a major cost to simplifying: tightly controlled experiences designed in the lab can often only provide indirect insights into the real-world analogs of those experiences.  This has led to a major hole in the memory theory literature.  On the one hand, the most well-developed theories of human memory have been informed by a long history of tightly controlled (but simplified relative to the real-world) experiments~\citep[e.g.\ list-learning, priming, reaction times; ][]{Kaha12}.  On the other hand, studies of real-world classroom learning have largely avoided drawing direct links with those theories, instead focusing on questions about applied topics such as creating effective assessment tools, increasing student engagement and motivation, and understanding what makes an effective teacher~\citep{Seid07}.

The general approach we take in our project is to use prior experimental and theoretical work on list learning as a foundation for optimizing STEM learning.  Specifically, \textbf{Module 1 is centered around developing approaches for improving participants' memories for lists of words-- a simple and well-studied task expected to involve neural mechanisms similar to those underlying STEM learning}.  To foreshadow our proposal, \textbf{in Module 2, we have designed a set of analogous experiments and analyses that adapt the approach in Module 1 to a more naturalistic learning setting (learning from online STEM training videos)}.  By mirroring our approaches in Modules 1 and 2, we hope to be able to draw parallels between the two sets of experiments and findings.  Finally, \textbf{in Module 3, we plan to replicate the EEG experiments in Modules 1 and 2 using a low-cost EEG system that our team developed.  We will also develop (and document) an open-source toolkit for sharing our software (and data).  In this way, Module 3 will serve to help bring our work into real-world classrooms, and to real-world students}.

\subsection*{Context and the structure of memory}
Remembering words on a studied list requires distinguishing the current list from the rest of one's experience.  To model this fundamental memory capability, cognitive scientists have posited the existence of a special representation, called \emph{context}, that is associated with each list.  According to early theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are composed of many features which fluctuate from moment to moment, slowly drifting through a multidimensional feature space.  During recall, this representation forms part of the retrieval cue, enabling us to distinguish list items from non-list items.  Understanding the role of context in memory processes is particularly important in self-cued memory tasks, such as \textit{free recall}, where the retrieval cue is ``context'' itself.  In free recall, participants study lists of items and are instructed to recall the items in any order they choose.

\begin{wrapfigure}[10]{r}{1.3in}
  \begin{center}
    \vspace{-33pt}
    \includegraphics[width=1.1in]{figs/fingerprint_schematic.pdf}
  \end{center}
  \vspace{-18pt}
  \caption{\footnotesize \textbf{Memory fingerprint.}  Reflects a participant's tendency to cluster recalls according to each stimulus feature dimension.}
  \label{fig:fingerprint}
\end{wrapfigure}

Over the past half-century, context-based models have enjoyed impressive success at explaining many stereotyped behaviors observed during free recall and other list-learning tasks \citep[e.g.][]{Este55a, RaaiShif80, GlenEtal83, HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, PolyEtalTulv, SedeEtal08, PolyEtal09, ShanHowa12}.  These phenomena include the well-known recency and primacy effects (superior recall of items from the end and, to a lesser extent, from the beginning of the study list), as well as semantic and temporal clustering effects \citep[for review see][]{KahaEtal08}.  The contiguity effect is an example of temporal clustering, which is perhaps the dominant form of organization in free recall.  This effect can be seen in the tendency for people to successively recall items that occupied neighboring positions in the study list.  For example, if a list contained the sub-sequence \textsc{``absence hollow pupil''} and the participant recalls the word \textsc{``hollow''}, it is far more likely that the next response will be either \textsc{``pupil''} or \textsc{``absence''} than some other list item~\citep{Kaha96}.  In addition, there is a strong forward bias in the contiguity effect: subjects make forward transitions (i.e., \textsc{``hollow''} followed by \textsc{``pupil''}) about twice as often as they make backward transitions, despite an overall tendency to begin recall at the end of the list.  There are also striking effects of semantic clustering \citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52, MannKaha12}, whereby the recall of a given item is more likely to be followed by recall of a similar or related item than a dissimilar or unrelated one.

In general, people organize memories for words along a wide variety of stimulus dimensions.  As captured by models like the \textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each word (e.g.\ the word's meaning, font size, font color, location on the screen, size of the object the word represents, etc.) are incorporated into the participant's mental context representation~\citep[also see][]{SmitVela01, MannEtal15}.  During a memory test, any of these features may serve as a memory cue, which in turn leads the participant to recall in succession words that share stimulus features.  We refer to a participant's particular tendency to organize their recalls according to each stimulus dimension as their \textit{memory fingerprint} (Fig.~\ref{fig:fingerprint}).  Formally, the memory fingerprint is a vector whose elements range between 0 and 1, where each element of the memory fingerprint reflects a stimulus feature.  A value close to 1 indicates that the participant nearly always orders their recalls according to the given stimulus dimension.  A value close to 0 indicates that the participant almost \textit{never} orders their recalls according to the given stimulus dimension; a value close to 0.5 indicates that the participant's recalls are random with respect to that stimulus dimension.  In Experiment 1 we propose a free recall variant where each stimulus has many well-defined and quantifiable features, enabling us to track how the ways people organize their memories (i.e., their memory fingerprints) change over time.  In Experiment 2 we use a behavior-driven approach to manipulate these processes, and in Experiment 3 we use electroencephalographic (EEG) data to manipulate these processes.  In this way, Experiment 1 is focused on elucidating the dynamics of this very specific form of learning, and Experiments 2 and 3 are focused on manipulating the learning process, with the goal of improving (or diminishing) learning efficiency.

\subsection*{The feature-rich free recall paradigm}

\begin{wrapfigure}[10]{r}{1.4in}
\vspace{-57pt}
  \begin{center}
    \includegraphics[width=1.2in]{figs/FRFR}
  \end{center}
\vspace{-16pt}
  \caption{\footnotesize \textbf{Feature-rich free recall.}  After studying lists comprised of words that vary along several feature dimensions, participants verbally recall words in any order (microphone icon).}
  \label{fig:exp1}
\end{wrapfigure}

Each of our Module 1 experiments is based on a variant of the classic free recall paradigm that we term \textit{feature-rich free recall}.   In feature-rich free recall, participants study 16 lists, each comprised of 16 words that vary along a number of stimulus dimensions: semantic category, object size, text color, text location, word length, and starting letter (Fig.~\ref{fig:exp1}). Each list contains four words from each of four different semantic categories; all other stimulus features are randomized.  Each word is presented for 2 seconds, followed by a 2 second pause (blank screen).  After studying each list, the participant has 60 seconds to recall as many words as they can from that list, in any order they choose.  Because each individual word is associated with several well-defined (and quantifiable) features, and because each list incorporates a diverse mix of feature values along each dimension, this allows us to evaluate participants' memory fingerprints in rich detail.

Our experimental paradigm incorporates the Google Cloud Speech API speech-to-text engine to automatically transcribe participants' verbal recalls into text.  This allows recalls to be transcribed in real time-- a distinguishing feature of the experiment that is critical for using participants' recalls of prior lists to adapt the presentation orders of future lists, as in Experiments 2 and 3.  (In typical verbal recall experiments the audio data must be parsed manually, precluding this sort of design.)  We recently published a study verifying that the computer-generated recall transcripts are sufficiently similar to human-generated transcripts to yield reliable data~\citep{ZimaEtal18}.

\subsection*{Experiment 1: tracking the dynamics of memory organization over successive learning trials}
% maybe this could benefit from reframing wrt to the variance idea?
Large portions of the memory theory and education literatures have focused on identifying learners' unique learning styles.  In the memory theory literature~\citep{SmitVela01, Kaha12, HealKaha14a, MannEtal15}, the focus has been on identifying behavioral signatures of how memories are organized (i.e., in the language of this proposal, identifying memory fingerprints).  In the education literature, the focus has been been on asking whether catering lessons to an individual's unique learning style can help them to learn more efficiently~\citep{PashEtal08}.  Each of these literatures effectively treats people's learning styles as a fixed point.  \textbf{In other words, the assumption underlying the majority of these studies is that each individual has their own unique learning style, and that learning style remains roughly consistent over time.  Experiment 1 is designed to explicitly test this assumption using the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}).}  Importantly, we will collect data from two populations: (a) members of the Dartmouth and surrounding community, with the experiment running in a tightly controlled laboratory setting, and (b) Mechanical Turk workers, via the Internet.  Taken together, we hope that these two populations of participants will allow us to generalize our findings to a diverse set of learners.

\paragraph{Design and procedure.}
We will use the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}) to study how people's memory fingerprints are influenced by presentation order.  The experiment will be divided into six manipulation conditions (one for each stimulus feature); in each of these manipulation conditions the words from the first 8 lists participants study will be sorted according to one of the stimulus dimensions.  For example, in the ``color'' condition, words from the first 8 lists will be sorted by font color; in the ``starting letter'' condition, words from the first 8 lists will be sorted alphabetically; etc.  The second 8 lists will be presented in a random order in every condition.  As described below, we will assess how participants' memories are affected by these order manipulations, both for their memories of words on first 8 lists where the manipulations are applied, as well as on the second 8 lists.  Participants' behaviors will be compared to a seventh (control) condition, where the presentation order is randomized for all 16 lists.

We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our prior work to estimate the relevant parameters~\citep{MannEtal11, MannEtal12}.  The analysis indicated that we will require 30 in-laboratory participants in each of the six manipulation conditions to achieve sufficient power.  Because all of our findings will be compared relative to the random control condition, we plan to run 60 in-laboratory participants in that condition.  Prior work suggests that Mechanical Turk data is more variable than data collected in the laboratory~\citep{CrumEtal13}, so in each condition we plan to collect data from twice as many Mechanical Turk workers as in-laboratory participants.  In all, across the 7 experimental conditions, we plan to run 720 participants in Experiment 1 (240 in the lab and 480 online through Mechanical Turk).

\paragraph{Key analyses.}
Experiment 1 is motivated by two fundamental questions about the dynamics of memory organization and recall.  \textbf{First, we want to understand how memory fingerprints vary across people and over time.}  For example, are people's memory fingerprints generally stable across lists?  Or, what are the similarities and differences across people (in how people organize their memories, and how their memory fingerprints change over time)?  Are those individual differences predictive of memory performance? \textbf{Second, we want to understand how people's memory fingerprints can be manipulated.}  Specifically, if lists are sorted according to a particular feature (e.g. the words' semantic categories), how does that influence people's tendencies to cluster their recalls of those words with respect to that feature?  And which of those tendencies persist during recalls of later (randomly ordered) lists?  For example, if we organize early lists alphabetically, are  participants more likely recall the words from those lists alphabetically as well?  More generally, might there be some way of influencing how people organize their memories, even after the influencing cue is no longer present?

% figures: pilot data displaying random-condition fingerprints, relationship between fingerprint variance and memory
\begin{wrapfigure}[10]{R}{4.0in}
  \vspace{-52pt}
  \begin{center}
    \includegraphics[width=4.0in]{figs/random_pilot.pdf}
  \end{center}
\vspace{-18pt}
  \caption{\footnotesize \textbf{Preliminary data for feature-rich free recall.}  \textbf{a.} Clustering scores along each feature dimensions (i.e. memory fingerprints).  Each dot represents the average clustering score along one dimension, for one participant. \textbf{b.} Across-subject correlation between fingerprint variability and memory performance.  \textbf{c.} Across-condition correlation between fingerprint variability and memory performance.}
  \label{fig:random_pilot}
\end{wrapfigure}

\paragraph{Experiment 1 preliminary data.}
We ran a preliminary study ($n = 172$) to measure participants' memory fingerprints under the in-laboratory conditions outlined in Experiment 1.  Using data from the control condition (where all lists were ordered randomly; $n = 45$), we computed the distribution of average memory fingerprints (across lists) from each participant (Fig.~\ref{fig:random_pilot}a).  The degree to which participants organize their memories with respect to each feature varies substantially across people (e.g.\ there is a wide spread of values covered across participants within each feature dimension).  In addition, some features are more likely to be leveraged in people's memories than others.  For example, participants are more likely to organize their recalls by category, size, and temporal proximity than by font color or word length.  We also examined how stable participants' memory fingerprints were across lists.  We found that participants whose memory fingerprints were more stable across lists (i.e., lower variance) remembered more words (Fig.~\ref{fig:random_pilot}b).  We also examined the manipulation conditions, where the first 8 lists participants studied were sorted according to one feature (category: $n=21$, color: $n=20$, screen location: $n=21$, first letter: $n=20$, word length: $n=23$, size: $n=22$).  Relative to randomly sorted lists, sorting by some features increased the average memory fingerprint variability and sorting by other features decreased the average memory fingerprint variability.  (We used a permutation-based procedure to compute the variability, which controlled for the distributions of feature values of the words participants remembered.)  We found that participants remembered more words in the conditions where the presentation orders led to lower fingerprint variability (Fig.~\ref{fig:random_pilot}c).  Taken together, these preliminary data suggest that participants who naturally exhibit more stable fingerprints remember more words, and fingerprint stability may be influenced by the order in which words are presented.  In turn, sorting the word presentations along different dimensions can enhance (or diminish) memory performance according to whether fingerprint variability is increased or decreased, respectively.

\subsection*{Experiment 2: adaptive list learning using (behavioral) memory fingerprints}
\begin{wrapfigure}[9]{R}{1.6in}
  \vspace{-25pt}
  \begin{center}
    \includegraphics[width=1.5in]{figs/adaptive_fr_flowchart.pdf}
  \end{center}
\vspace{-20pt}
  \caption{\footnotesize \textbf{AdaptiveFR.}  The average memory fingerprint is updated after recalling each list, and is used to determine the presentation order of the next list.}
  \label{fig:adaptiveFR_flowchart}
\end{wrapfigure}

Our pilot data from Experiment 1 suggest that the order in which words are studied affects how they are later recalled.  Further, certain orderings can lead to lower memory fingerprint variability, which in turn leads (on average) to better memory for the studied words.  Whereas Experiment 1 explores the memory effects of pre-defined order manipulations, in Experiment 2 we propose to adapt the presentation order of future lists, using behavior measured on previous lists.  \textbf{Our overarching goal is to explore whether memory models that are custom-fit to individual participants may be exploited (through presentation order manipulations) to enhance or diminish memory performance.}

\begin{wrapfigure}[24]{R}{2in}
  \vspace{-25pt}
  \begin{center}
    \includegraphics[width=2in]{figs/adaptive_pilot}
  \end{center}
\vspace{-20pt}
  \caption{\footnotesize \textbf{AdaptiveFR pilot data. a.} Fingerprint variability decreases when lists are sorted according to the average memory fingerprints from prior lists (stabilize condition).  \textbf{b.}  Stabilizing memory fingerprints improves recall specifically for middle-of-the-list words.}
  \label{fig:adaptive_pilot}
\end{wrapfigure}

\paragraph{Design and procedure.} We will use a modified version of the feature-rich free recall paradigm designed to enhance or diminish participants' memory performance by adaptively adjusting the presentation order of future lists using recall data from previous lists (\textit{AdaptiveFR}; Fig.~\ref{fig:adaptiveFR_flowchart}).  As in Experiment 1, participants in Experiment 2 will study 16 lists.  The first 4 lists will be sorted randomly; these are used to compute the participants' individualized memory fingerprints.  Each of the remaining 12 lists (grouped into three blocks of 4 lists) will be sorted according to one of three algorithms (every participant will study and recall lists sorted using all three algorithms, but in a block-randomized order).   In the \textit{stabilize} condition, the list words will be sorted into an order that matches the participant's average memory fingerprint from all prior lists.  For example, if the memory fingerprint indicates a strong tendency to cluster recalls by category, words from the same category will tend to appear at adjacent study positions on ``stabilize'' lists.  To the extent to which the presentation order influences recall order, the stabilize condition should reduce variability in participants' memory fingerprints (by influencing future memory fingerprints to match the current average fingerprint). In the \textit{destabilize} condition, words will be sorted in an attempt to maximize the fingerprint variability.  Specifically, we will generate 10,000 random permutations of each to-be-presented list, compute a memory fingerprint for each (treating each permutation as if it were a recall sequence), and select the permutation whose fingerprint is most different (in terms of Mahalanobis distance) from the average fingerprint from prior lists.  Finally, in the \textit{random} condition, words will be shuffled randomly; this condition will be used as a baseline for the others.  We also plan to run series of analogous yoked control conditions in a second set of participants, whereby memory fingerprints based on \textit{other} participants' recalls are used to sort future lists.

We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our pilot data as a guide (Fig.~\ref{fig:adaptive_pilot}).  The analysis indicates that we will require at least 33 in-laboratory participants to achieve sufficient power, which we plan to round up to 40 participants.  Similarly (following our logic from Experiment 1) we plan to run 80 Mechanical Turk workers in the online version of Experiment 2.  We will also duplicate the experiment using yoked controls, for a total of 80 in-laboratory and 160 Mechanical Turk participants.

\paragraph{Key analyses.}  The primary questions driving Experiment 2 concern how list order manipulations can influence memory.  \textbf{First, we want to understand whether we can adaptively enhance or diminish memory performance by increasing or decreasing the stability of people's memory fingerprints.}  Essentially, Experiment 2 is intended to serve as a strong test of our hypothesized causal link between the stability of people's memory fingerprints and their memory performance.  We will test whether our experimental manipulations can indeed influence fingerprint stability, and if so, whether those manipulations also impact memory performance.  \textbf{Second, we want to understand the extent to which these person-specific reordering manipulations are beneficial, relative to population-driven reordering manipulations.}  For example, if we customize the word presentation orders based on an individual's data, do we see stronger effects on their performance than if we had instead customized the lists based on the average data from other participants?

\paragraph{Pilot data.}  We ran a pilot study ($n = 12$) to test the feasibility of our approach.  We found that when word lists were sorted to be consistent with a participant's memory fingerprint (the stabilize condition), this led to lower fingerprint variability ($t(11) = 1.94, p = 0.06$; Fig.~\ref{fig:adaptive_pilot}a) and enhanced memory (Fig.~\ref{fig:adaptive_pilot}b).  Notably, this memory enhancement was specific to words from the middle of the list (positions 8--12; $t(11)=3.18, p=0.008$). This provides preliminary evidence that manipulations that increase the stability of participants' recall tendencies also improve memory.

\subsection*{Experiment 3: adaptive list learning using (neural) memory fingerprints}
The above experiments leverage participants' \textit{behaviors} (i.e., the sequences of recalls they make) to optimize learning by changing the order in which subsequent lists are presented.  In Experiment 3, we seek to optimize learning using \textit{neural} (EEG) data. \textbf{Our prior work shows that the neural patterns recorded while participants study a list of words may be used to predict their individual tendencies to cluster their recalls according to presentation order (temporal clustering; Fig.~\ref{fig:clust}a) or word meaning (semantic clustering; Fig.~\ref{fig:clust}b).}

% Logically, there are two sets of factors that can contribute to recall order.  The first set of factors are processes that occur at the time information is encoded.  The results summarized in Figure~\ref{fig:clust} show that some information about the order in which words on a studied list will later be recalled may be predicted from neural recordings taken at the time of the words are studied (encoded).  The second set of factors that can contribute to recall order occur at retrieval.  Any additional processes or stimuli that occur as the participant recalls the words could conceivably affect recall order.  For example, a sudden pang of hunger or boredom that happens to occur during the recall phase of the experiment might lead the participant to recall other words associated with those new thoughts.  The neural patterns recorded as the participant studies a list of words cannot be used to recover order effects driven by retrieval-specific processes.

\begin{wrapfigure}[12]{R}{2.5in}
\vspace{-32pt}
  \begin{center}
    \includegraphics[width=2.5in]{figs/neural_predictors_of_clustering.pdf}
  \end{center}
\vspace{-19pt}
  \caption{\footnotesize \textbf{Pilot data: neural signatures of memory fingerprints.}  Electrophysiological data recorded during study may be used to predict the degree to which a participant will later (\textbf{a.}) temporally cluster their recalls; panel adapted from~\citep{MannEtal11} or (\textbf{b.}) semantically clustering their recalls; panel adapted from~\citep{MannEtal12}.}
  \label{fig:clust}
\end{wrapfigure}

In Experiment 3, we ask whether these prior findings extend to additional dimensions of the memory fingerprint (i.e., how participants order their recalls of words).  Following from our prior work, we will record EEG activity (wavelet-derived mean spectral power in 50 log-spaced frequency bands, from each electrode) as each word is studied~\citep{MannEtal11, MannEtal12}.  We will also examine patterns of cross-frequency coupling and other dynamic connectivity patterns~\citep{MannEtal14b}.  We will then identify a neural predictor for each dimension of the memory fingerprint-- i.e., one neurally derived predictor each of how likely the word being studied will follow a word matching the same semantic category, or with a similar color, location, starting letter, etc.  Specifically, to obtain each predictor (e.g.\ of clustering recalls by color), we will consider the set of correctly recalled words (excluding the first).  For each recall in turn, we will ask: considering the dimension of interest (e.g.\ font color), how similar is each not-yet-recalled word to the previous recall?  We will use the distribution of similarity values to tag each correct recall with a percentile rank.  Regressing the neural activity recorded just before each word was studied against these percentile ranks will yield a neural predictor (classifier) of clustering by the given memory fingerprint dimension.  (This process must be repeated for each memory fingerprint dimension in turn.)  These neural predictors will be updated after each recall interval, and will be used during the study period of the next list to select each word to be presented next according to the ongoing neural recordings (Fig.~\ref{fig:adaptiveFR_eeg_order}).

\begin{wrapfigure}[11]{R}{1.7in}
\vspace{-46pt}
  \begin{center}
    \includegraphics[width=1.5in]{figs/adaptive_fr_EEG_flowchart.pdf}
  \end{center}
\vspace{-23pt}
  \caption{\footnotesize \textbf{AdaptiveFR with EEG.}  EEG recordings during study are used to update the memory fingerprint classifier after each batch of recalls.  The classifier is used to adapt the presentation order on a word-by-word basis when the next list is studied.}
  \label{fig:adaptiveFR_eeg_order}
\end{wrapfigure}

\paragraph{Design and procedure.} The experiment and parameters will be analogous to Experiment 2, except that we will use EEG data recorded as participants are studying the words (rather than behavioral data from participants' recalls) to estimate their memory fingerprints.  We plan to record the EEG data using an ActiChamp 160-electrode system (with a photosensor-based electrode mapping system, to allow us to perform source-localization analyses).  Participants will study a total of 16 lists of 16 words each; we will adaptively adjust the presentation orders of each list as described next.

The first four lists will always be sorted randomly.  We will use EEG recordings as participants study these lists to train classifiers to detect the neural signatures of clustering along each memory fingerprint dimension.  Note that, whereas in Experiment 2 the presentation order of the full list is determined before the first word is presented, in Experiment 3 we will be adaptively adjusting the presentation order on a word-by-word basis using classifiers applied to the ongoing EEG recordings.  In other words, we will use the EEG activity recorded during the inter-stimulus interval to predict how the next word will be organized in memory.  Then, using the just-presented word as a reference, we will select a word that is either \textit{consistent}, \textit{inconsistent}, or \textit{random} with respect to that predicted memory fingerprint (Fig.~\ref{fig:adaptiveFR_eeg_order}).  For example, in the \textit{consistent} condition, if the classifier predicts that the next word is likely to be organized by category and font color, then the next to-be-presented word will be chosen to match the category and font color of the just-presented word.  More generally, we will rank all yet-to-be-presented words according to how similar they are along each stimulus dimension (to the just-presented word).  We will then select a matching word by weighting these similarity rankings (using the neurally predicted memory fingerprint to define the weights).  In the \textit{inconsistent} condition, we will next present the word that \textit{least} matches the previous word, using the decoded fingerprint.  In this way, the neurally driven \textit{consistent} and \textit{inconsistent} conditions of Experiment 3 are analogous to the behaviorally driven \textit{stabilize} and \textit{destabilize} conditions of Experiment 2.  We also plan to run yoked control conditions, whereby the presentation orders participants exhibit are driven by \textit{other} participants' neural data.

Using our prior work (Fig.~\ref{fig:clust}) to approximate the relevant parameters, we ran a power analysis (power: 0.8, Type 1 error rate: 0.05) to estimate the number of participants we will need to achieve sufficient power.  The analysis indicates that we will require at least 42 participants in each experimental condition; we plan to run 50 participants using the neurally driven adaptive design described above, plus an additional 50 in the yoked control condition.

\paragraph{Key analyses.}  The overarching goal of Experiment 3 is to test the limits of how the fine-scale temporal dynamics of how people organize incoming information may be leveraged to enhance (or diminish) memory.  The temporal resolution of adaptive learning approaches is limited by the temporal resolution of the measurements on which the adaptive aspects are based.  Whereas recall behaviors may only be assessed periodically (e.g.\ in Experiment 2, during each recall period-- after the participant studies the list), the neural activity driving those behaviors may be measured and analyzed many times per second.  In this way, we seek to use Experiment 3 to answer deep questions about the temporal dynamics of how people organize their memories.  For example, do the same principles about fingerprint stability leading to enhanced memory that we discovered in our pilot studies (Figs.~\ref{fig:random_pilot}, \ref{fig:adaptive_pilot}) also apply to the moment-by-moment fluctuations we will estimate in Experiment 3?  And do people exhibit similar dynamics in those fluctuations within (or across) lists?  Or are those dynamics predictive of recall performance?  In addition to examining the neural estimates of memory fingerprint dynamics, we will also measure the improvement (or detriment) in memory performance across the different experimental conditions (analogous to our proposed analyses in Experiment 2).

Our second set of questions is more general, and is intended to parallel several ideas in Module 2.  Specifically, can we use neural data to uncover (and leverage) the fine-scale temporal dynamics of learning and memory?  In Module 1 (Experiment 3) our focus is on uncovering (and leveraging) the fine-scale temporal dynamics of list learning.  In Module 2 our focus is on uncovering (and leveraging) the fine-scale temporal dynamics of STEM learning.

### 2022-12-23 paper/supplement.tex
title tweak

@@
 
-\title{\textit{Supplementary materials for}: Carryover clustering effects in free recall reveal how prior experiences influence memory for new experiences}
+\title{\textit{Supplementary materials for}: Carryover effects in free recall reveal how prior experiences influence memories of new experiences}
 \author{Jeremy R.
@@
 Fitzpatrick\textsuperscript{1}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili Interactive\\\textsuperscript{3}Princeton University\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
-
-\bibliographystyle{apa}
+\date{}
 
@@
 \bibliographystyle{apa}
-\bibliography{cdl}
+\bibliography{CDL-bibliography/cdl}
 

### 2022-12-26 paper/main.tex
added note to intro about potential framing adjustments...
title tweak
working on introduction (backing up progress)

@@
 
-\title{Carryover effects in free recall reveal how prior experiences influence memory for new experiences}
+\title{Carryover effects in free recall reveal how prior experiences influence memories of new experiences}
 \author{Jeremy R.
@@
 
+%NOTE: should we start with the naturalistic framing?  or would it be more effective and direct to
+% jump right into free recall?  i've added some potentially useful text to boneyard.tex that could be
+% pasted in with some tweaking.  some of the above text could also be moved to the discussion section.
+
 % gap between "classic" free recall tasks and naturalistic (or real-world) memory tasks
+Our understanding of how we form situation models and schemas, and how they
+interact with our subjective experiences and memories, is constrained in
+part by substantial differences in how we study these processes.  Situation models
+and schemas are most often studied using ``naturalistic'' 
+stimuli such as narratives and movies~\citep{CITES}.  In contrast, our
+understanding of how we organize our memories has been most widely studied
+using more traditional paradigms like free recall of random word lists~\citep{Kaha12}.
+Because random word lists are unstructured by design, it is not clear if or how non-trivial
+situation models might apply to these stimuli.
 

### 2022-12-27 paper/main.tex
backing up progress on intro
backing up progress to intro

@@
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 \usepackage[utf8]{inputenc}
@@
 
-\begin{abstract}
-We perceive, interpet, and remember ongoing experiences through the lens of our prior experiences.
-Inferring that we are one type of situation versus another can lead us to interpret the same physical
-experience differently.  In turn, this can affect how we focus our attention, form expectations of what will happen next, 
-remember what is happening now, draw on our prior related experiences, and so on.  To study these phenomena,
-we asked participants to perform simple word list learning tasks.  Across different experimental conditions, we held the set of
-to-be-learned words constant, but we manipulated the orders in which the words were studied.  We found that 
-these order manipulations affected not only how the participants recalled the ordered lists, but also how they recalled later randomly
-ordered lists.  Our work shows how structure in our ongoing experiences can exert influence on how we remember unrelated
-subsequent experiences.
-\end{abstract}
+\begin{abstract} We perceive, interpet, and remember ongoing experiences
+through the lens of our prior experiences. Inferring that we are one type of
+situation versus another can lead us to interpret the same physical experience
+differently. In turn, this can affect how we focus our attention, form
+expectations of what will happen next, remember what is happening now, draw on
+our prior related experiences, and so on. To study these phenomena, we asked
+participants to perform simple word list learning tasks. Across different
+experimental conditions, we held the set of to-be-learned words constant, but
+we manipulated the orders in which the words were studied. We found that these
+order manipulations affected not only how the participants recalled the ordered
+lists, but also how they recalled later randomly ordered lists. Our work shows
+how structure in our ongoing experiences can exert influence on how we remember
+unrelated subsequent experiences. \end{abstract}
 
@@
 % the role of context and prior experience in memory
-Experience is subjective: different people who encounter identical physical experiences
-can take away very different meanings and memories.  One reason is that our subjective
-experiences in the moment are shaped in part the idiosyncratic prior experiences, memories,
-goals, thoughts, expectations, and emotions that we bring with us into the present moment.
-These factors collectively define a \textit{context} for our experiences~\citep{Mann20}.
 
-% situation models: forming expectations, predicting ambiguous future experiences
-The contexts we encounter help us to construct \textit{situation models}~\citep{RangRitc12, MannEtal15}
-or \textit{schemas}~\citep{MasiEtal22, BaldEtal18} that describe how experiences are likely to unfold
-based on our prior experiences with similar contextual cues.  For example, when we enter a sit-down restaurant,
-we might expect to be seated at a table, given a menu, and served food.  Priming someone
-to expect a particular situation or context can also influence how they resolve potentail ambiguities
-in their ongoing experiences, including ambiguous movies and narratives~\citep{YeshEtal17}.
+Experience is subjective: different people who encounter identical physical
+experiences can take away very different meanings and memories. One reason is
+that our subjective experiences in the moment are shaped in part the
+idiosyncratic prior experiences, memories, goals, thoughts, expectations, and
+emotions that we bring with us into the present moment. These factors
+collectively define a \textit{context} for our experiences~\citep{Mann20}. %
+situation models: forming expectations, predicting ambiguous future experiences
+The contexts we encounter help us to construct \textit{situation
+models}~\citep{RangRitc12, MannEtal15} or \textit{schemas}~\citep{MasiEtal22,
+BaldEtal18} that describe how experiences are likely to unfold based on our
+prior experiences with similar contextual cues. For example, when we enter a
+sit-down restaurant, we might expect to be seated at a table, given a menu, and
+served food. Priming someone to expect a particular situation or context can
+also influence how they resolve potentail ambiguities in their ongoing
+experiences, including ambiguous movies and narratives~\citep{YeshEtal17}.
 
-%NOTE: should we start with the naturalistic framing?  or would it be more effective and direct to
-% jump right into free recall?  i've added some potentially useful text to boneyard.tex that could be
-% pasted in with some tweaking.  some of the above text could also be moved to the discussion section.
+% gap between "classic" free recall tasks and naturalistic (or real-world)
+% memory tasks
 
-% gap between "classic" free recall tasks and naturalistic (or real-world) memory tasks
 Our understanding of how we form situation models and schemas, and how they
-interact with our subjective experiences and memories, is constrained in
-part by substantial differences in how we study these processes.  Situation models
-and schemas are most often studied using ``naturalistic'' 
-stimuli such as narratives and movies~\citep{CITES}.  In contrast, our
-understanding of how we organize our memories has been most widely studied
-using more traditional paradigms like free recall of random word lists~\citep{Kaha12}.
-Because random word lists are unstructured by design, it is not clear if or how non-trivial
-situation models might apply to these stimuli.
+interact with our subjective experiences and memories, is constrained in part
+by substantial differences in how we study these processes. Situation models
+and schemas are most often studied using ``naturalistic'' stimuli such as
+narratives and movies~\citep{ZwaaEtal95,ZwaaRadv98, NastEtal20}. In contrast,
+our understanding of how we organize our memories has been most widely studied
+using more traditional paradigms like free recall of random word
+lists~\citep{Kaha12}. In free recall, participants study lists of items and are
+instructed to recall the items in any order they choose. The orders in which
+words come to mind can provide insights into how participants have organized
+their memories of the studied words. Because random word lists are unstructured
+by design, it is not clear if or how non-trivial situation models might apply
+to these stimuli. Nevertheless, there are \textit{some} commonalities between
+memory for word lists and memory for real-world experiences.
+
+Like remembering real-world experiences, remembering words on a studied list
+requires distinguishing the current list from the rest of one's experience. To
+model this fundamental memory capability, cognitive scientists have posited the
+existence of a special representation, called \emph{context}, that is
+associated with each list. According to early
+theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are
+composed of many features which fluctuate from moment to moment, slowly
+drifting through a multidimensional feature space. During recall, this
+representation forms part of the retrieval cue, enabling us to distinguish list
+items from non-list items. Understanding the role of context in memory
+processes is particularly important in self-cued memory tasks, such as
+\textit{free recall}, where the retrieval cue is ``context'' itself.
+
+Over the past half-century, context-based models have enjoyed impressive
+success at explaining many stereotyped behaviors observed during free recall
+and other list-learning tasks~\citep{Este55a, RaaiShif80, GlenEtal83,
+HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, PolyEtalTulv, SedeEtal08,
+PolyEtal09, ShanHowa12}. These phenomena include the well-known recency and
+primacy effects (superior recall of items from the end and, to a lesser extent,
+from the beginning of the study list), as well as semantic and temporal
+clustering effects~\citep{KahaEtal08}. The contiguity effect is an example of
+temporal clustering, which is perhaps the dominant form of organization in free
+recall. This effect can be seen in the tendency for people to successively
+recall items that occupied neighboring positions in the study list. For
+example, if a list contained the sub-sequence \textsc{``absence hollow pupil''}
+and the participant recalls the word \textsc{``hollow''}, it is far more likely
+that the next response will be either \textsc{``pupil''} or
+\textsc{``absence''} than some other list item~\citep{Kaha96}. In addition,
+there is a strong forward bias in the contiguity effect: subjects make forward
+transitions (i.e., \textsc{``hollow''} followed by \textsc{``pupil''}) about
+twice as often as they make backward transitions, despite an overall tendency
+to begin recall at the end of the list. There are also striking effects of
+semantic clustering~\citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52,
+MannKaha12}, whereby the recall of a given item is more likely to be followed
+by recall of a similar or related item than a dissimilar or unrelated one.
+
+In general, people organize memories for words along a wide variety of stimulus
+dimensions. As captured by models like the \textit{Context Maintenance and
+Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each
+word (e.g.\ the word's meaning, font size, font color, location on the screen,
+size of the object the word represents, etc.) are incorporated into the
+participant's mental context representation~\citep{SmitVela01, MannEtal15, Mann20}.
+During a memory test, any of these features may serve as a memory cue, which in
+turn leads the participant to recall in succession words that share stimulus
+features.
+
+% link clustering to schemas...
+
+These clustering behaviors in ....
 
 % feature-rich free recall, basic manipulation conditions, preview of findings
+
+\begin{figure} 
+    \begin{center}
+        \includegraphics[width=0.4 \textwidth]{figures/FRFR} \end{center} \vspace{-16pt}
+
+        \caption{\textbf{Feature-rich free recall.} After studying lists
+comprised of words that vary along several feature dimensions, participants
+verbally recall words in any order (microphone icon).}
+
+        \label{fig:exp}
+\end{figure}
 

### 2022-12-29 paper/main.tex
checking in first draft of intro

@@
 MannKaha12}, whereby the recall of a given item is more likely to be followed
-by recall of a similar or related item than a dissimilar or unrelated one.
-
-In general, people organize memories for words along a wide variety of stimulus
+by recall of a similar or related item than a dissimilar or unrelated one. In
+general, people organize memories for words along a wide variety of stimulus
 dimensions. As captured by models like the \textit{Context Maintenance and
@@
 size of the object the word represents, etc.) are incorporated into the
-participant's mental context representation~\citep{SmitVela01, MannEtal15, Mann20}.
-During a memory test, any of these features may serve as a memory cue, which in
-turn leads the participant to recall in succession words that share stimulus
-features.
+participant's mental context representation~\citep{SmitVela01, MannEtal15,
+Mann20, MannEtal11, MannEtal12}. During a memory test, any of these features
+may serve as a memory cue, which in turn leads the participant to recall in
+succession words that share stimulus features.
 
@@
 
-These clustering behaviors in ....
+A key mystery is whether the sorts of situation models and schemas that people
+use to organize their memories of real-world experiences might map onto the
+clustering effects that reflect how people organize their memories for word
+lists. On one hand, situation models and clustering effects both reflect
+statistical regularities in ongoing experience. Our memory systems exploit
+these regularities when generating inferences about the unobserved past and
+yet-to-be-experienced future~\citep{XuEtal22, SchaTurk15, RangRitc12,
+BoweEtal79, MomeEtal17}. On the other hand, the rich structure of real-world
+experiences and other naturalistic stimuli that enable people to form deep and meaningful
+situation models and schemas have no obvious analog in simple word lists.  Often lists
+in free recall studies are explicitly \textit{designed} to be devoid of exploitable
+temporal structure, for example by sorting the words in a random order~\citep{Kaha12}.
 
 % feature-rich free recall, basic manipulation conditions, preview of findings
+We designed an experimental paradigm to explore how people organize their
+memories for simple stimuli (word lists) whose temporal properties change
+across different ``situations,'' analogous to how the content of real-world
+experiences change across different real-world situations. We asked
+participants to study and freely recall a series of word lists
+(Fig.~\ref{fig:exp}). Across the different conditions in the experiment, we
+varied the lists' presentation orders in different ways across lists. The
+studied items (words) were designed to vary along three general dimensions:
+semantic (word \textit{category}, and physical \textit{size} of the referent),
+lexicographic (word \textit{length} and \textit{first letter}), and visual
+(font \textit{color} and the onscreen \textit{location} of each word). In our
+main manipulation conditions, we asked participants to study and recall eight
+lists whose items were sorted by a target feature (e.g., word category). Next,
+we asked them to study and recall an additional eight lists whose items had the
+same features, but that were sorted in a random temporal order. We were
+interested in how these order manipulations affected participants' recall
+behaviors on early (sorted) lists, as well as how order manipulations on early
+lists affected recall behaviors on later (unsorted) lists. We used a series of
+control conditions as a baseline; in these control conditions all of the lists
+were sorted randomly, but we manipulated the presence or absence of the visual
+features. Finally, in an \textit{adaptive} experimental condition we used
+participants' recall behaviors on early lists to manipulate, in real-time, the
+presentation orders of subsequent lists. In this adaptive condition, we sought
+to identify potential commonalities within and across participants in how
+people organized their memories and how those organizational tendancies affect
+overall performance.
 

### 2022-12-31 paper/main.tex
working on results section; backing up progress on main and supplemental figures

@@
 
-%\newcommand{\demo}{S1}
+\newcommand{\dynamicsRandom}{S1}
+\newcommand{\dynamicsAdaptive}{S2}
+\newcommand{\fingerprintsRandom}{S3}
+\newcommand{\fingerprintsAdaptive}{S4}
+\newcommand{\recallInit}{S5}
 
@@
 
-\begin{figure}[tp]
-    \centering
+\begin{figure}[tp] \centering
 \includegraphics[width=\textwidth]{figures/recall_dynamics}
-    \caption{\textbf{Recall dynamics in free recall.}}
+
+\caption{\textbf{Recall dynamics in feature rich free recall (order
+manipulation conditions).} \textbf{Left panels.} The probabilities of
+initiating recall with each word are plotted as a function of presentation
+position. \textbf{Middle panels.} The conditional probabilities of recalling
+each word are plotted as a function of the relative position (Lag) to the words
+recalled just-prior. \textbf{Right panels.} The overall probabilities of
+recalling each word are plotted as a function of presentation position.
+\textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
+intervals (calculated across participants). Top panels display the recall
+dynamics for early (order manipulation) lists in each condition (color). Bottom
+panels display the recall dynamics for late (randomly ordered) lists.  See
+Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the random (control) and adaptive conditions.}
+
     \label{fig:recall-dynamics}
@@
 % figure: clustering effects
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprints}
+
+    \caption{\textbf{Memory ``fingerprints.'' (order manipulation conditions).}
+    The across-participant distributions of clustering scores for each feature
+    type ($x$-coordinate) are displayed for each experimental condition
+    (color), separately for order manipulation (early, top) and randomly
+    ordered (late, bottom) lists. See
+    Figures~\fingerprintsRandom~and~\fingerprintsAdaptive~for analogous plots
+    for the random (control) and adaptive conditions.}
+        \label{fig:fingerprints}
+    \end{figure}
 
 % figure: recall initiation
+Figure~\recallInit.
 

### 2022-12-31 paper/supplement.tex
working on results section; backing up progress on main and supplemental figures

@@
 \usepackage{placeins}
+\usepackage{rotating}
 
@@
 
-\newcommand{\demo}{1}
+\maketitle
 
-\maketitle
+\begin{figure}[p] \centering
+\includegraphics[width=\textwidth]{figures/recall_dynamics_random}
+
+\caption{\textbf{Recall dynamics in feature rich free recall (random conditions).} \textbf{Left panels.} The probabilities of
+initiating recall with each word are plotted as a function of presentation
+position. \textbf{Middle panels.} The conditional probabilities of recalling
+each word are plotted as a function of the relative position (Lag) to the words
+recalled just-prior. \textbf{Right panels.} The overall probabilities of
+recalling each word are plotted as a function of presentation position.
+\textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
+intervals (calculated across participants). Top panels display the recall
+dynamics for early (order manipulation) lists in each condition (color). Bottom
+panels display the recall dynamics for late (randomly ordered) lists.}
+
+    \label{fig:recall-dynamics-random}
+\end{figure}
+
+\begin{figure}[p] \centering
+    \includegraphics[width=\textwidth]{figures/recall_dynamics_adaptive}
+
+    \caption{\textbf{Recall dynamics in feature rich free recall (adaptive conditions).} \textbf{Left panels.} The probabilities of
+    initiating recall with each word are plotted as a function of presentation
+    position. \textbf{Middle panels.} The conditional probabilities of recalling
+    each word are plotted as a function of the relative position (Lag) to the words
+    recalled just-prior. \textbf{Right panels.} The overall probabilities of
+    recalling each word are plotted as a function of presentation position.
+    \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
+    intervals (calculated across participants). Condition is denoted by color.}
+
+        \label{fig:recall-dynamics-adaptive}
+    \end{figure}
+
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprints_random}
+
+    \caption{\textbf{Memory ``fingerprints.'' (random conditions).}
+    The across-participant distributions of clustering scores for each feature
+    type ($x$-coordinate) are displayed for each experimental condition
+    (color), separately for early (top) and late (bottom) lists.}
+        \label{fig:fingerprints-random}
+\end{figure}
+
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprints_adaptive}
+
+    \caption{\textbf{Memory ``fingerprints.'' (adaptive conditions).}
+    The across-participant distributions of clustering scores for each feature
+    type ($x$-coordinate) are displayed for each experimental condition
+    (color).}
+        \label{fig:fingerprints-adaptive}
+\end{figure}
+
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprints_adaptive}
+
+    \caption{\textbf{Memory ``fingerprints.'' (adaptive conditions).}
+    The across-participant distributions of clustering scores for each feature
+    type ($x$-coordinate) are displayed for each experimental condition
+    (color), separately for order manipulation (early, top) and randomly
+    ordered (late, bottom) lists.}
+        \label{fig:fingerprints-adaptive}
+\end{figure}
+
+\begin{sidewaysfigure}
+    \includegraphics[width=\textwidth]{figures/pnr_matrices}
+
+    \caption{\textbf{Probability of $n$\textsuperscript{th} recall matrices.}  Each sub-panel displays
+    the average probability of recalling the given word (Presentation position, $x$-coordinate)
+    at the given output position ($y$-coordinate); color denotes the probability.  \textbf{A. Random conditions.}  The top rows
+    display data from early (order manipulation) lists, the middle rows display data from late (randomly ordered)
+    lists, and the bottom rows display the differences between the matrices in the top and middle rows.  Panel columns denote experimental conditions.
+    \textbf{B. Order manipulation conditions.}  The matrices are displayed in the same format as those in Panel A.
+    Panel columns are organized by feature type (semantic, lexicographic, or visual).  \textbf{C. Adaptive conditions.}
+    The sub-panels are displayed in the same formats as Panels A and B, but here the matrices
+    and contrasts (indicated in $y$-axis labels) reflect different list manipulation conditions.}
+
+    \label{fig:pnr}
+\end{sidewaysfigure}
 

### 2022-12-31 paper/main.tex
minor typo fixes

@@
 
-    \caption{\textbf{Memory ``fingerprints.'' (order manipulation conditions).}
+    \caption{\textbf{Memory ``fingerprints'' (order manipulation conditions).}
     The across-participant distributions of clustering scores for each feature

### 2022-12-31 paper/supplement.tex
minor typo fixes

@@
 
-    \caption{\textbf{Memory ``fingerprints.'' (random conditions).}
+    \caption{\textbf{Memory ``fingerprints'' (random conditions).}
     The across-participant distributions of clustering scores for each feature
@@
 
-    \caption{\textbf{Memory ``fingerprints.'' (adaptive conditions).}
+    \caption{\textbf{Memory ``fingerprints'' (adaptive conditions).}
     The across-participant distributions of clustering scores for each feature

### 2023-01-02 paper/main.tex
working on clustering figures, corrected label errors
added clustering scatterplot figure

@@
 % figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
+\begin{figure}[tp] \centering
+\includegraphics[width=\textwidth]{figures/feature_clustering_vs_accuracy_and_contiguity}
+
+\caption{\textbf{Interactions between feature clustering, recall probability,
+and contiguity.} \textbf{A.} Recall probability versus feature clustering
+scores for order manipulation (early) lists. \textbf{B.} Recall probability
+versus feature clustering for randomly ordered (late) lists. \textbf{C.} Recall
+probability on late lists versus feature clustering on early lists. \textbf{D.}
+Recall probability on early lists versus feature clustering on late lists.
+\textbf{E.} Temporal clustering scores (contiguity) versus feature clustering
+scores on early lists. \textbf{F.} Temporal clustering scores versus feature
+clustering scores on late lists. \textbf{G.} Temporal clustering scores on late
+lists versus feature clustering scores on early lists. \textbf{H.} Temporal
+clustering scores on early lists versus feature clustering scores on late
+lists. \textbf{All panels.} Each dot in the main scatterplots denotes the
+average scores for one participant. The colored regression lines are computed
+across participants. The inset displays condition-averaged results, where each
+dot reflects a single condition and the regression line is computed across
+experimental conditions. All error ribbons denote bootstrap-estimated 95\%
+confidence intervals.} \label{fig:clustering-scatterplots}
+
+\end{figure}
 

### 2023-01-02 paper/main.tex
finished drafts of all main and supplemental figures

@@
 \newcommand{\dynamicsAdaptive}{S2}
-\newcommand{\fingerprintsRandom}{S3}
-\newcommand{\fingerprintsAdaptive}{S4}
-\newcommand{\recallInit}{S5}
+\newcommand{\accuracyByList}{S3}
+\newcommand{\clusterCorrs}{S4}
+\newcommand{\fingerprintsRandom}{S5}
+\newcommand{\fingerprintsAdaptive}{S6}
+\newcommand{\recallInit}{S7}
+\newcommand{\fingerprintTrajectoryRandom}{S8}
 
@@
 
+% figure: accuracy by list
+Figure~\accuracyByList.
+
+% figure: recall initiation
+Figure~\recallInit.
+
 % figure: clustering effects
@@
 
-% figure: recall initiation
-Figure~\recallInit.
+% figure: correlations between clustering scores
+Figure~\clusterCorrs.
 
@@
 % figure: fingerprint trajectories
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprint_trajectories}
 
-% figure: carryover effects: clustering early vs. clustering late
+    \caption{\textbf{Memory fingerprint dynamics (order manipulation
+    conditions).} \textbf{A.} Each column (and color) reflects an experimental
+    condition. In the top panels, each marker displays a 2D projection of the
+    (across-participant) average memory fingerprint for one list. Order
+    manipulation (early) lists are denoted by circles and randomly ordered
+    (late) lists are denoted by stars. All of the fingerprints (across all
+    conditions and lists) are projected into a common space. The bar plots in
+    the bottom panels display the Euclidean distances of the per-list memory
+    fingerprints to the list 0 fingerprint, for each condition. Error bars
+    denote bootstrap-estimated 95\% confidence intervals. The dotted vertical
+    lines denote the boundaries between early and late lists. \textbf{B.} In
+    this panel, the fingerprints for early (circle) and late (star) lists are
+    averaged across lists and participants before projecting the fingerprints
+    into a (new) 2D space. See Figure~\fingerprintTrajectoryRandom~for
+    analogous plots for the random (control) conditions. }
+    \label{fig:fingerprint-trajectories}
 
-% figure: clustering carryover vs. accuracy difference
+    \end{figure}
 
-% figure: clustering carryover vs. temporal clustering differences
+% figure: carryover effects: 
+%  - clustering early vs. clustering late
+%  - clustering carryover vs. accuracy difference
+%  - clustering carryover vs. temporal clustering differences
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/clustering_carryover}
+
+    \caption{\textbf{Feature clustering carryover effects.} \textbf{A.} Feature clustering scores for ordder manipulation (early) versus randomly ordered (late) lists.
+    \textbf{B.}  Accuracy differences (on early versus late lists) versus feature clustering ``carryover'' (defined as the differences between the average clustering scores on early and late lists).
+    \textbf{C.}  Temporal clustering differences (on early versus late lists) versus feature clustering carryover.  \textbf{All panels.} Each dot in the main scatterplots denotes the
+    average scores for one participant. The colored regression lines are computed
+    across participants. The inset displays condition-averaged results, where each
+    dot reflects a single condition and the regression line is computed across
+    experimental conditions. All error ribbons denote bootstrap-estimated 95\%
+    confidence intervals.} \label{fig:clustering-carryover}
+    \end{figure}
 

### 2023-01-02 paper/supplement.tex
finished drafts of all main and supplemental figures

@@
 \begin{figure}[tp] \centering
+    \includegraphics[width=0.75\textwidth]{figures/accuracy_by_list}
+
+    \caption{\textbf{Recall accuracy by study list number.}  Each panel displays the average recall accuracy (across participants) as a function of the number of studied lists for the
+    random (control) conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}.  The conditions are denoted by color.  Note that the first four
+    lists in all of the ``adaptive'' conditions were ordered randomly to compute a baseline fingerprint for each participant prior to initiating the adaptive ordering procedure.  All error
+    ribbons denote bootstrap-estimated 95\% confidence intervals, calculated across participants.}
+        \label{fig:accuracy-by-list}
+\end{figure}
+
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/clustering_correlations}
+
+    \caption{\textbf{Correlations between feature clustering scores (order manipulation conditions).}  Each column reflects one experimental condition.  The matrices in the top and middle
+    rows display across-participant correlations between clustering scores for each feature dimension (top: order manipulation lists; middle: randomly ordered lists).  The matrices
+    in the bottom rows display the differences between the top and middle rows.}
+        \label{fig:clustering-correlations}
+\end{figure}
+
+\begin{figure}[tp] \centering
     \includegraphics[width=\textwidth]{figures/fingerprints_random}
@@
 
-\begin{figure}[tp] \centering
-    \includegraphics[width=\textwidth]{figures/fingerprints_adaptive}
-
-    \caption{\textbf{Memory ``fingerprints.'' (adaptive conditions).}
-    The across-participant distributions of clustering scores for each feature
-    type ($x$-coordinate) are displayed for each experimental condition
-    (color), separately for order manipulation (early, top) and randomly
-    ordered (late, bottom) lists.}
-        \label{fig:fingerprints-adaptive}
-\end{figure}
-
 \begin{sidewaysfigure}
@@
 
-\newpage
-\renewcommand{\refname}{Supplementary references}
-\bibliographystyle{apa}
-\bibliography{CDL-bibliography/cdl}
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprint_trajectories_random}
+
+    \caption{\textbf{Memory fingerprint dynamics (random conditions).}
+    \textbf{A.} Each column (and color) reflects an experimental condition. In
+    the top panels, each marker displays a 2D projection of the
+    (across-participant) average memory fingerprint for one list. Early lists
+    are denoted by circles and late lists are denoted by stars. All of the
+    fingerprints (across all conditions and lists) are projected into a common
+    space. The bar plots in the bottom panels display the Euclidean distances
+    of the per-list memory fingerprints to the list 0 fingerprint, for each
+    condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
+    The dotted vertical lines denote the boundaries between early and late
+    lists. \textbf{B.} In this panel, the fingerprints for early (circle) and
+    late (star) lists are averaged across lists and participants before
+    projecting the fingerprints into a (new) 2D space.} \label{fig:fingerprint-trajectories-random}
+
+    \end{figure}
+
+% accuracy by list (list order effects)
+
+% correlations between different types of clustering
+
+%\newpage
+%\renewcommand{\refname}{Supplementary references}
+%\bibliographystyle{apa}
+%\bibliography{CDL-bibliography/cdl}
 

### 2023-01-03 paper/main.tex
added event boundary probability panel to figure 2

@@
 \caption{\textbf{Recall dynamics in feature rich free recall (order
-manipulation conditions).} \textbf{Left panels.} The probabilities of
-initiating recall with each word are plotted as a function of presentation
-position. \textbf{Middle panels.} The conditional probabilities of recalling
-each word are plotted as a function of the relative position (Lag) to the words
-recalled just-prior. \textbf{Right panels.} The overall probabilities of
-recalling each word are plotted as a function of presentation position.
-\textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
-intervals (calculated across participants). Top panels display the recall
-dynamics for early (order manipulation) lists in each condition (color). Bottom
-panels display the recall dynamics for late (randomly ordered) lists.  See
-Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the random (control) and adaptive conditions.}
+manipulation conditions).} \textbf{A.} Behavioral plots. \textbf{Left panels.}
+The probabilities of initiating recall with each word are plotted as a function
+of presentation position. \textbf{Middle panels.} The conditional probabilities
+of recalling each word are plotted as a function of the relative position (Lag)
+to the words recalled just-prior. \textbf{Right panels.} The overall
+probabilities of recalling each word are plotted as a function of presentation
+position. \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\%
+confidence intervals (calculated across participants). Top panels display the
+recall dynamics for early (order manipulation) lists in each condition (color).
+Bottom panels display the recall dynamics for late (randomly ordered) lists.
+See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
+random (control) and adaptive conditions. \textbf{B.} Proportion of event
+boundaries (see \textit{Methods}) for each condition's feature of focus,
+plotted as a function of presentation position.}
 

### 2023-01-03 paper/main.tex
added a notebook for analyzing demographic survey responses and compiling demographic information   - also added demographic survey responses to the data folder

@@
 \begin{figure} 
-    \begin{center}
-        \includegraphics[width=0.4 \textwidth]{figures/FRFR} \end{center} \vspace{-16pt}
+    \centering
+        \includegraphics[width=0.4 \textwidth]{figures/FRFR}
 
@@
 
+\begin{figure} 
+    \centering
+
+    \includegraphics[width=0.4 \textwidth]{figures/source/accuracy_adaptive}
+
+        \caption{\textbf{Recall performance (adaptive conditons).} The bars display the average probability of recall (taken across words, lists, and participants)
+        for lists from each adaptive condition.  Error bars denote bootstrap-estimated 95\% confidence intervals.  For additional details about
+        participants' behavior and performance during the adaptive conditions, see Figure~\dynamicsAdaptive.}
+
+    \label{fig:adaptive}
+\end{figure}
+
 \section*{Discussion}
@@
 \section*{Materials and methods}
+
 \subsection*{Participants}

### 2023-01-04 paper/main.tex
working on methods; backing up current progress

@@
 
+We enrolled a total of 491 Dartmouth undergraduate students across 11
+experimental conditions. The conditions included two primary controls (feature
+rich, reduced), two secondary controls (reduced (early), reduced (late)), six
+order manipulation conditions (category, size, length, first letter, color, and
+locaiton), and a final adaptive condition. Each of these conditions are
+described in the \textit{Experimental design} subsection below.
+
+Participants received course credit for enrolling in our study. We asked each
+participant to fill out a demographic survey that included information about
+their self-reported age, gender, ethnicity, race, education, vision, reading
+impairments, medications or recent injuries, coffee consumption on the day of
+testing, and level of alertness at the time of testing. All components of the
+demographics survey were optional. One participant elected not to fill out any
+part of the demographic survey, and all other participants report some or all
+of their requested demographic information.
+
+We aimed to run (to completion) at least 60 participants in each of the two
+primary control conditions and in the adaptive condition. In all other
+conditions we set a target enrollment of at least 30 participants. Because our
+data collection efforts were coordinated 12 researchers and multiple testing
+rooms and computers, it was not feasible for individual experimenters to know
+how many participants had been run in each experimental condition until the
+relevant databases were synchronized at the end of each working day. We also
+over-enrolled participants for each condition to help ensure that we met our
+minimum enrollment targets even if some participants dropped out of the study
+prematurely or did not show up for their testing session. This led us to exceed
+our target enrollments for several conditions.
+
+Participants were assigned to experimental conditions based loosely on their
+date of participation. (This aspect of our procedure helped us to more easily
+synchronize the experiment databases across multiple testing computers.)
+A complete report... %STOPPED HERE
+
 \subsection*{Experimental design}
 
+Our experiment is a variant of the classic free recall paradigm that we term
+\textit{feature-rich free recall}. In feature-rich free recall, participants
+study 16 lists, each comprised of 16 words that vary along a number of stimulus
+dimensions: semantic category, object size, text color, text location, word
+length, and starting letter (Fig.~\ref{fig:exp1}). Each list contains four
+words from each of four different semantic categories; all other stimulus
+features are randomized. After studying each list, the participant attempts to
+recall as many words as they can from that list, in any order they choose.
+Because each individual word is associated with several well-defined (and
+quantifiable) features, and because each list incorporates a diverse mix of
+feature values along each dimension, this allows us to evaluate participants'
+memory fingerprints in rich detail.
+
+Our experimental paradigm incorporates the Annyang and Google Cloud Speech API
+speech-to-text engines to automatically transcribe participants' verbal recalls
+into text. This allows recalls to be transcribed in real time-- a
+distinguishing feature of the experiment; in typical verbal recall experiments
+the audio data must be parsed manually. (We have run a pilot experiment to
+verify that the automatically transcribed recalls are sufficiently close to
+human-transcribed recalls to yield reliable data.) This real-time speech
+processing component of the paradigm plays an important role in Experiments
+3--5.
+
+\subsubsection*{Stimuli}
+
+\subsubsection*{Real-time speech-to-text processing}
+
+\subsubsection*{Order manipulation conditions}
+
+\paragraph{Feature dimensions.}
+
+\paragraph{Constructing feature-sorted lists.}
+
+\subsubsection*{Random conditions}
+
+\subsubsection*{Adaptive conditions}
+
+\paragraph{Online ``fingerprint'' analysis.}
+
+\paragraph{Ordering ``stabilize'' lists by an estimated fingerprint.}
+
+\paragraph{Ordering ``destabilize'' lists by an estimated fingerprint.}
+
 \subsection*{Analysis}
+
+\subsubsection*{Probability of first recall and probability of $n$\textsuperscript{th} recall}
+
+\subsubsection*{Lag conditional response probability}
+
+\subsubsection*{Computing clustering scores and memory fingerprints}
+
+\subsubsection*{Identifying event boundaries}
+
+\subsubsection*{Serial position curves and recall accuracy}
+
+\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
 

### 2023-01-05 paper/main.tex
working on summaries of demographic data, writing up methods
filled out participants section.   - the paper seems to be turning into something more appropriate for a long-form journal, so i've moved the methods section to after the into

@@
 
+\section*{Materials and methods}
+
+% need a summary statement...
+
+\subsection*{Participants}
+
+We enrolled a total of 491 Dartmouth undergraduate students across 11
+experimental conditions. The conditions included two primary controls (feature
+rich, reduced), two secondary controls (reduced (early), reduced (late)), six
+order manipulation conditions (category, size, length, first letter, color, and
+location), and a final adaptive condition. Each of these conditions are
+described in the \textit{Experimental design} subsection below.
+
+Participants received course credit for enrolling in our study. We asked each
+participant to fill out a demographic survey that included information about
+their self-reported age, gender, ethnicity, race, education, vision, reading
+impairments, medications or recent injuries, coffee consumption on the day of
+testing, and level of alertness at the time of testing. All components of the
+demographics survey were optional. One participant elected not to fill out any
+part of the demographic survey, and all other participants report some or all
+of their requested demographic information.
+
+We aimed to run (to completion) at least 60 participants in each of the two
+primary control conditions and in the adaptive condition. In all other
+conditions we set a target enrollment of at least 30 participants. Because our
+data collection efforts were coordinated 12 researchers and multiple testing
+rooms and computers, it was not feasible for individual experimenters to know
+how many participants had been run in each experimental condition until the
+relevant databases were synchronized at the end of each working day. We also
+over-enrolled participants for each condition to help ensure that we met our
+minimum enrollment targets even if some participants dropped out of the study
+prematurely or did not show up for their testing session. This led us to exceed
+our target enrollments for several conditions.
+
+Participants were assigned to experimental conditions based loosely on their
+date of participation. (This aspect of our procedure helped us to more easily
+synchronize the experiment databases across multiple testing computers.)  Of the 490 participants
+who opted to fill out the demographics survey, reported ages ranged from 17 to 31 years (mean: 19.1; standard deviation: 1.356).
+A total of 318 participants reported their gender as female, 170 as male, and 2 participants declined to report their gender.
+A total of 442 participants reported their ethnicity as ``not Hispanic or Latino,'' 39 as ``Hispanic or Latino,'' and 9 declined
+to report their ethnicity.  Participants reported their races as White (345 participants), Asian (120 participants), Black or African American (31 participants),
+American Indian or Alaska Native (11 particiapnts), Native Hawaiian or Other Pacific Islander (4 participants), Mixed race (3 participants), Middle Eastern (1 participant), and
+Arab (1 participant).  A total of 5 participants declined to report their race.  We note that several participants reported more than one of racial category.
+Participants reported their highest degrees achieved as ``Some college'' (359 participants), ``High school graduate'' (117 participants),
+``College graduate'' (7 participants), ``Some high school'' (5 participants), ``Doctorate'' (1 participant), and ``Master's degree'' (1 participant).
+A total of 482 participants reported no reading impairments, and 8 reported mild reading impairments such as mild dyslexia.  A total of 489
+participants reported having normal color vision and 1 participant reported that they were color blind.  A total of 482 participants
+reported taking no prescription medications and having no recent injuries; 4 participants reported having ADHD, 1 reported having dyslexia, 1 reported having
+allergies, 1 reported a recently torn ACL/MCL, and 1 reported a concussion from several months prior.  The participants reported consuming 0 -- 3 cups of coffee
+prior to the testing session (mean: 0.32 cups; standard deviation: 0.58 cups).  Participants reported their current level of alertness, and we converted
+their responses to numerical scores as follows: ``very sluggish'' (-2), ``a little sluggish'' (-1), ``neutral'' (0), ``a little alert'' (1), and ``very alert'' (2).
+Across all participants, the full range of alertness levels were reported (range: -2 -- 2; mean: 0.35; standard deviation: 0.89) .
+
+We dropped from our dataset the 1 participant who reported abnormal color vision, as well as 39 participants whose data
+were corrupted due to technical failures while running the experiment or during the daily database merges.  In total, this left usable data from
+452 participants, broken down by experimental condition as follows: feature rich (67 participants), reduced (61 participants), reduced (late) (41 participants), 
+reduced (early), (42 participants), category (30 participants), size (30 participants), length (30 participants), first letter (30 participants),
+color (31 participants), location (30 participants), and adaptive (60 participants).  The participant who declined to fill out their demographic
+survey participated in the location condition, and we verified verbally that they had normal color vision.
+
+\subsection*{Experimental design}
+
+Our experiment is a variant of the classic free recall paradigm that we term
+\textit{feature-rich free recall}. In feature-rich free recall, participants
+study 16 lists, each comprised of 16 words that vary along a number of stimulus
+dimensions: semantic category, object size, text color, text location, word
+length, and starting letter (Fig.~\ref{fig:exp}). Each list contains four
+words from each of four different semantic categories and two object sizes; all other stimulus
+features are randomized. After studying each list, the participant attempts to
+recall as many words as they can from that list, in any order they choose.
+Because each individual word is associated with several well-defined (and
+quantifiable) features, and because each list incorporates a diverse mix of
+feature values along each dimension, this allows us to evaluate participants'
+memory fingerprints in rich detail.
+
+Our experimental paradigm incorporates the Annyang and Google Cloud Speech API
+speech-to-text engines to automatically transcribe participants' verbal recalls
+into text. This allows recalls to be transcribed in real time-- a
+distinguishing feature of the experiment; in typical verbal recall experiments
+the audio data must be parsed manually. (We have run a pilot experiment to
+verify that the automatically transcribed recalls are sufficiently close to
+human-transcribed recalls to yield reliable data.) This real-time speech
+processing component of the paradigm plays an important role in Experiments
+3--5.
+
+\subsubsection*{Stimuli}
+
+\subsubsection*{Real-time speech-to-text processing}
+
+\subsubsection*{Order manipulation conditions}
+
+\paragraph{Feature dimensions.}
+
+\paragraph{Constructing feature-sorted lists.}
+
+\subsubsection*{Random conditions}
+
+\subsubsection*{Adaptive conditions}
+
+\paragraph{Online ``fingerprint'' analysis.}
+
+\paragraph{Ordering ``stabilize'' lists by an estimated fingerprint.}
+
+\paragraph{Ordering ``destabilize'' lists by an estimated fingerprint.}
+
+\subsection*{Analysis}
+
+\subsubsection*{Probability of first recall and probability of $n$\textsuperscript{th} recall}
+
+\subsubsection*{Lag conditional response probability}
+
+\subsubsection*{Computing clustering scores and memory fingerprints}
+
+\subsubsection*{Identifying event boundaries}
+
+\subsubsection*{Serial position curves and recall accuracy}
+
+\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
+
 \section*{Results}
@@
 
-\section*{Materials and methods}
-
-\subsection*{Participants}
-
-We enrolled a total of 491 Dartmouth undergraduate students across 11
-experimental conditions. The conditions included two primary controls (feature
-rich, reduced), two secondary controls (reduced (early), reduced (late)), six
-order manipulation conditions (category, size, length, first letter, color, and
-locaiton), and a final adaptive condition. Each of these conditions are
-described in the \textit{Experimental design} subsection below.
-
-Participants received course credit for enrolling in our study. We asked each
-participant to fill out a demographic survey that included information about
-their self-reported age, gender, ethnicity, race, education, vision, reading
-impairments, medications or recent injuries, coffee consumption on the day of
-testing, and level of alertness at the time of testing. All components of the
-demographics survey were optional. One participant elected not to fill out any
-part of the demographic survey, and all other participants report some or all
-of their requested demographic information.
-
-We aimed to run (to completion) at least 60 participants in each of the two
-primary control conditions and in the adaptive condition. In all other
-conditions we set a target enrollment of at least 30 participants. Because our
-data collection efforts were coordinated 12 researchers and multiple testing
-rooms and computers, it was not feasible for individual experimenters to know
-how many participants had been run in each experimental condition until the
-relevant databases were synchronized at the end of each working day. We also
-over-enrolled participants for each condition to help ensure that we met our
-minimum enrollment targets even if some participants dropped out of the study
-prematurely or did not show up for their testing session. This led us to exceed
-our target enrollments for several conditions.
-
-Participants were assigned to experimental conditions based loosely on their
-date of participation. (This aspect of our procedure helped us to more easily
-synchronize the experiment databases across multiple testing computers.)
-A complete report... %STOPPED HERE
-
-\subsection*{Experimental design}
-
-Our experiment is a variant of the classic free recall paradigm that we term
-\textit{feature-rich free recall}. In feature-rich free recall, participants
-study 16 lists, each comprised of 16 words that vary along a number of stimulus
-dimensions: semantic category, object size, text color, text location, word
-length, and starting letter (Fig.~\ref{fig:exp1}). Each list contains four
-words from each of four different semantic categories; all other stimulus
-features are randomized. After studying each list, the participant attempts to
-recall as many words as they can from that list, in any order they choose.
-Because each individual word is associated with several well-defined (and
-quantifiable) features, and because each list incorporates a diverse mix of
-feature values along each dimension, this allows us to evaluate participants'
-memory fingerprints in rich detail.
-
-Our experimental paradigm incorporates the Annyang and Google Cloud Speech API
-speech-to-text engines to automatically transcribe participants' verbal recalls
-into text. This allows recalls to be transcribed in real time-- a
-distinguishing feature of the experiment; in typical verbal recall experiments
-the audio data must be parsed manually. (We have run a pilot experiment to
-verify that the automatically transcribed recalls are sufficiently close to
-human-transcribed recalls to yield reliable data.) This real-time speech
-processing component of the paradigm plays an important role in Experiments
-3--5.
-
-\subsubsection*{Stimuli}
-
-\subsubsection*{Real-time speech-to-text processing}
-
-\subsubsection*{Order manipulation conditions}
-
-\paragraph{Feature dimensions.}
-
-\paragraph{Constructing feature-sorted lists.}
-
-\subsubsection*{Random conditions}
-
-\subsubsection*{Adaptive conditions}
-
-\paragraph{Online ``fingerprint'' analysis.}
-
-\paragraph{Ordering ``stabilize'' lists by an estimated fingerprint.}
-
-\paragraph{Ordering ``destabilize'' lists by an estimated fingerprint.}
-
-\subsection*{Analysis}
-
-\subsubsection*{Probability of first recall and probability of $n$\textsuperscript{th} recall}
-
-\subsubsection*{Lag conditional response probability}
-
-\subsubsection*{Computing clustering scores and memory fingerprints}
-
-\subsubsection*{Identifying event boundaries}
-
-\subsubsection*{Serial position curves and recall accuracy}
-
-\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
-
 \bibliographystyle{apa}

### 2023-01-05 paper/main.tex
backing up changes to participants, stimuli, and speech-to-text sections of materials and methods

@@
 
-\begin{figure} 
-    \centering
-        \includegraphics[width=0.4 \textwidth]{figures/FRFR}
-
-        \caption{\textbf{Feature-rich free recall.} After studying lists
-comprised of words that vary along several feature dimensions, participants
-verbally recall words in any order (microphone icon).}
-
-    \label{fig:exp}
-\end{figure}
-
 \section*{Materials and methods}
-
-% need a summary statement...
 
@@
 date of participation. (This aspect of our procedure helped us to more easily
-synchronize the experiment databases across multiple testing computers.)  Of the 490 participants
-who opted to fill out the demographics survey, reported ages ranged from 17 to 31 years (mean: 19.1; standard deviation: 1.356).
-A total of 318 participants reported their gender as female, 170 as male, and 2 participants declined to report their gender.
-A total of 442 participants reported their ethnicity as ``not Hispanic or Latino,'' 39 as ``Hispanic or Latino,'' and 9 declined
-to report their ethnicity.  Participants reported their races as White (345 participants), Asian (120 participants), Black or African American (31 participants),
-American Indian or Alaska Native (11 particiapnts), Native Hawaiian or Other Pacific Islander (4 participants), Mixed race (3 participants), Middle Eastern (1 participant), and
-Arab (1 participant).  A total of 5 participants declined to report their race.  We note that several participants reported more than one of racial category.
-Participants reported their highest degrees achieved as ``Some college'' (359 participants), ``High school graduate'' (117 participants),
-``College graduate'' (7 participants), ``Some high school'' (5 participants), ``Doctorate'' (1 participant), and ``Master's degree'' (1 participant).
-A total of 482 participants reported no reading impairments, and 8 reported mild reading impairments such as mild dyslexia.  A total of 489
-participants reported having normal color vision and 1 participant reported that they were color blind.  A total of 482 participants
-reported taking no prescription medications and having no recent injuries; 4 participants reported having ADHD, 1 reported having dyslexia, 1 reported having
-allergies, 1 reported a recently torn ACL/MCL, and 1 reported a concussion from several months prior.  The participants reported consuming 0 -- 3 cups of coffee
-prior to the testing session (mean: 0.32 cups; standard deviation: 0.58 cups).  Participants reported their current level of alertness, and we converted
-their responses to numerical scores as follows: ``very sluggish'' (-2), ``a little sluggish'' (-1), ``neutral'' (0), ``a little alert'' (1), and ``very alert'' (2).
-Across all participants, the full range of alertness levels were reported (range: -2 -- 2; mean: 0.35; standard deviation: 0.89) .
+synchronize the experiment databases across multiple testing computers.) Of the
+490 participants who opted to fill out the demographics survey, reported ages
+ranged from 17 to 31 years (mean: 19.1; standard deviation: 1.356). A total of
+318 participants reported their gender as female, 170 as male, and 2
+participants declined to report their gender. A total of 442 participants
+reported their ethnicity as ``not Hispanic or Latino,'' 39 as ``Hispanic or
+Latino,'' and 9 declined to report their ethnicity. Participants reported their
+races as White (345 participants), Asian (120 participants), Black or African
+American (31 participants), American Indian or Alaska Native (11 particiapnts),
+Native Hawaiian or Other Pacific Islander (4 participants), Mixed race (3
+participants), Middle Eastern (1 participant), and Arab (1 participant). A
+total of 5 participants declined to report their race. We note that several
+participants reported more than one of racial category. Participants reported
+their highest degrees achieved as ``Some college'' (359 participants), ``High
+school graduate'' (117 participants), ``College graduate'' (7 participants),
+``Some high school'' (5 participants), ``Doctorate'' (1 participant), and
+``Master's degree'' (1 participant). A total of 482 participants reported no
+reading impairments, and 8 reported mild reading impairments such as mild
+dyslexia. A total of 489 participants reported having normal color vision and 1
+participant reported that they were color blind. A total of 482 participants
+reported taking no prescription medications and having no recent injuries; 4
+participants reported having ADHD, 1 reported having dyslexia, 1 reported
+having allergies, 1 reported a recently torn ACL/MCL, and 1 reported a
+concussion from several months prior. The participants reported consuming 0 --
+3 cups of coffee prior to the testing session (mean: 0.32 cups; standard
+deviation: 0.58 cups). Participants reported their current level of alertness,
+and we converted their responses to numerical scores as follows: ``very
+sluggish'' (-2), ``a little sluggish'' (-1), ``neutral'' (0), ``a little
+alert'' (1), and ``very alert'' (2). Across all participants, the full range of
+alertness levels were reported (range: -2 -- 2; mean: 0.35; standard deviation:
+0.89) .
 
-We dropped from our dataset the 1 participant who reported abnormal color vision, as well as 39 participants whose data
-were corrupted due to technical failures while running the experiment or during the daily database merges.  In total, this left usable data from
-452 participants, broken down by experimental condition as follows: feature rich (67 participants), reduced (61 participants), reduced (late) (41 participants), 
-reduced (early), (42 participants), category (30 participants), size (30 participants), length (30 participants), first letter (30 participants),
-color (31 participants), location (30 participants), and adaptive (60 participants).  The participant who declined to fill out their demographic
-survey participated in the location condition, and we verified verbally that they had normal color vision.
+We dropped from our dataset the 1 participant who reported abnormal color
+vision, as well as 39 participants whose data were corrupted due to technical
+failures while running the experiment or during the daily database merges. In
+total, this left usable data from 452 participants, broken down by experimental
+condition as follows: feature rich (67 participants), reduced (61
+participants), reduced (late) (41 participants), reduced (early), (42
+participants), category (30 participants), size (30 participants), length (30
+participants), first letter (30 participants), color (31 participants),
+location (30 participants), and adaptive (60 participants). The participant who
+declined to fill out their demographic survey participated in the location
+condition, and we verified verbally that they had normal color vision.
 
@@
 dimensions: semantic category, object size, text color, text location, word
-length, and starting letter (Fig.~\ref{fig:exp}). Each list contains four
-words from each of four different semantic categories and two object sizes; all other stimulus
-features are randomized. After studying each list, the participant attempts to
-recall as many words as they can from that list, in any order they choose.
-Because each individual word is associated with several well-defined (and
-quantifiable) features, and because each list incorporates a diverse mix of
-feature values along each dimension, this allows us to evaluate participants'
-memory fingerprints in rich detail.
+length, and starting letter (Fig.~\ref{fig:exp}). Each list contains four words
+from each of four different semantic categories and two object sizes; all other
+stimulus features are randomized. After studying each list, the participant
+attempts to recall as many words as they can from that list, in any order they
+choose. Because each individual word is associated with several well-defined
+(and quantifiable) features, and because each list incorporates a diverse mix
+of feature values along each dimension, this allows us to evaluate
+participants' memory fingerprints in rich detail.
 
-Our experimental paradigm incorporates the Annyang and Google Cloud Speech API
-speech-to-text engines to automatically transcribe participants' verbal recalls
-into text. This allows recalls to be transcribed in real time-- a
-distinguishing feature of the experiment; in typical verbal recall experiments
-the audio data must be parsed manually. (We have run a pilot experiment to
-verify that the automatically transcribed recalls are sufficiently close to
-human-transcribed recalls to yield reliable data.) This real-time speech
-processing component of the paradigm plays an important role in Experiments
-3--5.
+\begin{figure} 
+    \centering
+        \includegraphics[width=0.75 \textwidth]{figures/FRFR}
+
+        \caption{\textbf{Feature-rich free recall.} After studying lists
+comprised of words that vary along several feature dimensions, participants
+verbally recall words in any order (microphone icon).}
+
+    \label{fig:exp}
+\end{figure}
 
@@
 
+Stimuli in our paradigm were 256 English words selected in a previous
+study~\citep{ZimaEtal18}. The words all referred to concrete nouns, and were
+chosen from 15 unique semantic categories: body parts, building-related,
+cities, clothing, countries, flowers, fruits, insects, instruments,
+kitchen-related, mammals, (US) states, tools, trees, and vegetables. We also
+tagged each word according to the approximate size of the object the word
+referred to. Words were labeled as ``small'' if the corresponding object was
+likely able to ``fit in a standard shoebox'' or ``large'' if the object was
+larger than a shoebox. Semantic categories varied in how many object sizes they
+reflected (mean number of different sizes per category: 1.33; standard
+deviation: 0.49). The numbers of words in each semantic category also varied
+from 12 -- 28 (mean number of words per category: 17.07; standard deviation
+number of words: 4.65). We also identified lexicographic features for each
+word, including the words' first letters and lengths (i.e., number of letters).
+Across all categories, all possible first letters were represented except for
+`Q' (average number of unique first letters per category: 11; standard
+deviation: 2 letters). Word lengths ranged from 3 -- 12 letters (average: 6.17
+letters; standard deviation: 2.06 letters).
+
+We assigned the categorized words into a total of 16 lists with several
+constraints. First, we required that each list contained words from exactly 4
+unique categories, each with exactly 4 examplars from each category. Second, we
+required that (across all words on the list) at least one instance of both
+object sizes were represented. On average, each category was represented in
+4.27 lists (standard deviation: 1.16 lists). Aside from these two constraints,
+we assigned each word to a unique list. After random assignment, each list
+contained words with an average of 11.13 unique starting letters (standard
+deviation: 1.15 letters) and an average word length of 6.17 letters (standard
+deviation: 0.34 letters).
+
+The above assignments of words to lists was performed once across all
+participants, such that every participant studied the same set of 16 lists. We
+randomized the study order of these lists across participants. For participants
+in some conditions, on some lists, we also randomly varied two additional
+visual features to each word: the presentation font color, and the word's
+onscreen location. These attributes were assigned independently for word (and
+for every participant) at the times the words were displayed onscreen. These
+visual features were varied for words in all lists and conditions except for
+the ``reduced'' condition (all lists), the first eight lists of the ``reduced
+(early)'' condition, and the last eight lists of the ``reduced (late)''
+condition. In these latter cases, words were all presented in black at the
+center of the experimental computer's display.
+
+To assign a random font color to each word, we selected three integers
+uniformly and at random between 0 and 255, corresponding to the red (r), green
+(g), and blue (b) color channels for that word. To assign random presentation
+locations to each word, we selected two floating point numbers uniformly at
+random (one for the word's horizontal $x$ coordinate and the other for its
+vertical $y$ coordinate).  The bounds of these coordinates were selected to cover the entire
+visible area of the display without cutting off any part of the words.  The words were shown
+on 27~in (diagonal) Retina 5K iMac displays (resolution: 5120 $\times$ 2880 pixels).
+
 \subsubsection*{Real-time speech-to-text processing}
+
+Our experimental paradigm incorporates the Google Cloud Speech API
+speech-to-text engine~\citep{HalpEtal16} to automatically transcribe
+participants' verbal recalls into text. This allows recalls to be transcribed
+in real time-- a distinguishing feature of the experiment; in typical verbal
+recall experiments the audio data must be parsed manually. In prior work, we
+used a similar experimental setup (equivalent to the ``reduced'' condition in
+the present study) to verify that the automatically transcribed recalls were
+sufficiently close to human-transcribed recalls to yield reliable
+data~\citep{ZimaEtal18}. This real-time speech processing component of the
+paradigm plays an important role in the ``adaptive'' condition of the
+experiment, as described below.
 

### 2023-01-05 paper/main.tex
added example lists from each experimental conditions (in the FRFR figure).  working on methods (still...).   - also added maddy as an author (yay!)

@@
 \author{Jeremy R.
-Manning\textsuperscript{1, *}, Andrew C.
-Heusser\textsuperscript{1, 2}, Kirsten Ziman\textsuperscript{1, 3},\\Emily Whitaker\textsuperscript{1}, and Paxton C.
-Fitzpatrick\textsuperscript{1}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili Interactive\\\textsuperscript{3}Princeton University\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
+Manning\textsuperscript{1, *}, Kirsten Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
+Fitzpatrick\textsuperscript{1}, Madeline R.
+Lee\textsuperscript{1}, and Andrew C.
+Heusser\textsuperscript{1, 3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton University\\\textsuperscript{3}Akili Interactive\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
 
@@
 alertness levels were reported (range: -2 -- 2; mean: 0.35; standard deviation:
-0.89) .
+0.89).
 
@@
 study 16 lists, each comprised of 16 words that vary along a number of stimulus
-dimensions: semantic category, object size, text color, text location, word
-length, and starting letter (Fig.~\ref{fig:exp}). Each list contains four words
-from each of four different semantic categories and two object sizes; all other
-stimulus features are randomized. After studying each list, the participant
-attempts to recall as many words as they can from that list, in any order they
-choose. Because each individual word is associated with several well-defined
-(and quantifiable) features, and because each list incorporates a diverse mix
-of feature values along each dimension, this allows us to evaluate
-participants' memory fingerprints in rich detail.
+dimensions (Fig.~\ref{fig:exp}). The stimulus dimensions include two semantic
+features related to the \textit{meanings} of the words (semantic category,
+referrent object size), two lexicographic features related to the
+\textit{letters} that make up the words (word length in number of letters,
+identity of the word's first letter), and two visual features that are
+independent of the words themselves (text color, presentation location). Each
+list contains four words from each of four different semantic categories and
+two object sizes; all other stimulus features are randomized. After studying
+each list, the participant attempts to recall as many words as they can from
+that list, in any order they choose. Because each individual word is associated
+with several well-defined (and quantifiable) features, and because each list
+incorporates a diverse mix of feature values along each dimension, this allows
+us to evaluate participants' memory fingerprints in rich detail.
 
-\begin{figure} 
+\begin{figure}[tp]
     \centering
-        \includegraphics[width=0.75 \textwidth]{figures/FRFR}
+        \includegraphics[width=\textwidth]{figures/FRFR}
 
-        \caption{\textbf{Feature-rich free recall.} After studying lists
-comprised of words that vary along several feature dimensions, participants
-verbally recall words in any order (microphone icon).}
+\caption{\textbf{Feature-rich free recall.} After studying lists comprised of
+words that vary along several feature dimensions, participants verbally recall
+words in any order (microphone icon). Each experimental condition manipulates
+word features and/or presentation orders within and/or across lists. The rows
+display representative (illustrated) examples of the first lists participants
+might encounter in each condition. The rectangles during the ``Presentation
+phase'' show illustrated screen captures during a series of word presentations.
+Each word appeared onscreen for 2 seconds, followed by 2 seconds of blank
+screen. The red microphone icons during the ``Recall'' phase denote the one
+minute verbal recall interval.  The labels on the right (and corresponding groupings on the left)
+denote experimental condition labels.}
 
@@
 
+\subsubsection*{Random conditions}
+
+We used four ``control'' conditions to evaluate and explore participants'
+baseline behaviors. We also used performance on these control conditions to
+help interpret performance in other ``manipulation'' conditions. Two control
+conditions served as ``anchorpoints.'' In the first anchorpoint condition,
+which we call the \textit{feature rich} condition, we randomly shuffled the
+presentation order (independently for each participant) of the words on each
+list. The second anchorpoint condition
+
 \subsubsection*{Order manipulation conditions}
 
-\paragraph{Feature dimensions.}
-
 \paragraph{Constructing feature-sorted lists.}
-
-\subsubsection*{Random conditions}
 

### 2023-01-05 paper/supplement.tex
added example lists from each experimental conditions (in the FRFR figure).  working on methods (still...).   - also added maddy as an author (yay!)

@@
 
+\doublespacing
+
 \newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits}
@@
 \author{Jeremy R.
-Manning\textsuperscript{1, *}, Andrew C.
-Heusser\textsuperscript{1, 2}, Kirsten Ziman\textsuperscript{1, 3},\\Emily Whitaker\textsuperscript{1}, and Paxton C.
-Fitzpatrick\textsuperscript{1}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili Interactive\\\textsuperscript{3}Princeton University\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
+Manning\textsuperscript{1, *}, Kirsten Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
+Fitzpatrick\textsuperscript{1}, Madeline R.
+Lee\textsuperscript{1}, and Andrew C.
+Heusser\textsuperscript{1, 3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton University\\\textsuperscript{3}Akili Interactive\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
+
 \date{}

### 2023-01-06 paper/main.tex
- finished "random conditions" description - added bryan as a co-author (yay!) - working on description of order manipulation conditions

@@
 
-\title{Carryover effects in free recall reveal how prior experiences influence memories of new experiences}
-\author{Jeremy R.
-Manning\textsuperscript{1, *}, Kirsten Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
-Fitzpatrick\textsuperscript{1}, Madeline R.
-Lee\textsuperscript{1}, and Andrew C.
-Heusser\textsuperscript{1, 3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton University\\\textsuperscript{3}Akili Interactive\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
+\title{Carryover effects in free recall reveal how prior experiences influence
+memories of new experiences}
+
+\author{Jeremy R. Manning\textsuperscript{1, *}, Kirsten
+Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
+Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Bryan J.
+Bollinger\textsuperscript{1},\\and Andrew C. Heusser\textsuperscript{1,
+3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
+University\\\textsuperscript{3}Akili
+Interactive\\\textsuperscript{*}Corresponding author:
+jeremy.r.manning@dartmouth.edu}
 
@@
 The above assignments of words to lists was performed once across all
-participants, such that every participant studied the same set of 16 lists. We
-randomized the study order of these lists across participants. For participants
-in some conditions, on some lists, we also randomly varied two additional
-visual features to each word: the presentation font color, and the word's
-onscreen location. These attributes were assigned independently for word (and
-for every participant) at the times the words were displayed onscreen. These
-visual features were varied for words in all lists and conditions except for
-the ``reduced'' condition (all lists), the first eight lists of the ``reduced
-(early)'' condition, and the last eight lists of the ``reduced (late)''
-condition. In these latter cases, words were all presented in black at the
-center of the experimental computer's display.
+participants, such that every participant studied the same set of 16 lists. In
+every condition we randomized the study order of these lists across
+participants. For participants in some conditions, on some lists, we also
+randomly varied two additional visual features to each word: the presentation
+font color, and the word's onscreen location. These attributes were assigned
+independently for word (and for every participant) at the times the words were
+displayed onscreen. These visual features were varied for words in all lists
+and conditions except for the ``reduced'' condition (all lists), the first
+eight lists of the ``reduced (early)'' condition, and the last eight lists of
+the ``reduced (late)'' condition. In these latter cases, words were all
+presented in black at the center of the experimental computer's display.
 
@@
 on 27~in (diagonal) Retina 5K iMac displays (resolution: 5120 $\times$ 2880 pixels).
+
+Most of the experimental manipulations we carried out entailed presenting or
+sorting the presented words differently on the first eight lists participants
+studied (which we call \textit{early} lists) versus on the final eight lists
+they studied (\textit{late} lists). Since every participant studied exactly 16
+lists, using this terminology every list was either ``early'' or ``late''
+depending on its order in the list study sequence.
 
@@
 
-\subsubsection*{Random conditions}
+\subsubsection*{Random conditions (Fig.~\ref{fig:exp}, top four rows)}
 
@@
 presentation order (independently for each participant) of the words on each
-list. The second anchorpoint condition
+list. In the second anchorpoint condition, which we call the \textit{reduced}
+condition, we randomized word presentations as in the feature rich condition.
+However, rather than assigning each word a random color and location, we
+instead displayed all of the words in black and at the center of the screen.
 
-\subsubsection*{Order manipulation conditions}
+In the \textit{reduced (early)} condition, we followed the ``reduced''
+procedure (presenting each word in black at the center of the screen) for early
+lists, and followed the ``feature rich'' procedure (presenting each word in a
+random color and location) for late lists. Finally, in the \textit{reduced
+(late)} condition, we followed the feature rich procedure for earlylists and
+the reduced procedure for late lists.
+
+\subsubsection*{Order manipulation conditions (Fig.~\ref{fig:exp}, middle six
+rows)}
+
+Each of six \textit{order manipulation} conditions used a different
+feature-based sorting procedure to order words on early lists, where each
+sorting procedure relied on one feature dimension. On late lists, the words
+were always presented in a randomized order (chosen anew for each participant).
+In all of the order manipulation conditions, we varied words' font colors and
+onscreen locations, as in the feature rich condition.
 

### 2023-01-06 paper/main.tex
draft of order manipulation conditions:   - added Allison as an author (yay!)   - added "stick" figure to illustrate stochastic word ordering procedure   - added descriptions + equations for order manipulation conditions

@@
 Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
-Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Bryan J.
-Bollinger\textsuperscript{1},\\and Andrew C. Heusser\textsuperscript{1,
+Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, 
+Allison M Frantz\textsuperscript{1}, \\Bryan J.
+Bollinger\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
 3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
@@
 feature-based sorting procedure to order words on early lists, where each
-sorting procedure relied on one feature dimension. On late lists, the words
-were always presented in a randomized order (chosen anew for each participant).
-In all of the order manipulation conditions, we varied words' font colors and
+sorting procedure relied on one relevant feature dimension. All of the
+irrelevant features varied freely across words on early lists, in that we did
+not consider irrelevant features in ordering the early lists. However, some
+features were correlated-- for example, some semantic categories of words
+referred to objects that tended to be a particular size, which means that
+category and size are not fully independent. On late lists, the words were
+always presented in a randomized order (chosen anew for each participant). In
+all of the order manipulation conditions, we varied words' font colors and
 onscreen locations, as in the feature rich condition.
 
-\paragraph{Constructing feature-sorted lists.}
+\paragraph{Defining feature-based distances.} Sorting words according to a
+given relevant feature requires first defining a distance function for
+quantifying the dissimilarity between each pair of features. This function
+varied according to the type of features. Semantic features (category and size)
+are \textit{categorical}. For these features, we defined a binary distance
+function: two words were considered to ``match'' (i.e., have a distance of 0)
+if their labels are the same (i.e., both from the same semantic category or
+both of the same size). If two words' labels were different for a given
+feature, we defined the words to have a distance of 1 for that feature.
+Lexicographic features (length and first letter) are \textit{discrete}. For
+these features we defined a discrete distance function. Specifically, we
+defined the distance between two words as either the absolute difference
+between their lengths, or the absolute distance between their starting letters
+in the English alphabet, respectively. For example, two words that started with
+the same letter would have a ``first letter'' distance of 0, and words starting
+with `J' and `A' would have a first letter distance of 9. Because words'
+lengths and letters' positions in the alphabet are always integers, these
+discrete distances always take on integer values. Finally, the visual features
+(color and location) are \textit{continuous} and \textit{multivariate}, in that
+each ``feature'' takes on multiple (positive) real values. We defined the
+``color'' and ``location'' distances between two words as the Euclidean
+distances between their $(r, g, b)$ color or $(x, y)$ location vectors,
+respectively. Therefore the color and location distance measures always take on
+positive real values (upper bounded at 441.67 for color, or 27~in for location,
+reflecting the distances between the corresponding maximally different
+vectors).
+
+\paragraph{Constructing feature-sorted lists.} Given a list of words, a
+relevant feature, and each word's value(s) for that feature, we developed a
+stochastic algorithm for (noisily) sorting the words. First, we choose a word
+uniformly at random from the set of candidates. Next, we compute the distances
+between the chosen word's feature(s) and the corresponding feature(s) of all
+yet-to-be-presented words. Third, we convert these distances (between the
+previously presented word's feature values, $a$, and the candidate word's
+feature values, $b$) to similarity scores: \begin{equation}
+\mathrm{similarity}(a, b) = \exp\{-\tau \cdot \mathrm{distance}(a, b)\},
+\label{eqn:similarity} \end{equation} where $\tau = 1$ in our implementation.
+We note that increasing the value of $\tau$ would amplify the influence of
+similarity on order, and decreasing the value of $\tau$ would diminish the
+influence of similarity on order. Also note that this approach requires $\tau >
+0$. Finally, we computed a set of normalized similarity values by dividing the
+similarities by their sum: \begin{equation}
+\mathrm{similarity}_{\mathrm{normalized}}(a, b) = \frac{\mathrm{similarity}(a,
+b)}{\sum_{i=1}^n \mathrm{similarity}(a, i)}, \end{equation} where in the
+demoniator, $i$ takes on each of the $n$ feature values of the to-be-presented
+words. The resulting set of normalized similarity scores sums to one.
+
+\begin{figure}[tp]
+    \centering
+        \includegraphics[width=\textwidth]{figures/stick}
+
+\caption{\textbf{Generating stochastic feature-sorted lists.} For a given
+feature dimension (e.g., color), we compute the similarity
+(Eqn.~\ref{eqn:similarity}) between the feature value(s) of the previous item,
+$x$, and all yet-to-be-presented items ($a$ -- $g$). Next, we normalize these
+similarity scores so that they sum to one. We lay in sequence a set of
+``sticks,'' one for each candidate item, whose lengths are equal to these
+normalized similarity scores. Note that the combined lengths of these sticks is
+one. To select the next to-be-presented item, we draw a random number, $r$,
+from the uniform distibution bounded between 0 and 1 (inclusive). The identity
+of the next item is given by the stick adjacent to an indicator that moves
+distance $r$ (starting from 0) along the sequence of sticks. In this case, the
+next to-be-presented item is $e$. Note that each item's chances of selection is
+proportional to its similarity to the previous item, along the given feature
+dimension.} \label{fig:stick} 
+\end{figure}
+
+As illustrated in Figure~\ref{fig:stick}, we use these normalized similarity
+scores to construct a sequence of ``sticks'' that we lay end to end in a line.
+Each of the $n$ sticks corresponds to a single to-be-presented word, and the
+stick lengths are proportional to the relative similarities between each word's
+feature value(s) and the feature value(s) of the just-presented word. We choose
+the next to-be-presented word by moving an indicator along the set of sticks,
+by a distance chosen uniformly at random on the interval $\left[0, 1\right]$.
+We select the word associated with the stick lying next to the indicator to be
+presented next. This process continues iteratively (re-computing the similarity
+scores and stochastically choosing the next to-be-presented word using the
+just-presented word) until all of the words have been presented. The result is
+an ordered list that tends to change gradually along the selected feature
+dimension.
 

### 2023-01-06 paper/main.tex
starting to work on description of the "adaptive" condition

@@
 
-\subsubsection*{Adaptive conditions}
+\subsubsection*{Adaptive condition}
+
+We designed the \textit{adaptive} experimental condition to study the effect on
+memory for information that matched (or mismatched) the ways participants
+``naturally'' organized their memories of the lists they studied. Like the
+other conditions, all participants in the adaptive condition studied a total of
+16 lists, in a randomized order. We varied the words' colors and locations for
+every word presentation, as in the feature rich and order manipulation
+conditions.
+
+All participants in the adaptive condition began the experiment by studying a
+set of four \textit{initialization} lists. Words and features on these lists
+were presented in a randomized order (computed independently for each
+participant). These initialization lists were used to estimate each
+participant's ``memory fingerprint,'' defined below. At a high level, a
+participant's memory fingerprint describes how they prioritize different
+semantic, lexicographic, and/or visual features when they organize their
+memories.
+
+Next, participants studied a sequence of 12 lists in three batches of 4 lists
+each. These batches came in three types: \textit{random}, \textit{stabilize},
+and \textit{destabilize}. The batch types determined how words on the lists in
+that batch were ordered. Lists in each batch were always presented
+consecutively (e.g., a participant might receive four random lists, followed by
+four stabilize lists, followed by four destabilize lists). The batch orders
+were evenly counterbalanced across participants: there are six possible
+orderings of the three batches, and 10 participants were randomly assigned to
+each ordering sub-condition.
+
+Lists in the random batches were sorted randomly (as on the initialization
+lists and in the feature rich condition). Lists in the stabilize and
+destabilize batches were sorted in ways that either matched or mismatched each
+participant's memory fingerprint, respectively. Our procedures for computing
+participants' memory fingerprints and ordering the stabilize and destabilize
+lists are described next.
+
+\paragraph*{Computing clustering scores and memory fingerprints.}
 
@@
 
-\subsubsection*{Computing clustering scores and memory fingerprints}
+\paragraph{Temporal clustering score.}
 

### 2023-01-06 paper/main.tex
filling in analysis descriptions...

@@
 
-\paragraph*{Computing clustering scores and memory fingerprints.}
+\paragraph*{Clustering scores and memory fingerprints.}
 
@@
 
-\subsection*{Analysis}
+\subsection*{Analyses}
 
-\subsubsection*{Probability of first recall and probability of $n$\textsuperscript{th} recall}
+\subsubsection*{Probability of $n$\textsuperscript{th} recall curves}
 
-\subsubsection*{Lag conditional response probability}
+Probability of first recall curves~\citep{AtkiShif68, PostPhil65, WelcBurn24}
+reflect the probability that an item will be recalled first, as a function of
+its serial position during encoding. To carry out this analysis, we initialized (for each participant)
+a number-of-lists (16) by number-of-words-per-list (16) matrix of zeros.
+Then, for each list, we found the index of the word that was recalled first, and we filled in that
+position in the matrix with a 1.  Finally, we averaged over the rows of the matrix to obtain a 1 by 16 array of
+probabilities, for each participant.  We used an analogous procedure to compute probabilility of $n$\textsuperscript{th}
+reacall curves for each participant.  Specifically, we filled in the corresponding matrices according to the $n$\textsuperscript{th}
+recall on each list that each participant made.  When a given participant had made fewer than $n$ recalls for a given list,
+we simply excluded that list from our analysis when computing that paritcipant's curve(s).
 
-\paragraph{Temporal clustering score.}
+\subsubsection*{Lag-conditional response probability curve}
+
+\paragraph*{Temporal clustering score.}
+
+\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
+
+\subsubsection{Serial position curve}
 
 \subsubsection*{Identifying event boundaries}
-
-\subsubsection*{Serial position curves and recall accuracy}
-
-\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
 

### 2023-01-06 paper/main.tex
added lag-CRP description

@@
 reflect the probability that an item will be recalled first, as a function of
-its serial position during encoding. To carry out this analysis, we initialized (for each participant)
-a number-of-lists (16) by number-of-words-per-list (16) matrix of zeros.
-Then, for each list, we found the index of the word that was recalled first, and we filled in that
-position in the matrix with a 1.  Finally, we averaged over the rows of the matrix to obtain a 1 by 16 array of
-probabilities, for each participant.  We used an analogous procedure to compute probabilility of $n$\textsuperscript{th}
-reacall curves for each participant.  Specifically, we filled in the corresponding matrices according to the $n$\textsuperscript{th}
-recall on each list that each participant made.  When a given participant had made fewer than $n$ recalls for a given list,
-we simply excluded that list from our analysis when computing that paritcipant's curve(s).
+its serial position during encoding. To carry out this analysis, we initialized
+(for each participant) a number-of-lists (16) by number-of-words-per-list (16)
+matrix of zeros. Then, for each list, we found the index of the word that was
+recalled first, and we filled in that position in the matrix with a 1. Finally,
+we averaged over the rows of the matrix to obtain a 1 by 16 array of
+probabilities, for each participant. We used an analogous procedure to compute
+probabilility of $n$\textsuperscript{th} reacall curves for each participant.
+Specifically, we filled in the corresponding matrices according to the
+$n$\textsuperscript{th} recall on each list that each participant made. When a
+given participant had made fewer than $n$ recalls for a given list, we simply
+excluded that list from our analysis when computing that paritcipant's
+curve(s).
 
 \subsubsection*{Lag-conditional response probability curve}
+
+The lag-conditional probability (lag-CRP) curve~\citep{Kaha96} reflects the
+probability of recalling a given item after the just-recalled item, as a
+function of their relative encoding positions (lag). In other words, a lag of 1
+indicates that a recalled item was presented immediately after the previously
+recalled item, and a lag of $−3$ indicates that a recalled item came 3 items
+before the previously recalled item. For each recall transition (following the
+first recall), we computed the lag between the just-recalled word's
+presentation position and the next-recalled word's presentation position. We
+computed the proportions of transitions (between successively recalled words)
+for each lag, normalizing for the total numbers of possible transitions. In
+carrying out this analysis, we excluded all incorrect recalls and successive
+repetitions (e.g., recalling the same word twice in a row).  This yielded, for each list,
+a 1 by number-of-lags ($-15$ to $+15$; 30 lags in total, excluding lags of 0) array of conditional probabilities.
+We averaged these probabilities across lists to obtain a single lag-CRP for each participant.
 

### 2023-01-09 paper/main.tex
backing up changes...
working on methods section:   - added descriptions of temporal clustering and SPC analyses

@@
 
-\paragraph*{Clustering scores and memory fingerprints.}
+\paragraph*{Clustering scores (uncorrected).}
+
+\paragraph*{Permutation-corrected clustering scores.}
+
+\paragraph*{Memory fingerprints.}
 
@@
 carrying out this analysis, we excluded all incorrect recalls and successive
-repetitions (e.g., recalling the same word twice in a row).  This yielded, for each list,
-a 1 by number-of-lags ($-15$ to $+15$; 30 lags in total, excluding lags of 0) array of conditional probabilities.
-We averaged these probabilities across lists to obtain a single lag-CRP for each participant.
+repetitions (e.g., recalling the same word twice in a row). This yielded, for
+each list, a 1 by number-of-lags ($-15$ to $+15$; 30 lags in total, excluding
+lags of 0) array of conditional probabilities. We averaged these probabilities
+across lists to obtain a single lag-CRP for each participant.
 
@@
 
+Temporal clustering describes a participant's tendency to organize their recall
+sequences by the learned items' encoding positions. For instance, if a
+participant recalled the episode events in the exact order they occurred (or in
+exact reverse order), this would yield a score of 1. If a participant recalled
+the events in random order, this would yield an expected score of 0.5. For each
+recall-event transition (and separately for each participant), we sorted all
+not-yet-recalled events according to their absolute lag (that is, distance away
+in the episode). We then computed the percentile rank of the next event the
+participant recalled. We took an average of these percentile ranks across all
+of the participant’s recalls to obtain a single (uncorrected) temporal clustering score for
+the participant.  Finally, as with the feature-based clustering scores, we applied
+our permutation correction procedure to obtain a corrected clustering score.
+
 \subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
 
-\subsubsection{Serial position curve}
+\subsubsection*{Serial position curve}
+
+Serial position curves~\citep{Murd62a} reflect the proportion of participants
+who remember each item as a function of the item’s serial position during
+encoding. For each participant, we initialized a number-of-lists (16) by number-of-words-per-list
+(16) matrix of zeros. Then, for each correct recall, we identified the
+presentation position of the word and entered a 1 into that position (row:
+list; column: presentation position) in the matrix. This resulted in a matrix
+whose entries indicated whether or not the words presented at each position, on
+each list, were recalled by the participant (depending on whether the
+corresponding entires were set to one or zero). Finally, we averaged over the
+rows of the matrix to yield a 1 by 16 array representing the proportion of
+words at each position that the participant remembered.
 

### 2023-01-10 paper/main.tex
added a description of feature clustering scores (uncorrected)

@@
 
-\paragraph*{Clustering scores (uncorrected).}
+\paragraph*{Feature clustering scores (uncorrected).}
 
-\paragraph*{Permutation-corrected clustering scores.}
+Feature clustering describes a participant’s tendency to recall similar
+presented items together in their recall sequences, where ``similarity''
+considers one given feature dimension (e.g., category, color, etc.). Computing
+the clustering score for one feature dimension starts by considering those
+feature values from the first word the participant recalled on the list. Next,
+we sort all not-yet-recalled words in ascending order according to their
+feature-based distance to the just-recalled item (see \textit{Defining
+feature-based distances}). We then compute the percentile rank of the observed
+next recall. We averaged these percentile ranks across all of the participant’s
+recalls for the current list to obtain a single uncorrected clustering score
+for the list, for the given feature dimension. We repeat this process for each
+feature dimension in turn to obtain a single uncorrected clustering score for
+each list, for each feature dimension.
+
+\paragraph*{Permutation-corrected feature clustering scores.}
 

### 2023-01-10 paper/main.tex
backing up methods additions...

@@
 
-Feature clustering describes a participant’s tendency to recall similar
+Feature clustering scores describe participants' tendencies to recall similar
 presented items together in their recall sequences, where ``similarity''
-considers one given feature dimension (e.g., category, color, etc.). Computing
-the clustering score for one feature dimension starts by considering those
-feature values from the first word the participant recalled on the list. Next,
-we sort all not-yet-recalled words in ascending order according to their
-feature-based distance to the just-recalled item (see \textit{Defining
-feature-based distances}). We then compute the percentile rank of the observed
-next recall. We averaged these percentile ranks across all of the participant’s
-recalls for the current list to obtain a single uncorrected clustering score
-for the list, for the given feature dimension. We repeat this process for each
-feature dimension in turn to obtain a single uncorrected clustering score for
-each list, for each feature dimension.
+considers one given feature dimension (e.g., category, color, etc.). We base
+our main approach to computing clustering scores on analogous temporal and semantic clustering scores
+developed by~\citep{PolyEtal09}. Computing the clustering score for one feature
+dimension starts by considering those feature values from the first word the
+participant recalled on the list. Next, we sort all not-yet-recalled words in
+ascending order according to their feature-based distance to the just-recalled
+item (see \textit{Defining feature-based distances}). We then compute the
+percentile rank of the observed next recall. We averaged these percentile ranks
+across all of the participant’s recalls for the current list to obtain a single
+uncorrected clustering score for the list, for the given feature dimension. We
+repeat this process for each feature dimension in turn to obtain a single
+uncorrected clustering score for each list, for each feature dimension.
 
 \paragraph*{Permutation-corrected feature clustering scores.}
+
+Suppose that two lists contain unequal numbers of items of each size. For
+example, suppose that list $A$ contains all ``large'' items, whereas list $B$
+contains an equal mix of ``large'' and ``small'' items. For a participant
+recalling list $A$, any correctly recalled item will necessarily match the size
+of the previous correctly recalled item. In other words, successively recalling
+several list $A$ items of the same size is essentially meaningless, since
+\textit{any} correctly recalled list $A$ word will be large. In contrast,
+successively recalling several list $B$ items \textit{could} be meaningful,
+since (early in the recall sequence) the yet-to-be-recalled items come from a
+mix of sizes. However, once all of the ``small'' items on list $B$ have been
+recalled, the best possible next matching recall will be a large item. And all
+subsequent correct recalls must also be large items-- so for those later
+recalls it becomes difficult to determine whether the participant is
+successively recalling ``large'' items because they are organizing their
+memories according to size, or (alternatively), whether they are simply
+recalling the yet-to-be-recalled items in a random order. In general, the
+precise order and blend of feature values expressed in a given list, the
+orders and numbers of correct recalls a participant makes, the number of itervening
+presentation positions between successive recalls, and so on, can all affect the
+range of clustering scores that are possible to observe for a given list.  The uncorrected
+clustering score therefore conflates participants' actual memory organization with
+other ``nuisance'' factors.
+
+Following our prior work~\citep{HeusEtal17}, we used a permutation-correction
+procedure to help isolate the behavioral aspects of clustering that we were
+most interested in. After computing the uncorrected clustering score (for the
+given list and observed recall sequence), we compute a ``null'' distribution of
+$n$ additional clustering scores after randomly shuffling the recall order (we
+use $n = 500$ in the present study). This null distribution represents an
+approximation of the range of clustering scores one might expect to observe by
+``chance,'' given that a hypothetical participant was \textit{not} truly
+clustering their recalls, but where the hypothetical participant studied and
+recalled exactly the same items (with the same features) as the true
+participant.  We define the permutation-corrected clustering score as the percentile rank of the
+observed uncorrected clustering score in this estimated null distribution.  In this way,
+a corrected score of 1 indicates that the observed score was greater than any clustering score
+one might expect by chance; in other words, good evidence that the participant was truly clustering their
+recalls along the given feature dimension.
 

### 2023-01-10 paper/main.tex
formatting tweaks...
finished a complete draft of the methods section   - note: the PDF isn't compiling correctly within VS Code...i'm not sure why...

@@
 \usepackage{lineno}
+\usepackage{amsopn}
 
@@
 screen. The red microphone icons during the ``Recall'' phase denote the one
-minute verbal recall interval.  The labels on the right (and corresponding groupings on the left)
-denote experimental condition labels.}
+minute verbal recall interval. The labels on the right (and corresponding
+groupings on the left) denote experimental condition labels.}
 
@@
 random (one for the word's horizontal $x$ coordinate and the other for its
-vertical $y$ coordinate).  The bounds of these coordinates were selected to cover the entire
-visible area of the display without cutting off any part of the words.  The words were shown
-on 27~in (diagonal) Retina 5K iMac displays (resolution: 5120 $\times$ 2880 pixels).
+vertical $y$ coordinate). The bounds of these coordinates were selected to
+cover the entire visible area of the display without cutting off any part of
+the words. The words were shown on 27~in (diagonal) Retina 5K iMac displays
+(resolution: 5120 $\times$ 2880 pixels).
 
@@
 dimension.} \label{fig:stick}
+
 \end{figure}
@@
 considers one given feature dimension (e.g., category, color, etc.). We base
-our main approach to computing clustering scores on analogous temporal and semantic clustering scores
-developed by~\citep{PolyEtal09}. Computing the clustering score for one feature
-dimension starts by considering those feature values from the first word the
-participant recalled on the list. Next, we sort all not-yet-recalled words in
-ascending order according to their feature-based distance to the just-recalled
-item (see \textit{Defining feature-based distances}). We then compute the
-percentile rank of the observed next recall. We averaged these percentile ranks
-across all of the participant’s recalls for the current list to obtain a single
-uncorrected clustering score for the list, for the given feature dimension. We
-repeat this process for each feature dimension in turn to obtain a single
-uncorrected clustering score for each list, for each feature dimension.
+our main approach to computing clustering scores on analogous temporal and
+semantic clustering scores developed by~\citep{PolyEtal09}. Computing the
+clustering score for one feature dimension starts by considering those feature
+values from the first word the participant recalled on the list. Next, we sort
+all not-yet-recalled words in ascending order according to their feature-based
+distance to the just-recalled item (see \textit{Defining feature-based
+distances}). We then compute the percentile rank of the observed next recall.
+We averaged these percentile ranks across all of the participant’s recalls for
+the current list to obtain a single uncorrected clustering score for the list,
+for the given feature dimension. We repeat this process for each feature
+dimension in turn to obtain a single uncorrected clustering score for each
+list, for each feature dimension.
+
+\paragraph*{Temporal clustering score (uncorrected).}
+
+Temporal clustering describes a participant's tendency to organize their recall
+sequences by the learned items' encoding positions. For instance, if a
+participant recalled the episode events in the exact order they occurred (or in
+exact reverse order), this would yield a score of 1. If a participant recalled
+the events in random order, this would yield an expected score of 0.5. For each
+recall-event transition (and separately for each participant), we sorted all
+not-yet-recalled events according to their absolute lag (that is, distance away
+in the episode). We then computed the percentile rank of the next event the
+participant recalled. We took an average of these percentile ranks across all
+of the participant’s recalls to obtain a single (uncorrected) temporal
+clustering score for the participant.
 
@@
 recalling the yet-to-be-recalled items in a random order. In general, the
-precise order and blend of feature values expressed in a given list, the
-orders and numbers of correct recalls a participant makes, the number of itervening
-presentation positions between successive recalls, and so on, can all affect the
-range of clustering scores that are possible to observe for a given list.  The uncorrected
-clustering score therefore conflates participants' actual memory organization with
-other ``nuisance'' factors.
+precise order and blend of feature values expressed in a given list, the orders
+and numbers of correct recalls a participant makes, the number of itervening
+presentation positions between successive recalls, and so on, can all affect
+the range of clustering scores that are possible to observe for a given list.
+The uncorrected clustering score therefore conflates participants' actual
+memory organization with other ``nuisance'' factors.
 
@@
 recalled exactly the same items (with the same features) as the true
-participant.  We define the permutation-corrected clustering score as the percentile rank of the
-observed uncorrected clustering score in this estimated null distribution.  In this way,
-a corrected score of 1 indicates that the observed score was greater than any clustering score
-one might expect by chance; in other words, good evidence that the participant was truly clustering their
-recalls along the given feature dimension.
+participant. We define the permutation-corrected clustering score as the
+percentile rank of the observed uncorrected clustering score in this estimated
+null distribution. In this way, a corrected score of 1 indicates that the
+observed score was greater than any clustering score one might expect by
+chance; in other words, good evidence that the participant was truly clustering
+their recalls along the given feature dimension.  We applied this correction procedure
+to all of the clustering scores (feature and temporal) reported in this paper.
 
@@
 
+We define each participant's \textit{memory fingerprint} as the set of their
+permutation-corrected clustering scores across all dimensions we tracked in our
+study, including their six feature-based clustering scores (category, size,
+length, first letter, color, and location) and their temporal clustering score.
+Conceptually, this memory fingerprint describes the participant's tendancies to
+order (and, presumably, organize in memory) the studied words along each
+dimension. To obtain stable estimates of these fingerprints for each
+participant, we averaged clustering scores across lists. We also tracked and
+characterized how participants' fingerprints changed across lists (e.g.,
+Figs.~\ref{fig:fingerprint-trajectories},~\fingerprintTrajectoryRandom).
+
 \paragraph{Online ``fingerprint'' analysis.}
 
-\paragraph{Ordering ``stabilize'' lists by an estimated fingerprint.}
+The presentation orders of some lists in the adaptive condition of our
+experiment (see \textit{Adaptive condition}) were sorted according to
+participants' \textit{current} memory fingerprint, estimated using all of the
+lists they had studied up to that point in the experiment. Because our
+experiment incorporated a speech-to-text component, all of the behavioral data
+for each participant could be analyzed just a few seconds after the conclusion
+of the recall intervals for each list. We used the \texttt{Quail} Python
+package~\citep{HeusEtal17} to apply speech-to-text alorithms to the just
+collected data, aggregate the data for the given participant, and estimate the
+participant's memory fingerprint using all of their available data up to that
+point in the experiment. Two aspects of our implementation are worth noting.
+First, because memory fingerprints are averaged across lists, the
+already-computed memory fingerprints for earliar lists could be cached and
+loaded as needed in future computations. This meant that our computations
+pertaining to updating our estimates of a participant's memory fingerprint only
+needed to consider data from the most recent list. Second, each element of the
+null distributions of uncorrected fingerprint scores (see
+\textit{Permutation-corrected feature clustering scores}) could be estimated
+independently from the others. This enabled us to make use of the testing
+computers' multi-core CPU archetectuers by elements of the null distributions
+in batches of eight (i.e., the number of CPU cores on each testing computer).
+Taken together, we were able to compress the fingerprint computations into just
+a few seconds of computing time. The combined processing time for the
+speech-to-text algorithm and fingerprint computations easily fit within the
+inter-list intervals, where participants typically paused before moving on to
+the next list.
 
-\paragraph{Ordering ``destabilize'' lists by an estimated fingerprint.}
+\paragraph{Ordering ``stabilize'' and ``destabilize'' lists by an estimated
+fingerprint.}
+
+In the adaptive condition of our experiment, the presentation orders for
+\textit{stabilize} and \textit{destabilize} lists were chosen to either
+maximally or minimally (respectively) comport with participants' memory
+fingerprints. Given a participant's memory fingerprint and a to-be-presented
+set of items, we designed a permutation-based procedure for ordering the items.
+First, we dropped from the participant's fingerprint the temporal clustering
+score. For the remaining feature dimensions, we arranged the clustering scores
+in the fingerprint into a template vector, $f$. Second, we computed $n = 2500$
+random permutations of the to-be-presented items. These permutations served as
+prospective presentation orders. We sought to select the specific order that
+most (or least) matched $f$. Third, for each random permutation, we computed
+the (permutation-corrected) ``fingerprint,'' treating the permutation as though
+it were a potential ``perfect'' recall sequence. (We did not include temporal
+clustering scores in these fingerprints.) This yielded a ``simulated fingerprint''
+vector, $\hat{f}_p$ for each permutation $p$.  We used these simulated fingerprints to
+select a specific permutation, $i$, that either maximized (for stabilized lists) or minimized (for destabilize lists)
+the correlation between $\hat{f}_i$ and $f$.
+
+\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
+
+Following some of our prior work~\citep{HeusEtal18a}, we use low-dimensional
+embeddings to help visualize how participants' memory fingerprints change
+across lists
+(Figs.~\ref{fig:fingerprint-trajectories}A,~\fingerprintTrajectoryRandom A). To
+compute a shared embedding space across participants and experimental
+conditions, we concatenated the full set of fingerprints (across all lists,
+participants, and experimental conditions) to create a large matrix with
+number-of-lists $\times$ number-of-participants rows and seven columns (one for
+each word feature dimension's clustering scores, plus an additional temporal
+clustering score column). We used principal components analysis to project the
+seven-dimensional observations into a two-dimensional space (using the two
+principal components that explained the most variance in the data). For two
+visualizations
+(Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B)
+we computed an additional set of two-dimensional embeddings for participants'
+\textit{average} fingerprints (i.e., across lists within a given group of
+lists-- early or late). For those visualizations we averaged each participant's
+rows (for the given group of lists) in the combined fingerprint matrix prior to
+projecting it into the shared two-dimensional space. This yielded a single
+two-dimensional coordinate for each \textit{participant} and \textit{list
+group}, rather than for each individual list. We used these embeddings solely
+for visualization. All statistical tests were carried out in the original
+(seven-dimensional) feature spaces.
 
@@
 
-\paragraph*{Temporal clustering score.}
-
-Temporal clustering describes a participant's tendency to organize their recall
-sequences by the learned items' encoding positions. For instance, if a
-participant recalled the episode events in the exact order they occurred (or in
-exact reverse order), this would yield a score of 1. If a participant recalled
-the events in random order, this would yield an expected score of 0.5. For each
-recall-event transition (and separately for each participant), we sorted all
-not-yet-recalled events according to their absolute lag (that is, distance away
-in the episode). We then computed the percentile rank of the next event the
-participant recalled. We took an average of these percentile ranks across all
-of the participant’s recalls to obtain a single (uncorrected) temporal clustering score for
-the participant.  Finally, as with the feature-based clustering scores, we applied
-our permutation correction procedure to obtain a corrected clustering score.
-
-\subsubsection*{Computing low-dimensional embeddings of memory fingerprints}
-
 \subsubsection*{Serial position curve}
@@
 who remember each item as a function of the item’s serial position during
-encoding. For each participant, we initialized a number-of-lists (16) by number-of-words-per-list
-(16) matrix of zeros. Then, for each correct recall, we identified the
-presentation position of the word and entered a 1 into that position (row:
-list; column: presentation position) in the matrix. This resulted in a matrix
-whose entries indicated whether or not the words presented at each position, on
-each list, were recalled by the participant (depending on whether the
-corresponding entires were set to one or zero). Finally, we averaged over the
-rows of the matrix to yield a 1 by 16 array representing the proportion of
-words at each position that the participant remembered.
+encoding. For each participant, we initialized a number-of-lists (16) by
+number-of-words-per-list (16) matrix of zeros. Then, for each correct recall,
+we identified the presentation position of the word and entered a 1 into that
+position (row: list; column: presentation position) in the matrix. This
+resulted in a matrix whose entries indicated whether or not the words presented
+at each position, on each list, were recalled by the participant (depending on
+whether the corresponding entires were set to one or zero). Finally, we
+averaged over the rows of the matrix to yield a 1 by 16 array representing the
+proportion of words at each position that the participant remembered.
 
 \subsubsection*{Identifying event boundaries}
+
+We used the distances between feature values for successively presented words
+(see \textit{Defining feature-based distances}) to estimate ``event
+boundaries'' where the feature values changed more than
+usual~\citep{EzzyDava11, MannEtal16,RadvCope06, SwalEtal09, SwalEtal11,
+DuBrDava16}. For each list, for each feature dimension, we computed the
+distribution of distances between the feature values for successively presented
+words. We defined event boundaries (e.g., Fig.~\ref{fig:recall-dynamics}B) as
+occuring between any successive pair of words whose distances along the given
+feature dimension were greater than one standard deviation above the mean for
+that list. Note that, because event boundaries are defined for each feature
+dimension, each individual list may contain several sets of event boundaries,
+each at different moments in the presentation sequence (depending on the
+feature dimension of interest).
 
@@
 
-    \caption{\textbf{Memory ``fingerprints'' (order manipulation conditions).}
-    The across-participant distributions of clustering scores for each feature
-    type ($x$-coordinate) are displayed for each experimental condition
-    (color), separately for order manipulation (early, top) and randomly
-    ordered (late, bottom) lists. See
-    Figures~\fingerprintsRandom~and~\fingerprintsAdaptive~for analogous plots
-    for the random (control) and adaptive conditions.}
+\caption{\textbf{Memory ``fingerprints'' (order manipulation conditions).} The
+across-participant distributions of clustering scores for each feature type
+($x$-coordinate) are displayed for each experimental condition (color),
+separately for order manipulation (early, top) and randomly ordered (late,
+bottom) lists. See Figures~\fingerprintsRandom~and~\fingerprintsAdaptive~for
+analogous plots for the random (control) and adaptive conditions.}
 \label{fig:fingerprints}
+
 \end{figure}
@@
 
-    \caption{\textbf{Feature clustering carryover effects.} \textbf{A.} Feature clustering scores for ordder manipulation (early) versus randomly ordered (late) lists.
-    \textbf{B.}  Accuracy differences (on early versus late lists) versus feature clustering ``carryover'' (defined as the differences between the average clustering scores on early and late lists).
-    \textbf{C.}  Temporal clustering differences (on early versus late lists) versus feature clustering carryover.  \textbf{All panels.} Each dot in the main scatterplots denotes the
-    average scores for one participant. The colored regression lines are computed
-    across participants. The inset displays condition-averaged results, where each
-    dot reflects a single condition and the regression line is computed across
-    experimental conditions. All error ribbons denote bootstrap-estimated 95\%
-    confidence intervals.} \label{fig:clustering-carryover}
+    \caption{\textbf{Feature clustering carryover effects.} \textbf{A.} Feature
+    clustering scores for ordder manipulation (early) versus randomly ordered
+    (late) lists. \textbf{B.} Accuracy differences (on early versus late lists)
+    versus feature clustering ``carryover'' (defined as the differences between
+    the average clustering scores on early and late lists). \textbf{C.}
+    Temporal clustering differences (on early versus late lists) versus feature
+    clustering carryover. \textbf{All panels.} Each dot in the main
+    scatterplots denotes the average scores for one participant. The colored
+    regression lines are computed across participants. The inset displays
+    condition-averaged results, where each dot reflects a single condition and
+    the regression line is computed across experimental conditions. All error
+    ribbons denote bootstrap-estimated 95\% confidence intervals.}
+    \label{fig:clustering-carryover}
+
 \end{figure}
@@
 
-        \caption{\textbf{Recall performance (adaptive conditons).} The bars display the average probability of recall (taken across words, lists, and participants)
-        for lists from each adaptive condition.  Error bars denote bootstrap-estimated 95\% confidence intervals.  For additional details about
-        participants' behavior and performance during the adaptive conditions, see Figure~\dynamicsAdaptive.}
+        \caption{\textbf{Recall performance (adaptive conditons).} The bars
+        display the average probability of recall (taken across words, lists,
+        and participants) for lists from each adaptive condition. Error bars
+        denote bootstrap-estimated 95\% confidence intervals. For additional
+        details about participants' behavior and performance during the
+        adaptive conditions, see Figure~\dynamicsAdaptive.}
 

### 2023-01-10 paper/main.tex
writing up some quick notes for results section (backing up progress)

@@
 
+- what do additonal visual features add?  (compare reduced vs. feature rich)
+- are the visual features "sticky"?  (compare feature rich vs. reduced (early), also reduced vs. reduced (early))
+- are impoverished stimuli "sticky"?  (compare feature rich vs. reduced (late), reduced vs. reduced (late), also reduced (early) vs. reduced (late))
+
+- are order effects "sticky"?  compare behavior on early vs. late lists for order manipulation condition
+  - does feature clustering on early lists correlate with recall on early (or late) lists?
+  - does feature clustering on late lists correlate with recall on early (or late) lists?
+  - (ditto, but replace "recall" with "temporal clustering")
+
+- for feature-rich lists, do order effects matter?
+  - recall + dynamics + organization on order-manipulation conditions vs. feature rich
+  - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
+
+- are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
+
 \begin{figure}[tp] \centering

### 2023-01-11 code/analyze.py
major code reorganization, started notebook for running stats

@@ -1,14 +1,26 @@
 import dill as pickle
+import numpy as np
+import pandas as pd
 
 import os
 import warnings
+import string
+import quail
 
 # noinspection PyProtectedMember
 from pathos.multiprocessing import ProcessingPool as Pool
 from multiprocessing import cpu_count
 from tqdm import tqdm
+from sklearn.decomposition import IncrementalPCA as PCA
 
-from dataloader import fetch_data, get_listgroups, datadir
+from dataloader import fetch_data, get_listgroups, datadir, feature_groupings
+
+results_file = 'analyzed_500_iter.pkl'
+
+random = ['feature rich', 'reduced (early)', 'reduced (late)', 'reduced']
+adaptive = ['adaptive']
+non_adaptive_exclude_random = ['category', 'size', 'length', 'first letter', 'color', 'location']
+
 
 def apply(egg, analysis, listgroup=None, **kwargs):
     warnings.simplefilter('ignore')
@@ -91,3 +103,455 @@ def analyze_data(analyses=['fingerprint', 'pfr', 'lagcrp', 'spc', 'accuracy'], d
             pickle.dump([results, analyses, listgroups], f)
 
     return results, analyses, listgroups
+
+
+results, analyses, listgroups = analyze_data(savefile=results_file)
+
+
+# hack for recovering fingerprint features (feature tags are lost in the pickling process)
+def recover_fingerprint_features(order_file, results):
+    if os.path.exists(order_file):
+        with open(order_file, 'rb') as f:
+            orders = pickle.load(f)
+    else:
+        data = fetch_data()
+        fingerprints = {x: d.analyze('fingerprint') for x, d in data.items()}
+        orders = {k: x.data.columns.tolist() for k, x in fingerprints.items()}
+
+        with open(order_file, 'wb') as f:
+            pickle.dump(orders, f)
+
+    for k, f in results['fingerprint'].items():
+        f.data.columns = orders[k]
+    
+    return orders
+
+
+def organize_by_listgroup(x, groups):
+    if type(x) is dict:        
+        return {k: organize_by_listgroup(v, groups[k]) if k in groups else organize_by_listgroup(v, groups) for k, v in x.items()}
+    
+    subjs = x.data.index.levels[0].values
+    lists = x.data.index.levels[1].values
+
+    if all([type(g) is list for g in groups]):
+        listgroups = {s: g for s, g in enumerate(groups)}
+    else:
+        listgroups = {s: groups for s in np.unique(subjs)}
+    
[…]

### 2023-01-11 code/dataloader.py

@@ -57,24 +57,6 @@ feature_groupings = {
     }
 
 
-colors = {
-    'feature rich': '#c0673c',
-    'reduced (early)': '#f15a22',
-    'reduced (late)': '#f7996c',
-    'reduced': '#fddac5',
-    'category': '#524fa1',
-    'size': '#aca7d3',
-    'length': '#cbdb2a',
-    'first letter': '#e8eeae',
-    'color': '#00a651',
-    'location': '#9bd3ae',
-    'random': '#407b8d',
-    'stabilize': '#00addc',
-    'destabilize': '#90d7ee',
-    'init': '#d3d3d3'
-}
-
-
 def grouping(feature):
     try:
         return [k for k, v in feature_groupings.items() if feature in v][0]

### 2023-01-11 code/plot.py

@@ -1,0 +1,561 @@
+import pandas as pd
+import numpy as np
+import seaborn as sns
+import dill as pickle
+
+import h5py
+import quail
+import os
+import warnings
+import string
+
+from tqdm import tqdm
+from matplotlib import pyplot as plt
+from matplotlib.ticker import MaxNLocator
+from sklearn.decomposition import IncrementalPCA as PCA
+
+
+figdir = os.path.join(os.path.split(os.getcwd())[0], 'paper', 'figures', 'source')
+if not os.path.exists(figdir):
+    os.makedirs(figdir)
+
+from dataloader import feature_groupings
+from analyze import analyze_data, filter, reorder_df, rename_features, mini_filter, sort_by_grouping, trajectorize, get_dists, listnum2group, results, analyses, listgroups
+
+
+colors = {
+    'feature rich': '#c0673c',
+    'reduced (early)': '#f15a22',
+    'reduced (late)': '#f7996c',
+    'reduced': '#fddac5',
+    'category': '#524fa1',
+    'size': '#aca7d3',
+    'length': '#cbdb2a',
+    'first letter': '#e8eeae',
+    'color': '#00a651',
+    'location': '#9bd3ae',
+    'random': '#407b8d',
+    'stabilize': '#00addc',
+    'destabilize': '#90d7ee',
+    'init': '#d3d3d3'
+}
+
+
+def combo_lineplot(x, include_conds='all', include_lists='all', fname=None, xlabel=None, ylabel=None, xlim=None, ylim=None, palette=None):
+    x, include_conds, include_lists = filter(x, include_conds, include_lists)
+
+    fig = plt.figure(figsize=(3, 2.5))
+    ax = plt.gca()
+    
+    for c in include_conds:
+        if palette is None:
+            if c.lower() in colors:
+                next_palette = [colors[c.lower()]]
+            else:
+                next_palette = [colors[i.lower()] for i in include_lists]
+        else:
+            next_palette = palette
+    
+        x[c].plot(ax=ax, palette=next_palette, legend=False)
+    
+    if xlabel is not None:
+        plt.xlabel(xlabel)
+    
+    if ylabel is not None:
+        plt.ylabel(ylabel)
+
+    if xlim is not None:
+        plt.xlim(xlim)
+    
+    if ylim is not None:
+        plt.ylim(ylim)
+    
+    ax.spines.right.set_visible(False)
+    ax.spines.top.set_visible(False)
+    ax.xaxis.set_major_locator(MaxNLocator(integer=True))
+
[…]

### 2023-01-11 paper/main.tex
typo fix in intro
major code reorganization, started notebook for running stats
- working on results section - added Campbell as an author (yay!)

@@
 Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
-Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, 
-Allison M Frantz\textsuperscript{1}, \\Bryan J.
-Bollinger\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
+Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Allison M
+Frantz\textsuperscript{1},\\Bryan J. Bollinger\textsuperscript{1}, Campbell E.
+Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
 3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
@@
 emotions that we bring with us into the present moment. These factors
-collectively define a \textit{context} for our experiences~\citep{Mann20}. %
-situation models: forming expectations, predicting ambiguous future experiences
+collectively define a \textit{context} for our experiences~\citep{Mann20}.
+
+% situation models: forming expectations, predicting ambiguous future experiences
+
 The contexts we encounter help us to construct \textit{situation
@@
 
+We sought to manipulate two aspects of how participants memorized sequences of
+word lists. First, we added two additional sources of visual variation to the
+individual word presentations: font color and onscreen location. Importantly,
+these visual features were independent of the meaning or semantic content of
+the words (e.g., word category, size of the refferent) and of the lexicographic
+properties of the word (e.g., word length, first letter). We wondered whether
+this additional word-independent information might facilitate recall (e.g., by
+providing new potential ways of organizing or retrieving memories of the
+studied words) or impair recall (e.g., by distracting participants). Second,
+our primary experimental manipulations entailed manipulating the orders in
+which words were studied (and how those orderings changed over time). We
+wondered whether presenting the same list of words in different orders (e.g.,
+sorted along one feature dimension versus another) might serve to influence how
+participants organized their memories of the words. We also wondered whether
+some order manipulations might be temporally ``sticky'' by influencing how
+\textit{future} lists were remembered.
+
 - what do additonal visual features add?  (compare reduced vs. feature rich)

### 2023-01-11 paper/main.tex
backing up progress on results section: finished describing results from "random" conditions

@@
 \usepackage{geometry}
-\usepackage[sort&compress, numbers, super]{natbib}
+\usepackage[sort&compress]{natbib}
 \usepackage{pxfonts}
@@
 
-- what do additonal visual features add?  (compare reduced vs. feature rich)
-- are the visual features "sticky"?  (compare feature rich vs. reduced (early), also reduced vs. reduced (early))
-- are impoverished stimuli "sticky"?  (compare feature rich vs. reduced (late), reduced vs. reduced (late), also reduced (early) vs. reduced (late))
+To obtain a clean preliminary estimate of the consequences on memory of
+randomly varying the font colors and locations of presented words (versus
+holding the font color fixed at black, and holding the display locations fixed
+at the center of the display) we compared participants' performance on the
+\textit{feature rich} and \textit{reduced} experimental conditions (see
+\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature rich
+condition the words' colors and locations varied randomly across words, and in
+the reduced condition words were always presented in black, at the center of
+the display. Aggregating across all lists for each participant, we found no
+difference in recall accuracy for feature rich versus reduced lists ($t(126) =
+-0.290, p = 0.772$). However, participants in the feature rich condition
+clustered their recalls substantially more along every dimension we examined
+(temporal clustering: $t(126) = 10.624, p < 0.001$; category clustering:
+$t(126) = 10.077, p < 0.001$; size clustering: $t(126) = 11.829, p < 0.001$;
+word length clustering: $t(126) = 10.639, p < 0.001$; first letter clustering:
+$t(126) = 7.775, p = 0.000$; see \textit{Permutation-corrected feature
+clustering scores} for more information about how we quantified each
+participant's clustering tendencies.) Taken together, these comparisons suggest
+that adding new features changes how participants organize their memories of
+studied words, even when those new features are independent of the words
+themselves and even when the new features vary randomly across words. We found
+no evidence that those additional uninformative features were distracting (in
+terms of their impact on memory performance), but they did affect participants'
+recall dynamics (measured via their clustering scores).
+
+We also wondered whether adding these irrelevant visual features to later lists
+(after the participants had already studied impoverished lists), or removing
+the visual features from later lists (after the participants had already
+studied visually diverse lists) might affect memory performance. In other
+words, we sought to test for potential effects of changing the ``richness'' of
+participants' experiences over time. All participants studied and recalled a
+total of 16 lists; we defined \textit{early} lists as the first eight lists and
+\textit{late} lists as the last eight lists each participant encountered. To
+help interpret our results, we compared participants' memories on early versus
+late lists in the above feature rich and reduced conditions. Participants in
+both conditions remembered more words on early versus late lists (feature rich:
+$t(66) = 4.553, p < 0.001$; reduced: $t(60) = 2.434, p = 0.018$). Participants
+in the feature rich (but not reduced) conditions exhibited more temporal
+clustering on early versus late lists (feature rich: $t(66) = 2.318, p =
+0.024$; reduced: $t(60) = 0.929, p = 0.357$). And participants in both
+conditions exhibited more semantic (category and size) clustering on early
+versus late lists (feature rich, category: $t(66) = 3.805, p < 0.001$; feature
+rich, size: $t(66) = 2.190, p = 0.032$; reduced, category: $t(60) = 2.856, p =
+0.006$; reduced, size: $t(60) = 2.947, p = 0.005$). Participants in the reduced
+(but not feature rich) conditions exhibited more lexicographic clustering on
+early versus late lists (feature rich, word length: $t(66) = 0.161, p = 0.872$;
+feature rich, first letter: $t(66) = 0.410, p = 0.683$; reduced, word length:
+$t(60) = 3.528, p = 0.001$; reduced, first letter: $t(60) = 2.275, p = 0.026$).
+Taken together, these comparisons suggest that even when the presence or
+absence of irrelevant visual features is stable across lists, participants
+still exhibit some differences in their performance and memory organization
+tendencies for early versus late lists.
+
+With these differences in mind, we next compared participants' memories on
+early versus late lists for two additional experimental conditions (see
+\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
+(early)} condition, we held the irrelevant visual features constant on early
+lists, but allowed them to vary randomly on late lists. In a \textit{reduced
+(late)} condition, we allowed the irrelevant visual features to vary randomly
+on early lists, but held them constant on late lists. Given our above findings
+that (a) participants tended to remember more words and exhibit stronger
+clustering effects on feature rich (versus reduced) lists, and (b) participants
+tended to remember more words and exhibit stronger clustering effects on early
+(versus late) lists, we expected these differences to be enhance in the reduced
+(early) condition and diminished in the reduced (late) condition. However, to
+our surprise, participants in \textit{neither} condition exhibited reliable
+early versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
+0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
+(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
+= 0.278$), nor feature based clustering (reduced (early), category: $t(41) =
+0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
+(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
+$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
+0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
+length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
+0.966, p = 0.340$). We hypothesized that adding or removing the irrelevant
+features was acting as a sort of ``event boundary'' between early and late
+lists. In prior work, we (and others) have found that memories formed just
+after event boundaries can be enhanced~\citep[e.g.,][]{MannEtal16}.
+
+% JRM: STOPPED HERE...
 

### 2023-01-11 code/analyze.py

@@ -12,6 +12,7 @@ from pathos.multiprocessing import ProcessingPool as Pool
 from multiprocessing import cpu_count
 from tqdm import tqdm
 from sklearn.decomposition import IncrementalPCA as PCA
+from scipy import stats
 
 from dataloader import fetch_data, get_listgroups, datadir, feature_groupings
 
@@ -127,6 +128,10 @@ def recover_fingerprint_features(order_file, results):
     return orders
 
 
+order_file = os.path.join(datadir, 'scratch', 'feature_order.pkl')
+orders = recover_fingerprint_features(order_file, results)
+
+
 def organize_by_listgroup(x, groups):
     if type(x) is dict:        
         return {k: organize_by_listgroup(v, groups[k]) if k in groups else organize_by_listgroup(v, groups) for k, v in x.items()}
@@ -146,6 +151,10 @@ def organize_by_listgroup(x, groups):
 
     return quail.FriedEgg(data=data, analysis=x.analysis, list_length=x.list_length, n_lists=x.n_lists, n_subjects=x.n_subjects,
                           position=x.position)
+
+
+results_by_list = results  # per-list results
+results = {a: organize_by_listgroup(results_by_list[a], listgroups) for a in results_by_list.keys()}  # averaged within each listgroup
 
 
 def select_conds(results, conds='all'):
@@ -555,3 +564,38 @@ def get_boundaries(n_stddev):
             pickle.dump([boundaries, accuracy_near_boundaries], f)
     
     return boundaries, accuracy_near_boundaries
+
+
+def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_sample=True):
+    x = x.data
+    y = y.data
+
+    if x_lists is not None:
+        if type(x_lists) is not list:
+            x_lists = [x_lists]
+        x = x.query('List in @x_lists')
+    if y_lists is not None:
+        if type(y_lists) is not list:
+            y_lists = [y_lists]
+        y = y.query('List in @y_lists')
+    
+    x = x.groupby('Subject').mean()
+    y = y.groupby('Subject').mean()
+
+    if x_col is not None:
+        x = x[x_col]
+    if y_col is not None:
+        y = y[y_col]
+
+    if independent_sample: 
+        tfun = stats.ttest_ind
+        df = len(x) + len(y) - 2
+    else:
+        tfun = stats.ttest_rel 
+        df = len(x) - 1
[…]

### 2023-01-12 paper/main.tex
added some additional stats for "random" conditions   - all of the additions involved comparing reduced (early) and reduced (late) to the reduced and feature rich conditions.

@@
 tended to remember more words and exhibit stronger clustering effects on early
-(versus late) lists, we expected these differences to be enhance in the reduced
-(early) condition and diminished in the reduced (late) condition. However, to
-our surprise, participants in \textit{neither} condition exhibited reliable
-early versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
-0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
-(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
-= 0.278$), nor feature based clustering (reduced (early), category: $t(41) =
-0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
-(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
-$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
-0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
-length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
-0.966, p = 0.340$). We hypothesized that adding or removing the irrelevant
-features was acting as a sort of ``event boundary'' between early and late
-lists. In prior work, we (and others) have found that memories formed just
-after event boundaries can be enhanced~\citep[e.g.,][]{MannEtal16}.
+(versus late) lists, we expected these early versus late differences to be
+enhanced in the reduced (early) condition and diminished in the reduced (late)
+condition. However, to our surprise, participants in \textit{neither} condition
+exhibited reliable early versus late differences in accuracy (reduced (early):
+$t(41) = 1.499, p = 0.141$; reduced (late): $t(40) = 1.462, p = 0.152$),
+temporal clustering (reduced (early): $t(41) = 0.998, p = 0.324$; reduced
+(late): $t(40) = 1.099, p = 0.278$), nor feature based clustering (reduced
+(early), category: $t(41) = 0.753, p = 0.456$; reduced (early), size: $t(41) =
+0.721, p = 0.475$; reduced (early), length: $t(41) = 0.493, p = 0.625$; reduced
+(early), first letter: $t(41) = 0.780, p = 0.440$; reduced (late), category:
+$t(40) = -0.086, p = 0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$;
+reduced (late), length: $t(40) = 1.476, p = 0.148$; reduced (late), first
+letter: $t(40) = 0.966, p = 0.340$). We hypothesized that adding or removing
+the irrelevant features was acting as a sort of ``event boundary'' between
+early and late lists. In prior work, we (and others) have found that memories
+formed just after event boundaries can be enhanced~\citep[e.g., due to less
+contextual interference between pre- and post-boundary items;][]{MannEtal16}.
+
+We found that \textit{adding} irrelevant visual features on later lists that
+had not been present on early lists (as in the reduced (early) condition)
+served to enhance recall performance relative to conditions where all lists had
+the same blends of features (accuracy for feature rich versus reduced (early):
+$t(107) = -2.230, p = 0.028$; reduced versus reduced (early): $t(101) = -2.045,
+p = 0.043$; also see Fig.~\accuracyByList A). However, \textit{subtracting}
+irrelevant visual features on later lists that \textit{had} been present on
+early lists (as in the reduced (late) condition) did not appear to impact
+recall performance (accuracy for feature rich versus reduced (late): $t(106) =
+-0.638, p = 0.525$; reduced versus reduced (late): $t(100) = -0.407, p =
+0.685$). These comparisons suggest that recall accuracy has a directional
+component (i.e., accuracy is affected differently by removing features later
+that had been present earlier versus adding features later that had
+\textit{not} been present earlier). In contrast, we found that participants
+exhibited more temporal and feature-based clustering when we added irrelevant
+visual features to \textit{any} lists (comparisons of clustering on feature
+rich and reduced lists are reported above; temporal clustering in reduced
+versus reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
+-9.780$, $p$s $< 0.001$; feature based clustering in reduced versus reduced
+(early) and reduced versus reduced (late) conditions: $t$s $\leq -5.443$, $p$s
+$< 0.001$). Temporal and feature-based clustering were not reliably different
+in the feature rich, reduced (early), and reduced (late) conditions (temporal
+clustering in feature rich versus reduced (early) and feature rich versus
+reduced (late) conditions: $t$s $\geq -1.434$, $p$s $\geq 0.154$; feature based
+clustering in feature rich versus reduced (early) and feature rich versus
+reduced (late) conditions: $t$s $\geq -1.359$, $p$s $> 0.177$).
+
+Taken together, our findings thus far suggest that adding item features that
+change over time, even when they vary randomly and independently of the items,
+can enhance participants' overall memory performance and can also enhance
+temporal and feature-based clustering. To the extent that the number of item
+features that vary from moment to moment approximates the ``richness'' of
+participants' experiences, our findings suggest that participants remember
+``richer'' stimuli better and organize richer stimuli more reliably in their
+memories. Next, we turn to examine the memory effects of varying the temporal
+ordering of different stimulus features while holding the features themselves
+constant. We hypothesized that changing the order in which participants were
+exposed to the words on a given list might enhance (or diminish) the relative
+influence of different features. For example, presenting a set of words
+alphabetically might enhance participants' attention to the studied items'
+first letters, whereas sorting the same list of words by semantic category
+might instead enhance participants' attention to the words' semantic
+attributes. Importantly, we expected these order manipulations to hold even
+when the variation in the total set of features (across words) was held
+constant across lists (e.g., unlike in the reduced (early) and reduced (late)
+conditions, where visual features were added or removed from a subset of the
+lists participants studied).
 
@@
 
-- are order effects "sticky"?  compare behavior on early vs. late lists for order manipulation condition
+- when lists are sorted by a given feature, how is memory performance affected (relative to feature rich -- early lists only)?
+  - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- early lists only
+
+- are order effects "sticky"?
+  - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
+  - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
+
+  - compare behavior on early vs. late lists for order manipulation condition
     - does feature clustering on early lists correlate with recall on early (or late) lists?

### 2023-01-12 code/analyze.py
working on order manipulation stats (backing up progress)

@@ -599,3 +599,41 @@ def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_
         print(f't({df}) = {result.statistic[0]:.3f}, p = {result.pvalue[0]:.3f}')
     except IndexError:
         print(f't({df}) = {result.statistic:.3f}, p = {result.pvalue:.3f}')
+
+
+def stack_fried_eggs(*args):
+    if len(args) == 0:
+        return None
+    elif len(args) == 1:
+        return args[0]
+    
+    data = [args[0].data]
+    analysis = args[0].analysis
+    list_length = args[0].list_length
+    n_lists = args[0].n_lists
+    n_subjects = args[0].n_subjects
+    position = args[0].position
+
+    idx_vals = args[0].data.index.values
+
+    for i in range(1, len(args)):
+        next_idx_vals = args[i].data.index.values
+        next_idx_vals = [[x[0] + n_subjects, x[1]] for x in next_idx_vals]
+
+        data.append(pd.DataFrame(args[i].data.values, index=pd.MultiIndex.from_tuples(next_idx_vals, names=args[i].data.index.names), columns=args[i].data.columns))
+        n_subjects += args[i].n_subjects
+
+        assert args[i].analysis == analysis, 'Analysis must be the same for all arguments'
+        assert args[i].list_length == list_length, 'List length must be the same for all arguments'
+        assert args[i].n_lists == n_lists, 'Number of lists must be the same for all arguments'
+        assert args[i].position == position, 'Position must be the same for all arguments'
+
+    return quail.FriedEgg(data=pd.concat(data, axis=0), analysis=analysis, list_length=list_length, n_lists=n_lists, n_subjects=n_subjects, position=position)
+
+
+def merge_results(results, groups):
+    merged_results = {}
+    for g in groups.keys():
+        next_fried_eggs = [results[c] for c in groups[g]]
+        merged_results[g] = stack_fried_eggs(*next_fried_eggs)
+    return merged_results

### 2023-01-12 paper/main.tex
working on results section-- writing up order manipulation results (early lists)

@@
 
+Across six order manipulation conditions, we sorted early lists by each feature
+dimension but randomly ordered the items on late lists (see \textit{Order
+manipulation conditions}; features: category, size, length, first letter,
+color, and location). Participants in the category-ordered condition showed an
+increase in memory performance on early lists (accuracy, relative to early
+feature rich lists; $t(95) = 3.034, p = 0.003$). Participants in the
+color-ordered condition also showed a trending increase in memory performance
+on early lists (again, relative to early feature rich lists: $t(96) = 1.850, p
+= 0.067$). Participants' performance on early lists in all of the other order
+manipulation conditions was indistinguishable from performnace on the early
+feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$). Participants in both of
+the semanticly ordered conditions exhibited stronger temporal clustering on
+early lists (versus early feature rich lists; category: $t(95) = 8.508, p <
+0.001$; size: $t(95) = 2.429, p = 0.017$). Participants in the length-ordered
+condition tended to exhibit \textit{less} temporal clustering on early lists
+relative to early feature rich lists ($t(95) = -1.666, p = 0.099$), whereas
+participants in the first letter-ordered condition exhibited stronger temporal
+clustering on early lists ($t(95) = 2.587, p = 0.011$). Participants in the
+visually ordered conditions exhibited more similar performance on early lists,
+relative to early feature rich lists (color: $t(96) = -1.064, p = 0.290$; we
+found a trending enhancement for participants in the location-ordered
+condition: $t(95) = 1.682, p = 0.096$). We also compared feature-based
+clustering on early lists across the order manipulation and feature rich
+conditions. Since results were similar across both semantic conditoins
+(category and size), both lexicographic conditions (length and first letter),
+and both visual conditions (color and location), here we aggregate data from
+conditions that manipulated each of these three feature groupings in our
+comparisons to simplify the presentation. On early lists, participants in the
+semantically ordered conditions exhibited stronger semantic clustering relative
+to participants in the feature rich condition (category: $t(125) = 2.524, p =
+0.013$; size:$t(125) = 3.510, p = 0.001$), but showed no reliable differences
+in lexicographic (length: $t(125) = 0.539, p = 0.591$; first letter: $t(125) =
+-0.587, p = 0.558$) or visual (color: $t(125) = -0.579, p = 0.564$; location:
+$t(125) = -0.346, p = 0.730$) clustering. Similarly, participants in the
+lexicographically ordered conditions exhibited stronger (relative to feature
+rich participants) lexicographic clustering (length: $t(125) = 3.426, p =
+0.001$; first letter: $t(125) = 3.236, p = 0.002$) on early lists, but showed
+no reliable differences in semantic (category: $t(125) = -1.078, p = 0.283$;
+size: $t(125) = -0.310, p = 0.757$) or visual (color: $t(125) = -0.209, p =
+0.835$; location: $t(125) = -0.004, p = 0.997$) clustering. And participants in
+the visually ordered conditions exhibited stronger visual clustering (again,
+relative to feature rich participants, and on early lists; color: $t(126) =
+2.099, p = 0.038$; location: $t(126) = 4.392, p = 0.000$), but showed now
+reliable differences in semantic (category: $t(126) = 0.204, p = 0.839$; size:
+$t(126) = -0.093, p = 0.926$) or lexicographic (length: $t(126) = 0.714, p =
+0.476$; first letter: $t(126) = 0.820, p = 0.414$) clustering. Taken together,
+these order manipulation results suggest several broad patterns. First, most of
+the order manipulations we carried out did \textit{not} reliably affect overall
+recall performance. Second, most of the order manipulations increased
+participants' tendencies to temporally cluster their recalls. Third, all of the
+order manipulations enhanced participants' clustering of each condition's
+target feature (i.e., semantic manipulations enhanced semantic clustering,
+lexicographic manipulations enhanced lexicographic clustering, and visual
+manipulations enhanced visual clustering) while leaving clustering along other
+feature dimensions roughly unchanged (i.e., semantic manipulations did not
+affect lexicographic or color clustering, and so on).
+
 % JRM: STOPPED HERE...

### 2023-01-12 paper/main.tex
backing up progress on results section...

@@
 
+%%%%%
+% - next paragraph: relative strengths of order manipulations (compare order manipulations to *other order manipulations* instead of feature rich)
+
+% - and then, another paragraph: stickiness-- compare results on late lists (order manip vs. feature rich, order manip vs. other order manip).
+
+% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
+%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
+%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
+
 % JRM: STOPPED HERE...
 
-- when lists are sorted by a given feature, how is memory performance affected (relative to feature rich -- early lists only)?
-  - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- early lists only
-
-- are order effects "sticky"?
-  - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
-  - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
-
-  - compare behavior on early vs. late lists for order manipulation condition
-    - does feature clustering on early lists correlate with recall on early (or late) lists?
-    - does feature clustering on late lists correlate with recall on early (or late) lists?
-    - (ditto, but replace "recall" with "temporal clustering")
-
-- for feature-rich lists, do order effects matter?
-  - recall + dynamics + organization on order-manipulation conditions vs. feature rich
-  - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
-
-- are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
+% - are order effects "sticky"?
+%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
+%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
+
+%   - compare behavior on early vs. late lists for order manipulation condition
+%     - does feature clustering on early lists correlate with recall on early (or late) lists?
+%     - does feature clustering on late lists correlate with recall on early (or late) lists?
+%     - (ditto, but replace "recall" with "temporal clustering")
+
+% - for feature-rich lists, do order effects matter?
+%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
+%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
+
+% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
 

### 2023-01-12 paper/main.tex
backing up progress on results section   - finished writing up all early list (order manipulation) results   - finished draft of group-level late list results

@@
 lists participants studied).
+
+\begin{figure}[tp] \centering
+\includegraphics[width=\textwidth]{figures/recall_dynamics}
+
+\caption{\textbf{Recall dynamics in feature rich free recall (order
+manipulation conditions).} \textbf{A.} Behavioral plots. \textbf{Left panels.}
+The probabilities of initiating recall with each word are plotted as a function
+of presentation position. \textbf{Middle panels.} The conditional probabilities
+of recalling each word are plotted as a function of the relative position (Lag)
+to the words recalled just-prior. \textbf{Right panels.} The overall
+probabilities of recalling each word are plotted as a function of presentation
+position. \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\%
+confidence intervals (calculated across participants). Top panels display the
+recall dynamics for early (order manipulation) lists in each condition (color).
+Bottom panels display the recall dynamics for late (randomly ordered) lists.
+See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
+random (control) and adaptive conditions. \textbf{B.} Proportion of event
+boundaries (see \textit{Methods}) for each condition's feature of focus,
+plotted as a function of presentation position.}
+
+    \label{fig:recall-dynamics}
+\end{figure}
 
@@
 0.476$; first letter: $t(126) = 0.820, p = 0.414$) clustering. Taken together,
-these order manipulation results suggest several broad patterns. First, most of
-the order manipulations we carried out did \textit{not} reliably affect overall
+these order manipulation results suggest several broad patterns
+(Figs.~\ref{fig:recall-dynamics}, \ref{fig:fingerprints}). First, most of the
+order manipulations we carried out did \textit{not} reliably affect overall
 recall performance. Second, most of the order manipulations increased
@@
 
-%%%%%
-% - next paragraph: relative strengths of order manipulations (compare order manipulations to *other order manipulations* instead of feature rich)
-
-% - and then, another paragraph: stickiness-- compare results on late lists (order manip vs. feature rich, order manip vs. other order manip).
-
-% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
-%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
-%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
-
-% JRM: STOPPED HERE...
-
-% - are order effects "sticky"?
-%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
-%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
-
-%   - compare behavior on early vs. late lists for order manipulation condition
-%     - does feature clustering on early lists correlate with recall on early (or late) lists?
-%     - does feature clustering on late lists correlate with recall on early (or late) lists?
-%     - (ditto, but replace "recall" with "temporal clustering")
-
-% - for feature-rich lists, do order effects matter?
-%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
-%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
-
-% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
-
-\begin{figure}[tp] \centering
-\includegraphics[width=\textwidth]{figures/recall_dynamics}
-
-\caption{\textbf{Recall dynamics in feature rich free recall (order
-manipulation conditions).} \textbf{A.} Behavioral plots. \textbf{Left panels.}
-The probabilities of initiating recall with each word are plotted as a function
-of presentation position. \textbf{Middle panels.} The conditional probabilities
-of recalling each word are plotted as a function of the relative position (Lag)
-to the words recalled just-prior. \textbf{Right panels.} The overall
-probabilities of recalling each word are plotted as a function of presentation
-position. \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\%
-confidence intervals (calculated across participants). Top panels display the
-recall dynamics for early (order manipulation) lists in each condition (color).
-Bottom panels display the recall dynamics for late (randomly ordered) lists.
-See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
-random (control) and adaptive conditions. \textbf{B.} Proportion of event
-boundaries (see \textit{Methods}) for each condition's feature of focus,
-plotted as a function of presentation position.}
-
-    \label{fig:recall-dynamics}
-\end{figure}
-
-% figure: accuracy by list
-Figure~\accuracyByList.
-
-% figure: recall initiation
-Figure~\recallInit.
-
-% figure: clustering effects
 \begin{figure}[tp] \centering
@@
 \end{figure}
+
+The above comparisons between memory performance on early lists in the order
+manipulation versus feature rich conditions highlight how sorted lists are
+remembered differently from random lists. We also wondered how sorting lists
+along each feature dimension influenced memory relative to sorting lists along
+the other feature dimensions. Participants trended towards remembering early
+lists that were sorted semantically better than lexicographically sorted lists
+($t(118) = 1.936, p = 0.055$). Participants also remembered visually sorted
+lists better than lexicographically sorted lists ($t(119) = 2.145, p = 0.034$).
+However, participants showed no reliable differences in recall performance on
+semantically versus visually sorted lists ($t(119) = 0.113, p = 0.910$).
+Participants temporally clustered semantically sorted lists more strongly than
+either lexicographically ($t(118) = 5.572, p < 0.001$) or visually ($t(119) =
+6.215, p < 0.001$) sorted lists, but did not show reliable differences in
+temporal clustering on lexicographically versus visually sorted lists ($t(119)
+= 0.189, p = 0.850$). Participants also showed reliably more semantic
+clustering on semantically sorted lists than lexicographically (category:
+$t(118) = 3.492, p = 0.001$, size: $t(118) = 3.972, p < 0.001$) or visually
+(category: $t(119) = 2.702, p = 0.008$, size: $t(119) = 4.230, p < 0.001$)
+sorted lists; more lexicographic clustering on lexicographically sorted lists
+than semantically (length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) =
+3.686, p = 0.000$) or visually (length: $t(119) = 3.024, p = 0.003$; first
+letter: $t(119) = 2.644, p = 0.009$) sorted lists; and more visual clustering
+on visually sorted lists than semantically (color: $t(119) = -2.659, p =
+0.009$; location: $t(119) = -4.604, p = 0.000$) or lexicographically (color:
+$t(119) = -2.366, p = 0.020$; location: $t(119) = -4.265, p < 0.001$) sorted
+lists. In summary, sorting lists by different features appeared to have
+slightly different effects on overall memory performance and temporal
+clustering, and people tended to cluster their recalls along a given feature
+dimension more when the studied lists were (versus were not) sorted along that
+dimension.
+
+Beyond affecting how we process and remember \textit{ongoing} experiences, what
+is happening to us now can also affect how we process and remember
+\textit{future} experiences. Within the framework of our study, we wondered: if
+early lists are sorted along different feature dimensions, might this affect
+how people remember later (random) lists? In exploring this question, we
+considered both group-level effects (i.e., effects that tended to be common
+across individuals) and participant-level effects (i.e., effect that were
+idiosyncratic across individuals).
+
+At the group level, there seemed to be almost no lingering impact of sorting
+early lists on memory for later lists. To simplify the presentation, we report
+these null results in aggregate across the three feature groupings. Relative to
+memory performance on late feature rich lists, participants' memory performance
+in all six order manipulation conditions showed no reliable differences
+(semantic: $t(125) = 0.487, p = 0.627$; lexicographic: $t(125) = 0.878, p =
+0.382$; visual: $t(126) = 1.437, p = 0.153$). Nor did we observe any reliable
+differences in temporal clustering on late lists (relative to late feature rich
+lists; semantic: $t(125) = 0.146, p = 0.884$; lexicographic: $t(125) = 0.923, p
+= 0.358$; visual: $t(126) = 0.525, p = 0.601$). Aside from a slightly increased
+tendency for participants to cluster words by their length on late visual order
+manipulation lists (more than late feature rich lists; $t(126) = 2.199, p =
+0.030$), we observed no reliable differences in any type of feature clustering
+on late order manipulation condition lists versus late feature rich lists
+($\|t\|$s $\leq 1.234, p$s $\geq 0.220$).
+
+% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
+%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
+%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
+
+% JRM: STOPPED HERE...
+
+% - are order effects "sticky"?
+%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
+%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
+
+%   - compare behavior on early vs. late lists for order manipulation condition
+%     - does feature clustering on early lists correlate with recall on early (or late) lists?
+%     - does feature clustering on late lists correlate with recall on early (or late) lists?
+%     - (ditto, but replace "recall" with "temporal clustering")
+
+% - for feature-rich lists, do order effects matter?
+%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
+%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
+
+% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
+
+% figure: accuracy by list
+Figure~\accuracyByList.
+
+% figure: recall initiation
+Figure~\recallInit.
+
+% figure: clustering effects
 

### 2023-01-13 paper/main.tex
working on results section:   - exploring individual differences in links between feature clustering, accuracy, and temporal clustering on early and late lists in the order manipulation conditions  - added a new barplot to more directly illustrate group-level trends  - also added some across-condition stats
Note: i haven't yet added the new stats to the text

@@
 
+\begin{figure}[tp] \centering
+    \includegraphics[width=0.7\textwidth]{figures/memory_perf_barchart_compare}
+
+\caption{\textbf{Recall probability and clustering scores on early and late
+lists.} The bar heights display the average (across participants) recall
+probabilities (\textbf{A.}), temporal clustering scores (\textbf{B.}), and
+feature clustering scores (\textbf{C.}) for early (gray) and late (gold) lists.
+For the feature rich bars (left), the feature clustering scores are averaged
+across features. For the order manipulation conditions, feature clustering
+scores are displayed for the focused-on feature for each condition (e.g.,
+category clustering scores are displayed for the category condition, and so
+on). All panels: error bars denote bootstrap-estimated 95\% confidence
+intervals. The horizontal dotted lines denote the average values (across all
+lists and participants) for the feature rich condition.
+\textbf} \label{fig:barplots}
+
+\end{figure}
+
 At the group level, there seemed to be almost no lingering impact of sorting

### 2023-01-13 code/analyze.py

@@ -637,3 +637,33 @@ def merge_results(results, groups):
         next_fried_eggs = [results[c] for c in groups[g]]
         merged_results[g] = stack_fried_eggs(*next_fried_eggs)
     return merged_results
+
+
+def create_clustering_df(results):
+    conds = ['feature rich', 'category', 'size', 'length', 'first letter', 'color', 'location']
+    decoder = {'category': 'category', 'size': 'size', 'length': 'wordLength', 'first letter': 'firstLetter', 'color': 'color', 'location': 'location'}
+
+    dfs = []
+    
+    n_subjects = 0
+
+    for c in conds:
+        idx_vals = results['fingerprint'][c].data.index.values
+        idx_vals = [[x[0] + n_subjects, x[1]] for x in idx_vals]
+
+        x = pd.DataFrame(index=pd.MultiIndex.from_tuples(idx_vals, names=results['fingerprint'][c].data.index.names))
+        x['Condition'] = c.capitalize()
+
+        if c == 'feature rich':
+            x['Feature clustering score'] = results['fingerprint'][c].data[list(decoder.values())].mean(axis=1).values
+        else:
+            x['Feature clustering score'] = results['fingerprint'][c].data[decoder[c]].values
+        
+        x['Temporal clustering score'] = results['fingerprint'][c].data['temporal'].values
+        x['Recall probability'] = results['accuracy'][c].data.values
+
+        dfs.append(x)
+
+        n_subjects += results['accuracy'][c].n_subjects
+    
+    return pd.concat(dfs).reset_index()

### 2023-01-13 code/plot.py

@@ -37,7 +37,9 @@ colors = {
     'random': '#407b8d',
     'stabilize': '#00addc',
     'destabilize': '#90d7ee',
-    'init': '#d3d3d3'
+    'init': '#d3d3d3',
+    'early': '#bdccd4',
+    'late': '#fdb913'
 }
 
 
@@ -555,6 +557,38 @@ def plot_boundary_density_maps(conds, bounds, behaviors, listgroups, width=3, he
 
             plot_helper(bounds[cond][feature], behaviors[cond], listgroups[cond], ax=axes[row, column], behavioral_column=behavioral_column[row], xlabel=use_xlabel, ylabel=use_ylabel, title=title, xlim=xlim, ylim=ylim, cmap=cmap)
     
+    if fname is not None:
+        fig.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
+    
+    return fig
+
+
+def barplot_helper(clustering_results, x='Condition', y=None, hue=None, palette=None, ax=None, fname=None, width=8, height=3, ylim=None, ref=None):
+    fig = plt.figure(figsize=(width, height))
+    ax = plt.gca()
+
+    sns.barplot(data=clustering_results, x=x, y=y, hue=hue, palette=palette, ax=ax)
+
+    xlim = plt.xlim()
+    if ref is not None:
+        y = clustering_results.query('Condition == @ref')[y].mean()
+        plt.plot(xlim, [y, y], '--', color='black', linewidth=1)
+        plt.xlim(xlim)
+
+    if ylim is not None:
+        plt.ylim(ylim)
+    
+    plt.legend([],[], frameon=False)
+
+    ax.spines.right.set_visible(False)
+    ax.spines.top.set_visible(False)
+
+    xlabel = ax.get_xlabel()
+    ylabel = ax.get_ylabel()
+
+    plt.xlabel(xlabel, fontsize=14)
+    plt.ylabel(ylabel, fontsize=14)
+
     if fname is not None:
         fig.savefig(os.path.join(figdir, fname + '.pdf'), bbox_inches='tight')
     

### 2023-01-13 paper/main.tex
added a paragraph about event boundaries (within list) to the results section.  also starting to write up indiivdual differences results...

@@
 random (control) and adaptive conditions. \textbf{B.} Proportion of event
-boundaries (see \textit{Methods}) for each condition's feature of focus,
-plotted as a function of presentation position.}
+boundaries (see \textit{Identifying event boundaries}) for each condition's
+feature of focus, plotted as a function of presentation position.}
 
@@
 these order manipulation results suggest several broad patterns
-(Figs.~\ref{fig:recall-dynamics}, \ref{fig:fingerprints}). First, most of the
+(Figs.~\ref{fig:recall-dynamics}A, \ref{fig:fingerprints}). First, most of the
 order manipulations we carried out did \textit{not} reliably affect overall
@@
 \end{figure}
+
+When we closely examined the sequences of words participants recalled in early
+order manipulated lists (Fig.~\ref{fig:recall-dynamics}A, top panel), we
+noticed several differences from the dynamics of participants' recalls of
+randomly ordered lists (Fig.~\dynamicsRandom). One striking difference is that
+participants in the category condition (dark purple curves,
+Fig.~\ref{fig:recall-dynamics}) most often initiated recall with the
+fourth-from-last item (\textit{Recall initiation}, top left panel), whereas
+participants who recalled randomly ordered lists tended to initiate recall with
+either the first or last list items (Fig.~\dynamicsRandom, top left panel). We
+hypothesized that the participants might be ``clumping'' their recalls into
+groups of items that shared category labels. Indeed, when we compared the
+positions of feature changes in the study sequence
+(Fig.~\ref{fig:recall-dynamics}B; see \textit{Identifying event boundaries})
+with the positions of items participants recalled first, we noticed a striking
+correspondence in both semantic conditions. Specifically, on category-ordered
+lists, the category labels changed every four items on average (dark purple
+peaks in Fig.~\ref{fig:recall-dynamics}B), and participants also seemed to
+display an increased tendency (relative to other order manipulation and random
+conditions) to initiate recall of category-ordered lists with items whose study
+positions were integer multiples of four. Similarly, for size-ordered lists,
+the size labels changed every eight items on average (light purple peaks in
+Fig.~\ref{fig:recall-dynamics}B), and participants also seemed to display an
+icnreased tendancy to initiate recall of size-ordered lists with items whose
+study positions were integer multiples of eight. A second striking difference
+is that participants in the category condition exhibited a much steeper lag-CRP
+(Fig.~\ref{fig:recall-dynamics}A, top middle panel) than participants in other
+conditions. (This is another expression of participants' increased tendencies
+to temporally cluster their recalls on category-ordered lists, as we reported
+above.) Taken together, these order-specific idiosyncracies suggest a
+hierarchical set of influences on participants' memories. At longer timescales,
+``event boundaries'' (to use the term loosely) can be induced across lists by
+adding or removing irrelevant visual features. At shorter timescales, ``event
+boundaries'' can be induced across items (within a single list) by adjusting
+how item features change throughout the list.
 
@@
 
-% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
-%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
-%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
-
-% JRM: STOPPED HERE...
-
-% - are order effects "sticky"?
-%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
-%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
-
-%   - compare behavior on early vs. late lists for order manipulation condition
-%     - does feature clustering on early lists correlate with recall on early (or late) lists?
-%     - does feature clustering on late lists correlate with recall on early (or late) lists?
-%     - (ditto, but replace "recall" with "temporal clustering")
-
-% - for feature-rich lists, do order effects matter?
-%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
-%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
-
-% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
-
-% figure: accuracy by list
-Figure~\accuracyByList.
-
-% figure: recall initiation
-Figure~\recallInit.
-
-% figure: clustering effects
-
-% figure: correlations between clustering scores
-Figure~\clusterCorrs.
-
-% figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
 \begin{figure}[tp] \centering
@@
 
+When we examined the data at the level of individual participants
+(Fig.~\ref{fig:clustering-scatterplots}), a clearer story emerged.
+
+% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
+%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
+%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
+
+% JRM: STOPPED HERE...
+
+% - are order effects "sticky"?
+%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
+%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
+
+%   - compare behavior on early vs. late lists for order manipulation condition
+%     - does feature clustering on early lists correlate with recall on early (or late) lists?
+%     - does feature clustering on late lists correlate with recall on early (or late) lists?
+%     - (ditto, but replace "recall" with "temporal clustering")
+
+% - for feature-rich lists, do order effects matter?
+%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
+%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
+
+% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
+
+% figure: accuracy by list
+Figure~\accuracyByList.
+
+% figure: recall initiation
+Figure~\recallInit.
+
+% figure: clustering effects
+
+% figure: correlations between clustering scores
+Figure~\clusterCorrs.
+
+% figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
+
 % figure: fingerprint trajectories

### 2023-01-13 paper/main.tex
backing up description of fig 6 results

@@
 When we examined the data at the level of individual participants
-(Fig.~\ref{fig:clustering-scatterplots}), a clearer story emerged.
+(Fig.~\ref{fig:clustering-scatterplots}), a clearer story emerged. Within each
+order manipulation condition, participants exhibited a range of feature
+clustering scores, on both early and late lists
+(Fig.~\ref{fig:clustering-scatterplots}A, B). Across every order manipulation
+condition, participants who exhibited stronger feature clustering (for their
+condition's manipulated feature) recalled more words. This trend held overall
+across conditions and participants (early: $r(179) = 0.537, p < 0.001$; late: $
+r(179) = 0.492, p = 0.000$) as well as for each condition individually for
+early ($r$s $\geq 0.386$, all $p$s $\leq 0.035$) and late ($r$s $\geq 0.462$,
+all $p$s $\leq 0.010$) lists. We found no evidence of a condition-level trend;
+for example the conditions where participants tended to show stronger
+clustering scores were not correlated with the conditions where participants
+remembered more words (early: $r(4) = 0.526, p = 0.284$; late: $r(4) = -0.257,
+p = 0.623$; see insets of panels A and B). We observed carryover associations
+between feature clustering and recall performance
+(Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who showed stronger
+feature clustering on early lists tended to recall more items on late lists
+(across conditions: $r(179) = 0.492, p < 0.001$; all conditions individually:
+$r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who recalled more items
+on early lists also tended to show stronger feature clustering on late lists
+(across conditions: $r(179) = 0.280, p < 0.001$; all non-visual conditions:
+$r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29) = 0.298, p = 0.103$;
+location: $r(28) = 0.354, p = 0.055$). Neither of these effects showed
+condition-level trends (early feature clustering versus late recall
+probability: $r(4) = -0.299, p = 0.565$; early recall probability versus late
+feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for associations
+between feature clustering and temporal clustering. Across every order
+manipulation condition, participants who exhibited stronger feature clustering
+also exhibited stronger temporal clustering. For early lists
+(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}E), this trend held
+overall ($r(179) = 0.924, p < 0.001$), for each condition individually (all
+$r$s $\geq 0.822$, all $p$s $< 0.001$), and across conditions ($r(4) = 0.964, p
+= 0.002$). For late lists
+(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}F), the results were more
+variable (overall: $r(179) = 0.348, p = 0.000$; all non-visual conditions: $r$s
+$\geq 0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453, p = 0.011$;
+location: $ r(28) = 0.190, p = 0.314$; across-conditions: $r(4) = -0.036, p =
+0.945$). While less robust than the carryover associations between feature
+clustering and recall performance, we also observed some carryover associations
+between feature clustering and temporal clustering
+(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}G, H). Participants who
+showed stronger feature clustering on early lists trended towards showing
+stronger temporal clustering on later lists (overall: $r(179) = 0.301, p <
+0.001$; for individual conditions: all $r$s $\geq 0.297$, all $p$s $\leq
+0.111$; across conditions: $ r(4) = 0.107, p = 0.840$). And participants who
+showed stronger temporal clustering on early lists trended towards showing
+stronger feature clustering on later lists (overall: $r(179) = 0.579, p <
+0.001$; all non-visual conditions: $r$s $\geq 0.323$, all $p$s $\leq 0.082$;
+visual conditions: $r$s $\geq 0.089$, all $p$s $\leq 0.632$; across conditions:
+$r(4) = 0.916, p = 0.010$).
 

### 2023-01-13 paper/main.tex
finished writing up individual difference results (up through fig. 8).   - note: i moved the "fingerprint trajectories" figure (now fig. 6) earlier in the text, since i thought it fit better in the "group results" descriptions

--- 2023-01-13 paper/main.tex
@@
 noticed several differences from the dynamics of participants' recalls of
-randomly ordered lists (Fig.~\dynamicsRandom). One striking difference is that
+randomly ordered lists (Figs.~\dynamicsRandom,~\recallInit). One striking difference is that
 participants in the category condition (dark purple curves,
@@
 
+We also looked for more subtle group-level patterns. For example, perhaps
+sorting early lists by one feature dimension could affect how participants
+cluster \textit{other} features (on early and/or late lists) as well. We
+defined participants' \textit{memory fingerprints} as the set of temporal and
+feature clustering scores. A participant's memory fingerprint describes how
+they tend to retrieve memories of the studied items, perhaps searching through
+several feature spaces (or along several representational dimensions). To gain
+insights into the dynamics of how participants' clustering scores tended to
+change over time, we computed the average (across participants) fingerprint
+from each list, from each order manipulation condition
+(Fig.~\ref{fig:fingerprint-trajectories}). We projected these fingerprints into
+a two-dimensional space to help visualize the dynamics (top panels; see
+\textit{Computing low-dimensional embeddings of memory fingerprints}). We found
+that participants' average fingerprints tended to remain relatively stable on
+early lists, and exhibited a ``jump'' to another stable state on later lists.
+The sizes of these jumps varied somewhat across conditions (the Euclidean
+distances between fingerprints in their original high dimensional spaces are
+displayed in the bottom panels). We also averaged the fingerprints across early
+and late lists, respectively, for each condition
+(Fig.~\ref{fig:fingerprint-trajectories}B). We found that participants'
+fingerprints on early lists seem to be influenced by the order manipulations on
+those lists (see the locations of the circles in
+Fig.~\ref{fig:fingerprint-trajectories}B). There also seemed to be some
+consistency across different features within a broader type. For example, both
+semantic feature conditions (category and size; purple markers) diverge in a
+similar direction from the group; both lexicographic feature conditions (length
+and first letter; yellow markers) diverge in a similar direction; and both
+visiual conditions (color and location; green) also diverge in a similar
+direction. But on late lists, participants' fingerprints seem to return to a
+common state that is roughly shared across conditions (i.e., the stars in that
+panel are clumped together).
+
+\begin{figure}[tp] \centering
+    \includegraphics[width=\textwidth]{figures/fingerprint_trajectories}
+
+    \caption{\textbf{Memory fingerprint dynamics (order manipulation
+    conditions).} \textbf{A.} Each column (and color) reflects an experimental
+    condition. In the top panels, each marker displays a 2D projection of the
+    (across-participant) average memory fingerprint for one list. Order
+    manipulation (early) lists are denoted by circles and randomly ordered
+    (late) lists are denoted by stars. All of the fingerprints (across all
+    conditions and lists) are projected into a common space. The bar plots in
+    the bottom panels display the Euclidean distances of the per-list memory
+    fingerprints to the list 0 fingerprint, for each condition. Error bars
+    denote bootstrap-estimated 95\% confidence intervals. The dotted vertical
+    lines denote the boundaries between early and late lists. \textbf{B.} In
+    this panel, the fingerprints for early (circle) and late (star) lists are
+    averaged across lists and participants before projecting the fingerprints
+    into a (new) 2D space. See Figure~\fingerprintTrajectoryRandom~for
+    analogous plots for the random (control) conditions. }
+    \label{fig:fingerprint-trajectories}
+
+    \end{figure}
+
+When we examined the data at the level of individual participants
+(Figs.~\ref{fig:clustering-scatterplots} and \ref{fig:clustering-carryover}), a
+clearer story emerged. Within each order manipulation condition, participants
+exhibited a range of feature clustering scores, on both early and late lists
+(Fig.~\ref{fig:clustering-scatterplots}A, B). Across every order manipulation
+condition, participants who exhibited stronger feature clustering (for their
+condition's manipulated feature) recalled more words. This trend held overall
+across conditions and participants (early: $r(179) = 0.537, p < 0.001$; late: $
+r(179) = 0.492, p = 0.000$) as well as for each condition individually for
+early ($r$s $\geq 0.386$, all $p$s $\leq 0.035$) and late ($r$s $\geq 0.462$,
+all $p$s $\leq 0.010$) lists. We found no evidence of a condition-level trend;
+for example the conditions where participants tended to show stronger
+clustering scores were not correlated with the conditions where participants
+remembered more words (early: $r(4) = 0.526, p = 0.284$; late: $r(4) = -0.257,
+p = 0.623$; see insets of panels A and B). We observed carryover associations
+between feature clustering and recall performance
+(Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who showed stronger
+feature clustering on early lists tended to recall more items on late lists
+(across conditions: $r(179) = 0.492, p < 0.001$; all conditions individually:
+$r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who recalled more items
+on early lists also tended to show stronger feature clustering on late lists
+(across conditions: $r(179) = 0.280, p < 0.001$; all non-visual conditions:
+$r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29) = 0.298, p = 0.103$;
+location: $r(28) = 0.354, p = 0.055$). Neither of these effects showed
+condition-level trends (early feature clustering versus late recall
+probability: $r(4) = -0.299, p = 0.565$; early recall probability versus late
+feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for associations
+between feature clustering and temporal clustering. Across every order
+manipulation condition, participants who exhibited stronger feature clustering
+also exhibited stronger temporal clustering. For early lists
+(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}E), this trend held
+overall ($r(179) = 0.924, p < 0.001$), for each condition individually (all
+$r$s $\geq 0.822$, all $p$s $< 0.001$), and across conditions ($r(4) = 0.964, p
+= 0.002$). For late lists
+(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}F), the results were more
+variable (overall: $r(179) = 0.348, p = 0.000$; all non-visual conditions: $r$s
+$\geq 0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453, p = 0.011$;
+location: $ r(28) = 0.190, p = 0.314$; across-conditions: $r(4) = -0.036, p =
+0.945$). While less robust than the carryover associations between feature
+clustering and recall performance, we also observed some carryover associations
+between feature clustering and temporal clustering
+(Fig.~\ref{fig:clustering-scatterplots}G, H). Participants who
+showed stronger feature clustering on early lists trended towards showing
+stronger temporal clustering on later lists (overall: $r(179) = 0.301, p <
+0.001$; for individual conditions: all $r$s $\geq 0.297$, all $p$s $\leq
+0.111$; across conditions: $ r(4) = 0.107, p = 0.840$). And participants who
+showed stronger temporal clustering on early lists trended towards showing
+stronger feature clustering on later lists (overall: $r(179) = 0.579, p <
+0.001$; all non-visual conditions: $r$s $\geq 0.323$, all $p$s $\leq 0.082$;
+visual conditions: $r$s $\geq 0.089$, all $p$s $\leq 0.632$; across conditions:
+$r(4) = 0.916, p = 0.010$). Taken together, the results displayed in
+Figure~\ref{fig:clustering-scatterplots} show that participants who were more
+sensitive to the order manipulations (i.e., participants who showed stronger
+feature clustering for their condition's feature on early lists) remembered
+more words and showed stronger temporal clustering. These associations also
+appeared to carry over across lists, even when the items on later lists were
+presented in a random order.
+
 \begin{figure}[tp] \centering
@@
 
-When we examined the data at the level of individual participants
-(Fig.~\ref{fig:clustering-scatterplots}), a clearer story emerged.
-
-% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
-%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
-%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
-
-% JRM: STOPPED HERE...
-
-% - are order effects "sticky"?
-%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
-%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
-
-%   - compare behavior on early vs. late lists for order manipulation condition
-%     - does feature clustering on early lists correlate with recall on early (or late) lists?
-%     - does feature clustering on late lists correlate with recall on early (or late) lists?
-%     - (ditto, but replace "recall" with "temporal clustering")
-
-% - for feature-rich lists, do order effects matter?
-%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
-%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
-
-% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
-
-% figure: accuracy by list
-Figure~\accuracyByList.
-
-% figure: recall initiation
-Figure~\recallInit.
-
-% figure: clustering effects
-
-% figure: correlations between clustering scores
-Figure~\clusterCorrs.
-
-% figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
-
-% figure: fingerprint trajectories
-\begin{figure}[tp] \centering
-    \includegraphics[width=\textwidth]{figures/fingerprint_trajectories}
-
-    \caption{\textbf{Memory fingerprint dynamics (order manipulation
-    conditions).} \textbf{A.} Each column (and color) reflects an experimental
-    condition. In the top panels, each marker displays a 2D projection of the
-    (across-participant) average memory fingerprint for one list. Order
-    manipulation (early) lists are denoted by circles and randomly ordered
-    (late) lists are denoted by stars. All of the fingerprints (across all
-    conditions and lists) are projected into a common space. The bar plots in
-    the bottom panels display the Euclidean distances of the per-list memory
-    fingerprints to the list 0 fingerprint, for each condition. Error bars
-    denote bootstrap-estimated 95\% confidence intervals. The dotted vertical
-    lines denote the boundaries between early and late lists. \textbf{B.} In
-    this panel, the fingerprints for early (circle) and late (star) lists are
-    averaged across lists and participants before projecting the fingerprints
-    into a (new) 2D space. See Figure~\fingerprintTrajectoryRandom~for
-    analogous plots for the random (control) conditions. }
-    \label{fig:fingerprint-trajectories}
-
-    \end{figure}
-
-% figure: carryover effects: 
-%  - clustering early vs. clustering late
-%  - clustering carryover vs. accuracy difference
-%  - clustering carryover vs. temporal clustering differences
+If participants show different sensitivities to order manipulations, how do
+their behaviors carry over to later lists? We found that participants who
+showed strong feature clustering on early lists often tended to show strong
+feature clustering on late lists (Fig.~\ref{fig:clustering-carryover}A; overall
+across participants and conditions: $r(179) = 0.592, p < 0.001$; non-visual
+feature conditions: all $r$s $\geq 0.350$, all $p$s $\leq 0.058$; color: $r(29)
+= -0.071, p = 0.704$; location: $r(28) = 0.032, p = 0.868$; across conditions:
+$r(4) = 0.934, p = 0.006$). Although participants tended to show weaker feature
+clustering on late lists (Fig.~\ref{fig:fingerprint-trajectories}) on
+\textit{average}, the associations between early and late lists for individual
+participants suggests that some influence of early order manipulations may
+linger on late lists. We found that participants who exhibited larger carryover
+in feature clustering (i.e., continued to show strong feature clustering on
+late lists) for the semantic order manipulations (but not other manipulations)
+also tended to show a larger improvement in recall
+(Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.378, p < 0.001$;
+category: $r(28) = 0.419, p = 0.021$; size: $r(28) = 0.737, p < 0.001$;
+non-semantic conditions: all $r$s $\leq 0.252$, all $p$s $\geq 0.179$; across
+conditions: $r(4) = 0.773, p = 0.072$) on late lists, relative to early lists.
+Participants who exhibited larger carryover in feature clustering also tended
+to show stronger temporal clustering on late lists (relative to early lists)
+for all but the category condition (Fig.~\ref{fig:clustering-carryover}C;
+overall: $r(179) = 0.434, p < 0.001$; category: $r(28) = 0.229, p = 0.223$; all
+non-category conditions: all $r$s $\geq 0.448$, all $p$s $\leq 0.012$; across
+conditions: $r(4) = 0.598, p = 0.210$). We suggest two potential
+interpretations of these findings. First, it is possible that some participants
+are more ``maliable'' or ``adaptable'' with respect to how they organize
+incoming information. When presented with list of items sorted along
+\textit{any} feature dimension, they will simply adopt that feature as a
+dominant dimension for organizing those items and subsequent (randomly ordered)
+items. This flexibility in memory organization might afford such participants a
+memory advantage, explaining their strong recall performance. An alternative
+interpretation is that each participant comes into our study with a
+``preferred'' way of organizing incoming information. If they happen to be
+assigned to an order manipulation condition that matches their preferences,
+then they will appear to be ``sensitive'' to the order manipulation and also
+exhibit a high degree of carryover in feature clustering from early to late
+lists. These participants might demonstrate strong recall performance not
+because of their inherantly superior memory abilities, but rather because the
+specific condition they were assigned to happened to be especially easy for
+them, given their pre-experimental tendancies.
+
 \begin{figure}[tp] \centering
@@
 
+% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
+%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
+%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
+
+% JRM: STOPPED HERE...
+
+% - are order effects "sticky"?
+%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
+%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
+
+%   - compare behavior on early vs. late lists for order manipulation condition
+%     - does feature clustering on early lists correlate with recall on early (or late) lists?
+%     - does feature clustering on late lists correlate with recall on early (or late) lists?
+%     - (ditto, but replace "recall" with "temporal clustering")
+
+% - for feature-rich lists, do order effects matter?
+%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
+%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
+
+% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
+
+% figure: accuracy by list
+Figure~\accuracyByList.
+
+% figure: recall initiation
+Figure~\recallInit.
+
+% figure: clustering effects
+
+% figure: correlations between clustering scores
+Figure~\clusterCorrs.
+
+% figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
+
+% figure: fingerprint trajectories
+
+% figure: carryover effects: 
+%  - clustering early vs. clustering late
+%  - clustering carryover vs. accuracy difference
+%  - clustering carryover vs. temporal clustering differences
+
 \begin{figure} 

### 2023-01-13 paper/main.tex
finished complete draft of results section (with all figures and stats)

@@
 non-category conditions: all $r$s $\geq 0.448$, all $p$s $\leq 0.012$; across
-conditions: $r(4) = 0.598, p = 0.210$). We suggest two potential
-interpretations of these findings. First, it is possible that some participants
-are more ``maliable'' or ``adaptable'' with respect to how they organize
-incoming information. When presented with list of items sorted along
-\textit{any} feature dimension, they will simply adopt that feature as a
-dominant dimension for organizing those items and subsequent (randomly ordered)
-items. This flexibility in memory organization might afford such participants a
-memory advantage, explaining their strong recall performance. An alternative
-interpretation is that each participant comes into our study with a
-``preferred'' way of organizing incoming information. If they happen to be
-assigned to an order manipulation condition that matches their preferences,
-then they will appear to be ``sensitive'' to the order manipulation and also
-exhibit a high degree of carryover in feature clustering from early to late
-lists. These participants might demonstrate strong recall performance not
-because of their inherantly superior memory abilities, but rather because the
-specific condition they were assigned to happened to be especially easy for
-them, given their pre-experimental tendancies.
+conditions: $r(4) = 0.598, p = 0.210$).
 
@@
 
-% - the key: individual differences.  some participants are more affected by order manips than others.  for those participants, order manips are sticky and affect recall behavior more.
-%   - one subtlety: explain fig 5 and 7 scatterplots (and inlays)-- some *conditions* are overall more affective; pattern is more complicated than can be identified with pairwise comparisons
-%   - implication re: individual diffs: adaptive condition-- stabilize should improve performance, destabilize should worsen performance
-
-% JRM: STOPPED HERE...
-
-% - are order effects "sticky"?
-%   - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
-%   - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only
-
-%   - compare behavior on early vs. late lists for order manipulation condition
-%     - does feature clustering on early lists correlate with recall on early (or late) lists?
-%     - does feature clustering on late lists correlate with recall on early (or late) lists?
-%     - (ditto, but replace "recall" with "temporal clustering")
-
-% - for feature-rich lists, do order effects matter?
-%   - recall + dynamics + organization on order-manipulation conditions vs. feature rich
-%   - fingerprint trajectories: how much do fingerprints change over time, are they sensitive to order manipulations?
-
-% - are fingerprints maleable?  how does match between fingerprint + presentation order affect recall performance (adaptive conditions)?
-
-% figure: accuracy by list
-Figure~\accuracyByList.
-
-% figure: recall initiation
-Figure~\recallInit.
-
-% figure: clustering effects
-
-% figure: correlations between clustering scores
-Figure~\clusterCorrs.
-
-% figure: feature clustering vs. accuracy, feature clustering vs. temporal clustering
-
-% figure: fingerprint trajectories
-
-% figure: carryover effects: 
-%  - clustering early vs. clustering late
-%  - clustering carryover vs. accuracy difference
-%  - clustering carryover vs. temporal clustering differences
+We suggest two potential interpretations of these findings. First, it is
+possible that some participants are more ``maliable'' or ``adaptable'' with
+respect to how they organize incoming information. When presented with list of
+items sorted along \textit{any} feature dimension, they will simply adopt that
+feature as a dominant dimension for organizing those items and subsequent
+(randomly ordered) items. This flexibility in memory organization might afford
+such participants a memory advantage, explaining their strong recall
+performance. An alternative interpretation is that each participant comes into
+our study with a ``preferred'' way of organizing incoming information. If they
+happen to be assigned to an order manipulation condition that matches their
+preferences, then they will appear to be ``sensitive'' to the order
+manipulation and also exhibit a high degree of carryover in feature clustering
+from early to late lists. These participants might demonstrate strong recall
+performance not because of their inherantly superior memory abilities, but
+rather because the specific condition they were assigned to happened to be
+especially easy for them, given their pre-experimental tendancies. To help
+distinguish between these interpretations, we designed an \textit{adaptive}
+experimental condition (see \textit{Adaptive condition}). The primary
+manipulation in the adaptive condition is that participants each experience
+three key types of lists. On \textit{random} lists, words are ordered randomly
+(as in the feature rich condition). On \textit{stabilize} lists, the
+presentation order is adjusted to be maximally similar to the current estimate
+of the participant's memory fingerprint (see \textit{Online “fingerprint”
+analysis}). Third, on \textit{destabilize} lists, the presentation is adjusted
+to be \textit{minimally} similar to the current estimate of the participant's
+memory fingerprint (see \textit{Ordering ``stabilize'' and ``destabilize''
+lists by an estimated fingerprint}). The orders in which participants
+experienced each type of list were counterbalanced across participants to help
+reduce the influence of potential list order effects. Because the presentation
+orders on stabilize and destabilize lists are adjusted to best match each
+participant's (potentially unique) memory fingerprint, the adaptive condition
+removes uncertainty about whether participants' assigned conditions might just
+``happen'' to match their preferred ways or organizing their memories.
 
@@
 
-    \includegraphics[width=0.4 \textwidth]{figures/source/accuracy_adaptive}
-
-        \caption{\textbf{Recall performance (adaptive conditons).} The bars
-        display the average probability of recall (taken across words, lists,
-        and participants) for lists from each adaptive condition. Error bars
-        denote bootstrap-estimated 95\% confidence intervals. For additional
-        details about participants' behavior and performance during the
-        adaptive conditions, see Figure~\dynamicsAdaptive.}
+    \includegraphics[width=\textwidth]{figures/adaptive_results}
+
+        \caption{\textbf{Adaptive free recall.} \textbf{A.} Average probability
+        of recall (taken across words, lists, and participants) for lists from
+        each adaptive condition. \textbf{B.} Average temporal clustering scores
+        for lists from each adaptive condition. \textbf{C.} Recall probability
+        versus temporal clustering scores by participant (main panel; each
+        participant contributes one dot per condition) and averaged within
+        condition (inset; each dot represents a single condition). All panels:
+        Error bars and ribbons denote bootstrap-estimated 95\% confidence
+        intervals. For additional details about participants' behavior and
+        performance during the adaptive conditions, see
+        Figure~\dynamicsAdaptive.}
 
@@
 \end{figure}
+
+Overall, we found that participants tended to be better at remembering words on
+stabilize lists relative to words on random ($t(59) = 1.740, p = 0.087$) or
+destabilize ($t(59) = 1.714, p = 0.092$) lists (Fig.~\ref{fig:adaptive}A).
+Participants showed no reliable differences in their memory performance on
+destabilize versus random lists ($t(59) = -0.249, p = 0.804$). Participants
+also exhibited stronger temporal clustering on stabilize lists, relative to
+random ($t(59) = 3.554, p = 0.001$) and destabilize ($t(59) = 4.045, p <
+0.001$) lists (Fig.~\ref{fig:adaptive}B). We found no reliable differences in
+temporal clustering for items on random versus destabilize lists ($t(59) =
+-0.781, p = 0.438$).
+
+As in the other experimental manipulations, participants in the adaptive
+condition exhibited substantial variability with respect to their overall
+memory performance and their clustering tendencies (Fig.~\ref{fig:adaptive}C).
+We found that individual participants who exhibited strong temporal clustering
+scores also tended to recall more items. This held across subjects, aggregating
+across all list types ($r(178) = 0.721, p < 0.001$), and for each list type
+individually (all $r$s $\geq 0.683$, all $p$s $\leq 0.001$). Taken together,
+the results from the adaptive condition suggest that each participant comes
+into the experiment with their own unique memory organization tendencies, as
+characterized by their memory fingerprint. When participants study lists whose
+items come pre-sorted according to their unique preferences, they tend to remember
+more and show stronger temporal clustering.
 

### 2023-01-14 paper/main.tex
added a new panel to figure 9, showing how fingerprints become more stable on "stabilize" and "random" lists, and less stable on "destabilize" lists

@@
         participant contributes one dot per condition) and averaged within
-        condition (inset; each dot represents a single condition). All panels:
+        condition (inset; each dot represents a single condition). \textbf{D.}
+        Per-list correlations between the current list's fingerprint and the
+        average fingerprint computed from all previous lists. The normalized
+        list numbers ($x$-axis) denote the number of lists of the same type
+        that the participant had experienced at the time of the current list.
+        All panels: Colors denote the sorting type (condition) for each list.
         Error bars and ribbons denote bootstrap-estimated 95\% confidence
@@
 
+Participants' fingerprints on stabilize and random lists tended to become
+(numerically) slightly more similar to their average fingerprints computed from the
+previous lists they had experienced, and their fingerprints on destabilize
+lists tended to become numerically less similar (Fig.~\ref{fig:adaptive}D).
 Overall, we found that participants tended to be better at remembering words on

### 2023-01-15 paper/main.tex
working on discussion, added some text and notes

@@
 
-% connections to prior work: context effects, priming, situation models
-
-% implications for adaptive learning, education and training
+We asked participants to study and freely recall word lists. The words on each
+list (and the set of lists) were held constant across participants. For each
+word, we considered (and manipulated) two semantic features (category and size)
+that reflected aspects of the \textit{meanings} of the words, along with two
+lexicographic features (word length and first letter), which reflected aspects
+of the words' \textit{letters}. These semantic and lexicographic features are
+intrinsic to each word. We also considered and manipulated two additional
+visual features (color and location) that affected the \textit{appearance} of
+each studied item, but could be varied independently of the words' identities.
+Across different experimental conditions, we manipulated how the visual
+features varied across words (within each list), along with the orders of each
+list's words. Although participants' task (verbally recalling as many words as
+possible, in any order, within one minute) remained constant across all of
+these conditions, and although the set of words they studied on each list
+remained constant, our manipulations substantially affected participants'
+memories. The impact of some of the manipulations also affected how
+participants remembered \textit{future} lists that were sorted randomly.
+
+% specific findings:
+%   reduced (early/late) vs. feature rich and reduced: event boundaries
+%   order manipulations: lingering effects
+%   adaptive condition: people have unique preferences, and matching them can improve their memory and affect how the remembered info is organized
+
+% connections to prior work: 
+%     - context effects: 
+%     - priming: 
+%     - situation models:
+
+% implications for:
+%     - theory: flexibility in how we organize information, bridge between "classic" and naturalistic experiments
+%     - applications: adaptive learning, education and training
+
+% concluding remarks:
+%    - what are the limits of human learning?
+how much does what we remember depend on how we experience it?
+our expectations, strategies, situation models, etc.
+shape how our experiences are remembered.
+%    - but those aspects of our memory are not fixed: when we are exposed to the same experience in a new way, it can change how we remember it and how we
+%      remember *future* experiences
 

### 2023-01-15 paper/main.tex
quick round of edits through end of methods

@@
 
-\title{Carryover effects in free recall reveal how prior experiences influence
-memories of new experiences}
+\title{Carryover effects in free recall reveal how past experiences influence
+memories of future experiences}
 
@@
 \begin{abstract} We perceive, interpet, and remember ongoing experiences
-through the lens of our prior experiences. Inferring that we are one type of
+through the lens of our prior experiences. Inferring that we are in one type of
 situation versus another can lead us to interpret the same physical experience
 differently. In turn, this can affect how we focus our attention, form
-expectations of what will happen next, remember what is happening now, draw on
+expectations about what will happen next, remember what is happening now, draw on
 our prior related experiences, and so on. To study these phenomena, we asked
@@
 served food. Priming someone to expect a particular situation or context can
-also influence how they resolve potentail ambiguities in their ongoing
+also influence how they resolve potential ambiguities in their ongoing
 experiences, including ambiguous movies and narratives~\citep{YeshEtal17}.
@@
 using more traditional paradigms like free recall of random word
-lists~\citep{Kaha12}. In free recall, participants study lists of items and are
+lists~\citep{Kaha12, Kaha20}. In free recall, participants study lists of items and are
 instructed to recall the items in any order they choose. The orders in which
@@
 requires distinguishing the current list from the rest of one's experience. To
-model this fundamental memory capability, cognitive scientists have posited the
-existence of a special representation, called \emph{context}, that is
-associated with each list. According to early
-theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are
+model this fundamental memory capability, cognitive scientists have posited a
+special context representation that is associated with each list. According to
+early theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are
 composed of many features which fluctuate from moment to moment, slowly
@@
 \textit{free recall}, where the retrieval cue is ``context'' itself.
+Conceptually, the same general processes might be said to describe how
+real-world contexts evolve during natural experiences. However, this is still
+an open area of study~\citep{Mann20, Mann21a}.
 
@@
 and other list-learning tasks~\citep{Este55a, RaaiShif80, GlenEtal83,
-HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, PolyEtalTulv, SedeEtal08,
-PolyEtal09, ShanHowa12}. These phenomena include the well-known recency and
-primacy effects (superior recall of items from the end and, to a lesser extent,
-from the beginning of the study list), as well as semantic and temporal
-clustering effects~\citep{KahaEtal08}. The contiguity effect is an example of
-temporal clustering, which is perhaps the dominant form of organization in free
-recall. This effect can be seen in the tendency for people to successively
-recall items that occupied neighboring positions in the study list. For
-example, if a list contained the sub-sequence \textsc{``absence hollow pupil''}
-and the participant recalls the word \textsc{``hollow''}, it is far more likely
-that the next response will be either \textsc{``pupil''} or
-\textsc{``absence''} than some other list item~\citep{Kaha96}. In addition,
-there is a strong forward bias in the contiguity effect: subjects make forward
-transitions (i.e., \textsc{``hollow''} followed by \textsc{``pupil''}) about
-twice as often as they make backward transitions, despite an overall tendency
-to begin recall at the end of the list. There are also striking effects of
-semantic clustering~\citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52,
-MannKaha12}, whereby the recall of a given item is more likely to be followed
-by recall of a similar or related item than a dissimilar or unrelated one. In
-general, people organize memories for words along a wide variety of stimulus
-dimensions. As captured by models like the \textit{Context Maintenance and
-Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each
-word (e.g.\ the word's meaning, font size, font color, location on the screen,
-size of the object the word represents, etc.) are incorporated into the
-participant's mental context representation~\citep{SmitVela01, MannEtal15,
-Mann20, MannEtal11, MannEtal12}. During a memory test, any of these features
-may serve as a memory cue, which in turn leads the participant to recall in
-succession words that share stimulus features.
+HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, SedeEtal08, PolyEtal09,
+ShanHowa12}. These phenomena include the well-known recency and primacy effects
+(superior recall of items from the end and, to a lesser extent, from the
+beginning of the study list), as well as semantic and temporal clustering
+effects~\citep{KahaEtal08b}. The contiguity effect is an example of temporal
+clustering, which is perhaps the dominant form of organization in free recall.
+This effect can be seen in the tendency for people to successively recall items
+that occupied neighboring positions in the study list~\citep{Kaha96}. There are
+also striking effects of semantic clustering~\citep{RomnEtal93, Bous53,
+BousEtal54, JenkRuss52, MannKaha12}, whereby the recall of a given item is more
+likely to be followed by recall of a similar or related item than a dissimilar
+or unrelated one. In general, people organize memories for words along a wide
+variety of stimulus dimensions. As formalized by models like the \textit{Context
+Maintenance and Retrieval Model}~\citep{PolyEtal09}, the stimulus features
+associated with each word (e.g.\ the word's meaning, font size, font color,
+location on the screen, size of the object the word represents, etc.) are
+incorporated into the participant's mental context
+representation~\citep{SmitVela01, MannEtal15, Mann20, MannEtal11, MannEtal12}.
+During a memory test, any of these features may serve as a memory cue, which in
+turn leads the participant to recall in succession words that share stimulus
+features.
 
@@
 
-A key mystery is whether the sorts of situation models and schemas that people
+A key mystery is whether (and how) the sorts of situation models and schemas that people
 use to organize their memories of real-world experiences might map onto the
@@
 lists. On one hand, situation models and clustering effects both reflect
-statistical regularities in ongoing experience. Our memory systems exploit
+statistical regularities in ongoing experiences. Our memory systems exploit
 these regularities when generating inferences about the unobserved past and
@@
 participants' recall behaviors on early lists to manipulate, in real-time, the
-presentation orders of subsequent lists. In this adaptive condition, we sought
-to identify potential commonalities within and across participants in how
-people organized their memories and how those organizational tendancies affect
-overall performance.
+presentation orders of subsequent lists. In this adaptive condition we varied
+the agreement between how participants preferred to organize their memories of
+the studied items versus the orders in which the items were presented.
 
@@
 Participants received course credit for enrolling in our study. We asked each
-participant to fill out a demographic survey that included information about
-their self-reported age, gender, ethnicity, race, education, vision, reading
+participant to fill out a demographic survey that included questions about
+their age, gender, ethnicity, race, education, vision, reading
 impairments, medications or recent injuries, coffee consumption on the day of
@@
 demographics survey were optional. One participant elected not to fill out any
-part of the demographic survey, and all other participants report some or all
-of their requested demographic information.
+part of the demographic survey, and all other participants answered some or all
+of the survey questions.
 
 We aimed to run (to completion) at least 60 participants in each of the two
-primary control conditions and in the adaptive condition. In all other
+primary control conditions and in the adaptive condition. In all of the other
 conditions we set a target enrollment of at least 30 participants. Because our
-data collection efforts were coordinated 12 researchers and multiple testing
-rooms and computers, it was not feasible for individual experimenters to know
-how many participants had been run in each experimental condition until the
-relevant databases were synchronized at the end of each working day. We also
-over-enrolled participants for each condition to help ensure that we met our
-minimum enrollment targets even if some participants dropped out of the study
-prematurely or did not show up for their testing session. This led us to exceed
-our target enrollments for several conditions.
+data collection procedures entailed the coordinated efforts of 12 researchers
+and multiple testing rooms and computers, it was not feasible for individual
+experimenters to know how many participants had been run in each experimental
+condition until the relevant databases were synchronized at the end of each
+working day. We also over-enrolled participants for each condition to help
+ensure that we met our minimum enrollment targets even if some participants
+dropped out of the study prematurely or did not show up for their testing
+session. This led us to exceed our target enrollments for several conditions.
+Nevertheless, we analyze all viable data in the present paper.
 
@@
 490 participants who opted to fill out the demographics survey, reported ages
-ranged from 17 to 31 years (mean: 19.1; standard deviation: 1.356). A total of
-318 participants reported their gender as female, 170 as male, and 2
-participants declined to report their gender. A total of 442 participants
+ranged from 17 to 31 years (mean: 19.1 years; standard deviation: 1.356 years).
+A total of 318 participants reported their gender as female, 170 as male, and
+two participants declined to report their gender. A total of 442 participants
 reported their ethnicity as ``not Hispanic or Latino,'' 39 as ``Hispanic or
-Latino,'' and 9 declined to report their ethnicity. Participants reported their
-races as White (345 participants), Asian (120 participants), Black or African
-American (31 participants), American Indian or Alaska Native (11 particiapnts),
-Native Hawaiian or Other Pacific Islander (4 participants), Mixed race (3
-participants), Middle Eastern (1 participant), and Arab (1 participant). A
-total of 5 participants declined to report their race. We note that several
-participants reported more than one of racial category. Participants reported
-their highest degrees achieved as ``Some college'' (359 participants), ``High
-school graduate'' (117 participants), ``College graduate'' (7 participants),
-``Some high school'' (5 participants), ``Doctorate'' (1 participant), and
-``Master's degree'' (1 participant). A total of 482 participants reported no
-reading impairments, and 8 reported mild reading impairments such as mild
-dyslexia. A total of 489 participants reported having normal color vision and 1
-participant reported that they were color blind. A total of 482 participants
-reported taking no prescription medications and having no recent injuries; 4
-participants reported having ADHD, 1 reported having dyslexia, 1 reported
-having allergies, 1 reported a recently torn ACL/MCL, and 1 reported a
-concussion from several months prior. The participants reported consuming 0 --
-3 cups of coffee prior to the testing session (mean: 0.32 cups; standard
-deviation: 0.58 cups). Participants reported their current level of alertness,
-and we converted their responses to numerical scores as follows: ``very
-sluggish'' (-2), ``a little sluggish'' (-1), ``neutral'' (0), ``a little
-alert'' (1), and ``very alert'' (2). Across all participants, the full range of
-alertness levels were reported (range: -2 -- 2; mean: 0.35; standard deviation:
-0.89).
-
-We dropped from our dataset the 1 participant who reported abnormal color
-vision, as well as 39 participants whose data were corrupted due to technical
-failures while running the experiment or during the daily database merges. In
-total, this left usable data from 452 participants, broken down by experimental
-condition as follows: feature rich (67 participants), reduced (61
+Latino,'' and nine declined to report their ethnicity. Participants reported
+their races as White (345 participants), Asian (120 participants), Black or
+African American (31 participants), American Indian or Alaska Native (11
+particiapnts), Native Hawaiian or Other Pacific Islander (four participants),
+Mixed race (three participants), Middle Eastern (one participant), and Arab
+(one participant). A total of five participants declined to report their race.
+We note that several participants reported more than one of racial category.
+Participants reported their highest degrees achieved as ``Some college'' (359
+participants), ``High school graduate'' (117 participants), ``College
+graduate'' (seven participants), ``Some high school'' (five participants),
+``Doctorate'' (one participant), and ``Master's degree'' (one participant). A
+total of 482 participants reported no reading impairments, and eight reported
+having mild reading impairments. A total of 489 participants reported having
+normal color vision and one participant reported that they were red-green color
+blind. A total of 482 participants reported taking no prescription medications
+and having no recent injuries; four participants reported having ADHD, one reported
+having dyslexia, one reported having allergies, one reported a recently torn
+ACL/MCL, and one reported a concussion from several months prior. The
+participants reported consuming 0 -- 3 cups of coffee prior to the testing
+session (mean: 0.32 cups; standard deviation: 0.58 cups). Participants reported
+their current level of alertness, and we converted their responses to numerical
+scores as follows: ``very sluggish'' (-2), ``a little sluggish'' (-1),
+``neutral'' (0), ``a little alert'' (1), and ``very alert'' (2). Across all
+participants, the full range of alertness levels were reported (range: -2 -- 2;
+mean: 0.35; standard deviation: 0.89).
+
+We dropped from our dataset the one participant who reported having abnormal
+color vision, as well as 39 participants whose data were corrupted due to
+technical failures while running the experiment or during the daily database
+merges. In total, this left usable data from 452 participants, broken down by
+experimental condition as follows: feature rich (67 participants), reduced (61
 participants), reduced (late) (41 participants), reduced (early), (42
@@
 declined to fill out their demographic survey participated in the location
-condition, and we verified verbally that they had normal color vision.
+condition, and we verified verbally that they had normal color vision and no
+significant reading impairments.
 
@@
 incorporates a diverse mix of feature values along each dimension, this allows
-us to evaluate participants' memory fingerprints in rich detail.
+us to estimate which features participants are considering or leveraging in
+organizing their memories.
 
@@
 
-Stimuli in our paradigm were 256 English words selected in a previous
+The stimuli in our paradigm were 256 English words selected in a previous
 study~\citep{ZimaEtal18}. The words all referred to concrete nouns, and were
@@
 participants. For participants in some conditions, on some lists, we also
-randomly varied two additional visual features to each word: the presentation
-font color, and the word's onscreen location. These attributes were assigned
-independently for word (and for every participant) at the times the words were
-displayed onscreen. These visual features were varied for words in all lists
-and conditions except for the ``reduced'' condition (all lists), the first
-eight lists of the ``reduced (early)'' condition, and the last eight lists of
-the ``reduced (late)'' condition. In these latter cases, words were all
-presented in black at the center of the experimental computer's display.
-
-To assign a random font color to each word, we selected three integers
-uniformly and at random between 0 and 255, corresponding to the red (r), green
-(g), and blue (b) color channels for that word. To assign random presentation
-locations to each word, we selected two floating point numbers uniformly at
-random (one for the word's horizontal $x$ coordinate and the other for its
-vertical $y$ coordinate). The bounds of these coordinates were selected to
-cover the entire visible area of the display without cutting off any part of
+randomly varied two additional visual features associated with each word: the
+presentation font color, and the word's onscreen location. These attributes
+were assigned independently for each word (and for every participant). These
+visual features were varied for words in all lists and conditions except for
+the ``reduced'' condition (all lists), the first eight lists of the ``reduced
+(early)'' condition, and the last eight lists of the ``reduced (late)''
+condition. In these latter cases, words were all presented in black at the
+center of the experimental computer's display.
+
+To select a random font color for each word, we drew three integers uniformly
+and at random from the interval $\[0, 255\]$, corresponding to the red (r),
+green (g), and blue (b) color channels for that word. To assign random
+presentation locations to each word, we selected two floating point numbers
+uniformly at random (one for the word's horizontal $x$ coordinate and the other
+for its vertical $y$ coordinate). The bounds of these coordinates were selected
+to cover the entire visible area of the display without cutting off any part of
 the words. The words were shown on 27~in (diagonal) Retina 5K iMac displays
@@
 they studied (\textit{late} lists). Since every participant studied exactly 16
-lists, using this terminology every list was either ``early'' or ``late''
-depending on its order in the list study sequence.
+lists, every list was either ``early'' or ``late'' depending on its order in
+the list study sequence.
 
@@
 in real time-- a distinguishing feature of the experiment; in typical verbal
-recall experiments the audio data must be parsed manually. In prior work, we
-used a similar experimental setup (equivalent to the ``reduced'' condition in
-the present study) to verify that the automatically transcribed recalls were
-sufficiently close to human-transcribed recalls to yield reliable
+recall experiments the audio data must be parsed and transcribed manually. In
+prior work, we used a similar experimental setup (equivalent to the ``reduced''
+condition in the present study) to verify that the automatically transcribed
+recalls were sufficiently close to human-transcribed recalls to yield reliable
 data~\citep{ZimaEtal18}. This real-time speech processing component of the
@@
 random color and location) for late lists. Finally, in the \textit{reduced
-(late)} condition, we followed the feature rich procedure for earlylists and
+(late)} condition, we followed the feature rich procedure for early lists and
 the reduced procedure for late lists.
@@
 features were correlated-- for example, some semantic categories of words
-referred to objects that tended to be a particular size, which means that
-category and size are not fully independent. On late lists, the words were
+referred to objects that tended to be a particular size, which meant that
+category and size were not fully independent. On late lists, the words were
 always presented in a randomized order (chosen anew for each participant). In
@@
 the same letter would have a ``first letter'' distance of 0, and words starting
-with `J' and `A' would have a first letter distance of 9. Because words'
-lengths and letters' positions in the alphabet are always integers, these
-discrete distances always take on integer values. Finally, the visual features
-(color and location) are \textit{continuous} and \textit{multivariate}, in that
-each ``feature'' takes on multiple (positive) real values. We defined the
-``color'' and ``location'' distances between two words as the Euclidean
-distances between their $(r, g, b)$ color or $(x, y)$ location vectors,
-respectively. Therefore the color and location distance measures always take on
-positive real values (upper bounded at 441.67 for color, or 27~in for location,
-reflecting the distances between the corresponding maximally different
-vectors).
+with `J' and `A' respectively would have a first letter distance of 9. Because
+words' lengths and letters' positions in the alphabet are always integers,
+these discrete distances always take on integer values. Finally, the visual
+features (color and location) are \textit{continuous} and
+\textit{multivariate}, in that each ``feature'' takes on multiple (positive)
+real values. We defined the ``color'' and ``location'' distances between two
+words as the Euclidean distances between their $(r, g, b)$ color or $(x, y)$
+location vectors, respectively. Therefore the color and location distance
+measures always take on positive real values (upper-bounded at 441.67 for
+color, or 27~in for location, reflecting the distances between the
+corresponding maximally different vectors).
 
@@
 relevant feature, and each word's value(s) for that feature, we developed a
-stochastic algorithm for (noisily) sorting the words. First, we choose a word
-uniformly at random from the set of candidates. Next, we compute the distances
-between the chosen word's feature(s) and the corresponding feature(s) of all
-yet-to-be-presented words. Third, we convert these distances (between the
-previously presented word's feature values, $a$, and the candidate word's
-feature values, $b$) to similarity scores: \begin{equation}
-\mathrm{similarity}(a, b) = \exp\{-\tau \cdot \mathrm{distance}(a, b)\},
-\label{eqn:similarity} \end{equation} where $\tau = 1$ in our implementation.
-We note that increasing the value of $\tau$ would amplify the influence of
-similarity on order, and decreasing the value of $\tau$ would diminish the
-influence of similarity on order. Also note that this approach requires $\tau >
-0$. Finally, we computed a set of normalized similarity values by dividing the
-similarities by their sum: \begin{equation}
+stochastic algorithm for (noisily) sorting the words. The stochastic aspect of
+our sorting procedure enabled us to obtain unique lists for each participant.
+First, we choose a word uniformly at random from the set of candidates. Next,
+we compute the distances between the chosen word's feature(s) and the
+corresponding feature(s) of all yet-to-be-presented words. Third, we convert
+these distances (between the previously presented word's feature values, $a$,
+and the candidate word's feature values, $b$) to similarity scores:
+\begin{equation} \mathrm{similarity}(a, b) = \exp\{-\tau \cdot
+\mathrm{distance}(a, b)\}, \label{eqn:similarity} \end{equation} where $\tau =
+1$ in our implementation. We note that increasing the value of $\tau$ would
+amplify the influence of similarity on order, and decreasing the value of
+$\tau$ would diminish the influence of similarity on order. Also note that this
+approach requires $\tau > 0$. Finally, we computed a set of normalized
+similarity values by dividing the similarities by their sum: \begin{equation}
 \mathrm{similarity}_{\mathrm{normalized}}(a, b) = \frac{\mathrm{similarity}(a,
@@
 $x$, and all yet-to-be-presented items ($a$ -- $g$). Next, we normalize these
-similarity scores so that they sum to one. We lay in sequence a set of
+similarity scores so that they sum to one. We lay, in sequence, a set of
 ``sticks,'' one for each candidate item, whose lengths are equal to these
@@
 We designed the \textit{adaptive} experimental condition to study the effect on
-memory for information that matched (or mismatched) the ways participants
-``naturally'' organized their memories of the lists they studied. Like the
-other conditions, all participants in the adaptive condition studied a total of
-16 lists, in a randomized order. We varied the words' colors and locations for
-every word presentation, as in the feature rich and order manipulation
-conditions.
+memory of lists that matched (or mismatched) the ways participants
+``naturally'' organized their memories. Like the other conditions, all
+participants in the adaptive condition studied a total of 16 lists, in a
+randomized order. We varied the words' colors and locations for every word
+presentation, as in the feature rich and order manipulation conditions.
 
@@
 participant's ``memory fingerprint,'' defined below. At a high level, a
-participant's memory fingerprint describes how they prioritize different
-semantic, lexicographic, and/or visual features when they organize their
-memories.
-
-Next, participants studied a sequence of 12 lists in three batches of 4 lists
-each. These batches came in three types: \textit{random}, \textit{stabilize},
-and \textit{destabilize}. The batch types determined how words on the lists in
-that batch were ordered. Lists in each batch were always presented
-consecutively (e.g., a participant might receive four random lists, followed by
-four stabilize lists, followed by four destabilize lists). The batch orders
-were evenly counterbalanced across participants: there are six possible
-orderings of the three batches, and 10 participants were randomly assigned to
-each ordering sub-condition.
+participant's memory fingerprint describes how they prioritize or consider
+different semantic, lexicographic, and/or visual features when they organize
+their memories.
+
+Next, participants studied a sequence of 12 lists in three batches of four
+lists each. These batches came in three types: \textit{random},
+\textit{stabilize}, and \textit{destabilize}. The batch types determined how
+words on the lists in that batch were ordered. Lists in each batch were always
+presented consecutively (e.g., a participant might receive four random lists,
+followed by four stabilize lists, followed by four destabilize lists). The
+batch orders were evenly counterbalanced across participants: there are six
+possible orderings of the three batches, and 10 participants were randomly
+assigned to each ordering sub-condition.
 
@@
 destabilize batches were sorted in ways that either matched or mismatched each
-participant's memory fingerprint, respectively. Our procedures for computing
+participant's memory fingerprint, respectively. Our procedures for estimating
 participants' memory fingerprints and ordering the stabilize and destabilize
@@
 our main approach to computing clustering scores on analogous temporal and
-semantic clustering scores developed by~\citep{PolyEtal09}. Computing the
-clustering score for one feature dimension starts by considering those feature
-values from the first word the participant recalled on the list. Next, we sort
-all not-yet-recalled words in ascending order according to their feature-based
-distance to the just-recalled item (see \textit{Defining feature-based
-distances}). We then compute the percentile rank of the observed next recall.
-We averaged these percentile ranks across all of the participant’s recalls for
-the current list to obtain a single uncorrected clustering score for the list,
-for the given feature dimension. We repeat this process for each feature
-dimension in turn to obtain a single uncorrected clustering score for each
-list, for each feature dimension.
+semantic clustering scores developed by~\cite{PolyEtal09}. Computing the
+clustering score for one feature dimension starts by considering the
+corresponding feature values from the first word the participant recalled
+correctly from the just-studied list. Next, we sort all not-yet-recalled words
+in ascending order according to their feature-based distance to the
+just-recalled item (see \textit{Defining feature-based distances}). We then
+compute the percentile rank of the observed next recall. We average these
+percentile ranks across all of the participant’s recalls for the current list
+to obtain a single uncorrected clustering score for the list, for the given
+feature dimension. We repeated this process for each feature dimension in turn
+to obtain a single uncorrected clustering score for each list, for each feature
+dimension.
 
@@
 sequences by the learned items' encoding positions. For instance, if a
-participant recalled the episode events in the exact order they occurred (or in
-exact reverse order), this would yield a score of 1. If a participant recalled
-the events in random order, this would yield an expected score of 0.5. For each
-recall-event transition (and separately for each participant), we sorted all
-not-yet-recalled events according to their absolute lag (that is, distance away
-in the episode). We then computed the percentile rank of the next event the
+participant recalled the lists' words in the exact order they were presented
+(or in exact reverse order), this would yield a score of 1. If a participant
+recalled the words in random order, this would yield an expected score of 0.5.
+For each recall transition (and separately for each participant), we sorted all
+not-yet-recalled words according to their absolute lag (that is, distance away
+in the list). We then computed the percentile rank of the next word the
 participant recalled. We took an average of these percentile ranks across all
@@
 since (early in the recall sequence) the yet-to-be-recalled items come from a
-mix of sizes. However, once all of the ``small'' items on list $B$ have been
+mix of sizes. However, once all of the small items on list $B$ have been
 recalled, the best possible next matching recall will be a large item. And all
@@
 recalls it becomes difficult to determine whether the participant is
-successively recalling ``large'' items because they are organizing their
+successively recalling large items because they are organizing their
 memories according to size, or (alternatively), whether they are simply
@@
 the range of clustering scores that are possible to observe for a given list.
-The uncorrected clustering score therefore conflates participants' actual
+An uncorrected clustering score therefore conflates participants' actual
 memory organization with other ``nuisance'' factors.
 
-Following our prior work~\citep{HeusEtal17}, we used a permutation-correction
-procedure to help isolate the behavioral aspects of clustering that we were
-most interested in. After computing the uncorrected clustering score (for the
-given list and observed recall sequence), we compute a ``null'' distribution of
-$n$ additional clustering scores after randomly shuffling the recall order (we
-use $n = 500$ in the present study). This null distribution represents an
-approximation of the range of clustering scores one might expect to observe by
-``chance,'' given that a hypothetical participant was \textit{not} truly
-clustering their recalls, but where the hypothetical participant studied and
-recalled exactly the same items (with the same features) as the true
-participant. We define the permutation-corrected clustering score as the
-percentile rank of the observed uncorrected clustering score in this estimated
-null distribution. In this way, a corrected score of 1 indicates that the
-observed score was greater than any clustering score one might expect by
-chance; in other words, good evidence that the participant was truly clustering
-their recalls along the given feature dimension.  We applied this correction procedure
-to all of the clustering scores (feature and temporal) reported in this paper.
+Following our prior work~\citep{HeusEtal17}, we used a permutation-based
+correction procedure to help isolate the behavioral aspects of clustering that
+we were most interested in. After computing the uncorrected clustering score
+(for the given list and observed recall sequence), we compute a ``null''
+distribution of $n$ additional clustering scores after randomly shuffling the
+order of the recalled words (we use $n = 500$ in the present study). This null
+distribution represents an approximation of the range of clustering scores one
+might expect to observe by ``chance,'' given that a hypothetical participant
+was \textit{not} truly clustering their recalls, but where the hypothetical
+participant still studied and recalled exactly the same items (with the same
+features) as the true participant. We define the \textit{permutation-corrected
+clustering score} as the percentile rank of the observed uncorrected clustering
+score in this estimated null distribution. In this way, a corrected score of 1
+indicates that the observed score was greater than any clustering score one
+might expect by chance; in other words, good evidence that the participant was
+truly clustering their recalls along the given feature dimension. We applied
+this correction procedure to all of the clustering scores (feature and
+temporal) reported in this paper.
 
@@
 length, first letter, color, and location) and their temporal clustering score.
-Conceptually, this memory fingerprint describes the participant's tendancies to
-order (and, presumably, organize in memory) the studied words along each
-dimension. To obtain stable estimates of these fingerprints for each
-participant, we averaged clustering scores across lists. We also tracked and
-characterized how participants' fingerprints changed across lists (e.g.,
+Conceptually, a participant's memory fingerprint describes their tendancy to
+order in their recall sequences (and, presumably, organize in memory) the
+studied words along each dimension. To obtain stable estimates of these
+fingerprints for each participant, we averaged clustering scores across lists.
+We also tracked and characterized how participants' fingerprints changed across
+lists (e.g.,
 Figs.~\ref{fig:fingerprint-trajectories},~\fingerprintTrajectoryRandom).
@@
 of the recall intervals for each list. We used the \texttt{Quail} Python
-package~\citep{HeusEtal17} to apply speech-to-text alorithms to the just
-collected data, aggregate the data for the given participant, and estimate the
-participant's memory fingerprint using all of their available data up to that
-point in the experiment. Two aspects of our implementation are worth noting.
-First, because memory fingerprints are averaged across lists, the
-already-computed memory fingerprints for earliar lists could be cached and
-loaded as needed in future computations. This meant that our computations
-pertaining to updating our estimates of a participant's memory fingerprint only
-needed to consider data from the most recent list. Second, each element of the
-null distributions of uncorrected fingerprint scores (see
-\textit{Permutation-corrected feature clustering scores}) could be estimated
-independently from the others. This enabled us to make use of the testing
-computers' multi-core CPU archetectuers by elements of the null distributions
-in batches of eight (i.e., the number of CPU cores on each testing computer).
-Taken together, we were able to compress the fingerprint computations into just
-a few seconds of computing time. The combined processing time for the
-speech-to-text algorithm and fingerprint computations easily fit within the
-inter-list intervals, where participants typically paused before moving on to
-the next list.
+package~\citep{HeusEtal17} to apply speech-to-text alorithms to the
+just-collected data, aggregate the data for the given participant, and estimate
+the participant's memory fingerprint using all of their available data up to
+that point in the experiment. Two aspects of our implementation are worth
+noting. First, because memory fingerprints are computed independently for each
+list and then averaged across lists, the already-computed memory fingerprints
+for earlier lists could be cached and loaded as needed in future computations.
+This meant that our computations pertaining to updating our estimate of a
+participant's memory fingerprint only needed to consider data from the most
+recent list. Second, each element of the null distributions of uncorrected
+fingerprint scores (see \textit{Permutation-corrected feature clustering
+scores}) could be estimated independently from the others. This enabled us to
+make use of the testing computers' multi-core CPU architectures by elements of
+the null distributions in batches of eight (i.e., the number of CPU cores on
+each testing computer). Taken together, we were able to compress the relevant
+computations into just a few seconds of computing time. The combined processing
+time for the speech-to-text algorithm, fingerprint computations, and
+permutation-based ordering procedure (described next) easily fit within the
+inter-list intervals, where participants paused for a self-paced break before
+moving on to study and recall the next list.
 
@@
 random permutations of the to-be-presented items. These permutations served as
-prospective presentation orders. We sought to select the specific order that
-most (or least) matched $f$. Third, for each random permutation, we computed
-the (permutation-corrected) ``fingerprint,'' treating the permutation as though
-it were a potential ``perfect'' recall sequence. (We did not include temporal
-clustering scores in these fingerprints.) This yielded a ``simulated fingerprint''
-vector, $\hat{f}_p$ for each permutation $p$.  We used these simulated fingerprints to
-select a specific permutation, $i$, that either maximized (for stabilized lists) or minimized (for destabilize lists)
-the correlation between $\hat{f}_i$ and $f$.
+candidate presentation orders. We sought to select the specific order that most
+(or least) matched $f$. Third, for each random permutation, we computed the
+(permutation-corrected) ``fingerprint,'' treating the permutation as though it
+were a potential ``perfect'' recall sequence. (We did not include temporal
+clustering scores in these fingerprints.) This yielded a ``simulated
+fingerprint'' vector, $\hat{f}_p$ for each permutation $p$. We used these
+simulated fingerprints to select a specific permutation, $i$, that either
+maximized (for stabilize lists) or minimized (for destabilize lists) the
+correlation between $\hat{f}_i$ and $f$.
 
@@
 
-Following some of our prior work~\citep{HeusEtal18a}, we use low-dimensional
-embeddings to help visualize how participants' memory fingerprints change
-across lists
+\textbf{JRM NOTE: REMINDER TO CHECK THIS PARAGRAPH AGAINST ANALYSIS CODE FOR
+ACCURACY...} Following some of our prior work~\citep{HeusEtal18a, HeusEtal21},
+we use low-dimensional embeddings to help visualize how participants' memory
+fingerprints change across lists
 (Figs.~\ref{fig:fingerprint-trajectories}A,~\fingerprintTrajectoryRandom A). To
@@
 number-of-lists $\times$ number-of-participants rows and seven columns (one for
-each word feature dimension's clustering scores, plus an additional temporal
-clustering score column). We used principal components analysis to project the
-seven-dimensional observations into a two-dimensional space (using the two
-principal components that explained the most variance in the data). For two
-visualizations
+each feature clustering score, plus an additional temporal clustering score
+column). We used principal components analysis to project the seven-dimensional
+observations into a two-dimensional space (using the two principal components
+that explained the most variance in the data). For two visualizations
 (Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B)
@@
 \textit{average} fingerprints (i.e., across lists within a given group of
-lists-- early or late). For those visualizations we averaged each participant's
-rows (for the given group of lists) in the combined fingerprint matrix prior to
-projecting it into the shared two-dimensional space. This yielded a single
-two-dimensional coordinate for each \textit{participant} and \textit{list
-group}, rather than for each individual list. We used these embeddings solely
-for visualization. All statistical tests were carried out in the original
-(seven-dimensional) feature spaces.
+lists-- early or late). For those visualizations we averaged across the rows
+(for each condition and group of lists) in the combined fingerprint matrix
+prior to projecting it into the shared two-dimensional space. This yielded a
+single two-dimensional coordinate for each \textit{list group}, rather than for
+each individual list. We used these embeddings solely for visualization. All
+statistical tests were carried out in the original (seven-dimensional) feature
+spaces.
 
@@
 indicates that a recalled item was presented immediately after the previously
-recalled item, and a lag of $−3$ indicates that a recalled item came 3 items
+recalled item, and a lag of $-3$ indicates that a recalled item came three items
 before the previously recalled item. For each recall transition (following the
@@
 Serial position curves~\citep{Murd62a} reflect the proportion of participants
-who remember each item as a function of the item’s serial position during
+who remember each item as a function of the items' serial positions during
 encoding. For each participant, we initialized a number-of-lists (16) by

### 2023-01-16 paper/main.tex
backing up progress on discussion section   - added a more detailed summary and overview of the key results

@@
 
-We sought to manipulate two aspects of how participants memorized sequences of
-word lists. First, we added two additional sources of visual variation to the
-individual word presentations: font color and onscreen location. Importantly,
-these visual features were independent of the meaning or semantic content of
-the words (e.g., word category, size of the refferent) and of the lexicographic
-properties of the word (e.g., word length, first letter). We wondered whether
-this additional word-independent information might facilitate recall (e.g., by
-providing new potential ways of organizing or retrieving memories of the
-studied words) or impair recall (e.g., by distracting participants). Second,
-our primary experimental manipulations entailed manipulating the orders in
-which words were studied (and how those orderings changed over time). We
-wondered whether presenting the same list of words in different orders (e.g.,
-sorted along one feature dimension versus another) might serve to influence how
-participants organized their memories of the words. We also wondered whether
-some order manipulations might be temporally ``sticky'' by influencing how
-\textit{future} lists were remembered.
+While holding the set of words (and the assignments of words to lists)
+constant, we manipulated two aspects of participants' experiences of studying
+each list. We sought to understand the effects of these manipulations on
+participants' memories for the studied words. First, we added two additional
+sources of visual variation to the individual word presentations: font color
+and onscreen location. Importantly, these visual features were independent of
+the meaning or semantic content of the words (e.g., word category, size of the
+refferent, etc.) and of the lexicographic properties of the words (e.g., word length,
+first letter, etc.). We wondered whether this additional word-independent information
+might facilitate recall (e.g., by providing new potential ways of organizing or
+retrieving memories of the studied words) or impair recall (e.g., by
+distracting participants with irrelevant information). Second, we manipulated
+the orders in which words were studied (and how those orderings changed over
+time). We wondered whether presenting the same list of words with different
+appearances (e.g., by manipulating font size and onscreen location) or in
+different orders (e.g., sorted along one feature dimension versus another)
+might serve to influence how participants organized their memories of the
+words. We also wondered whether some order manipulations might be temporally
+``sticky'' by influencing how \textit{future} lists were remembered.
 
@@
 word length clustering: $t(126) = 10.639, p < 0.001$; first letter clustering:
-$t(126) = 7.775, p = 0.000$; see \textit{Permutation-corrected feature
+$t(126) = 7.775, p < 0.001$; see \textit{Permutation-corrected feature
 clustering scores} for more information about how we quantified each
@@
 We asked participants to study and freely recall word lists. The words on each
-list (and the set of lists) were held constant across participants. For each
-word, we considered (and manipulated) two semantic features (category and size)
-that reflected aspects of the \textit{meanings} of the words, along with two
-lexicographic features (word length and first letter), which reflected aspects
-of the words' \textit{letters}. These semantic and lexicographic features are
-intrinsic to each word. We also considered and manipulated two additional
-visual features (color and location) that affected the \textit{appearance} of
-each studied item, but could be varied independently of the words' identities.
-Across different experimental conditions, we manipulated how the visual
-features varied across words (within each list), along with the orders of each
-list's words. Although participants' task (verbally recalling as many words as
-possible, in any order, within one minute) remained constant across all of
-these conditions, and although the set of words they studied on each list
-remained constant, our manipulations substantially affected participants'
-memories. The impact of some of the manipulations also affected how
-participants remembered \textit{future} lists that were sorted randomly.
+list (and the total set of lists) were held constant across participants. For
+each word, we considered (and manipulated) two semantic features (category and
+size) that reflected aspects of the \textit{meanings} of the words, along with
+two lexicographic features (word length and first letter), which reflected
+aspects of the words' \textit{letters}. These semantic and lexicographic
+features are intrinsic to each word. We also considered and manipulated two
+additional visual features (color and location) that affected the
+\textit{appearance} of each studied item, but could be varied independently of
+the words' identities. Across different experimental conditions, we manipulated
+how the visual features varied across words (within each list), along with the
+orders of each list's words. Although participants' task (verbally recalling as
+many words as possible, in any order, within one minute) remained constant
+across all of these conditions, and although the set of words they studied on
+each list remained constant, our manipulations substantially affected
+participants' memories. The impact of some of the manipulations also affected
+how participants remembered \textit{future} lists that were sorted randomly.
 
@@
 %   adaptive condition: people have unique preferences, and matching them can improve their memory and affect how the remembered info is organized
+
+We found that participants in our feature rich condition (where we varied
+words' appearances) recalled similar proportions of words to participants in a
+reduced condition (where appearance was held constant across words). However,
+varying the words' appearances led participants to exhibit much more temporal
+and feature-based clustering. This suggests that even seemingly irrelevant
+elements of our experiences can affect how we remember them.
+
+When we held the within-list variability in participants' visual experiences
+fixed across lists (in the feature rich and reduced conditions), they
+remembered more words on early versus late lists. On feature rich lists, they
+also showed stronger clustering on early versus late lists. However, when we
+\textit{varied} participants' visual experiences across lists (in the ``reduced
+(early)'' and ``reduced (late)'' conditions), these early versus late accuracy
+and clustering differences disappeared. Abruptly changing how irrelevant visual
+features change across words seems to act as a sort of ``event boundary'' that
+partially resets how participants process and remember post-boundary lists.
+Within-list clustering also increases in these manipulations, suggesting that
+the "within-event" words are being more tightly associated with each other.
+
+When we held the visual features constant on early lists, but then varied
+words' appearances on later lists (i.e., the reduced (early) condition), this improved
+participants' overall memory performance. However, this impact was directional:
+when we \textit{removed} visual features on late lists that had been present on
+early lists (i.e., the reduced (late) condition), we saw no memory improvement.
+
+When we (stochastically) sorted early lists along different feature dimensions,
+we found several impacts on participants' memories. Sorting early lists
+semantically (by word category) enhanced participants' memories for those
+lists, but the effects on performance of sorting along other feature dimensions
+were inconclusive. However, each order manipulation substantially affected how
+participants \textit{organized} their memories of words from the ordered lists.
+When we sorted lists semantically participants displayed stronger semantic
+clustering; when we sorted lists lexicographically they displayed stronger
+lexicographic clustering; and when we sorted lists visually they displayed
+stronger visual clustering. Clustering along the unmanipulated feature
+dimensions in each of these cases was unchanged.
+
+The order manipulations we examined also appeared to induce, in some cases, a
+tendancy to ``clump'' similar words within a list. This was most apparent on
+semantically ordered lists, where the probability of initiating recall with a
+given word seemed to follow groupings defined by feature changepoints.
+
+We also examined the impact of early list order manipulations on memory for
+late lists. At the group level, we found little evidence for lingering
+``carryover'' effects of these manipulations; participants in the order
+manipulation conditions showed similar memory performance and clustering on
+late lists to participants in the corresponding control (feature rich)
+condition. At the level of individual participants, however, we found several
+meaningful patterns.
+
+Participants who showed stronger feature clustering on early (order
+manipulated) lists tended to better remember late (randomly ordered) lists.
+Participants who remembered early lists better also tended to show stronger
+feature clustering (along their condition's feature dimension) on late lists
+(even though the words on those late lists were presented in a random order).
+We also observed some (weaker) carryover effects of temporal clustering.
+Participants who showed stronger feature clustering (along their condition's
+feature dimension) on early lists tended to show stronger temporal clustering
+on late lists. And participants who showed stronger temporal clustering on
+early lists also tended to show stronger feature clustering on late lists.
+Essentially, these order manipulations appeared to affect each participant
+differently. Some participants were sensitive to our manipulations, and those
+particpants showed stronger impacts on their memory performance for the ordered
+lists as well as future (random) lists. Other participants appeared relatively
+insensitive to our manipulations, and those participants showed little
+carryover effects on late lists.
+
+These results at the individual participant level suggested to us that either
+(a) some participants were more sensitive to \textit{any} order manipulation,
+or (b) some participants might be more (or less) sensitive to manipulations
+along \textit{particular} (e.g., preferred) feature dimensions. To help
+distinguish between these possibilities, we designed an adaptive manipulation
+whereby we attempted to manipulate whether participants studied words in an
+order that matched (or mismatched) our estimate of how they would cluster or
+organize the studied words in memory (i.e., their idiosyncratic memory
+fingerprint). We found that when we presented words in orders that were
+consistent with participants' memory fingerprints, they remembered more words
+overall and showed stronger temporal clustering. This comports well with the
+second possibility described above. Specifically, each participant seems to
+bring into the experiment their own idiosynchratic preferences and strategies
+for organizing the words in their memories. When we presented the words in an
+order consistent with each participant's idiosyncratic strategies, their memory
+performance improved. This might indicate that the participants were spending
+less cognitive effort ``reorganizing'' the incoming words on those lists, which
+freed up resources to devote to encoding processes instead.
 

### 2023-01-16 paper/main.tex
backing up draft of context effects discussion section

@@
 
+\subsection*{Recap: visual feature manipulations}
+
 We found that participants in our feature rich condition (where we varied
@@
 early lists (i.e., the reduced (late) condition), we saw no memory improvement.
+
+\subsection*{Recap: order manipulations}
 
@@
 % connections to prior work: 
-%     - context effects: 
-%     - priming: 
+%     - context effects: role of irrelevant features, temporal context and order effects
+%     - priming: being "primed" to 
 %     - situation models:
+
+\subsection*{Context effects on memory performance and organization}
+
+In everyday experience, each moment's unique blend of contextual features
+(where we are, who we are with, what else we are thinking of at the time, what
+else we experience nearby in time, etc.) plays an important role in how we
+interpret, experience, and remember that moment, and how we relate it to our
+other experiences~\citep[e.g., for review see ][]{Mann20}. What are the
+analogues of real-world contexts in laboratory tasks like the free recall
+paradigm employed in our study? In general, modern formal accounts of free
+recall~\citep{Kaha20} describe context as comprising a mix of (a) features
+pertaining to or associated with each item and (b) other items and thoughts
+experienced nearby in time, e.g., that might still be ``lingering'' in the
+participant's thoughts at the time they study the item. Item features can
+include semantic properties (i.e., features related to the item's meaning),
+lexicographic properties (i.e., features related to the item's letters),
+sensory properties (i.e., feature related to the item's appearance, sound,
+smell, etc.), emotional proprties (i.e., features related to how meaningful the
+item is, whether the item evokes positive or negative feelings, etc.),
+utility-related properties (e.g., features that describe how an item might be
+used or incorporated into a particular task or situation), and more.
+Essentially any aspect of the participant's experience that can be
+characterized, measured, or otherwise described can be considered to influence
+the participant's mental context at the moment they experience that item.
+Temporally proximal features include aspects of the participant's internal or
+external experience that are \textit{not} specifically occuring at the moment
+they encounter an item, but that nonetheless influence how they process the
+item. Thoughts related to percepts, goals, expectations, other experiences, and
+so on that might have been cued (directly or indirectly) by the participant's
+recent experiences prior to the current moment all fall into this category.
+Internally driven mental states, such as thinking about an experience unrelated
+to the experiment, also fall into this category.
+
+Contextual features need not be intentionally or consciously perceived by the
+participant to affect memory, nor do they need to be relevant to the task
+instructions or the participant's goals. Incidental factors such as font
+color~\citep{JonePyc14}, background color~\citep{IsarIsar07}, inter-stimulus
+images~\citep{GersEtal13, MannEtal16, ChiuEtal21}, background
+sounds~\citep{BeamJone10, SahaSmit14}, secondary tasks~\citep{MasiSaha14,
+PolyEtal09}, and more can all impact how the participant remembers, and
+organizes in memory, lists of studied items.
+
+Consistent with this prior work, we found that participants are sensitive to
+task-irrelevant visual features. We also found that changing the dynamics of
+those task-irrelevant visual features (in the reduced (early) and reduced
+(late) conditions) \textit{also} affects participants' memories. This suggests
+that it is not only the contextual features themselves that affect memory, but
+also the \textit{dymamics} of context-- i.e., how the contextual features
+associated with each item change over time.
+
+\subsection*{Priming effects on memory performance and organization}
+
+\subsection*{Expectation, event boundaries, and situation models}
 
@@
 %     - applications: adaptive learning, education and training
+
+\subsection*{Theoretical implications}
+
+\subsection*{Potential applications}
 
@@
 
+\subsection*{Concluding remarks}
+
 \bibliographystyle{apa}

### 2023-01-16 paper/main.tex
added keywords, backing up progress on discussion

@@
 
-\title{Carryover effects in free recall reveal how past experiences influence
+\title{\Large Carryover effects in free recall reveal how past experiences influence
 memories of future experiences}
@@
 
-\begin{abstract} We perceive, interpet, and remember ongoing experiences
+\begin{abstract} \footnotesize{We perceive, interpet, and remember ongoing experiences
 through the lens of our prior experiences. Inferring that we are in one type of
@@
 how structure in our ongoing experiences can exert influence on how we remember
-unrelated subsequent experiences. \end{abstract}
+unrelated subsequent experiences.
+
+\textbf{Keywords: episodic memory, free recall, incidental features, implicit priming, temporal order}}
+
+\end{abstract}
 

### 2023-01-16 paper/main.tex
added some notes on priming (discussion section)
added a discussion section on priming effects

@@
 
-In everyday experience, each moment's unique blend of contextual features
+In real-world experience, each moment's unique blend of contextual features
 (where we are, who we are with, what else we are thinking of at the time, what
@@
 
+% explicit priming (resolve ambiguous stimuli, e.g., green eyes study)
+% negative priming: supressing previously irrelevant stimuli or features hurts participants' recall for those stimuli when they're used on subsequent lists
+% implicit priming (most like here): patterns in prior experiences sets expectations about what is to come, which can affect memory for upcoming stimuli
+%    - participants in the reduced ({early, late}) conditions picked up on irrelevant features in early lists and when those features changed their memories for subsequent lists were affected.
+%    - participants in the order manipulation conditions picked up on structure in temporal order on early lists, and that affected how they remembered later lists
+
+When our ongoing experiences are ambiguous, we can draw on our past
+experiences, expectations, and other real, perceived, or inferred cues to help
+resolve the ambiguities. We may also be overtly or covertly ``primed'' to
+influence how we are likely to resolve ambiuities. For example, before
+listening to a story with several equally plausable interpretations, providing
+participants with ``background'' information beforehand can lead them towards
+one interpretation versus another~\citep{YeshEtal17}. More broadly, our
+conscious and unconscious biases and preferences can influence not only how we
+interpret high-level ambuities, but even how we process low-level sensory
+information~\citep{KataEtal23}.
+
+In more simplified scenarios, like list learning paradigms, the stimuli and
+tasks participants encounter before studying a given list can influence what
+and how they remember. For example, when participants are directed to supress,
+disregard, or ignore ``distracting'' stimuli early on in an experiment,
+participants often tend to remember those stimuli less well when they are
+re-used as to-be-remembered targets later on in the experiment~\citep{Tipp85}.
+In general, participants' memories can be influenced by a wide range of
+positive and negative factors~\citep{BaloEtal92, ClayChat89, Donn88,
+FlexTulv82, GottEtal12, HuanEtal04, Hube08, HubeEtal01, McNa94, Neel77, Rabi86,
+TulvScha91, WatkEtal92, WiggMart98}.
+
+The order manipulation conditions in our experiment show that participants can
+also be primed to pick up on more subtle statistical structure in their
+experiences, like the dynamics of how the presentation orders of stimuli vary
+along particular feature dimensions. These order manipulations affected not
+only how participants remembered the manipulated lists, but also how they
+remembered \textit{future} lists with different (randomized) temporal
+properties.
+
 \subsection*{Expectation, event boundaries, and situation models}

### 2023-01-17 paper/main.tex
added a discussion section on expectation, event boundaries, and situation models

@@
 these regularities when generating inferences about the unobserved past and
-yet-to-be-experienced future~\citep{XuEtal22, SchaTurk15, RangRitc12,
+yet-to-be-experienced future~\citep{XuEtal23, SchaTurk15, RangRitc12,
 BoweEtal79, MomeEtal17}. On the other hand, the rich structure of real-world
@@
 
+Our findings that participants' current and future memory behaviors are
+sensitive to manipulations in which features change over time, and how features
+change across items and lists, suggest parallels with studies on how we form
+expectations and predictions, segment our continuous experiences into discrete
+events, and make sense of different scenarios and situations. Each of these
+real-world cognitive phenomena entail identifying statistical regularities in
+our experiences, and exploiting those regularities to gain insight, form
+inferences, organize or interpret memories, and so on. Our past experiences
+enable us to predict what is likely to happen in the future, given what
+happened ``next'' in our previous experiences that were similar to
+now~\citep{Mann20, EichFort09, BarrEtal20, Brig12, ChowEtal16, GlucEtal02,
+GoldEtal21, GrifStey03, JonePash07, KimEtal14, TamiThor18, XuEtal23}.
+
+When our expectations are violated, such as when our observations disagree with
+our predictions, we may perceive the ``rules'' or ``situation'' to have
+changed. \textit{Event boundaries} denote abrupts change in the state of our
+experience, for example when we transition from one situation to
+another~\citep{RadvZack17, ZwaaRadv98}. Crossing an event boundary can impair
+our memory for pre-boundary information and enhance our memory for
+post-boundary information~\citep{RadvCope06, SahaKell02, MannEtal16}. Event
+boundaries are also tightly associated with the notion of \textit{situation
+models} and \textit{schemas}-- mental frameworks for organizing our
+understanding about the rules of how we and others are likely to behave, how
+events are likely to unfold over time, how different elements are likely to
+interact, and so on. For example, a situation model pertaining to a particular
+restaurant might set our expectations about what we are likely to experience
+when we visit that restaurant (e.g., what the building will look like, how it
+will smell when we enter, how crowded the restaurant is likely to be, the
+sounds we are likely to hear, etc.). Similarly, we might learn a schema
+describing how events are likely to unfold \textit{across} any sit-down
+restaurant-- e.g., open the door, wait to be seated, receive a menu, decide
+what to order, place the order, and so on. Situation models and schemas can
+help us to generalize across our experiences, and to generate expectations
+about how new experiences are likely to unfold. When those expectations are
+violated, we can perceive ourselves to have crossed into a new situation.
+
+In our study, we found that abruptly changing the ``rules'' about how the
+visual appearances of words are determined, or about the orders in which words
+are studied, can lead participants to behave similarly to what one might expect
+upon crossing an event boundary. Adding in variability in font color and
+presentation locations for words on late lists, after those visual features had
+been held constant on early lists, led participants to remember more words on
+those later lists. (We found the same pattern when we varied words' colors and
+locations on early lists but held them constant on late lists.) One potential
+explanation is that participants perceive an ``event boundary'' to have
+occurred when they encounter the first ``late'' list. According to contextual
+change accounts of memory across event boundaries~\citep[e.g.,][]{SahaKell02},
+this could help to explain why participants in the reduced (early) and reduced
+(late) conditions exhibited better overall memory performance. Specifically,
+their memory for late list items could benefit from less interference from
+early list items, and the contextual features associated with late list items
+(after the ``event boundary'') might serve as more specific recall cues for
+those late items (relative to if the boundary had not occurred).
+
 % implications for:

### 2023-01-17 paper/main.tex
wrote draft of theoretical implications (discussion) section

@@
 
+Although most modern formal theories of episodic memory have been developed and
+tested to explain memory for list learning tasks~\citep{Kaha20}, a number of
+recent studies suggest some substantial differences between memory for lists
+versus naturalistic stimuli (e.g., real-world experiences, narratives, films,
+etc.)~\citep{NastEtal20, HeusEtal21, Mann21a, LeeEtal20}. One reason is that
+naturalistic stimuli are often much more engaging than the highly simplified
+list learning tasks typically employed in the psychological laboratory, perhaps
+leading participants to pay more attention, exert more effort, and stay more
+consistently motivated to perform well~\citep{NastEtal20}. Another reason is
+that the temporal unfoldings of events and occurances in naturalistic stimuli
+tend to be much more meaningful than the temporal unfoldings of items on
+typical lists used in laboratory memory tasks. Real-world events exhibit
+important associations at a broad range of timescales. For example, an early
+detail in a detective story may prove to be a clue to solving the mystery later
+on. Further, what happens in one moment typically carries some predictive
+information about what came before or after~\citep{XuEtal23}. In contrast, the
+lists used in laboratory memory tasks are most often ordered randomly, by design,
+to \textit{remove} meaningful temporal structure in the
+stimulus~\citep{Kaha12}.
+
+On one hand, naturalistic stimuli provide a potential means of understanding
+how our memory systems function in the circumstances we most often encounter in
+our everyday lives. This implies that, to understand how memory works in the
+``real world,'' we should study memory for stimuli that reflect the relevant
+statistical structure of real-world experiences. On the other hand,
+naturalistic stimuli can be difficult to precisely characterize or model,
+making it difficult to distinguish whether specific behavioral trends follow
+from fundamental workings of our memory systems, from some aspect of the
+stimulus, or from idiosyncratic interactions or inteference between our memory
+systems and the stimulus. This challenge implies that, to understand the
+fundamental nature of memory in its ``pure'' form, we should study memory for
+highly simplified stimuli that can provide relatively unbiased (compared with
+real-world experiences) measures of the relevant patterns and tendancies.
+
+The experiment we report in this paper was designed to help bridge some of the
+gap between naturalistic tasks and more traditional list learning tasks. We had
+people study word lists similar to those used in classic memory studies, but we
+also systematically varied the lists' ``richness'' (by adding or removing
+visual features) and temporal structure (through order manipulations that
+varied over time and across experimental conditions). We found that
+participants' memory behaviors were sensitive to these manipulations. Some of
+the manipulations led to changes that were common across people (e.g., more
+temporal clustering when words' appearances were varied; enhanced memory for
+lists following an ``event boundary;'' more feature clustering on
+order-manipulated lists; etc.). Other manipulations led to changes that were
+idiosyncratic (especially carryover effects from order manipulations; e.g.,
+participants who remembered more words on early order-manipulated lists tended
+to show stronger feature clustering for their condition's feature dimension on
+late randomly ordered lists, etc.). We also found that participants remembered
+more words from lists that were sorted to align with their idiosyncratic
+clustering preferences. Taken together, our results suggest that our memories
+are suceptable to external influences (i.e., to the statistical structure of
+ongoing experiences), but the effects of past experiences on future memory are
+largely idiosyncratic across people.
+
 \subsection*{Potential applications}

### 2023-01-17 paper/main.tex
finished complete draft of discussion section

@@
 
+Every participant in our study encountered exactly the same words, split into
+exactly the same lists. But participants' memory performance, the orders in
+which they recalled the words, and the effects of early list manipulations on
+later lists, varied according to how we presented the to-be-remembered words.
+
+Our findings raise a number of exciting questions. For example, how far might
+these manipulations be extended? In other words, might there be more
+sophisticated or clever feature or order manipulations that one might implement
+to have stronger impacts on memory? Are there limits to how much impact (on
+memory performance and/or organization) these sorts of manipulations can have?
+Are those limits universal across people, or are there individual differences
+(based on prior experiences, natural strategies, neuroanatomy, etc.) that
+impose person-specific limits on the potential impact of presentation-level
+manipulations on memory?
+
+Our findings indicate that the way word lists are presented affects how people
+remember them. To the extent that word list memory reflects memory processes
+that are relevant to real-world experiences, one could imagine potential
+real-world applications of our findings. For example, we found that
+participants remembered more words when the presentation order agreed with
+their memory fingerprints. If analogous fingerprints could be estimated for
+classroom content, perhaps they could be utilized manually by teachers, or even
+by automated content presentation systems, to optimize how and what students
+remember.
+
 % concluding remarks:
@@
 
+Our work raises deep questiong about the fundamental nature of human learning.
+What are the limits of our memory systems? How much does what we remember (and
+how we remember) depend on how we learn or experience the to-be-remembered
+content? We know that our expectations, strategies, situation models learned
+through prior experiences, and more, collectively shape how our experiences are
+remembered. But those aspects of our memory are not fixed: when we are exposed
+to the same experience in a new way, it can change how we remember
+that experience, and also how we remember, process, or perceive
+\textit{future} experiences.
+
 \bibliographystyle{apa}

### 2023-01-17 paper/main.tex
added author contributions, data and code availability, and acknowledgements statements

@@
 
+\subsection*{Author contributions}
+
+Conceptualization: JRM and ACH. Methodology: JRM and ACH. Software: JRM, KZ,
+PCF, CEF, and ACH. Analysis: JRM, PCF, and ACH. Data collection: KZ, EW, PCF,
+MRL, AMF, BJB, and CEF. Data curation and management: KZ, EW, PCF, MRL, ACH.
+Writing (original draft): JRM. Writing (review and editing): KZ, EW, PCF, MRL,
+AMF, BJB, CEF, and ACH. Supervision: JRM and ACH. Project administration: KZ,
+EW, and PCF. Funding acquisition: JRM.
+
+\subsection*{Data and code availability}
+
+All of the data analyzed in this manuscript, along with all of the code for
+carrying out the analyses may be found at
+https://github.com/ContextLab/FRFR-analyses. Code for running the non-adaptive
+expermental conditions may be found at
+https://github.com/Context\-Lab/efficient-learning-code. Code for running the
+adaptive experimental condition may be found at
+https://github.com/ContextLab/adaptiveFR. We have also released an associated
+Python toolbox for analyzing free recall data, which may be found at
+https://cdl-quail.read\-the\-docs.io/\-en/\-latest/.
+
+\subsection*{Acknowledgements}
+
+We acknowledge useful discussions, assistance in setting up an earlier version
+of this study, and assistance with some of the data collection efforts from
+Rachel Chacko, Joseph Finkelstein, Sheherzad Mohydin, Lucy Owen, Gal Perlman,
+Darya Romanova, Jake Rost, Jessica Tin, Marisol Tracy, and Peter Tran. Our work
+was supported in part by NSF CAREER Award Number 2145172 to JRM. The content
+is solely the responsibility of the authors and does not necessarily represent
+the official views of our supporting organizations. The funders had no role in
+study design, data collection and analysis, decision to publish, or preparation
+of the manuscript.
+
 \bibliographystyle{apa}

### 2023-01-17 paper/main.tex
spell check and some compilation error fixes

@@
 African American (31 participants), American Indian or Alaska Native (11
-particiapnts), Native Hawaiian or Other Pacific Islander (four participants),
+participants), Native Hawaiian or Other Pacific Islander (four participants),
 Mixed race (three participants), Middle Eastern (one participant), and Arab
@@
 features related to the \textit{meanings} of the words (semantic category,
-referrent object size), two lexicographic features related to the
+referent object size), two lexicographic features related to the
 \textit{letters} that make up the words (word length in number of letters,
@@
 constraints. First, we required that each list contained words from exactly 4
-unique categories, each with exactly 4 examplars from each category. Second, we
+unique categories, each with exactly 4 exemplars from each category. Second, we
 required that (across all words on the list) at least one instance of both
@@
 To select a random font color for each word, we drew three integers uniformly
-and at random from the interval $\[0, 255\]$, corresponding to the red (r),
+and at random from the interval $\left[0, 255\right]$, corresponding to the red (r),
 green (g), and blue (b) color channels for that word. To assign random
@@
 help interpret performance in other ``manipulation'' conditions. Two control
-conditions served as ``anchorpoints.'' In the first anchorpoint condition,
+conditions served as ``anchor points.'' In the first anchor point condition,
 which we call the \textit{feature rich} condition, we randomly shuffled the
 presentation order (independently for each participant) of the words on each
-list. In the second anchorpoint condition, which we call the \textit{reduced}
+list. In the second anchor point condition, which we call the \textit{reduced}
 condition, we randomized word presentations as in the feature rich condition.
@@
 b)}{\sum_{i=1}^n \mathrm{similarity}(a, i)}, \end{equation} where in the
-demoniator, $i$ takes on each of the $n$ feature values of the to-be-presented
+denominator, $i$ takes on each of the $n$ feature values of the to-be-presented
 words. The resulting set of normalized similarity scores sums to one.
@@
 one. To select the next to-be-presented item, we draw a random number, $r$,
-from the uniform distibution bounded between 0 and 1 (inclusive). The identity
+from the uniform distribution bounded between 0 and 1 (inclusive). The identity
 of the next item is given by the stick adjacent to an indicator that moves
@@
 precise order and blend of feature values expressed in a given list, the orders
-and numbers of correct recalls a participant makes, the number of itervening
+and numbers of correct recalls a participant makes, the number of intervening
 presentation positions between successive recalls, and so on, can all affect
@@
 length, first letter, color, and location) and their temporal clustering score.
-Conceptually, a participant's memory fingerprint describes their tendancy to
+Conceptually, a participant's memory fingerprint describes their tendency to
 order in their recall sequences (and, presumably, organize in memory) the
@@
 of the recall intervals for each list. We used the \texttt{Quail} Python
-package~\citep{HeusEtal17} to apply speech-to-text alorithms to the
+package~\citep{HeusEtal17} to apply speech-to-text algorithms to the
 just-collected data, aggregate the data for the given participant, and estimate
@@
 probabilities, for each participant. We used an analogous procedure to compute
-probabilility of $n$\textsuperscript{th} reacall curves for each participant.
+probability of $n$\textsuperscript{th} recall curves for each participant.
 Specifically, we filled in the corresponding matrices according to the
@@
 given participant had made fewer than $n$ recalls for a given list, we simply
-excluded that list from our analysis when computing that paritcipant's
+excluded that list from our analysis when computing that participant's
 curve(s).
@@
 words. We defined event boundaries (e.g., Fig.~\ref{fig:recall-dynamics}B) as
-occuring between any successive pair of words whose distances along the given
+occurring between any successive pair of words whose distances along the given
 feature dimension were greater than one standard deviation above the mean for
@@
 the meaning or semantic content of the words (e.g., word category, size of the
-refferent, etc.) and of the lexicographic properties of the words (e.g., word length,
+referent, etc.) and of the lexicographic properties of the words (e.g., word length,
 first letter, etc.). We wondered whether this additional word-independent information
@@
 = 0.067$). Participants' performance on early lists in all of the other order
-manipulation conditions was indistinguishable from performnace on the early
+manipulation conditions was indistinguishable from performance on the early
 feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$). Participants in both of
-the semanticly ordered conditions exhibited stronger temporal clustering on
+the semantically ordered conditions exhibited stronger temporal clustering on
 early lists (versus early feature rich lists; category: $t(95) = 8.508, p <
@@
 clustering on early lists across the order manipulation and feature rich
-conditions. Since results were similar across both semantic conditoins
+conditions. Since results were similar across both semantic conditions
 (category and size), both lexicographic conditions (length and first letter),
@@
 Fig.~\ref{fig:recall-dynamics}B), and participants also seemed to display an
-icnreased tendancy to initiate recall of size-ordered lists with items whose
+increased tendency to initiate recall of size-ordered lists with items whose
 study positions were integer multiples of eight. A second striking difference
@@
 to temporally cluster their recalls on category-ordered lists, as we reported
-above.) Taken together, these order-specific idiosyncracies suggest a
+above.) Taken together, these order-specific idiosyncrasies suggest a
 hierarchical set of influences on participants' memories. At longer timescales,
@@
 and first letter; yellow markers) diverge in a similar direction; and both
-visiual conditions (color and location; green) also diverge in a similar
+visual conditions (color and location; green) also diverge in a similar
 direction. But on late lists, participants' fingerprints seem to return to a
@@
 also exhibited stronger temporal clustering. For early lists
-(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}E), this trend held
+(Fig.~\ref{fig:clustering-scatterplots}E), this trend held
 overall ($r(179) = 0.924, p < 0.001$), for each condition individually (all
@@
 = 0.002$). For late lists
-(Fig.~\ref{fig:clusteringfig:clustering-scatterplots}F), the results were more
+(Fig.~\ref{fig:clustering-scatterplots}F), the results were more
 variable (overall: $r(179) = 0.348, p = 0.000$; all non-visual conditions: $r$s
@@
     \caption{\textbf{Feature clustering carryover effects.} \textbf{A.} Feature
-    clustering scores for ordder manipulation (early) versus randomly ordered
+    clustering scores for order manipulation (early) versus randomly ordered
     (late) lists. \textbf{B.} Accuracy differences (on early versus late lists)
@@
 We suggest two potential interpretations of these findings. First, it is
-possible that some participants are more ``maliable'' or ``adaptable'' with
+possible that some participants are more ``malleable'' or ``adaptable'' with
 respect to how they organize incoming information. When presented with list of
@@
 from early to late lists. These participants might demonstrate strong recall
-performance not because of their inherantly superior memory abilities, but
+performance not because of their inherently superior memory abilities, but
 rather because the specific condition they were assigned to happened to be
-especially easy for them, given their pre-experimental tendancies. To help
+especially easy for them, given their pre-experimental tendencies. To help
 distinguish between these interpretations, we designed an \textit{adaptive}
@@
 The order manipulations we examined also appeared to induce, in some cases, a
-tendancy to ``clump'' similar words within a list. This was most apparent on
+tendency to ``clump'' similar words within a list. This was most apparent on
 semantically ordered lists, where the probability of initiating recall with a
-given word seemed to follow groupings defined by feature changepoints.
+given word seemed to follow groupings defined by feature change points.
 
@@
 differently. Some participants were sensitive to our manipulations, and those
-particpants showed stronger impacts on their memory performance for the ordered
+participants showed stronger impacts on their memory performance for the ordered
 lists as well as future (random) lists. Other participants appeared relatively
@@
 second possibility described above. Specifically, each participant seems to
-bring into the experiment their own idiosynchratic preferences and strategies
+bring into the experiment their own idiosyncratic preferences and strategies
 for organizing the words in their memories. When we presented the words in an
@@
 sensory properties (i.e., feature related to the item's appearance, sound,
-smell, etc.), emotional proprties (i.e., features related to how meaningful the
+smell, etc.), emotional properties (i.e., features related to how meaningful the
 item is, whether the item evokes positive or negative feelings, etc.),
@@
 Temporally proximal features include aspects of the participant's internal or
-external experience that are \textit{not} specifically occuring at the moment
+external experience that are \textit{not} specifically occurring at the moment
 they encounter an item, but that nonetheless influence how they process the
@@
 that it is not only the contextual features themselves that affect memory, but
-also the \textit{dymamics} of context-- i.e., how the contextual features
+also the \textit{dynamics} of context-- i.e., how the contextual features
 associated with each item change over time.
@@
 % explicit priming (resolve ambiguous stimuli, e.g., green eyes study)
-% negative priming: supressing previously irrelevant stimuli or features hurts participants' recall for those stimuli when they're used on subsequent lists
+% negative priming: suppressing previously irrelevant stimuli or features hurts participants' recall for those stimuli when they're used on subsequent lists
 % implicit priming (most like here): patterns in prior experiences sets expectations about what is to come, which can affect memory for upcoming stimuli
@@
 resolve the ambiguities. We may also be overtly or covertly ``primed'' to
-influence how we are likely to resolve ambiuities. For example, before
-listening to a story with several equally plausable interpretations, providing
+influence how we are likely to resolve ambiguities. For example, before
+listening to a story with several equally plausible interpretations, providing
 participants with ``background'' information beforehand can lead them towards
@@
 conscious and unconscious biases and preferences can influence not only how we
-interpret high-level ambuities, but even how we process low-level sensory
+interpret high-level ambiguities, but even how we process low-level sensory
 information~\citep{KataEtal23}.
@@
 tasks participants encounter before studying a given list can influence what
-and how they remember. For example, when participants are directed to supress,
+and how they remember. For example, when participants are directed to suppress,
 disregard, or ignore ``distracting'' stimuli early on in an experiment,
@@
 consistently motivated to perform well~\citep{NastEtal20}. Another reason is
-that the temporal unfoldings of events and occurances in naturalistic stimuli
+that the temporal unfoldings of events and occurrences in naturalistic stimuli
 tend to be much more meaningful than the temporal unfoldings of items on
@@
 from fundamental workings of our memory systems, from some aspect of the
-stimulus, or from idiosyncratic interactions or inteference between our memory
+stimulus, or from idiosyncratic interactions or interference between our memory
 systems and the stimulus. This challenge implies that, to understand the
@@
 highly simplified stimuli that can provide relatively unbiased (compared with
-real-world experiences) measures of the relevant patterns and tendancies.
+real-world experiences) measures of the relevant patterns and tendencies.
 
@@
 clustering preferences. Taken together, our results suggest that our memories
-are suceptable to external influences (i.e., to the statistical structure of
+are susceptible to external influences (i.e., to the statistical structure of
 ongoing experiences), but the effects of past experiences on future memory are
@@
 
-Our work raises deep questiong about the fundamental nature of human learning.
+Our work raises deep questions about the fundamental nature of human learning.
 What are the limits of our memory systems? How much does what we remember (and
@@
 https://github.com/ContextLab/FRFR-analyses. Code for running the non-adaptive
-expermental conditions may be found at
+experimental conditions may be found at
 https://github.com/Context\-Lab/efficient-learning-code. Code for running the

### 2023-01-17 paper/main.tex
corrected error in methods section (describing how embeddings were computed)

@@
 
-\textbf{JRM NOTE: REMINDER TO CHECK THIS PARAGRAPH AGAINST ANALYSIS CODE FOR
-ACCURACY...} Following some of our prior work~\citep{HeusEtal18a, HeusEtal21},
-we use low-dimensional embeddings to help visualize how participants' memory
+Following some of our prior work~\citep{HeusEtal18a, HeusEtal21}, we use
+low-dimensional embeddings to help visualize how participants' memory
 fingerprints change across lists
@@
 compute a shared embedding space across participants and experimental
-conditions, we concatenated the full set of fingerprints (across all lists,
-participants, and experimental conditions) to create a large matrix with
-number-of-lists $\times$ number-of-participants rows and seven columns (one for
-each feature clustering score, plus an additional temporal clustering score
-column). We used principal components analysis to project the seven-dimensional
-observations into a two-dimensional space (using the two principal components
-that explained the most variance in the data). For two visualizations
+conditions, we concatenated the full set of fingerprints (across all list
+groupings, participants, and experimental conditions) to create a large matrix
+with number-of-list-groupings $\times$ number-of-participants rows and seven
+columns (one for each feature clustering score, plus an additional temporal
+clustering score column). We used principal components analysis to project the
+seven-dimensional observations into a two-dimensional space (using the two
+principal components that explained the most variance in the data). For two
+visualizations
 (Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B)
-we computed an additional set of two-dimensional embeddings for participants'
-\textit{average} fingerprints (i.e., across lists within a given group of
-lists-- early or late). For those visualizations we averaged across the rows
-(for each condition and group of lists) in the combined fingerprint matrix
-prior to projecting it into the shared two-dimensional space. This yielded a
-single two-dimensional coordinate for each \textit{list group}, rather than for
-each individual list. We used these embeddings solely for visualization. All
+we computed an additional set of two-dimensional embeddings for the
+\textit{average} fingerprints across lists within a given list grouping (i.e.,
+early or late). For those visualizations, we averaged across the rows (for each
+condition and group of lists) in the combined fingerprint matrix prior to
+projecting it into the shared two-dimensional space. This yielded a single
+two-dimensional coordinate for each \textit{list group}, rather than for each
+individual list. We used these embeddings solely for visualization. All
 statistical tests were carried out in the original (seven-dimensional) feature

### 2023-01-17 paper/main.tex
inputting a round of written edits

@@
 
-\title{\Large Carryover effects in free recall reveal how past experiences influence
-memories of future experiences}
+\title{\Large Feature and order manipulations in a free recall task affect memory for current and future lists}
 
@@
 
-\begin{abstract} \footnotesize{We perceive, interpet, and remember ongoing experiences
-through the lens of our prior experiences. Inferring that we are in one type of
-situation versus another can lead us to interpret the same physical experience
-differently. In turn, this can affect how we focus our attention, form
-expectations about what will happen next, remember what is happening now, draw on
-our prior related experiences, and so on. To study these phenomena, we asked
-participants to perform simple word list learning tasks. Across different
+\begin{abstract} \footnotesize{We perceive, interpet, and remember ongoing
+experiences through the lens of our prior experiences. Inferring that we are in
+one type of situation versus another can lead us to interpret the same physical
+experience differently. In turn, this can affect how we focus our attention,
+form expectations about what will happen next, remember what is happening now,
+draw on our prior related experiences, and so on. To study these phenomena, we
+asked participants to perform simple word list learning tasks. Across different
 experimental conditions, we held the set of to-be-learned words constant, but
-we manipulated the orders in which the words were studied. We found that these
-order manipulations affected not only how the participants recalled the ordered
-lists, but also how they recalled later randomly ordered lists. Our work shows
-how structure in our ongoing experiences can exert influence on how we remember
-unrelated subsequent experiences.
+we manipulated how irrelevant visual features changed across words and lists,
+along with the orders in which the words were studied. We found that these
+manipulations affected not only how the participants recalled the manipulated
+lists, but also how they recalled later (random) lists. Our work shows how
+structure in our ongoing experiences can exert influence over how we remember
+our current experiences and unrelated subsequent experiences.
 
@@
 The contexts we encounter help us to construct \textit{situation
-models}~\citep{RangRitc12, MannEtal15} or \textit{schemas}~\citep{MasiEtal22,
-BaldEtal18} that describe how experiences are likely to unfold based on our
-prior experiences with similar contextual cues. For example, when we enter a
-sit-down restaurant, we might expect to be seated at a table, given a menu, and
-served food. Priming someone to expect a particular situation or context can
-also influence how they resolve potential ambiguities in their ongoing
-experiences, including ambiguous movies and narratives~\citep{YeshEtal17}.
+models}~\citep{RangRitc12, MannEtal15, RadvCope06, ZwaaEtal95} or
+\textit{schemas}~\citep{MasiEtal22, BaldEtal18} that describe how experiences
+are likely to unfold based on our prior experiences with similar contextual
+cues. For example, when we enter a sit-down restaurant, we might expect to be
+seated at a table, given a menu, and served food. Priming someone to expect a
+particular situation or context can also influence how they resolve potential
+ambiguities in their ongoing experiences, including in ambiguous movies and
+narratives~\citep{YeshEtal17}.
 
@@
 narratives and movies~\citep{ZwaaEtal95,ZwaaRadv98, NastEtal20}. In contrast,
-our understanding of how we organize our memories has been most widely studied
-using more traditional paradigms like free recall of random word
-lists~\citep{Kaha12, Kaha20}. In free recall, participants study lists of items and are
-instructed to recall the items in any order they choose. The orders in which
-words come to mind can provide insights into how participants have organized
-their memories of the studied words. Because random word lists are unstructured
-by design, it is not clear if or how non-trivial situation models might apply
-to these stimuli. Nevertheless, there are \textit{some} commonalities between
-memory for word lists and memory for real-world experiences.
+our understanding of how we organize our memories has been most widely informed
+by more traditional paradigms like free recall of random word
+lists~\citep{Kaha12, Kaha20}. In free recall, participants study lists of items
+and are instructed to recall the items in any order they choose. The orders in
+which words come to mind can provide insights into how participants have
+organized their memories of the studied words. Because random word lists are
+unstructured by design, it is not clear if or how non-trivial situation models
+might apply to these stimuli. Nevertheless, there are \textit{some}
+commonalities between memory for word lists and memory for real-world
+experiences.
 
@@
 items from non-list items. Understanding the role of context in memory
-processes is particularly important in self-cued memory tasks, such as
-\textit{free recall}, where the retrieval cue is ``context'' itself.
-Conceptually, the same general processes might be said to describe how
-real-world contexts evolve during natural experiences. However, this is still
-an open area of study~\citep{Mann20, Mann21a}.
+processes is particularly important in self-cued memory tasks, such as free
+recall, where the retrieval cue is ``context'' itself. Conceptually, the same
+general processes might be said to describe how real-world contexts evolve
+during natural experiences. However, this is still an open area of
+study~\citep{Mann20, Mann21a}.
 
@@
 This effect can be seen in the tendency for people to successively recall items
-that occupied neighboring positions in the study list~\citep{Kaha96}. There are
-also striking effects of semantic clustering~\citep{RomnEtal93, Bous53,
+that occupied neighboring positions in the studied list~\citep{Kaha96}. There
+are also striking effects of semantic clustering~\citep{RomnEtal93, Bous53,
 BousEtal54, JenkRuss52, MannKaha12}, whereby the recall of a given item is more
@@
 or unrelated one. In general, people organize memories for words along a wide
-variety of stimulus dimensions. As formalized by models like the \textit{Context
-Maintenance and Retrieval Model}~\citep{PolyEtal09}, the stimulus features
-associated with each word (e.g.\ the word's meaning, font size, font color,
-location on the screen, size of the object the word represents, etc.) are
-incorporated into the participant's mental context
+variety of stimulus dimensions. As formalized by models like the
+\textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the
+stimulus features associated with each word (e.g.\ the word's meaning, font
+size, font color, location on the screen, size of the object the word
+represents, etc.) are incorporated into the participant's mental context
 representation~\citep{SmitVela01, MannEtal15, Mann20, MannEtal11, MannEtal12}.
@@
 (Fig.~\ref{fig:exp}). Across the different conditions in the experiment, we
-varied the lists' presentation orders in different ways across lists. The
-studied items (words) were designed to vary along three general dimensions:
-semantic (word \textit{category}, and physical \textit{size} of the referent),
-lexicographic (word \textit{length} and \textit{first letter}), and visual
-(font \textit{color} and the onscreen \textit{location} of each word). In our
-main manipulation conditions, we asked participants to study and recall eight
-lists whose items were sorted by a target feature (e.g., word category). Next,
-we asked them to study and recall an additional eight lists whose items had the
-same features, but that were sorted in a random temporal order. We were
-interested in how these order manipulations affected participants' recall
-behaviors on early (sorted) lists, as well as how order manipulations on early
-lists affected recall behaviors on later (unsorted) lists. We used a series of
-control conditions as a baseline; in these control conditions all of the lists
-were sorted randomly, but we manipulated the presence or absence of the visual
-features. Finally, in an \textit{adaptive} experimental condition we used
-participants' recall behaviors on early lists to manipulate, in real-time, the
-presentation orders of subsequent lists. In this adaptive condition we varied
-the agreement between how participants preferred to organize their memories of
-the studied items versus the orders in which the items were presented.
+varied the lists' appearances and presentation orders in different ways across
+lists. The studied items (words) were designed to vary along three general
+dimensions: semantic (word \textit{category}, and physical \textit{size} of the
+referent), lexicographic (word \textit{length} and \textit{first letter}), and
+visual (font \textit{color} and the onscreen \textit{location} of each word).
+In two conditions, we manipulated whether the words' appearances were fixed or
+variable within each list. In six manipulation conditions, we asked
+participants to study and recall eight lists whose items were sorted by a
+target feature (e.g., word category). Next, we asked them to study and recall
+an additional eight lists whose items had the same features, but that were
+sorted in a random temporal order. We were interested in how these 
+manipulations affected participants' recall behaviors on early (manipulated) lists,
+as well as how order manipulations on early lists affected recall behaviors on
+later (random) lists. We used two control conditions as a baseline;
+in these control conditions all of the lists were sorted randomly, but we
+manipulated the presence or absence of the visual features. Finally, in an
+\textit{adaptive} experimental condition we used participants' recall behaviors
+on early lists to manipulate, in real-time, the presentation orders of
+subsequent lists. In this adaptive condition we varied the agreement between
+how participants preferred to organize their memories of the studied items
+versus the orders in which the items were presented.
 
@@
 We enrolled a total of 491 Dartmouth undergraduate students across 11
-experimental conditions. The conditions included two primary controls (feature
-rich, reduced), two secondary controls (reduced (early), reduced (late)), six
-order manipulation conditions (category, size, length, first letter, color, and
-location), and a final adaptive condition. Each of these conditions are
-described in the \textit{Experimental design} subsection below.
+experimental conditions. The conditions included two controls (feature rich,
+reduced), two visual manipulation conditions (reduced (early) and reduced
+(late)), six order manipulation conditions (category, size, length, first
+letter, color, and location), and a final adaptive condition. Each of these
+conditions are described in the \textit{Experimental design} subsection below.
 
@@
 (one participant). A total of five participants declined to report their race.
-We note that several participants reported more than one of racial category.
-Participants reported their highest degrees achieved as ``Some college'' (359
-participants), ``High school graduate'' (117 participants), ``College
-graduate'' (seven participants), ``Some high school'' (five participants),
-``Doctorate'' (one participant), and ``Master's degree'' (one participant). A
-total of 482 participants reported no reading impairments, and eight reported
-having mild reading impairments. A total of 489 participants reported having
-normal color vision and one participant reported that they were red-green color
-blind. A total of 482 participants reported taking no prescription medications
-and having no recent injuries; four participants reported having ADHD, one reported
-having dyslexia, one reported having allergies, one reported a recently torn
-ACL/MCL, and one reported a concussion from several months prior. The
-participants reported consuming 0 -- 3 cups of coffee prior to the testing
-session (mean: 0.32 cups; standard deviation: 0.58 cups). Participants reported
-their current level of alertness, and we converted their responses to numerical
-scores as follows: ``very sluggish'' (-2), ``a little sluggish'' (-1),
-``neutral'' (0), ``a little alert'' (1), and ``very alert'' (2). Across all
-participants, the full range of alertness levels were reported (range: -2 -- 2;
-mean: 0.35; standard deviation: 0.89).
+We note that several participants reported more than one of the above racial
+categories. Participants reported their highest degrees achieved as ``Some
+college'' (359 participants), ``High school graduate'' (117 participants),
+``College graduate'' (seven participants), ``Some high school'' (five
+participants), ``Doctorate'' (one participant), and ``Master's degree'' (one
+participant). A total of 482 participants reported no reading impairments, and
+eight reported having mild reading impairments. A total of 489 participants
+reported having normal color vision and one participant reported that they were
+red-green color blind. A total of 482 participants reported taking no
+prescription medications and having no recent injuries; four participants
+reported having ADHD, one reported having dyslexia, one reported having
+allergies, one reported a recently torn ACL/MCL, and one reported a concussion
+from several months prior. The participants reported consuming 0 -- 3 cups of
+coffee prior to the testing session (mean: 0.32 cups; standard deviation: 0.58
+cups). Participants reported their current level of alertness, and we converted
+their responses to numerical scores as follows: ``very sluggish'' (-2), ``a
+little sluggish'' (-1), ``neutral'' (0), ``a little alert'' (1), and ``very
+alert'' (2). Across all participants, the full range of alertness levels were
+reported (range: -2 -- 2; mean: 0.35; standard deviation: 0.89).
 
@@
 experimental condition as follows: feature rich (67 participants), reduced (61
-participants), reduced (late) (41 participants), reduced (early), (42
+participants), reduced (early), (42 participants), reduced (late) (41
 participants), category (30 participants), size (30 participants), length (30
@@
 word features and/or presentation orders within and/or across lists. The rows
-display representative (illustrated) examples of the first lists participants
-might encounter in each condition. The rectangles during the ``Presentation
-phase'' show illustrated screen captures during a series of word presentations.
-Each word appeared onscreen for 2 seconds, followed by 2 seconds of blank
-screen. The red microphone icons during the ``Recall'' phase denote the one
-minute verbal recall interval. The labels on the right (and corresponding
-groupings on the left) denote experimental condition labels.}
+display representative (illustrated) examples of items from the first lists
+participants might encounter in each condition. The rectangles during the
+``Presentation phase'' show illustrated screen captures during a series of word
+presentations. Each word appeared onscreen for 2 seconds, followed by 2 seconds
+of blank screen. The red microphone icons during the ``Recall'' phase denote
+the one minute verbal recall interval. The labels on the right (and
+corresponding groupings on the left) denote experimental condition labels.}
 
@@
 We assigned the categorized words into a total of 16 lists with several
-constraints. First, we required that each list contained words from exactly 4
-unique categories, each with exactly 4 exemplars from each category. Second, we
-required that (across all words on the list) at least one instance of both
-object sizes were represented. On average, each category was represented in
-4.27 lists (standard deviation: 1.16 lists). Aside from these two constraints,
-we assigned each word to a unique list. After random assignment, each list
-contained words with an average of 11.13 unique starting letters (standard
-deviation: 1.15 letters) and an average word length of 6.17 letters (standard
-deviation: 0.34 letters).
+constraints. First, we required that each list contained words from exactly
+four unique categories, each with exactly four exemplars from each category.
+Second, we required that (across all words on the list) at least one instance
+of both object sizes were represented. On average, each category was
+represented in 4.27 lists (standard deviation: 1.16 lists). Aside from these
+two constraints, we assigned each word to a unique list. After random
+assignment, each list contained words with an average of 11.13 unique starting
+letters (standard deviation: 1.15 letters) and an average word length of 6.17
+letters (standard deviation: 0.34 letters).
 
@@
 To select a random font color for each word, we drew three integers uniformly
-and at random from the interval $\left[0, 255\right]$, corresponding to the red (r),
-green (g), and blue (b) color channels for that word. To assign random
+and at random from the interval $\left[0, 255\right]$, corresponding to the red
+(r), green (g), and blue (b) color channels for that word. To assign random
 presentation locations to each word, we selected two floating point numbers
-uniformly at random (one for the word's horizontal $x$ coordinate and the other
-for its vertical $y$ coordinate). The bounds of these coordinates were selected
-to cover the entire visible area of the display without cutting off any part of
-the words. The words were shown on 27~in (diagonal) Retina 5K iMac displays
-(resolution: 5120 $\times$ 2880 pixels).
+uniformly and at random (one for the word's horizontal $x$ coordinate and the
+other for its vertical $y$ coordinate). The bounds of these coordinates were
+selected to cover the entire visible area of the display without cutting off
+any part of the words. The words were shown on 27~in (diagonal) Retina 5K iMac
+displays (resolution: 5120 $\times$ 2880 pixels).
 
@@
 
-We used four ``control'' conditions to evaluate and explore participants'
+We used two ``control'' conditions to evaluate and explore participants'
 baseline behaviors. We also used performance on these control conditions to
-help interpret performance in other ``manipulation'' conditions. Two control
-conditions served as ``anchor points.'' In the first anchor point condition,
-which we call the \textit{feature rich} condition, we randomly shuffled the
-presentation order (independently for each participant) of the words on each
-list. In the second anchor point condition, which we call the \textit{reduced}
-condition, we randomized word presentations as in the feature rich condition.
-However, rather than assigning each word a random color and location, we
-instead displayed all of the words in black and at the center of the screen.
-
-In the \textit{reduced (early)} condition, we followed the ``reduced''
-procedure (presenting each word in black at the center of the screen) for early
-lists, and followed the ``feature rich'' procedure (presenting each word in a
-random color and location) for late lists. Finally, in the \textit{reduced
-(late)} condition, we followed the feature rich procedure for early lists and
-the reduced procedure for late lists.
+help interpret performance in other ``manipulation'' conditions. In the first
+control condition, which we call the \textit{feature rich} condition, we
+randomly shuffled the presentation order (independently for each participant)
+of the words on each list. In the second control condition, which we call the
+\textit{reduced} condition, we randomized word presentations as in the feature
+rich condition. However, rather than assigning each word a random color and
+location, we instead displayed all of the words in black and at the center of
+the screen.
+
+We also designed two conditions where we varied the words' visual appearnaces
+across lists. In the \textit{reduced (early)} condition, we followed the
+``reduced'' procedure (presenting each word in black at the center of the
+screen) for early lists, and followed the ``feature rich'' procedure
+(presenting each word in a random color and location) for late lists. Finally,
+in the \textit{reduced (late)} condition, we followed the feature rich
+procedure for early lists and the reduced procedure for late lists.
 
@@
 quantifying the dissimilarity between each pair of features. This function
-varied according to the type of features. Semantic features (category and size)
-are \textit{categorical}. For these features, we defined a binary distance
-function: two words were considered to ``match'' (i.e., have a distance of 0)
-if their labels are the same (i.e., both from the same semantic category or
-both of the same size). If two words' labels were different for a given
-feature, we defined the words to have a distance of 1 for that feature.
+varied according to the type of feature under consideration. Semantic features
+(category and size) are \textit{categorical}. For these features, we defined a
+binary distance function: two words were considered to ``match'' (i.e., have a
+distance of 0) if their labels were the same (i.e., both from the same semantic
+category or both of the same size). If two words' labels were different for a
+given feature, we defined the words to have a distance of 1 for that feature.
 Lexicographic features (length and first letter) are \textit{discrete}. For
@@
 location vectors, respectively. Therefore the color and location distance
-measures always take on positive real values (upper-bounded at 441.67 for
+measures always take on non-negative real values (upper-bounded at 441.67 for
 color, or 27~in for location, reflecting the distances between the
@@
 stochastic algorithm for (noisily) sorting the words. The stochastic aspect of
-our sorting procedure enabled us to obtain unique lists for each participant.
-First, we choose a word uniformly at random from the set of candidates. Next,
-we compute the distances between the chosen word's feature(s) and the
-corresponding feature(s) of all yet-to-be-presented words. Third, we convert
-these distances (between the previously presented word's feature values, $a$,
-and the candidate word's feature values, $b$) to similarity scores:
-\begin{equation} \mathrm{similarity}(a, b) = \exp\{-\tau \cdot
+our sorting procedure enabled us to obtain unique orderings for each
+participant. First, we choose a word uniformly and at random from the set of
+candidates. Next, we compute the distances between the chosen word's feature(s)
+and the corresponding feature(s) of all yet-to-be-presented words. Third, we
+convert these distances (between the previously presented word's feature
+values, $a$, and the candidate word's feature values, $b$) to similarity
+scores: \begin{equation} \mathrm{similarity}(a, b) = \exp\{-\tau \cdot
 \mathrm{distance}(a, b)\}, \label{eqn:similarity} \end{equation} where $\tau =
@@
 denominator, $i$ takes on each of the $n$ feature values of the to-be-presented
-words. The resulting set of normalized similarity scores sums to one.
+words. The resulting set of normalized similarity scores sums to 1.
 
@@
 $x$, and all yet-to-be-presented items ($a$ -- $g$). Next, we normalize these
-similarity scores so that they sum to one. We lay, in sequence, a set of
+similarity scores so that they sum to 1. We lay, in sequence, a set of
 ``sticks,'' one for each candidate item, whose lengths are equal to these
-normalized similarity scores. Note that the combined lengths of these sticks is
-one. To select the next to-be-presented item, we draw a random number, $r$,
-from the uniform distribution bounded between 0 and 1 (inclusive). The identity
-of the next item is given by the stick adjacent to an indicator that moves
-distance $r$ (starting from 0) along the sequence of sticks. In this case, the
-next to-be-presented item is $e$. Note that each item's chances of selection is
-proportional to its similarity to the previous item, along the given feature
-dimension.} \label{fig:stick}
+normalized similarity scores. To select the next to-be-presented item, we draw
+a random number, $r$, from the uniform distribution bounded between 0 and 1
+(inclusive). The identity of the next item is given by the stick adjacent to an
+indicator that moves distance $r$ (starting from 0) along the sequence of
+sticks. In this case, the next to-be-presented item is $e$. Note that each
+item's chances of selection is proportional to its similarity to the previous
+item, along the given feature dimension.} \label{fig:stick}
 
@@
 the next to-be-presented word by moving an indicator along the set of sticks,
-by a distance chosen uniformly at random on the interval $\left[0, 1\right]$.
-We select the word associated with the stick lying next to the indicator to be
-presented next. This process continues iteratively (re-computing the similarity
-scores and stochastically choosing the next to-be-presented word using the
-just-presented word) until all of the words have been presented. The result is
-an ordered list that tends to change gradually along the selected feature
-dimension.
+by a distance chosen uniformly and at random on the interval $\left[0,
+1\right]$. We select the word associated with the stick lying next to the
+indicator to be presented next. This process continues iteratively
+(re-computing the similarity scores and stochastically choosing the next
+to-be-presented word using the just-presented word) until all of the words have
+been presented. The result is an ordered list that tends to change gradually
+along the selected feature dimension.
 
@@
 (or in exact reverse order), this would yield a score of 1. If a participant
-recalled the words in random order, this would yield an expected score of 0.5.
-For each recall transition (and separately for each participant), we sorted all
-not-yet-recalled words according to their absolute lag (that is, distance away
-in the list). We then computed the percentile rank of the next word the
-participant recalled. We took an average of these percentile ranks across all
-of the participant’s recalls to obtain a single (uncorrected) temporal
-clustering score for the participant.
+recalled the words in a random order, this would yield an expected score of
+0.5. For each recall transition (and separately for each participant), we
+sorted all not-yet-recalled words according to their absolute lag (that is,
+distance away in the list). We then computed the percentile rank of the next
+word the participant recalled. We took an average of these percentile ranks
+across all of the participant’s recalls to obtain a single (uncorrected)
+temporal clustering score for the participant.
 
@@
 \textit{any} correctly recalled list $A$ word will be large. In contrast,
-successively recalling several list $B$ items \textit{could} be meaningful,
-since (early in the recall sequence) the yet-to-be-recalled items come from a
-mix of sizes. However, once all of the small items on list $B$ have been
-recalled, the best possible next matching recall will be a large item. And all
-subsequent correct recalls must also be large items-- so for those later
-recalls it becomes difficult to determine whether the participant is
-successively recalling large items because they are organizing their
+successively recalling several list $B$ items of the same size \textit{could}
+be meaningful, since (early in the recall sequence) the yet-to-be-recalled
+items come from a mix of sizes. However, once all of the small items on list
+$B$ have been recalled, the best possible next matching recall will be a large
+item. And all subsequent correct recalls must also be large items-- so for
+those later recalls it becomes difficult to determine whether the participant
+is successively recalling large items because they are organizing their
 memories according to size, or (alternatively), whether they are simply
@@
 the range of clustering scores that are possible to observe for a given list.
-An uncorrected clustering score therefore conflates participants' actual
-memory organization with other ``nuisance'' factors.
+An uncorrected clustering score therefore conflates participants' actual memory
+organization with other ``nuisance'' factors.
 
@@
 scores}) could be estimated independently from the others. This enabled us to
-make use of the testing computers' multi-core CPU architectures by elements of
-the null distributions in batches of eight (i.e., the number of CPU cores on
-each testing computer). Taken together, we were able to compress the relevant
-computations into just a few seconds of computing time. The combined processing
-time for the speech-to-text algorithm, fingerprint computations, and
-permutation-based ordering procedure (described next) easily fit within the
-inter-list intervals, where participants paused for a self-paced break before
-moving on to study and recall the next list.
+make use of the testing computers' multi-core CPU architectures by considering
+(in parallel) elements of the null distributions in batches of eight (i.e., the
+number of CPU cores on each testing computer). Taken together, we were able to
+compress the relevant computations into just a few seconds of computing time.
+The combined processing time for the speech-to-text algorithm, fingerprint
+computations, and permutation-based ordering procedure (described next) easily
+fit within the inter-list intervals, where participants paused for a self-paced
+break before moving on to study and recall the next list.
 
@@
 compute a shared embedding space across participants and experimental
-conditions, we concatenated the full set of fingerprints (across all list
-groupings, participants, and experimental conditions) to create a large matrix
-with number-of-list-groupings $\times$ number-of-participants rows and seven
-columns (one for each feature clustering score, plus an additional temporal
-clustering score column). We used principal components analysis to project the
-seven-dimensional observations into a two-dimensional space (using the two
-principal components that explained the most variance in the data). For two
-visualizations
+conditions, we concatenated the full set of across-participant average
+fingerprints (for all lists and experimental conditions) to create a large
+matrix with number-of-lists (16) $\times$ number-of-conditions (10, encluding
+the adaptive condition) rows and seven columns (one for each feature clustering
+score, plus an additional temporal clustering score column). We used principal
+components analysis to project the seven-dimensional observations into a
+two-dimensional space (using the two principal components that explained the
+most variance in the data). For two visualizations
 (Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B)
@@
 projecting it into the shared two-dimensional space. This yielded a single
-two-dimensional coordinate for each \textit{list group}, rather than for each
-individual list. We used these embeddings solely for visualization. All
-statistical tests were carried out in the original (seven-dimensional) feature
-spaces.
+two-dimensional coordinate for each \textit{list group} (in each condition),
+rather than for each individual list. We used these embeddings solely for
+visualization. All statistical tests were carried out in the original
+(seven-dimensional) feature spaces.
 
@@
 (for each participant) a number-of-lists (16) by number-of-words-per-list (16)
-matrix of zeros. Then, for each list, we found the index of the word that was
+matrix of 0s. Then, for each list, we found the index of the word that was
 recalled first, and we filled in that position in the matrix with a 1. Finally,
@@
 excluded that list from our analysis when computing that participant's
-curve(s).
+curve(s). The probability of first recall curve corresponds to a special case
+where $n = 1$.
 
@@
 indicates that a recalled item was presented immediately after the previously
-recalled item, and a lag of $-3$ indicates that a recalled item came three items
-before the previously recalled item. For each recall transition (following the
-first recall), we computed the lag between the just-recalled word's
-presentation position and the next-recalled word's presentation position. We
-computed the proportions of transitions (between successively recalled words)
-for each lag, normalizing for the total numbers of possible transitions. In
-carrying out this analysis, we excluded all incorrect recalls and successive
-repetitions (e.g., recalling the same word twice in a row). This yielded, for
-each list, a 1 by number-of-lags ($-15$ to $+15$; 30 lags in total, excluding
-lags of 0) array of conditional probabilities. We averaged these probabilities
-across lists to obtain a single lag-CRP for each participant.
+recalled item, and a lag of $-3$ indicates that a recalled item came three
+items before the previously recalled item. For each recall transition
+(following the first recall), we computed the lag between the just-recalled
+word's presentation position and the next-recalled word's presentation
+position. We computed the proportions of transitions (between successively
+recalled words) for each lag, normalizing for the total numbers of possible
+transitions. In carrying out this analysis, we excluded all incorrect recalls
+and successive repetitions (e.g., recalling the same word twice in a row). This
+yielded, for each list, a 1 by number-of-lags ($-15$ to $+15$; 30 lags in
+total, excluding lags of 0) array of conditional probabilities. We averaged
+these probabilities across lists to obtain a single lag-CRP for each
+participant. Because transitions at large absolute lags are rare, these curves
+are typically displayed using range restrictions~\citep{Kaha12}.
 
@@
 encoding. For each participant, we initialized a number-of-lists (16) by
-number-of-words-per-list (16) matrix of zeros. Then, for each correct recall,
-we identified the presentation position of the word and entered a 1 into that
+number-of-words-per-list (16) matrix of 0s. Then, for each correct recall, we
+identified the presentation position of the word and entered a 1 into that
 position (row: list; column: presentation position) in the matrix. This
@@
 at each position, on each list, were recalled by the participant (depending on
-whether the corresponding entires were set to one or zero). Finally, we
-averaged over the rows of the matrix to yield a 1 by 16 array representing the
+whether the corresponding entires were set to 1 or 0). Finally, we averaged
+over the rows of the matrix to yield a 1 by 16 array representing the
 proportion of words at each position that the participant remembered.
@@
 temporal clustering (reduced (early): $t(41) = 0.998, p = 0.324$; reduced
-(late): $t(40) = 1.099, p = 0.278$), nor feature based clustering (reduced
+(late): $t(40) = 1.099, p = 0.278$), nor feature-based clustering (reduced
 (early), category: $t(41) = 0.753, p = 0.456$; reduced (early), size: $t(41) =
@@
 visual features to \textit{any} lists (comparisons of clustering on feature
-rich and reduced lists are reported above; temporal clustering in reduced
+rich anversusd reduced lists are reported above; temporal clustering in reduced
 versus reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
--9.780$, $p$s $< 0.001$; feature based clustering in reduced versus reduced
+-9.780$, $p$s $< 0.001$; feature-based clustering in reduced versus reduced
 (early) and reduced versus reduced (late) conditions: $t$s $\leq -5.443$, $p$s
@@
 clustering in feature rich versus reduced (early) and feature rich versus
-reduced (late) conditions: $t$s $\geq -1.434$, $p$s $\geq 0.154$; feature based
+reduced (late) conditions: $t$s $\geq -1.434$, $p$s $\geq 0.154$; feature-based
 clustering in feature rich versus reduced (early) and feature rich versus
@@
 ordering of different stimulus features while holding the features themselves
-constant. We hypothesized that changing the order in which participants were
+constant. We hypothesized that changing the orders in which participants were
 exposed to the words on a given list might enhance (or diminish) the relative
@@
 constant across lists (e.g., unlike in the reduced (early) and reduced (late)
-conditions, where visual features were added or removed from a subset of the
-lists participants studied).
+conditions, where variations in visual features were added or removed from a
+subset of the lists participants studied).
 
@@
 See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
-random (control) and adaptive conditions. \textbf{B.} Proportion of event
-boundaries (see \textit{Identifying event boundaries}) for each condition's
-feature of focus, plotted as a function of presentation position.}
+random and adaptive conditions. \textbf{B.} Proportion of event boundaries (see
+\textit{Identifying event boundaries}) for each condition's feature of focus,
+plotted as a function of presentation position.}
 
@@
 
-Across six order manipulation conditions, we sorted early lists by each feature
-dimension but randomly ordered the items on late lists (see \textit{Order
-manipulation conditions}; features: category, size, length, first letter,
-color, and location). Participants in the category-ordered condition showed an
-increase in memory performance on early lists (accuracy, relative to early
-feature rich lists; $t(95) = 3.034, p = 0.003$). Participants in the
+Across each of six order manipulation conditions, we sorted early lists by one
+feature dimension but randomly ordered the items on late lists (see
+\textit{Order manipulation conditions}; features: category, size, length, first
+letter, color, and location). Participants in the category-ordered condition
+showed an increase in memory performance on early lists (accuracy, relative to
+early feature rich lists; $t(95) = 3.034, p = 0.003$). Participants in the
 color-ordered condition also showed a trending increase in memory performance
 on early lists (again, relative to early feature rich lists: $t(96) = 1.850, p
-= 0.067$). Participants' performance on early lists in all of the other order
-manipulation conditions was indistinguishable from performance on the early
+= 0.067$). Participants' performances on early lists in all of the other order
+manipulation conditions were indistinguishable from performance on the early
 feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$). Participants in both of
@@
 clustering on early lists across the order manipulation and feature rich
-conditions. Since results were similar across both semantic conditions
+conditions. Since these results were similar across both semantic conditions
 (category and size), both lexicographic conditions (length and first letter),
@@
 conditions that manipulated each of these three feature groupings in our
-comparisons to simplify the presentation. On early lists, participants in the
+comparisons, to simplify the presentation. On early lists, participants in the
 semantically ordered conditions exhibited stronger semantic clustering relative
@@
 relative to feature rich participants, and on early lists; color: $t(126) =
-2.099, p = 0.038$; location: $t(126) = 4.392, p = 0.000$), but showed now
+2.099, p = 0.038$; location: $t(126) = 4.392, p < 0.001$), but showed now
 reliable differences in semantic (category: $t(126) = 0.204, p = 0.839$; size:
@@
 feature dimensions roughly unchanged (i.e., semantic manipulations did not
-affect lexicographic or color clustering, and so on).
+affect lexicographic or visual clustering, and so on).
 
@@
 bottom) lists. See Figures~\fingerprintsRandom~and~\fingerprintsAdaptive~for
-analogous plots for the random (control) and adaptive conditions.}
+analogous plots for the random and adaptive conditions.}
 \label{fig:fingerprints}
@@
 
-When we closely examined the sequences of words participants recalled in early
-order manipulated lists (Fig.~\ref{fig:recall-dynamics}A, top panel), we
+When we closely examined the sequences of words participants recalled from
+early order-manipulated lists (Fig.~\ref{fig:recall-dynamics}A, top panel), we
 noticed several differences from the dynamics of participants' recalls of
-randomly ordered lists (Figs.~\dynamicsRandom,~\recallInit). One striking difference is that
-participants in the category condition (dark purple curves,
+randomly ordered lists (Figs.~\dynamicsRandom,~\recallInit). One difference is
+that participants in the category condition (dark purple curves,
 Fig.~\ref{fig:recall-dynamics}) most often initiated recall with the
@@
 lists better than lexicographically sorted lists ($t(119) = 2.145, p = 0.034$).
-However, participants showed no reliable differences in recall performance on
-semantically versus visually sorted lists ($t(119) = 0.113, p = 0.910$).
-Participants temporally clustered semantically sorted lists more strongly than
-either lexicographically ($t(118) = 5.572, p < 0.001$) or visually ($t(119) =
-6.215, p < 0.001$) sorted lists, but did not show reliable differences in
-temporal clustering on lexicographically versus visually sorted lists ($t(119)
-= 0.189, p = 0.850$). Participants also showed reliably more semantic
-clustering on semantically sorted lists than lexicographically (category:
-$t(118) = 3.492, p = 0.001$, size: $t(118) = 3.972, p < 0.001$) or visually
-(category: $t(119) = 2.702, p = 0.008$, size: $t(119) = 4.230, p < 0.001$)
-sorted lists; more lexicographic clustering on lexicographically sorted lists
-than semantically (length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) =
-3.686, p = 0.000$) or visually (length: $t(119) = 3.024, p = 0.003$; first
-letter: $t(119) = 2.644, p = 0.009$) sorted lists; and more visual clustering
-on visually sorted lists than semantically (color: $t(119) = -2.659, p =
-0.009$; location: $t(119) = -4.604, p = 0.000$) or lexicographically (color:
-$t(119) = -2.366, p = 0.020$; location: $t(119) = -4.265, p < 0.001$) sorted
-lists. In summary, sorting lists by different features appeared to have
-slightly different effects on overall memory performance and temporal
-clustering, and people tended to cluster their recalls along a given feature
-dimension more when the studied lists were (versus were not) sorted along that
-dimension.
+However, participants showed no reliable differences in recall for semantically
+versus visually sorted lists ($t(119) = 0.113, p = 0.910$). Participants
+temporally clustered semantically sorted lists more strongly than either
+lexicographically ($t(118) = 5.572, p < 0.001$) or visually ($t(119) = 6.215, p
+< 0.001$) sorted lists, but did not show reliable differences in temporal
+clustering on lexicographically versus visually sorted lists ($t(119) = 0.189,
+p = 0.850$). Participants also showed reliably more semantic clustering on
+semantically sorted lists than lexicographically (category: $t(118) = 3.492, p
+= 0.001$, size: $t(118) = 3.972, p < 0.001$) or visually (category: $t(119) =
+2.702, p = 0.008$, size: $t(119) = 4.230, p < 0.001$) sorted lists; more
+lexicographic clustering on lexicographically sorted lists than semantically
+(length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) = 3.686, p =
+0.000$) or visually (length: $t(119) = 3.024, p = 0.003$; first letter: $t(119)
+= 2.644, p = 0.009$) sorted lists; and more visual clustering on visually
+sorted lists than semantically (color: $t(119) = -2.659, p = 0.009$; location:
+$t(119) = -4.604, p < 0.001$) or lexicographically (color: $t(119) = -2.366, p
+= 0.020$; location: $t(119) = -4.265, p < 0.001$) sorted lists. In summary,
+sorting lists by different features appeared to have slightly different effects
+on overall memory performance and temporal clustering, and people tended to
+cluster their recalls along a given feature dimension more when the studied
+lists were (versus were not) sorted along that dimension.
 
@@
 cluster \textit{other} features (on early and/or late lists) as well. We
-defined participants' \textit{memory fingerprints} as the set of temporal and
-feature clustering scores. A participant's memory fingerprint describes how
+defined participants' \textit{memory fingerprints} as the set of their temporal
+and feature clustering scores. A participant's memory fingerprint describes how
 they tend to retrieve memories of the studied items, perhaps searching through
@@
 (Fig.~\ref{fig:fingerprint-trajectories}B). We found that participants'
-fingerprints on early lists seem to be influenced by the order manipulations on
-those lists (see the locations of the circles in
+fingerprints on early lists seem to be influenced by the order manipulations
+for those lists (see the locations of the circles in
 Fig.~\ref{fig:fingerprint-trajectories}B). There also seemed to be some
@@
     into a (new) 2D space. See Figure~\fingerprintTrajectoryRandom~for
-    analogous plots for the random (control) conditions. }
+    analogous plots for the random conditions. }
     \label{fig:fingerprint-trajectories}
@@
 across conditions and participants (early: $r(179) = 0.537, p < 0.001$; late: $
-r(179) = 0.492, p = 0.000$) as well as for each condition individually for
+r(179) = 0.492, p < 0.001$) as well as for each condition individually for
 early ($r$s $\geq 0.386$, all $p$s $\leq 0.035$) and late ($r$s $\geq 0.462$,
@@
 (Fig.~\ref{fig:clustering-scatterplots}F), the results were more
-variable (overall: $r(179) = 0.348, p = 0.000$; all non-visual conditions: $r$s
+variable (overall: $r(179) = 0.348, p < 0.001$; all non-visual conditions: $r$s
 $\geq 0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453, p = 0.011$;
@@
 how the visual features varied across words (within each list), along with the
-orders of each list's words. Although participants' task (verbally recalling as
-many words as possible, in any order, within one minute) remained constant
-across all of these conditions, and although the set of words they studied on
-each list remained constant, our manipulations substantially affected
-participants' memories. The impact of some of the manipulations also affected
-how participants remembered \textit{future} lists that were sorted randomly.
+orders of each list's words. Although the participants' task (verbally
+recalling as many words as possible, in any order, within one minute) remained
+constant across all of these conditions, and although the set of words they
+studied on each list remained constant, our manipulations substantially
+affected participants' memories. The impact of some of the manipulations also
+affected how participants remembered \textit{future} lists that were sorted
+randomly.
 
@@
 and clustering differences disappeared. Abruptly changing how irrelevant visual
-features change across words seems to act as a sort of ``event boundary'' that
-partially resets how participants process and remember post-boundary lists.
-Within-list clustering also increases in these manipulations, suggesting that
-the "within-event" words are being more tightly associated with each other.
+features varied across words seemed to act as a sort of ``event boundary'' that
+partially reset how participants processed and remembered post-boundary lists.
+Within-list clustering also increased in these manipulations, suggesting that
+the "within-event" words were being more tightly associated with each other.
 
@@
 for organizing the words in their memories. When we presented the words in an
-order consistent with each participant's idiosyncratic strategies, their memory
-performance improved. This might indicate that the participants were spending
-less cognitive effort ``reorganizing'' the incoming words on those lists, which
-freed up resources to devote to encoding processes instead.
+order consistent with each participant's idiosyncratic fingerprint, their
+memory performance improved. This might indicate that the participants were
+spending less cognitive effort ``reorganizing'' the incoming words on those
+lists, which freed up resources to devote to encoding processes instead.
 
@@
 sounds~\citep{BeamJone10, SahaSmit14}, secondary tasks~\citep{MasiSaha14,
-PolyEtal09}, and more can all impact how the participant remembers, and
-organizes in memory, lists of studied items.
-
-Consistent with this prior work, we found that participants are sensitive to
+PolyEtal09}, and more can all impact how participants remember, and organize in
+memory, lists of studied items.
+
+Consistent with this prior work, we found that participants were sensitive to
 task-irrelevant visual features. We also found that changing the dynamics of
 those task-irrelevant visual features (in the reduced (early) and reduced
-(late) conditions) \textit{also} affects participants' memories. This suggests
+(late) conditions) \textit{also} affected participants' memories. This suggests
 that it is not only the contextual features themselves that affect memory, but
@@
 re-used as to-be-remembered targets later on in the experiment~\citep{Tipp85}.
-In general, participants' memories can be influenced by a wide range of
-positive and negative factors~\citep{BaloEtal92, ClayChat89, Donn88,
-FlexTulv82, GottEtal12, HuanEtal04, Hube08, HubeEtal01, McNa94, Neel77, Rabi86,
-TulvScha91, WatkEtal92, WiggMart98}.
+In general, participants' memories can be influenced by exposing them to a wide
+range of positive and negative priming factors before they encounter the
+to-be-remembered information~\citep{BaloEtal92, ClayChat89, Donn88, FlexTulv82,
+GottEtal12, HuanEtal04, Hube08, HubeEtal01, McNa94, Neel77, Rabi86, TulvScha91,
+WatkEtal92, WiggMart98}.
 
@@
 will smell when we enter, how crowded the restaurant is likely to be, the
-sounds we are likely to hear, etc.). Similarly, we might learn a schema
-describing how events are likely to unfold \textit{across} any sit-down
-restaurant-- e.g., open the door, wait to be seated, receive a menu, decide
-what to order, place the order, and so on. Situation models and schemas can
-help us to generalize across our experiences, and to generate expectations
-about how new experiences are likely to unfold. When those expectations are
-violated, we can perceive ourselves to have crossed into a new situation.
+sounds we are likely to hear, etc.). Similarly, as mentioned in the
+\textit{Introduction}, we might learn a schema describing how events are likely
+to unfold \textit{across} any sit-down restaurant-- e.g., open the door, wait
+to be seated, receive a menu, decide what to order, place the order, and so on.
+Situation models and schemas can help us to generalize across our experiences,
+and to generate expectations about how new experiences are likely to unfold.
+When those expectations are violated, we can perceive ourselves to have crossed
+into a new situation.
 
@@
 visual appearances of words are determined, or about the orders in which words
-are studied, can lead participants to behave similarly to what one might expect
-upon crossing an event boundary. Adding in variability in font color and
+are presented, can lead participants to behave similarly to what one might
+expect upon crossing an event boundary. Adding in variability in font color and
 presentation locations for words on late lists, after those visual features had
 been held constant on early lists, led participants to remember more words on
-those later lists. (We found the same pattern when we varied words' colors and
-locations on early lists but held them constant on late lists.) One potential
-explanation is that participants perceive an ``event boundary'' to have
-occurred when they encounter the first ``late'' list. According to contextual
-change accounts of memory across event boundaries~\citep[e.g.,][]{SahaKell02},
-this could help to explain why participants in the reduced (early) and reduced
-(late) conditions exhibited better overall memory performance. Specifically,
-their memory for late list items could benefit from less interference from
-early list items, and the contextual features associated with late list items
-(after the ``event boundary'') might serve as more specific recall cues for
-those late items (relative to if the boundary had not occurred).
+those later lists. One potential explanation is that participants perceive an
+``event boundary'' to have occurred when they encounter the first ``late''
+list. According to contextual change accounts of memory across event
+boundaries~\citep[e.g.,][]{SahaKell02}, this could help to explain why
+participants in the reduced (early) and reduced (late) conditions exhibited
+better overall memory performance. Specifically, their memory for late list
+items could benefit from less interference from early list items, and the
+contextual features associated with late list items (after the ``event
+boundary'') might serve as more specific recall cues for those late items
+(relative to if the boundary had not occurred).
 
@@
 recent studies suggest some substantial differences between memory for lists
-versus naturalistic stimuli (e.g., real-world experiences, narratives, films,
-etc.)~\citep{NastEtal20, HeusEtal21, Mann21a, LeeEtal20}. One reason is that
+versus naturalistic stimuli~\citep[e.g., real-world experiences, narratives,
+films, etc.;][]{NastEtal20, HeusEtal21, Mann21a, LeeEtal20}. One reason is that
 naturalistic stimuli are often much more engaging than the highly simplified
@@
 information about what came before or after~\citep{XuEtal23}. In contrast, the
-lists used in laboratory memory tasks are most often ordered randomly, by design,
-to \textit{remove} meaningful temporal structure in the
+lists used in laboratory memory tasks are most often ordered randomly, by
+design, to \textit{remove} meaningful temporal structure in the
 stimulus~\citep{Kaha12}.
@@
 from fundamental workings of our memory systems, from some aspect of the
-stimulus, or from idiosyncratic interactions or interference between our memory
-systems and the stimulus. This challenge implies that, to understand the
-fundamental nature of memory in its ``pure'' form, we should study memory for
-highly simplified stimuli that can provide relatively unbiased (compared with
-real-world experiences) measures of the relevant patterns and tendencies.
-
-The experiment we report in this paper was designed to help bridge some of the
+stimulus, or from idiosyncratic interactions or interference between
+participants' memory systems and the stimulus. This challenge implies that, to
+understand the fundamental nature of memory in its ``pure'' form, we should
+study memory for highly simplified stimuli that can provide relatively unbiased
+(compared with real-world experiences) measures of the relevant patterns and
+tendencies.
+
+The experiment we report in this paper was designed to help bridge some of this
 gap between naturalistic tasks and more traditional list learning tasks. We had
@@
 to show stronger feature clustering for their condition's feature dimension on
-late randomly ordered lists, etc.). We also found that participants remembered
+late randomly ordered lists; etc.). We also found that participants remembered
 more words from lists that were sorted to align with their idiosyncratic
@@
 
-Our findings indicate that the way word lists are presented affects how people
+Our findings indicate that the ways word lists are presented affects how people
 remember them. To the extent that word list memory reflects memory processes
@@
 experimental conditions may be found at
-https://github.com/Context\-Lab/efficient-learning-code. Code for running the
+https://github.com/Con\-text\-Lab/efficient-learning-code. Code for running the
 adaptive experimental condition may be found at
@@
 
-We acknowledge useful discussions, assistance in setting up an earlier version
-of this study, and assistance with some of the data collection efforts from
-Rachel Chacko, Joseph Finkelstein, Sheherzad Mohydin, Lucy Owen, Gal Perlman,
-Darya Romanova, Jake Rost, Jessica Tin, Marisol Tracy, and Peter Tran. Our work
-was supported in part by NSF CAREER Award Number 2145172 to JRM. The content
-is solely the responsibility of the authors and does not necessarily represent
-the official views of our supporting organizations. The funders had no role in
-study design, data collection and analysis, decision to publish, or preparation
-of the manuscript.
+We acknowledge useful discussions, assistance in setting up an earlier
+(unpublished) version of this study, and assistance with some of the data
+collection efforts from Rachel Chacko, Joseph Finkelstein, Sheherzad Mohydin,
+Lucy Owen, Gal Perlman, Darya Romanova, Jake Rost, Jessica Tin, Marisol Tracy,
+and Peter Tran. Our work was supported in part by NSF CAREER Award Number
+2145172 to JRM. The content is solely the responsibility of the authors and
+does not necessarily represent the official views of our supporting
+organizations. The funders had no role in study design, data collection and
+analysis, decision to publish, or preparation of the manuscript.
 

### 2023-01-17 paper/supplement.tex
- finished "random conditions" description - added bryan as a co-author (yay!) - working on description of order manipulation conditions
draft of order manipulation conditions:   - added Allison as an author (yay!)   - added "stick" figure to illustrate stochastic word ordering procedure   - added descriptions + equations for order manipulation conditions
- working on results section - added Campbell as an author (yay!)
quick round of edits through end of methods
backing up progress on discussion section   - added a more detailed summary and overview of the key results
inputting a round of wri

@@
 
-\title{\textit{Supplementary materials for}: Carryover effects in free recall reveal how prior experiences influence memories of new experiences}
-\author{Jeremy R.
-Manning\textsuperscript{1, *}, Kirsten Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
-Fitzpatrick\textsuperscript{1}, Madeline R.
-Lee\textsuperscript{1}, and Andrew C.
-Heusser\textsuperscript{1, 3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton University\\\textsuperscript{3}Akili Interactive\\\textsuperscript{*}Corresponding author: jeremy.r.manning@dartmouth.edu}
+\title{\textit{Supplementary materials for}: Feature and order manipulations in
+a free recall task affect memory for current and future lists}
+
+\author{Jeremy R. Manning\textsuperscript{1, *}, Kirsten
+Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C. 
+Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Allison M
+Frantz\textsuperscript{1},\\Bryan J. Bollinger\textsuperscript{1}, Campbell E.
+Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
+3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
+University\\\textsuperscript{3}Akili
+Interactive\\\textsuperscript{*}Corresponding author:
+jeremy.r.manning@dartmouth.edu}
 
@@
     \caption{\textbf{Recall accuracy by study list number.}  Each panel displays the average recall accuracy (across participants) as a function of the number of studied lists for the
-    random (control) conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}.  The conditions are denoted by color.  Note that the first four
+    random conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}.  The conditions are denoted by color.  Note that the first four
     lists in all of the ``adaptive'' conditions were ordered randomly to compute a baseline fingerprint for each participant prior to initiating the adaptive ordering procedure.  All error

### 2023-01-23 paper/boneyard.tex

@@
 
-Memory researchers, like cognitive neuropsychologists in many domains, have historically taken a ``less is more'' approach.
-They follow the logic that, given a memory topic of interest (e.g.\ memory for real-world autobiographical events), we can ask: what is the simplest experimental paradigm that we might reasonably expect to involve a shared or analogous set of behavioral tendencies and brain structures?
-There are many benefits to taking a simplified approach to studying complex cognitive phenomena.
-By composing a set of experimental circumstances where every aspect of the stimuli and timing is carefully controlled by the experimentalist, and where the behaviors a participant can exhibit are highly constrained, it becomes possible to systematically link specific experimental observations to their corresponding neural and behavioral mechanisms~\citep{OReiMuna00}.
-However, there is also a major cost to simplifying: tightly controlled experiences designed in the lab can often only provide indirect insights into the real-world analogs of those experiences.
-This has led to a major hole in the memory theory literature.
-On the one hand, the most well-developed theories of human memory have been informed by a long history of tightly controlled (but simplified relative to the real-world) experiments~\citep[e.g.\ list-learning, priming, reaction times; ][]{Kaha12}.
-On the other hand, studies of real-world classroom learning have largely avoided drawing direct links with those theories, instead focusing on questions about applied topics such as creating effective assessment tools, increasing student engagement and motivation, and understanding what makes an effective teacher~\citep{Seid07}.
+Memory researchers, like cognitive neuropsychologists in many domains, have
+historically taken a ``less is more'' approach. They follow the logic that,
+given a memory topic of interest (e.g.\ memory for real-world autobiographical
+events), we can ask: what is the simplest experimental paradigm that we might
+reasonably expect to involve a shared or analogous set of behavioral tendencies
+and brain structures? There are many benefits to taking a simplified approach
+to studying complex cognitive phenomena. By composing a set of experimental
+circumstances where every aspect of the stimuli and timing is carefully
+controlled by the experimentalist, and where the behaviors a participant can
+exhibit are highly constrained, it becomes possible to systematically link
+specific experimental observations to their corresponding neural and behavioral
+mechanisms~\citep{OReiMuna00}. However, there is also a major cost to
+simplifying: tightly controlled experiences designed in the lab can often only
+provide indirect insights into the real-world analogs of those experiences.
+This has led to a major hole in the memory theory literature. On the one hand,
+the most well-developed theories of human memory have been informed by a long
+history of tightly controlled (but simplified relative to the real-world)
+experiments~\citep[e.g.\ list-learning, priming, reaction times; ][]{Kaha12}.
+On the other hand, studies of real-world classroom learning have largely
+avoided drawing direct links with those theories, instead focusing on questions
+about applied topics such as creating effective assessment tools, increasing
+student engagement and motivation, and understanding what makes an effective
+teacher~\citep{Seid07}.
 
-The general approach we take in our project is to use prior experimental and theoretical work on list learning as a foundation for optimizing STEM learning.
-Specifically, \textbf{Module 1 is centered around developing approaches for improving participants' memories for lists of words-- a simple and well-studied task expected to involve neural mechanisms similar to those underlying STEM learning}.
-To foreshadow our proposal, \textbf{in Module 2, we have designed a set of analogous experiments and analyses that adapt the approach in Module 1 to a more naturalistic learning setting (learning from online STEM training videos)}.
-By mirroring our approaches in Modules 1 and 2, we hope to be able to draw parallels between the two sets of experiments and findings.
-Finally, \textbf{in Module 3, we plan to replicate the EEG experiments in Modules 1 and 2 using a low-cost EEG system that our team developed.
-We will also develop (and document) an open-source toolkit for sharing our software (and data).
-In this way, Module 3 will serve to help bring our work into real-world classrooms, and to real-world students}.
+The general approach we take in our project is to use prior experimental and
+theoretical work on list learning as a foundation for optimizing STEM learning.
+Specifically, \textbf{Module 1 is centered around developing approaches for
+improving participants' memories for lists of words-- a simple and well-studied
+task expected to involve neural mechanisms similar to those underlying STEM
+learning}. To foreshadow our proposal, \textbf{in Module 2, we have designed a
+set of analogous experiments and analyses that adapt the approach in Module 1
+to a more naturalistic learning setting (learning from online STEM training
+videos)}. By mirroring our approaches in Modules 1 and 2, we hope to be able to
+draw parallels between the two sets of experiments and findings. Finally,
+\textbf{in Module 3, we plan to replicate the EEG experiments in Modules 1 and
+2 using a low-cost EEG system that our team developed. We will also develop
+(and document) an open-source toolkit for sharing our software (and data). In
+this way, Module 3 will serve to help bring our work into real-world
+classrooms, and to real-world students}.
 
 \subsection*{Context and the structure of memory}
-Remembering words on a studied list requires distinguishing the current list from the rest of one's experience.
-To model this fundamental memory capability, cognitive scientists have posited the existence of a special representation, called \emph{context}, that is associated with each list.
-According to early theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are composed of many features which fluctuate from moment to moment, slowly drifting through a multidimensional feature space.
-During recall, this representation forms part of the retrieval cue, enabling us to distinguish list items from non-list items.
-Understanding the role of context in memory processes is particularly important in self-cued memory tasks, such as \textit{free recall}, where the retrieval cue is ``context'' itself.
-In free recall, participants study lists of items and are instructed to recall the items in any order they choose.
+
+Remembering words on a studied list requires distinguishing the current list
+from the rest of one's experience. To model this fundamental memory capability,
+cognitive scientists have posited the existence of a special representation,
+called \emph{context}, that is associated with each list. According to early
+theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are
+composed of many features which fluctuate from moment to moment, slowly
+drifting through a multidimensional feature space. During recall, this
+representation forms part of the retrieval cue, enabling us to distinguish list
+items from non-list items. Understanding the role of context in memory
+processes is particularly important in self-cued memory tasks, such as
+\textit{free recall}, where the retrieval cue is ``context'' itself. In free
+recall, participants study lists of items and are instructed to recall the
+items in any order they choose.
 
@@
 
-Over the past half-century, context-based models have enjoyed impressive success at explaining many stereotyped behaviors observed during free recall and other list-learning tasks \citep[e.g.][]{Este55a, RaaiShif80, GlenEtal83, HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, PolyEtalTulv, SedeEtal08, PolyEtal09, ShanHowa12}.
-These phenomena include the well-known recency and primacy effects (superior recall of items from the end and, to a lesser extent, from the beginning of the study list), as well as semantic and temporal clustering effects \citep[for review see][]{KahaEtal08}.
-The contiguity effect is an example of temporal clustering, which is perhaps the dominant form of organization in free recall.
-This effect can be seen in the tendency for people to successively recall items that occupied neighboring positions in the study list.
-For example, if a list contained the sub-sequence \textsc{``absence hollow pupil''} and the participant recalls the word \textsc{``hollow''}, it is far more likely that the next response will be either \textsc{``pupil''} or \textsc{``absence''} than some other list item~\citep{Kaha96}.
-In addition, there is a strong forward bias in the contiguity effect: subjects make forward transitions (i.e., \textsc{``hollow''} followed by \textsc{``pupil''}) about twice as often as they make backward transitions, despite an overall tendency to begin recall at the end of the list.
-There are also striking effects of semantic clustering \citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52, MannKaha12}, whereby the recall of a given item is more likely to be followed by recall of a similar or related item than a dissimilar or unrelated one.
+Over the past half-century, context-based models have enjoyed impressive
+success at explaining many stereotyped behaviors observed during free recall
+and other list-learning tasks \citep[e.g.][]{Este55a, RaaiShif80, GlenEtal83,
+HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, PolyEtalTulv, SedeEtal08,
+PolyEtal09, ShanHowa12}. These phenomena include the well-known recency and
+primacy effects (superior recall of items from the end and, to a lesser extent,
+from the beginning of the study list), as well as semantic and temporal
+clustering effects \citep[for review see][]{KahaEtal08}. The contiguity effect
+is an example of temporal clustering, which is perhaps the dominant form of
+organization in free recall. This effect can be seen in the tendency for people
+to successively recall items that occupied neighboring positions in the study
+list. For example, if a list contained the sub-sequence \textsc{``absence
+hollow pupil''} and the participant recalls the word \textsc{``hollow''}, it is
+far more likely that the next response will be either \textsc{``pupil''} or
+\textsc{``absence''} than some other list item~\citep{Kaha96}. In addition,
+there is a strong forward bias in the contiguity effect: subjects make forward
+transitions (i.e., \textsc{``hollow''} followed by \textsc{``pupil''}) about
+twice as often as they make backward transitions, despite an overall tendency
+to begin recall at the end of the list. There are also striking effects of
+semantic clustering \citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52,
+MannKaha12}, whereby the recall of a given item is more likely to be followed
+by recall of a similar or related item than a dissimilar or unrelated one.
 
-In general, people organize memories for words along a wide variety of stimulus dimensions.
-As captured by models like the \textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each word (e.g.\ the word's meaning, font size, font color, location on the screen, size of the object the word represents, etc.) are incorporated into the participant's mental context representation~\citep[also see][]{SmitVela01, MannEtal15}.
-During a memory test, any of these features may serve as a memory cue, which in turn leads the participant to recall in succession words that share stimulus features.
-We refer to a participant's particular tendency to organize their recalls according to each stimulus dimension as their \textit{memory fingerprint} (Fig.~\ref{fig:fingerprint}).
-Formally, the memory fingerprint is a vector whose elements range between 0 and 1, where each element of the memory fingerprint reflects a stimulus feature.
-A value close to 1 indicates that the participant nearly always orders their recalls according to the given stimulus dimension.
-A value close to 0 indicates that the participant almost \textit{never} orders their recalls according to the given stimulus dimension; a value close to 0.5 indicates that the participant's recalls are random with respect to that stimulus dimension.
-In Experiment 1 we propose a free recall variant where each stimulus has many well-defined and quantifiable features, enabling us to track how the ways people organize their memories (i.e., their memory fingerprints) change over time.
-In Experiment 2 we use a behavior-driven approach to manipulate these processes, and in Experiment 3 we use electroencephalographic (EEG) data to manipulate these processes.
-In this way, Experiment 1 is focused on elucidating the dynamics of this very specific form of learning, and Experiments 2 and 3 are focused on manipulating the learning process, with the goal of improving (or diminishing) learning efficiency.
+In general, people organize memories for words along a wide variety of stimulus
+dimensions. As captured by models like the \textit{Context Maintenance and
+Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each
+word (e.g.\ the word's meaning, font size, font color, location on the screen,
+size of the object the word represents, etc.) are incorporated into the
+participant's mental context representation~\citep[also see][]{SmitVela01,
+MannEtal15}. During a memory test, any of these features may serve as a memory
+cue, which in turn leads the participant to recall in succession words that
+share stimulus features. We refer to a participant's particular tendency to
+organize their recalls according to each stimulus dimension as their
+\textit{memory fingerprint} (Fig.~\ref{fig:fingerprint}). Formally, the memory
+fingerprint is a vector whose elements range between 0 and 1, where each
+element of the memory fingerprint reflects a stimulus feature. A value close to
+1 indicates that the participant nearly always orders their recalls according
+to the given stimulus dimension. A value close to 0 indicates that the
+participant almost \textit{never} orders their recalls according to the given
+stimulus dimension; a value close to 0.5 indicates that the participant's
+recalls are random with respect to that stimulus dimension. In Experiment 1 we
+propose a free recall variant where each stimulus has many well-defined and
+quantifiable features, enabling us to track how the ways people organize their
+memories (i.e., their memory fingerprints) change over time. In Experiment 2 we
+use a behavior-driven approach to manipulate these processes, and in Experiment
+3 we use electroencephalographic (EEG) data to manipulate these processes. In
+this way, Experiment 1 is focused on elucidating the dynamics of this very
+specific form of learning, and Experiments 2 and 3 are focused on manipulating
+the learning process, with the goal of improving (or diminishing) learning
+efficiency.
 
@@
 
-\begin{wrapfigure}[10]{r}{1.4in}
-\vspace{-57pt}
-  \begin{center}
-    \includegraphics[width=1.2in]{figs/FRFR}
-  \end{center}
-\vspace{-16pt}
-  \caption{\footnotesize \textbf{Feature-rich free recall.}  After studying lists comprised of words that vary along several feature dimensions, participants verbally recall words in any order (microphone icon).}
-  \label{fig:exp1}
+\begin{wrapfigure}[10]{r}{1.4in} \vspace{-57pt} \begin{center}
+\includegraphics[width=1.2in]{figs/FRFR} \end{center} \vspace{-16pt}
+\caption{\footnotesize \textbf{Feature-rich free recall.} After studying lists
+comprised of words that vary along several feature dimensions, participants
+verbally recall words in any order (microphone icon).} \label{fig:exp1}
 \end{wrapfigure}
 
-Each of our Module 1 experiments is based on a variant of the classic free recall paradigm that we term \textit{feature-rich free recall}.
-In feature-rich free recall, participants study 16 lists, each comprised of 16 words that vary along a number of stimulus dimensions: semantic category, object size, text color, text location, word length, and starting letter (Fig.~\ref{fig:exp1}).
-Each list contains four words from each of four different semantic categories; all other stimulus features are randomized.
-Each word is presented for 2 seconds, followed by a 2 second pause (blank screen).
-After studying each list, the participant has 60 seconds to recall as many words as they can from that list, in any order they choose.
-Because each individual word is associated with several well-defined (and quantifiable) features, and because each list incorporates a diverse mix of feature values along each dimension, this allows us to evaluate participants' memory fingerprints in rich detail.
+Each of our Module 1 experiments is based on a variant of the classic free
+recall paradigm that we term \textit{feature-rich free recall}. In feature-rich
+free recall, participants study 16 lists, each comprised of 16 words that vary
+along a number of stimulus dimensions: semantic category, object size, text
+color, text location, word length, and starting letter (Fig.~\ref{fig:exp1}).
+Each list contains four words from each of four different semantic categories;
+all other stimulus features are randomized. Each word is presented for 2
+seconds, followed by a 2 second pause (blank screen). After studying each list,
+the participant has 60 seconds to recall as many words as they can from that
+list, in any order they choose. Because each individual word is associated with
+several well-defined (and quantifiable) features, and because each list
+incorporates a diverse mix of feature values along each dimension, this allows
+us to evaluate participants' memory fingerprints in rich detail.
 
-Our experimental paradigm incorporates the Google Cloud Speech API speech-to-text engine to automatically transcribe participants' verbal recalls into text.
-This allows recalls to be transcribed in real time-- a distinguishing feature of the experiment that is critical for using participants' recalls of prior lists to adapt the presentation orders of future lists, as in Experiments 2 and 3.
-(In typical verbal recall experiments the audio data must be parsed manually, precluding this sort of design.)  We recently published a study verifying that the computer-generated recall transcripts are sufficiently similar to human-generated transcripts to yield reliable data~\citep{ZimaEtal18}.
-
-\subsection*{Experiment 1: tracking the dynamics of memory organization over successive learning trials}
-% maybe this could benefit from reframing wrt to the variance idea?
-Large portions of the memory theory and education literatures have focused on identifying learners' unique learning styles.
-In the memory theory literature~\citep{SmitVela01, Kaha12, HealKaha14a, MannEtal15}, the focus has been on identifying behavioral signatures of how memories are organized (i.e., in the language of this proposal, identifying memory fingerprints).
-In the education literature, the focus has been been on asking whether catering lessons to an individual's unique learning style can help them to learn more efficiently~\citep{PashEtal08}.
-Each of these literatures effectively treats people's learning styles as a fixed point.
-\textbf{In other words, the assumption underlying the majority of these studies is that each individual has their own unique learning style, and that learning style remains roughly consistent over time.
-Experiment 1 is designed to explicitly test this assumption using the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}).}  Importantly, we will collect data from two populations: (a) members of the Dartmouth and surrounding community, with the experiment running in a tightly controlled laboratory setting, and (b) Mechanical Turk workers, via the Internet.
-Taken together, we hope that these two populations of participants will allow us to generalize our findings to a diverse set of learners.
+Our experimental paradigm incorporates the Google Cloud Speech API
+speech-to-text engine to automatically transcribe participants' verbal recalls
+into text. This allows recalls to be transcribed in real time-- a
+distinguishing feature of the experiment that is critical for using
+participants' recalls of prior lists to adapt the presentation orders of future
+lists, as in Experiments 2 and 3. (In typical verbal recall experiments the
+audio data must be parsed manually, precluding this sort of design.) We
+recently published a study verifying that the computer-generated recall
+transcripts are sufficiently similar to human-generated transcripts to yield
+reliable data~\citep{ZimaEtal18}. \subsection*{Experiment 1: tracking the
+dynamics of memory organization over successive learning trials} % maybe this
+could benefit from reframing wrt to the variance idea? Large portions of the
+memory theory and education literatures have focused on identifying learners'
+unique learning styles. In the memory theory literature~\citep{SmitVela01,
+Kaha12, HealKaha14a, MannEtal15}, the focus has been on identifying behavioral
+signatures of how memories are organized (i.e., in the language of this
+proposal, identifying memory fingerprints). In the education literature, the
+focus has been been on asking whether catering lessons to an individual's
+unique learning style can help them to learn more
+efficiently~\citep{PashEtal08}. Each of these literatures effectively treats
+people's learning styles as a fixed point. \textbf{In other words, the
+assumption underlying the majority of these studies is that each individual has
+their own unique learning style, and that learning style remains roughly
+consistent over time. Experiment 1 is designed to explicitly test this
+assumption using the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}).}
+Importantly, we will collect data from two populations: (a) members of the
+Dartmouth and surrounding community, with the experiment running in a tightly
+controlled laboratory setting, and (b) Mechanical Turk workers, via the
+Internet. Taken together, we hope that these two populations of participants
+will allow us to generalize our findings to a diverse set of learners.
 
 \paragraph{Design and procedure.}
-We will use the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}) to study how people's memory fingerprints are influenced by presentation order.
-The experiment will be divided into six manipulation conditions (one for each stimulus feature); in each of these manipulation conditions the words from the first 8 lists participants study will be sorted according to one of the stimulus dimensions.
-For example, in the ``color'' condition, words from the first 8 lists will be sorted by font color; in the ``starting letter'' condition, words from the first 8 lists will be sorted alphabetically; etc.
-The second 8 lists will be presented in a random order in every condition.
-As described below, we will assess how participants' memories are affected by these order manipulations, both for their memories of words on first 8 lists where the manipulations are applied, as well as on the second 8 lists.
-Participants' behaviors will be compared to a seventh (control) condition, where the presentation order is randomized for all 16 lists.
 
-We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our prior work to estimate the relevant parameters~\citep{MannEtal11, MannEtal12}.
-The analysis indicated that we will require 30 in-laboratory participants in each of the six manipulation conditions to achieve sufficient power.
-Because all of our findings will be compared relative to the random control condition, we plan to run 60 in-laboratory participants in that condition.
-Prior work suggests that Mechanical Turk data is more variable than data collected in the laboratory~\citep{CrumEtal13}, so in each condition we plan to collect data from twice as many Mechanical Turk workers as in-laboratory participants.
-In all, across the 7 experimental conditions, we plan to run 720 participants in Experiment 1 (240 in the lab and 480 online through Mechanical Turk).
+We will use the feature-rich free recall paradigm (Fig.~\ref{fig:exp1}) to
+study how people's memory fingerprints are influenced by presentation order.
+The experiment will be divided into six manipulation conditions (one for each
+stimulus feature); in each of these manipulation conditions the words from the
+first 8 lists participants study will be sorted according to one of the
+stimulus dimensions. For example, in the ``color'' condition, words from the
+first 8 lists will be sorted by font color; in the ``starting letter''
+condition, words from the first 8 lists will be sorted alphabetically; etc. The
+second 8 lists will be presented in a random order in every condition. As
+described below, we will assess how participants' memories are affected by
+these order manipulations, both for their memories of words on first 8 lists
+where the manipulations are applied, as well as on the second 8 lists.
+Participants' behaviors will be compared to a seventh (control) condition,
+where the presentation order is randomized for all 16 lists.
+
+We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our
+prior work to estimate the relevant parameters~\citep{MannEtal11, MannEtal12}.
+The analysis indicated that we will require 30 in-laboratory participants in
+each of the six manipulation conditions to achieve sufficient power. Because
+all of our findings will be compared relative to the random control condition,
+we plan to run 60 in-laboratory participants in that condition. Prior work
+suggests that Mechanical Turk data is more variable than data collected in the
+laboratory~\citep{CrumEtal13}, so in each condition we plan to collect data
+from twice as many Mechanical Turk workers as in-laboratory participants. In
+all, across the 7 experimental conditions, we plan to run 720 participants in
+Experiment 1 (240 in the lab and 480 online through Mechanical Turk).
 
 \paragraph{Key analyses.}
-Experiment 1 is motivated by two fundamental questions about the dynamics of memory organization and recall.
-\textbf{First, we want to understand how memory fingerprints vary across people and over time.}  For example, are people's memory fingerprints generally stable across lists?
-Or, what are the similarities and differences across people (in how people organize their memories, and how their memory fingerprints change over time)?
-Are those individual differences predictive of memory performance?
-\textbf{Second, we want to understand how people's memory fingerprints can be manipulated.}  Specifically, if lists are sorted according to a particular feature (e.g.
-the words' semantic categories), how does that influence people's tendencies to cluster their recalls of those words with respect to that feature?
-And which of those tendencies persist during recalls of later (randomly ordered) lists?
-For example, if we organize early lists alphabetically, are  participants more likely recall the words from those lists alphabetically as well?
-More generally, might there be some way of influencing how people organize their memories, even after the influencing cue is no longer present?
 
-% figures: pilot data displaying random-condition fingerprints, relationship between fingerprint variance and memory
-\begin{wrapfigure}[10]{R}{4.0in}
-  \vspace{-52pt}
-  \begin{center}
-    \includegraphics[width=4.0in]{figs/random_pilot.pdf}
-  \end{center}
-\vspace{-18pt}
-  \caption{\footnotesize \textbf{Preliminary data for feature-rich free recall.}  \textbf{a.} Clustering scores along each feature dimensions (i.e.
-memory fingerprints).
-Each dot represents the average clustering score along one dimension, for one participant.
-\textbf{b.} Across-subject correlation between fingerprint variability and memory performance.
-\textbf{c.} Across-condition correlation between fingerprint variability and memory performance.}
-  \label{fig:random_pilot}
-\end{wrapfigure}
+Experiment 1 is motivated by two fundamental questions about the dynamics of
+memory organization and recall. \textbf{First, we want to understand how memory
+fingerprints vary across people and over time.} For example, are people's
+memory fingerprints generally stable across lists? Or, what are the
+similarities and differences across people (in how people organize their
+memories, and how their memory fingerprints change over time)? Are those
+individual differences predictive of memory performance? \textbf{Second, we
+want to understand how people's memory fingerprints can be manipulated.}
+Specifically, if lists are sorted according to a particular feature (e.g. the
+words' semantic categories), how does that influence people's tendencies to
+cluster their recalls of those words with respect to that feature? And which of
+those tendencies persist during recalls of later (randomly ordered) lists? For
+example, if we organize early lists alphabetically, are participants more
+likely recall the words from those lists alphabetically as well? More
+generally, might there be some way of influencing how people organize their
+memories, even after the influencing cue is no longer present?
+
+% figures: pilot data displaying random-condition fingerprints, relationship
+between fingerprint variance and memory \begin{wrapfigure}[10]{R}{4.0in}
+\vspace{-52pt} \begin{center}
+\includegraphics[width=4.0in]{figs/random_pilot.pdf} \end{center}
+\vspace{-18pt} \caption{\footnotesize \textbf{Preliminary data for feature-rich
+free recall.} \textbf{a.} Clustering scores along each feature dimensions (i.e.
+memory fingerprints). Each dot represents the average clustering score along
+one dimension, for one participant. \textbf{b.} Across-subject correlation
+between fingerprint variability and memory performance. \textbf{c.}
+Across-condition correlation between fingerprint variability and memory
+performance.} \label{fig:random_pilot} \end{wrapfigure}
 
 \paragraph{Experiment 1 preliminary data.}
-We ran a preliminary study ($n = 172$) to measure participants' memory fingerprints under the in-laboratory conditions outlined in Experiment 1.
-Using data from the control condition (where all lists were ordered randomly; $n = 45$), we computed the distribution of average memory fingerprints (across lists) from each participant (Fig.~\ref{fig:random_pilot}a).
-The degree to which participants organize their memories with respect to each feature varies substantially across people (e.g.\ there is a wide spread of values covered across participants within each feature dimension).
-In addition, some features are more likely to be leveraged in people's memories than others.
-For example, participants are more likely to organize their recalls by category, size, and temporal proximity than by font color or word length.
-We also examined how stable participants' memory fingerprints were across lists.
-We found that participants whose memory fingerprints were more stable across lists (i.e., lower variance) remembered more words (Fig.~\ref{fig:random_pilot}b).
-We also examined the manipulation conditions, where the first 8 lists participants studied were sorted according to one feature (category: $n=21$, color: $n=20$, screen location: $n=21$, first letter: $n=20$, word length: $n=23$, size: $n=22$).
-Relative to randomly sorted lists, sorting by some features increased the average memory fingerprint variability and sorting by other features decreased the average memory fingerprint variability.
-(We used a permutation-based procedure to compute the variability, which controlled for the distributions of feature values of the words participants remembered.)  We found that participants remembered more words in the conditions where the presentation orders led to lower fingerprint variability (Fig.~\ref{fig:random_pilot}c).
-Taken together, these preliminary data suggest that participants who naturally exhibit more stable fingerprints remember more words, and fingerprint stability may be influenced by the order in which words are presented.
-In turn, sorting the word presentations along different dimensions can enhance (or diminish) memory performance according to whether fingerprint variability is increased or decreased, respectively.
+
+We ran a preliminary study ($n = 172$) to measure participants' memory
+fingerprints under the in-laboratory conditions outlined in Experiment 1. Using
+data from the control condition (where all lists were ordered randomly; $n =
+45$), we computed the distribution of average memory fingerprints (across
+lists) from each participant (Fig.~\ref{fig:random_pilot}a). The degree to
+which participants organize their memories with respect to each feature varies
+substantially across people (e.g.\ there is a wide spread of values covered
+across participants within each feature dimension). In addition, some features
+are more likely to be leveraged in people's memories than others. For example,
+participants are more likely to organize their recalls by category, size, and
+temporal proximity than by font color or word length. We also examined how
+stable participants' memory fingerprints were across lists. We found that
+participants whose memory fingerprints were more stable across lists (i.e.,
+lower variance) remembered more words (Fig.~\ref{fig:random_pilot}b). We also
+examined the manipulation conditions, where the first 8 lists participants
+studied were sorted according to one feature (category: $n=21$, color: $n=20$,
+screen location: $n=21$, first letter: $n=20$, word length: $n=23$, size:
+$n=22$). Relative to randomly sorted lists, sorting by some features increased
+the average memory fingerprint variability and sorting by other features
+decreased the average memory fingerprint variability. (We used a
+permutation-based procedure to compute the variability, which controlled for
+the distributions of feature values of the words participants remembered.) We
+found that participants remembered more words in the conditions where the
+presentation orders led to lower fingerprint variability
+(Fig.~\ref{fig:random_pilot}c). Taken together, these preliminary data suggest
+that participants who naturally exhibit more stable fingerprints remember more
+words, and fingerprint stability may be influenced by the order in which words
+are presented. In turn, sorting the word presentations along different
+dimensions can enhance (or diminish) memory performance according to whether
+fingerprint variability is increased or decreased, respectively.
 
@@
 
-Our pilot data from Experiment 1 suggest that the order in which words are studied affects how they are later recalled.
-Further, certain orderings can lead to lower memory fingerprint variability, which in turn leads (on average) to better memory for the studied words.
-Whereas Experiment 1 explores the memory effects of pre-defined order manipulations, in Experiment 2 we propose to adapt the presentation order of future lists, using behavior measured on previous lists.
-\textbf{Our overarching goal is to explore whether memory models that are custom-fit to individual participants may be exploited (through presentation order manipulations) to enhance or diminish memory performance.}
+Our pilot data from Experiment 1 suggest that the order in which words are
+studied affects how they are later recalled. Further, certain orderings can
+lead to lower memory fingerprint variability, which in turn leads (on average)
+to better memory for the studied words. Whereas Experiment 1 explores the
+memory effects of pre-defined order manipulations, in Experiment 2 we propose
+to adapt the presentation order of future lists, using behavior measured on
+previous lists. \textbf{Our overarching goal is to explore whether memory
+models that are custom-fit to individual participants may be exploited (through
+presentation order manipulations) to enhance or diminish memory performance.}
 
@@
 
-We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our pilot data as a guide (Fig.~\ref{fig:adaptive_pilot}).
-The analysis indicates that we will require at least 33 in-laboratory participants to achieve sufficient power, which we plan to round up to 40 participants.
-Similarly (following our logic from Experiment 1) we plan to run 80 Mechanical Turk workers in the online version of Experiment 2.
-We will also duplicate the experiment using yoked controls, for a total of 80 in-laboratory and 160 Mechanical Turk participants.
+We carried out a power analysis (power: 0.8, Type 1 error rate: 0.05) using our
+pilot data as a guide (Fig.~\ref{fig:adaptive_pilot}). The analysis indicates
+that we will require at least 33 in-laboratory participants to achieve
+sufficient power, which we plan to round up to 40 participants. Similarly
+(following our logic from Experiment 1) we plan to run 80 Mechanical Turk
+workers in the online version of Experiment 2. We will also duplicate the
+experiment using yoked controls, for a total of 80 in-laboratory and 160
+Mechanical Turk participants.
 
@@
 \subsection*{Experiment 3: adaptive list learning using (neural) memory fingerprints}
-The above experiments leverage participants' \textit{behaviors} (i.e., the sequences of recalls they make) to optimize learning by changing the order in which subsequent lists are presented.
-In Experiment 3, we seek to optimize learning using \textit{neural} (EEG) data.
-\textbf{Our prior work shows that the neural patterns recorded while participants study a list of words may be used to predict their individual tendencies to cluster their recalls according to presentation order (temporal clustering; Fig.~\ref{fig:clust}a) or word meaning (semantic clustering; Fig.~\ref{fig:clust}b).}
+
+The above experiments leverage participants' \textit{behaviors} (i.e., the
+sequences of recalls they make) to optimize learning by changing the order in
+which subsequent lists are presented. In Experiment 3, we seek to optimize
+learning using \textit{neural} (EEG) data. \textbf{Our prior work shows that
+the neural patterns recorded while participants study a list of words may be
+used to predict their individual tendencies to cluster their recalls according
+to presentation order (temporal clustering; Fig.~\ref{fig:clust}a) or word
+meaning (semantic clustering; Fig.~\ref{fig:clust}b).}
 
@@
 
-\paragraph{Design and procedure.} The experiment and parameters will be analogous to Experiment 2, except that we will use EEG data recorded as participants are studying the words (rather than behavioral data from participants' recalls) to estimate their memory fingerprints.
-We plan to record the EEG data using an ActiChamp 160-electrode system (with a photosensor-based electrode mapping system, to allow us to perform source-localization analyses).
-Participants will study a total of 16 lists of 16 words each; we will adaptively adjust the presentation orders of each list as described next.
+\paragraph{Design and procedure.}
 
-The first four lists will always be sorted randomly.
-We will use EEG recordings as participants study these lists to train classifiers to detect the neural signatures of clustering along each memory fingerprint dimension.
-Note that, whereas in Experiment 2 the presentation order of the full list is determined before the first word is presented, in Experiment 3 we will be adaptively adjusting the presentation order on a word-by-word basis using classifiers applied to the ongoing EEG recordings.
-In other words, we will use the EEG activity recorded during the inter-stimulus interval to predict how the next word will be organized in memory.
-Then, using the just-presented word as a reference, we will select a word that is either \textit{consistent}, \textit{inconsistent}, or \textit{random} with respect to that predicted memory fingerprint (Fig.~\ref{fig:adaptiveFR_eeg_order}).
-For example, in the \textit{consistent} condition, if the classifier predicts that the next word is likely to be organized by category and font color, then the next to-be-presented word will be chosen to match the category and font color of the just-presented word.
-More generally, we will rank all yet-to-be-presented words according to how similar they are along each stimulus dimension (to the just-presented word).
-We will then select a matching word by weighting these similarity rankings (using the neurally predicted memory fingerprint to define the weights).
-In the \textit{inconsistent} condition, we will next present the word that \textit{least} matches the previous word, using the decoded fingerprint.
-In this way, the neurally driven \textit{consistent} and \textit{inconsistent} conditions of Experiment 3 are analogous to the behaviorally driven \textit{stabilize} and \textit{destabilize} conditions of Experiment 2.
-We also plan to run yoked control conditions, whereby the presentation orders participants exhibit are driven by \textit{other} participants' neural data.
+The experiment and parameters will be analogous to Experiment 2, except that we
+will use EEG data recorded as participants are studying the words (rather than
+behavioral data from participants' recalls) to estimate their memory
+fingerprints. We plan to record the EEG data using an ActiChamp 160-electrode
+system (with a photosensor-based electrode mapping system, to allow us to
+perform source-localization analyses). Participants will study a total of 16
+lists of 16 words each; we will adaptively adjust the presentation orders of
+each list as described next.
 
-Using our prior work (Fig.~\ref{fig:clust}) to approximate the relevant parameters, we ran a power analysis (power: 0.8, Type 1 error rate: 0.05) to estimate the number of participants we will need to achieve sufficient power.
-The analysis indicates that we will require at least 42 participants in each experimental condition; we plan to run 50 participants using the neurally driven adaptive design described above, plus an additional 50 in the yoked control condition.
+The first four lists will always be sorted randomly. We will use EEG recordings
+as participants study these lists to train classifiers to detect the neural
+signatures of clustering along each memory fingerprint dimension. Note that,
+whereas in Experiment 2 the presentation order of the full list is determined
+before the first word is presented, in Experiment 3 we will be adaptively
+adjusting the presentation order on a word-by-word basis using classifiers
+applied to the ongoing EEG recordings. In other words, we will use the EEG
+activity recorded during the inter-stimulus interval to predict how the next
+word will be organized in memory. Then, using the just-presented word as a
+reference, we will select a word that is either \textit{consistent},
+\textit{inconsistent}, or \textit{random} with respect to that predicted memory
+fingerprint (Fig.~\ref{fig:adaptiveFR_eeg_order}). For example, in the
+\textit{consistent} condition, if the classifier predicts that the next word is
+likely to be organized by category and font color, then the next
+to-be-presented word will be chosen to match the category and font color of the
+just-presented word. More generally, we will rank all yet-to-be-presented words
+according to how similar they are along each stimulus dimension (to the
+just-presented word). We will then select a matching word by weighting these
+similarity rankings (using the neurally predicted memory fingerprint to define
+the weights). In the \textit{inconsistent} condition, we will next present the
+word that \textit{least} matches the previous word, using the decoded
+fingerprint. In this way, the neurally driven \textit{consistent} and
+\textit{inconsistent} conditions of Experiment 3 are analogous to the
+behaviorally driven \textit{stabilize} and \textit{destabilize} conditions of
+Experiment 2. We also plan to run yoked control conditions, whereby the
+presentation orders participants exhibit are driven by \textit{other}
+participants' neural data.
 
-\paragraph{Key analyses.}  The overarching goal of Experiment 3 is to test the limits of how the fine-scale temporal dynamics of how people organize incoming information may be leveraged to enhance (or diminish) memory.
-The temporal resolution of adaptive learning approaches is limited by the temporal resolution of the measurements on which the adaptive aspects are based.
-Whereas recall behaviors may only be assessed periodically (e.g.\ in Experiment 2, during each recall period-- after the participant studies the list), the neural activity driving those behaviors may be measured and analyzed many times per second.
-In this way, we seek to use Experiment 3 to answer deep questions about the temporal dynamics of how people organize their memories.
-For example, do the same principles about fingerprint stability leading to enhanced memory that we discovered in our pilot studies (Figs.~\ref{fig:random_pilot}, \ref{fig:adaptive_pilot}) also apply to the moment-by-moment fluctuations we will estimate in Experiment 3?
-And do people exhibit similar dynamics in those fluctuations within (or across) lists?
-Or are those dynamics predictive of recall performance?
-In addition to examining the neural estimates of memory fingerprint dynamics, we will also measure the improvement (or detriment) in memory performance across the different experimental conditions (analogous to our proposed analyses in Experiment 2).
+Using our prior work (Fig.~\ref{fig:clust}) to approximate the relevant
+parameters, we ran a power analysis (power: 0.8, Type 1 error rate: 0.05) to
+estimate the number of participants we will need to achieve sufficient power.
+The analysis indicates that we will require at least 42 participants in each
+experimental condition; we plan to run 50 participants using the neurally
+driven adaptive design described above, plus an additional 50 in the yoked
+control condition.
 
-Our second set of questions is more general, and is intended to parallel several ideas in Module 2.
-Specifically, can we use neural data to uncover (and leverage) the fine-scale temporal dynamics of learning and memory?
-In Module 1 (Experiment 3) our focus is on uncovering (and leveraging) the fine-scale temporal dynamics of list learning.
-In Module 2 our focus is on uncovering (and leveraging) the fine-scale temporal dynamics of STEM learning.
+\paragraph{Key analyses.} The overarching goal of Experiment 3 is to test the
+limits of how the fine-scale temporal dynamics of how people organize incoming
+information may be leveraged to enhance (or diminish) memory. The temporal
+resolution of adaptive learning approaches is limited by the temporal
+resolution of the measurements on which the adaptive aspects are based. Whereas
+recall behaviors may only be assessed periodically (e.g.\ in Experiment 2,
+during each recall period-- after the participant studies the list), the neural
+activity driving those behaviors may be measured and analyzed many times per
+second. In this way, we seek to use Experiment 3 to answer deep questions about
+the temporal dynamics of how people organize their memories. For example, do
+the same principles about fingerprint stability leading to enhanced memory that
+we discovered in our pilot studies (Figs.~\ref{fig:random_pilot},
+\ref{fig:adaptive_pilot}) also apply to the moment-by-moment fluctuations we
+will estimate in Experiment 3? And do people exhibit similar dynamics in those
+fluctuations within (or across) lists? Or are those dynamics predictive of
+recall performance? In addition to examining the neural estimates of memory
+fingerprint dynamics, we will also measure the improvement (or detriment) in
+memory performance across the different experimental conditions (analogous to
+our proposed analyses in Experiment 2).
+
+Our second set of questions is more general, and is intended to parallel
+several ideas in Module 2. Specifically, can we use neural data to uncover (and
+leverage) the fine-scale temporal dynamics of learning and memory? In Module 1
+(Experiment 3) our focus is on uncovering (and leveraging) the fine-scale
+temporal dynamics of list learning. In Module 2 our focus is on uncovering (and
+leveraging) the fine-scale temporal dynamics of STEM learning.

### 2023-01-24 paper/main.tex
added darya as a co-author, removed kirsten as a co-author (by her request, per email on 1/23/23)

@@
 
-\author{Jeremy R. Manning\textsuperscript{1, *}, Kirsten
-Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C.
-Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Allison M
-Frantz\textsuperscript{1},\\Bryan J. Bollinger\textsuperscript{1}, Campbell E.
-Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
-3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
-University\\\textsuperscript{3}Akili
+\author{Jeremy R. Manning\textsuperscript{1, *}, Emily
+Whitaker\textsuperscript{1}, Paxton C. Fitzpatrick\textsuperscript{1},
+\\Madeline R. Lee\textsuperscript{1}, Allison M. Frantz\textsuperscript{1},
+Bryan J. Bollinger\textsuperscript{1},\\Darya Romanova\textsuperscript{1},
+Campbell E. Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
+2}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili
 Interactive\\\textsuperscript{*}Corresponding author:
@@
 
-Conceptualization: JRM and ACH. Methodology: JRM and ACH. Software: JRM, KZ,
-PCF, CEF, and ACH. Analysis: JRM, PCF, and ACH. Data collection: KZ, EW, PCF,
-MRL, AMF, BJB, and CEF. Data curation and management: KZ, EW, PCF, MRL, ACH.
-Writing (original draft): JRM. Writing (review and editing): KZ, EW, PCF, MRL,
-AMF, BJB, CEF, and ACH. Supervision: JRM and ACH. Project administration: KZ,
-EW, and PCF. Funding acquisition: JRM.
+Conceptualization: JRM and ACH. Methodology: JRM and ACH. Software: JRM, PCF,
+CEF, and ACH. Analysis: JRM, PCF, and ACH. Data collection: EW, PCF, MRL, AMF,
+BJB, DR, and CEF. Data curation and management: EW, PCF, MRL, ACH. Writing
+(original draft): JRM. Writing (review and editing): EW, PCF, MRL, AMF, BJB,
+DR, CEF, and ACH. Supervision: JRM and ACH. Project administration: EW, and
+PCF. Funding acquisition: JRM.
 
@@
 collection efforts from Rachel Chacko, Joseph Finkelstein, Sheherzad Mohydin,
-Lucy Owen, Gal Perlman, Darya Romanova, Jake Rost, Jessica Tin, Marisol Tracy,
-and Peter Tran. Our work was supported in part by NSF CAREER Award Number
+Lucy Owen, Gal Perlman, Jake Rost, Jessica Tin, Marisol Tracy,
+Peter Tran, and Kirsten Ziman. Our work was supported in part by NSF CAREER Award Number
 2145172 to JRM. The content is solely the responsibility of the authors and

### 2023-01-24 paper/supplement.tex
added darya as a co-author, removed kirsten as a co-author (by her request, per email on 1/23/23)

@@
 
-\author{Jeremy R. Manning\textsuperscript{1, *}, Kirsten
-Ziman\textsuperscript{1, 2}, Emily Whitaker\textsuperscript{1},\\Paxton C. 
-Fitzpatrick\textsuperscript{1}, Madeline R. Lee\textsuperscript{1}, Allison M
-Frantz\textsuperscript{1},\\Bryan J. Bollinger\textsuperscript{1}, Campbell E.
-Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
-3}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Princeton
-University\\\textsuperscript{3}Akili
+\author{Jeremy R. Manning\textsuperscript{1, *}, Emily
+Whitaker\textsuperscript{1}, Paxton C. Fitzpatrick\textsuperscript{1},
+\\Madeline R. Lee\textsuperscript{1}, Allison M. Frantz\textsuperscript{1},
+Bryan J. Bollinger\textsuperscript{1},\\Darya Romanova\textsuperscript{1},
+Campbell E. Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
+2}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili
 Interactive\\\textsuperscript{*}Corresponding author:

### 2023-01-24 paper/main.tex
entered edits from BJB

@@
 
-\begin{abstract} \footnotesize{We perceive, interpet, and remember ongoing
+\begin{abstract} \footnotesize{We perceive, interpret, and remember ongoing
 experiences through the lens of our prior experiences. Inferring that we are in
@@
 experiences can take away very different meanings and memories. One reason is
-that our subjective experiences in the moment are shaped in part the
+that our subjective experiences in the moment are shaped in part by the
 idiosyncratic prior experiences, memories, goals, thoughts, expectations, and
@@
 experiences and other naturalistic stimuli that enable people to form deep and meaningful
-situation models and schemas have no obvious analog in simple word lists.  Often lists
+situation models and schemas have no obvious analog in simple word lists.  Often, lists
 in free recall studies are explicitly \textit{designed} to be devoid of exploitable
-temporal structure, for example by sorting the words in a random order~\citep{Kaha12}.
+temporal structure, for example, by sorting the words in a random order~\citep{Kaha12}.
 
@@
 on early lists to manipulate, in real-time, the presentation orders of
-subsequent lists. In this adaptive condition we varied the agreement between
+subsequent lists. In this adaptive condition, we varied the agreement between
 how participants preferred to organize their memories of the studied items
@@
 letter, color, and location), and a final adaptive condition. Each of these
-conditions are described in the \textit{Experimental design} subsection below.
+conditions is described in the \textit{Experimental design} subsection below.
 
@@
 primary control conditions and in the adaptive condition. In all of the other
-conditions we set a target enrollment of at least 30 participants. Because our
+conditions, we set a target enrollment of at least 30 participants. Because our
 data collection procedures entailed the coordinated efforts of 12 researchers
@@
 words as the Euclidean distances between their $(r, g, b)$ color or $(x, y)$
-location vectors, respectively. Therefore the color and location distance
+location vectors, respectively. Therefore, the color and location distance
 measures always take on non-negative real values (upper-bounded at 441.67 for
@@
 values, $a$, and the candidate word's feature values, $b$) to similarity
-scores: \begin{equation} \mathrm{similarity}(a, b) = \exp\{-\tau \cdot
-\mathrm{distance}(a, b)\}, \label{eqn:similarity} \end{equation} where $\tau =
-1$ in our implementation. We note that increasing the value of $\tau$ would
-amplify the influence of similarity on order, and decreasing the value of
-$\tau$ would diminish the influence of similarity on order. Also note that this
-approach requires $\tau > 0$. Finally, we computed a set of normalized
-similarity values by dividing the similarities by their sum: \begin{equation}
+scores:
+
+\begin{equation}
+ \mathrm{similarity}(a, b) = \exp\{-\tau \cdot
+\mathrm{distance}(a, b)\}, \label{eqn:similarity} 
+\end{equation} 
+where $\tau = 1$ in our implementation. We note that increasing the value of
+$\tau$ would amplify the influence of similarity on order, and decreasing the
+value of $\tau$ would diminish the influence of similarity on order. Also note
+that this approach requires $\tau > 0$. Finally, we computed a set of
+normalized similarity values by dividing the similarities by their sum:
+
+\begin{equation}
 \mathrm{similarity}_{\mathrm{normalized}}(a, b) = \frac{\mathrm{similarity}(a,
-b)}{\sum_{i=1}^n \mathrm{similarity}(a, i)}, \end{equation} where in the
-denominator, $i$ takes on each of the $n$ feature values of the to-be-presented
-words. The resulting set of normalized similarity scores sums to 1.
+b)}{\sum_{i=1}^n \mathrm{similarity}(a, i)}, 
+\end{equation} 
+where in the denominator, $i$ takes on each of the $n$ feature values of the
+to-be-presented words. The resulting set of normalized similarity scores sums
+to 1.
 
@@
 considered both group-level effects (i.e., effects that tended to be common
-across individuals) and participant-level effects (i.e., effect that were
+across individuals) and participant-level effects (i.e., effects that were
 idiosyncratic across individuals).
@@
 participants \textit{organized} their memories of words from the ordered lists.
-When we sorted lists semantically participants displayed stronger semantic
-clustering; when we sorted lists lexicographically they displayed stronger
-lexicographic clustering; and when we sorted lists visually they displayed
+When we sorted lists semantically, participants displayed stronger semantic
+clustering; when we sorted lists lexicographically, they displayed stronger
+lexicographic clustering; and when we sorted lists visually, they displayed
 stronger visual clustering. Clustering along the unmanipulated feature
@@
 our predictions, we may perceive the ``rules'' or ``situation'' to have
-changed. \textit{Event boundaries} denote abrupts change in the state of our
-experience, for example when we transition from one situation to
+changed. \textit{Event boundaries} denote abrupt changes in the state of our
+experience, for example, when we transition from one situation to
 another~\citep{RadvZack17, ZwaaRadv98}. Crossing an event boundary can impair

### 2023-01-24 paper/main.tex
entered ECW edits

@@
 
-\author{Jeremy R. Manning\textsuperscript{1, *}, Emily
+\author{Jeremy R. Manning\textsuperscript{1, *}, Emily C.
 Whitaker\textsuperscript{1}, Paxton C. Fitzpatrick\textsuperscript{1},
@@
 organized their memories of the studied words. Because random word lists are
-unstructured by design, it is not clear if or how non-trivial situation models
+unstructured by design, it is not clear if, or how, non-trivial situation models
 might apply to these stimuli. Nevertheless, there are \textit{some}
@@
 
-Over the past half-century, context-based models have enjoyed impressive
-success at explaining many stereotyped behaviors observed during free recall
-and other list-learning tasks~\citep{Este55a, RaaiShif80, GlenEtal83,
-HowaKaha02a, SiroEtal05, KimbEtal07, PolyKaha08, SedeEtal08, PolyEtal09,
-ShanHowa12}. These phenomena include the well-known recency and primacy effects
-(superior recall of items from the end and, to a lesser extent, from the
-beginning of the study list), as well as semantic and temporal clustering
-effects~\citep{KahaEtal08b}. The contiguity effect is an example of temporal
-clustering, which is perhaps the dominant form of organization in free recall.
-This effect can be seen in the tendency for people to successively recall items
-that occupied neighboring positions in the studied list~\citep{Kaha96}. There
-are also striking effects of semantic clustering~\citep{RomnEtal93, Bous53,
-BousEtal54, JenkRuss52, MannKaha12}, whereby the recall of a given item is more
-likely to be followed by recall of a similar or related item than a dissimilar
-or unrelated one. In general, people organize memories for words along a wide
-variety of stimulus dimensions. As formalized by models like the
-\textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the
-stimulus features associated with each word (e.g.\ the word's meaning, font
-size, font color, location on the screen, size of the object the word
-represents, etc.) are incorporated into the participant's mental context
+Over the past half-century, context-based models have had impressive success at
+explaining many stereotyped behaviors observed during free recall and other
+list-learning tasks~\citep{Este55a, RaaiShif80, GlenEtal83, HowaKaha02a,
+SiroEtal05, KimbEtal07, PolyKaha08, SedeEtal08, PolyEtal09, ShanHowa12}. These
+phenomena include the well-known recency and primacy effects (superior recall
+of items from the end and, to a lesser extent, from the beginning of the study
+list), as well as semantic and temporal clustering effects~\citep{KahaEtal08b}.
+The contiguity effect is an example of temporal clustering, which is perhaps
+the dominant form of organization in free recall. This effect can be seen in
+the tendency for people to successively recall items that occupied neighboring
+positions in the studied list~\citep{Kaha96}. There are also striking effects
+of semantic clustering~\citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52,
+MannKaha12}, whereby the recall of a given item is more likely to be followed
+by recall of a similar or related item than a dissimilar or unrelated one. In
+general, people organize memories for words along a wide variety of stimulus
+dimensions. As formalized by models like the \textit{Context Maintenance and
+Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each
+word (e.g.\ the word's meaning, size of the object the word represents, the
+letters that make up the word, font size, font color, location on the screen,
+etc.) are incorporated into the participant's mental context
 representation~\citep{SmitVela01, MannEtal15, Mann20, MannEtal11, MannEtal12}.
@@
 
-Participants received course credit for enrolling in our study. We asked each
-participant to fill out a demographic survey that included questions about
-their age, gender, ethnicity, race, education, vision, reading
-impairments, medications or recent injuries, coffee consumption on the day of
-testing, and level of alertness at the time of testing. All components of the
-demographics survey were optional. One participant elected not to fill out any
-part of the demographic survey, and all other participants answered some or all
-of the survey questions.
+Participants either received course credit or a one-time \$10 payment for
+enrolling in our study. We asked each participant to fill out a demographic
+survey that included questions about their age, gender, ethnicity, race,
+education, vision, reading impairments, medications or recent injuries, coffee
+consumption on the day of testing, and level of alertness at the time of
+testing. All components of the demographics survey were optional. One
+participant elected not to fill out any part of the demographic survey, and all
+other participants answered some or all of the survey questions.
 
@@
 participant. First, we choose a word uniformly and at random from the set of
-candidates. Next, we compute the distances between the chosen word's feature(s)
+candidates. Second, we compute the distances between the chosen word's feature(s)
 and the corresponding feature(s) of all yet-to-be-presented words. Third, we
@@
 $B$ have been recalled, the best possible next matching recall will be a large
-item. And all subsequent correct recalls must also be large items-- so for
+item. All subsequent correct recalls must also be large items-- so for
 those later recalls it becomes difficult to determine whether the participant
@@
 recalling the yet-to-be-recalled items in a random order. In general, the
-precise order and blend of feature values expressed in a given list, the orders
-and numbers of correct recalls a participant makes, the number of intervening
+precise order and blend of feature values expressed in a given list, the order
+and number of correct recalls a participant makes, the number of intervening
 presentation positions between successive recalls, and so on, can all affect
@@
 0.685$). These comparisons suggest that recall accuracy has a directional
-component (i.e., accuracy is affected differently by removing features later
+component: accuracy is affected differently by removing features later
 that had been present earlier versus adding features later that had
-\textit{not} been present earlier). In contrast, we found that participants
+\textit{not} been present earlier. In contrast, we found that participants
 exhibited more temporal and feature-based clustering when we added irrelevant
@@
 sorting lists by different features appeared to have slightly different effects
-on overall memory performance and temporal clustering, and people tended to
+on overall memory performance and temporal clustering.  Participants also tended to
 cluster their recalls along a given feature dimension more when the studied
@@
 Conceptualization: JRM and ACH. Methodology: JRM and ACH. Software: JRM, PCF,
-CEF, and ACH. Analysis: JRM, PCF, and ACH. Data collection: EW, PCF, MRL, AMF,
-BJB, DR, and CEF. Data curation and management: EW, PCF, MRL, ACH. Writing
-(original draft): JRM. Writing (review and editing): EW, PCF, MRL, AMF, BJB,
-DR, CEF, and ACH. Supervision: JRM and ACH. Project administration: EW, and
+CEF, and ACH. Analysis: JRM, PCF, and ACH. Data collection: ECW, PCF, MRL, AMF,
+BJB, DR, and CEF. Data curation and management: ECW, PCF, MRL, and ACH. Writing
+(original draft): JRM. Writing (review and editing): ECW, PCF, MRL, AMF, BJB,
+DR, CEF, and ACH. Supervision: JRM and ACH. Project administration: ECW and
 PCF. Funding acquisition: JRM.

### 2023-01-24 paper/supplement.tex
entered ECW edits

@@
     \caption{\textbf{Recall accuracy by study list number.}  Each panel displays the average recall accuracy (across participants) as a function of the number of studied lists for the
-    random conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}.  The conditions are denoted by color.  Note that the first four
+    random conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}).  The conditions are denoted by color.  Note that the first four
     lists in all of the ``adaptive'' conditions were ordered randomly to compute a baseline fingerprint for each participant prior to initiating the adaptive ordering procedure.  All error

### 2023-01-24 paper/main.tex
entered CEF and DR comments

@@
 
-We also designed two conditions where we varied the words' visual appearnaces
+We also designed two conditions where we varied the words' visual appearances
 across lists. In the \textit{reduced (early)} condition, we followed the
@@
 visual features to \textit{any} lists (comparisons of clustering on feature
-rich anversusd reduced lists are reported above; temporal clustering in reduced
+rich versus reduced lists are reported above; temporal clustering in reduced
 versus reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
@@
 relative to feature rich participants, and on early lists; color: $t(126) =
-2.099, p = 0.038$; location: $t(126) = 4.392, p < 0.001$), but showed now
+2.099, p = 0.038$; location: $t(126) = 4.392, p < 0.001$), but showed no
 reliable differences in semantic (category: $t(126) = 0.204, p = 0.839$; size:
@@
 removes uncertainty about whether participants' assigned conditions might just
-``happen'' to match their preferred ways or organizing their memories.
+``happen'' to match their preferred ways of organizing their memories.
 

### 2023-01-24 paper/main.tex
entered PCF edits

@@
 draw on our prior related experiences, and so on. To study these phenomena, we
-asked participants to perform simple word list learning tasks. Across different
+asked participants to perform simple word list-learning tasks. Across different
 experimental conditions, we held the set of to-be-learned words constant, but
-we manipulated how irrelevant visual features changed across words and lists,
+we manipulated how incidental visual features changed across words and lists,
 along with the orders in which the words were studied. We found that these
 manipulations affected not only how the participants recalled the manipulated
-lists, but also how they recalled later (random) lists. Our work shows how
-structure in our ongoing experiences can exert influence over how we remember
-our current experiences and unrelated subsequent experiences.
-
-\textbf{Keywords: episodic memory, free recall, incidental features, implicit priming, temporal order}}
+lists, but also how they recalled later (randomly ordered) lists. Our work
+shows how structure in our ongoing experiences can influence how we remember
+both our current experiences and unrelated subsequent experiences.
+
+\textbf{Keywords: episodic memory, free recall, incidental features, implicit
+priming, temporal order}}
 
@@
 experiences can take away very different meanings and memories. One reason is
-that our subjective experiences in the moment are shaped in part by the
+that our moment-by-moment subjective experiences are shaped in part by the
 idiosyncratic prior experiences, memories, goals, thoughts, expectations, and
@@
 The contexts we encounter help us to construct \textit{situation
-models}~\citep{RangRitc12, MannEtal15, RadvCope06, ZwaaEtal95} or
-\textit{schemas}~\citep{MasiEtal22, BaldEtal18} that describe how experiences
-are likely to unfold based on our prior experiences with similar contextual
-cues. For example, when we enter a sit-down restaurant, we might expect to be
-seated at a table, given a menu, and served food. Priming someone to expect a
-particular situation or context can also influence how they resolve potential
-ambiguities in their ongoing experiences, including in ambiguous movies and
-narratives~\citep{YeshEtal17}.
+models}~\citep{RangRitc12, MannEtal15, RadvCope06, ZwaaEtal95, ZwaaRadv98} or
+\textit{schemas}~\citep{MasiEtal22, BaldEtal18, TseEtal07} that describe how
+experiences are likely to unfold based on our prior experiences with similar
+contextual cues. For example, when we enter a sit-down restaurant, we might
+expect to be seated at a table, given a menu, and served food. Priming someone
+to expect a particular situation or context can also influence how they resolve
+potential ambiguities in their ongoing experiences, including in ambiguous
+movies and narratives~\citep{YeshEtal17, RissEtal03}.
 
@@
 processes is particularly important in self-cued memory tasks, such as free
-recall, where the retrieval cue is ``context'' itself. Conceptually, the same
-general processes might be said to describe how real-world contexts evolve
-during natural experiences. However, this is still an open area of
-study~\citep{Mann20, Mann21a}.
+recall, where the retrieval cue is ``context'' itself~\citep{HowaKaha02a}.
+Conceptually, the same general processes might be said to describe how
+real-world contexts evolve during natural experiences. However, this is still
+an open area of study~\citep{Mann20, Mann21a}.
 
@@
 SiroEtal05, KimbEtal07, PolyKaha08, SedeEtal08, PolyEtal09, ShanHowa12}. These
-phenomena include the well-known recency and primacy effects (superior recall
+phenomena include the well known recency and primacy effects (superior recall
 of items from the end and, to a lesser extent, from the beginning of the study
-list), as well as semantic and temporal clustering effects~\citep{KahaEtal08b}.
-The contiguity effect is an example of temporal clustering, which is perhaps
-the dominant form of organization in free recall. This effect can be seen in
-the tendency for people to successively recall items that occupied neighboring
-positions in the studied list~\citep{Kaha96}. There are also striking effects
-of semantic clustering~\citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52,
-MannKaha12}, whereby the recall of a given item is more likely to be followed
-by recall of a similar or related item than a dissimilar or unrelated one. In
-general, people organize memories for words along a wide variety of stimulus
-dimensions. As formalized by models like the \textit{Context Maintenance and
-Retrieval Model}~\citep{PolyEtal09}, the stimulus features associated with each
-word (e.g.\ the word's meaning, size of the object the word represents, the
-letters that make up the word, font size, font color, location on the screen,
-etc.) are incorporated into the participant's mental context
-representation~\citep{SmitVela01, MannEtal15, Mann20, MannEtal11, MannEtal12}.
-During a memory test, any of these features may serve as a memory cue, which in
-turn leads the participant to recall in succession words that share stimulus
-features.
+list), as well as semantic and temporal clustering effects~\citep{KahaEtal08b,
+HowaKaha02b}. The contiguity effect is an example of temporal clustering, which
+is perhaps the dominant form of organization in free recall. This effect can be
+seen in people's tendencies to successively recall items that occupied
+neighboring positions in the studied list~\citep{Kaha96}. There are also
+striking effects of semantic clustering~\citep{RomnEtal93, Bous53, BousEtal54,
+JenkRuss52, MannKaha12}, whereby the recall of a given item is more likely to
+be followed by recall of a similar or related item than a dissimilar or
+unrelated one. In general, people organize memories for words along a wide
+variety of stimulus dimensions. As formalized by models like the
+\textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the
+stimulus features associated with each word (e.g.\ the word's meaning, size of
+the object the word represents, the letters that make up the word, font size,
+font color, location on the screen, etc.) are incorporated into the
+participant's mental context representation~\citep{SmitVela01, MannEtal11,
+MannEtal12, MannEtal15, Mann20}. During a memory test, any of these features
+may serve as a memory cue, which in turn leads the participant to recall in
+succession words that share stimulus features.
 
@@
 
-A key mystery is whether (and how) the sorts of situation models and schemas that people
-use to organize their memories of real-world experiences might map onto the
-clustering effects that reflect how people organize their memories for word
-lists. On one hand, situation models and clustering effects both reflect
+A key mystery is whether (and how) the sorts of situation models and schemas
+that people use to organize their memories of real-world experiences might map
+onto the clustering effects that reflect how people organize their memories for
+word lists. On one hand, both situation models and clustering effects reflect
 statistical regularities in ongoing experiences. Our memory systems exploit
@@
 BoweEtal79, MomeEtal17}. On the other hand, the rich structure of real-world
-experiences and other naturalistic stimuli that enable people to form deep and meaningful
-situation models and schemas have no obvious analog in simple word lists.  Often, lists
-in free recall studies are explicitly \textit{designed} to be devoid of exploitable
-temporal structure, for example, by sorting the words in a random order~\citep{Kaha12}.
+experiences and other naturalistic stimuli that enable people to form deep and
+meaningful situation models and schemas have no obvious analog in simple word
+lists. Often, lists in free recall studies are explicitly \textit{designed} to
+be devoid of exploitable temporal structure, for example, by sorting the words
+in a random order~\citep{Kaha12}.
 
@@
 participants to study and freely recall a series of word lists
-(Fig.~\ref{fig:exp}). Across the different conditions in the experiment, we
-varied the lists' appearances and presentation orders in different ways across
-lists. The studied items (words) were designed to vary along three general
-dimensions: semantic (word \textit{category}, and physical \textit{size} of the
-referent), lexicographic (word \textit{length} and \textit{first letter}), and
-visual (font \textit{color} and the onscreen \textit{location} of each word).
-In two conditions, we manipulated whether the words' appearances were fixed or
-variable within each list. In six manipulation conditions, we asked
-participants to study and recall eight lists whose items were sorted by a
-target feature (e.g., word category). Next, we asked them to study and recall
-an additional eight lists whose items had the same features, but that were
-sorted in a random temporal order. We were interested in how these 
-manipulations affected participants' recall behaviors on early (manipulated) lists,
-as well as how order manipulations on early lists affected recall behaviors on
-later (random) lists. We used two control conditions as a baseline;
-in these control conditions all of the lists were sorted randomly, but we
-manipulated the presence or absence of the visual features. Finally, in an
-\textit{adaptive} experimental condition we used participants' recall behaviors
-on early lists to manipulate, in real-time, the presentation orders of
-subsequent lists. In this adaptive condition, we varied the agreement between
-how participants preferred to organize their memories of the studied items
-versus the orders in which the items were presented.
+(Fig.~\ref{fig:exp}). In the different conditions in our experiment, we varied
+the lists' appearances and presentation orders in different ways. The studied
+items (words) were designed to vary along three general dimensions: semantic
+(word \textit{category}, and physical \textit{size} of the referent),
+lexicographic (word \textit{length} and \textit{first letter}), and visual
+(font \textit{color} and the onscreen \textit{location} of each word). We used
+two control conditions as a baseline; in these control conditions all of the
+lists were sorted randomly, but we manipulated the presence or absence of the
+visual features. In two conditions, we manipulated whether the words'
+appearances were fixed or variable within each list. In six manipulation
+conditions, we asked participants to first study and recall eight lists whose
+items were sorted by a target feature (e.g., word category), and then study and
+recall an additional eight lists whose items had the same features, but that
+were sorted in a random temporal order. We were interested in how these
+manipulations affected participants' recall behaviors on early (manipulated)
+lists, as well as how order manipulations on early lists affected recall
+behaviors on later (randomly ordered) lists. Finally, in an \textit{adaptive}
+experimental condition we used participants' recall behaviors on early lists to
+manipulate, in real-time, the presentation orders of subsequent lists. In this
+adaptive condition, we varied the agreement between how participants preferred
+to organize their memories of the studied items versus the orders in which the
+items were presented.
 
@@
 
-We enrolled a total of 491 Dartmouth undergraduate students across 11
+We enrolled a total of 491 members of the Dartmouth College community across 11
 experimental conditions. The conditions included two controls (feature rich,
-reduced), two visual manipulation conditions (reduced (early) and reduced
-(late)), six order manipulation conditions (category, size, length, first
+reduced), two visual manipulation conditions [reduced (early) and reduced
+(late)], six order manipulation conditions (category, size, length, first
 letter, color, and location), and a final adaptive condition. Each of these
@@
 allergies, one reported a recently torn ACL/MCL, and one reported a concussion
-from several months prior. The participants reported consuming 0 -- 3 cups of
+from several months prior. The participants reported consuming 0--3 cups of
 coffee prior to the testing session (mean: 0.32 cups; standard deviation: 0.58
@@
 alert'' (2). Across all participants, the full range of alertness levels were
-reported (range: -2 -- 2; mean: 0.35; standard deviation: 0.89).
+reported (range: -2--2; mean: 0.35; standard deviation: 0.89).
 
 We dropped from our dataset the one participant who reported having abnormal
-color vision, as well as 39 participants whose data were corrupted due to
+color vision, as well as 38 participants whose data were corrupted due to
 technical failures while running the experiment or during the daily database
@@
 Our experiment is a variant of the classic free recall paradigm that we term
-\textit{feature-rich free recall}. In feature-rich free recall, participants
-study 16 lists, each comprised of 16 words that vary along a number of stimulus
-dimensions (Fig.~\ref{fig:exp}). The stimulus dimensions include two semantic
-features related to the \textit{meanings} of the words (semantic category,
-referent object size), two lexicographic features related to the
-\textit{letters} that make up the words (word length in number of letters,
-identity of the word's first letter), and two visual features that are
+``\textit{feature-rich free recall}.'' In feature-rich free recall,
+participants study 16 lists, each comprised of 16 words that vary along a
+number of stimulus dimensions (Fig.~\ref{fig:exp}). The stimulus dimensions
+include two semantic features related to the \textit{meanings} of the words
+(semantic category, referent object size), two lexicographic features related
+to the \textit{letters} that make up the words (word length in number of
+letters, identity of the word's first letter), and two visual features that are
 independent of the words themselves (text color, presentation location). Each
-list contains four words from each of four different semantic categories and
-two object sizes; all other stimulus features are randomized. After studying
-each list, the participant attempts to recall as many words as they can from
-that list, in any order they choose. Because each individual word is associated
-with several well-defined (and quantifiable) features, and because each list
-incorporates a diverse mix of feature values along each dimension, this allows
-us to estimate which features participants are considering or leveraging in
-organizing their memories.
+list contains four words from each of four different semantic categories, with
+two object sizes reflected across all of the words. After studying each list,
+the participant attempts to recall as many words as they can from that list, in
+any order they choose. Because each individual word is associated with several
+well defined (and quantifiable) features, and because each list incorporates a
+diverse mix of feature values along each dimension, this allows us to estimate
+which features participants are considering or leveraging in organizing their
+memories.
 
@@
 word features and/or presentation orders within and/or across lists. The rows
-display representative (illustrated) examples of items from the first lists
+display representative (illustrated) examples of items from the first list
 participants might encounter in each condition. The rectangles during the
@@
 likely able to ``fit in a standard shoebox'' or ``large'' if the object was
-larger than a shoebox. Semantic categories varied in how many object sizes they
-reflected (mean number of different sizes per category: 1.33; standard
-deviation: 0.49). The numbers of words in each semantic category also varied
-from 12 -- 28 (mean number of words per category: 17.07; standard deviation
-number of words: 4.65). We also identified lexicographic features for each
-word, including the words' first letters and lengths (i.e., number of letters).
-Across all categories, all possible first letters were represented except for
-`Q' (average number of unique first letters per category: 11; standard
-deviation: 2 letters). Word lengths ranged from 3 -- 12 letters (average: 6.17
-letters; standard deviation: 2.06 letters).
+larger than a shoebox. Most semantic categories comprised words that reflected
+both ``small'' and ``large'' object sizes, but several included only one or the
+other (e.g., all countries, US states, and cities are larger than a shoebox;
+mean number of different sizes per category: 1.33; standard deviation: 0.49).
+The numbers of words in each semantic category also varied from 12--28 (mean
+number of words per category: 17.07; standard deviation number of words: 4.65).
+We also identified lexicographic features for each word, including the words'
+first letters and lengths (i.e., number of letters). Across all categories, all
+possible first letters were represented except for `Q' (average number of
+unique first letters per category: 11; standard deviation: 2 letters). Word
+lengths ranged from 3--12 letters (average: 6.17 letters; standard deviation:
+2.06 letters).
 
@@
 every condition we randomized the study order of these lists across
-participants. For participants in some conditions, on some lists, we also
-randomly varied two additional visual features associated with each word: the
-presentation font color, and the word's onscreen location. These attributes
-were assigned independently for each word (and for every participant). These
-visual features were varied for words in all lists and conditions except for
-the ``reduced'' condition (all lists), the first eight lists of the ``reduced
-(early)'' condition, and the last eight lists of the ``reduced (late)''
-condition. In these latter cases, words were all presented in black at the
-center of the experimental computer's display.
+participants. For participants in most conditions, on some or all of the lists,
+we also randomly varied two additional visual features associated with each
+word: the presentation font color, and the word's onscreen location. These
+attributes were assigned independently for each word (and for every
+participant). These visual features were varied for words in all lists and
+conditions except for the ``reduced'' condition (all lists), the first eight
+lists of the ``reduced (early)'' condition, and the last eight lists of the
+``reduced (late)'' condition. In these latter cases, words were all presented
+in black at the center of the experimental computer's display.
 
@@
 presentation locations to each word, we selected two floating point numbers
-uniformly and at random (one for the word's horizontal $x$ coordinate and the
-other for its vertical $y$ coordinate). The bounds of these coordinates were
+uniformly and at random (one for the word's horizontal $x$-coordinate and the
+other for its vertical $y$-coordinate). The bounds of these coordinates were
 selected to cover the entire visible area of the display without cutting off
-any part of the words. The words were shown on 27~in (diagonal) Retina 5K iMac
+any part of the words. The words were shown on 27-in (diagonal) Retina 5K iMac
 displays (resolution: 5120 $\times$ 2880 pixels).
@@
 participants' verbal recalls into text. This allows recalls to be transcribed
-in real time-- a distinguishing feature of the experiment; in typical verbal
-recall experiments the audio data must be parsed and transcribed manually. In
+in real time---a distinguishing feature of the experiment; in typical verbal
+recall experiments, the audio data must be parsed and transcribed manually. In
 prior work, we used a similar experimental setup (equivalent to the ``reduced''
@@
 irrelevant features varied freely across words on early lists, in that we did
-not consider irrelevant features in ordering the early lists. However, some
-features were correlated-- for example, some semantic categories of words
-referred to objects that tended to be a particular size, which meant that
+not consider irrelevant features in ordering the early lists. However, we note
+that some features were correlated---for example, some semantic categories of
+words referred to objects that tended to be a particular size, which meant that
 category and size were not fully independent. On late lists, the words were
@@
 in the English alphabet, respectively. For example, two words that started with
-the same letter would have a ``first letter'' distance of 0, and words starting
-with `J' and `A' respectively would have a first letter distance of 9. Because
-words' lengths and letters' positions in the alphabet are always integers,
-these discrete distances always take on integer values. Finally, the visual
-features (color and location) are \textit{continuous} and
-\textit{multivariate}, in that each ``feature'' takes on multiple (positive)
-real values. We defined the ``color'' and ``location'' distances between two
-words as the Euclidean distances between their $(r, g, b)$ color or $(x, y)$
-location vectors, respectively. Therefore, the color and location distance
-measures always take on non-negative real values (upper-bounded at 441.67 for
-color, or 27~in for location, reflecting the distances between the
+the same letter would have a ``first letter'' distance of 0, and a pair of
+words starting with `J' and `A' would have a first letter distance of 9.
+Because words' lengths and letters' positions in the alphabet are always
+integers, these discrete distances always take on integer values. Finally, the
+visual features (color and location) are \textit{continuous} and
+\textit{multivariate}, in that each ``feature'' is defined by multiple
+(positive) real values. We defined the ``color'' and ``location'' distances
+between two words as the Euclidean distances between their $(r, g, b)$ color or
+$(x, y)$ location vectors, respectively. Therefore, the color and location
+distance measures always take on non-negative real values (upper-bounded at
+441.67 for color, or 27~in for location, reflecting the distances between the
 corresponding maximally different vectors).
@@
 participant. First, we choose a word uniformly and at random from the set of
-candidates. Second, we compute the distances between the chosen word's feature(s)
-and the corresponding feature(s) of all yet-to-be-presented words. Third, we
-convert these distances (between the previously presented word's feature
-values, $a$, and the candidate word's feature values, $b$) to similarity
-scores:
+words on the to-be-presented list. Second, we compute the distances between the
+chosen word's feature(s) and the corresponding feature(s) of all
+yet-to-be-presented words. Third, we convert these distances (between the
+previously presented word's feature values, $a$, and the candidate word's
+feature values, $b$) to similarity scores:
 
@@
 (Eqn.~\ref{eqn:similarity}) between the feature value(s) of the previous item,
-$x$, and all yet-to-be-presented items ($a$ -- $g$). Next, we normalize these
+$x$, and all yet-to-be-presented items ($a$--$g$). Next, we normalize these
 similarity scores so that they sum to 1. We lay, in sequence, a set of
@@
 item's chances of selection is proportional to its similarity to the previous
-item, along the given feature dimension.} \label{fig:stick}
+item, along the given feature dimension (e.g., color).} \label{fig:stick}
 
@@
 been presented. The result is an ordered list that tends to change gradually
-along the selected feature dimension.
+along the selected feature dimension (for example ``sorted'' lists, see
+Fig.~\ref{fig:exp}, \textit{Order manipulation} lists).
 
@@
 $B$ have been recalled, the best possible next matching recall will be a large
-item. All subsequent correct recalls must also be large items-- so for
+item. All subsequent correct recalls must also be large items---so for
 those later recalls it becomes difficult to determine whether the participant
@@
 indicates that the observed score was greater than any clustering score one
-might expect by chance; in other words, good evidence that the participant was
+might expect by chance---in other words, good evidence that the participant was
 truly clustering their recalls along the given feature dimension. We applied
@@
 candidate presentation orders. We sought to select the specific order that most
-(or least) matched $f$. Third, for each random permutation, we computed the
-(permutation-corrected) ``fingerprint,'' treating the permutation as though it
-were a potential ``perfect'' recall sequence. (We did not include temporal
+(or least) closely matched $f$. Third, for each random permutation, we computed
+the (permutation-corrected) ``fingerprint,'' treating the permutation as though
+it were a potential ``perfect'' recall sequence. (We did not include temporal
 clustering scores in these fingerprints.) This yielded a ``simulated
@@
 
-Following some of our prior work~\citep{HeusEtal18a, HeusEtal21}, we use
-low-dimensional embeddings to help visualize how participants' memory
+Following some of our prior work~\citep{HeusEtal18a, HeusEtal21, MannEtal22},
+we use low-dimensional embeddings to help visualize how participants' memory
 fingerprints change across lists
@@
 most variance in the data). For two visualizations
-(Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B)
+(Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B),
 we computed an additional set of two-dimensional embeddings for the
@@
 
-The lag-conditional probability (lag-CRP) curve~\citep{Kaha96} reflects the
-probability of recalling a given item after the just-recalled item, as a
-function of their relative encoding positions (lag). In other words, a lag of 1
-indicates that a recalled item was presented immediately after the previously
-recalled item, and a lag of $-3$ indicates that a recalled item came three
-items before the previously recalled item. For each recall transition
+The lag-conditional response probability (lag-CRP) curve~\citep{Kaha96}
+reflects the probability of recalling a given item after the just-recalled
+item, as a function of their relative encoding positions (lag). In other words,
+a lag of 1 indicates that a recalled item was presented immediately after the
+previously recalled item, and a lag of $-3$ indicates that a recalled item came
+three items before the previously recalled item. For each recall transition
 (following the first recall), we computed the lag between the just-recalled
@@
 transitions. In carrying out this analysis, we excluded all incorrect recalls
-and successive repetitions (e.g., recalling the same word twice in a row). This
+and successive repetitions (i.e., recalling the same word twice in a row). This
 yielded, for each list, a 1 by number-of-lags ($-15$ to $+15$; 30 lags in
@@
 the display. Aggregating across all lists for each participant, we found no
-difference in recall accuracy for feature rich versus reduced lists ($t(126) =
--0.290, p = 0.772$). However, participants in the feature rich condition
-clustered their recalls substantially more along every dimension we examined
-(temporal clustering: $t(126) = 10.624, p < 0.001$; category clustering:
-$t(126) = 10.077, p < 0.001$; size clustering: $t(126) = 11.829, p < 0.001$;
-word length clustering: $t(126) = 10.639, p < 0.001$; first letter clustering:
-$t(126) = 7.775, p < 0.001$; see \textit{Permutation-corrected feature
-clustering scores} for more information about how we quantified each
-participant's clustering tendencies.) Taken together, these comparisons suggest
-that adding new features changes how participants organize their memories of
-studied words, even when those new features are independent of the words
-themselves and even when the new features vary randomly across words. We found
-no evidence that those additional uninformative features were distracting (in
-terms of their impact on memory performance), but they did affect participants'
-recall dynamics (measured via their clustering scores).
-
-We also wondered whether adding these irrelevant visual features to later lists
+difference in recall accuracy (i.e., the proportions of correctly recalled
+words) for feature rich versus reduced lists ($t(126) = -0.290, p = 0.772$).
+However, participants in the feature rich condition clustered their recalls
+substantially more along every dimension we examined (temporal clustering:
+$t(126) = 10.624, p < 0.001$; semantic category clustering: $t(126) = 10.077, p
+< 0.001$; size clustering: $t(126) = 11.829, p < 0.001$; word length
+clustering: $t(126) = 10.639, p < 0.001$; first letter clustering: $t(126) =
+7.775, p < 0.001$; see \textit{Permutation-corrected feature clustering scores}
+for more information about how we quantified each participant's clustering
+tendencies.) Taken together, these comparisons suggest that adding new features
+changes how participants organize their memories of studied words, even when
+those new features are independent of the words themselves and even when the
+new features vary randomly across words. We found no evidence that those
+additional uninformative features were distracting (in terms of their impact on
+memory performance), but they did affect participants' recall dynamics
+(measured via their clustering scores).
+
+We also wondered whether adding these incidental visual features to later lists
 (after the participants had already studied impoverished lists), or removing
@@
 Taken together, these comparisons suggest that even when the presence or
-absence of irrelevant visual features is stable across lists, participants
+absence of incidental visual features is stable across lists, participants
 still exhibit some differences in their performance and memory organization
@@
 formed just after event boundaries can be enhanced~\citep[e.g., due to less
-contextual interference between pre- and post-boundary items;][]{MannEtal16}.
+contextual interference between pre- and post-boundary items;][]{MannEtal16,
+PettEtal16, GoldEtal17, FlorEtal17}.
 
@@
 to participants in the feature rich condition (category: $t(125) = 2.524, p =
-0.013$; size:$t(125) = 3.510, p = 0.001$), but showed no reliable differences
+0.013$; size: $t(125) = 3.510, p = 0.001$), but showed no reliable differences
 in lexicographic (length: $t(125) = 0.539, p = 0.591$; first letter: $t(125) =
@@
 defined participants' \textit{memory fingerprints} as the set of their temporal
-and feature clustering scores. A participant's memory fingerprint describes how
-they tend to retrieve memories of the studied items, perhaps searching through
-several feature spaces (or along several representational dimensions). To gain
-insights into the dynamics of how participants' clustering scores tended to
-change over time, we computed the average (across participants) fingerprint
-from each list, from each order manipulation condition
-(Fig.~\ref{fig:fingerprint-trajectories}). We projected these fingerprints into
-a two-dimensional space to help visualize the dynamics (top panels; see
-\textit{Computing low-dimensional embeddings of memory fingerprints}). We found
-that participants' average fingerprints tended to remain relatively stable on
-early lists, and exhibited a ``jump'' to another stable state on later lists.
-The sizes of these jumps varied somewhat across conditions (the Euclidean
-distances between fingerprints in their original high dimensional spaces are
-displayed in the bottom panels). We also averaged the fingerprints across early
-and late lists, respectively, for each condition
+and feature clustering scores (see \textit{Memory fingerprints}). A
+participant's memory fingerprint describes how they tend to retrieve memories
+of the studied items, perhaps searching through several feature spaces (or
+along several representational dimensions). To gain insights into the dynamics
+of how participants' clustering scores tended to change over time, we computed
+the average (across participants) fingerprint from each list, from each order
+manipulation condition (Fig.~\ref{fig:fingerprint-trajectories}). We projected
+these fingerprints into a two-dimensional space to help visualize the dynamics
+(top panels; see \textit{Computing low-dimensional embeddings of memory
+fingerprints}). We found that participants' average fingerprints tended to
+remain relatively stable on early lists, and exhibited a ``jump'' to another
+stable state on later lists. The sizes of these jumps varied somewhat across
+conditions (the Euclidean distances between fingerprints in their original high
+dimensional spaces are displayed in the bottom panels). We also averaged the
+fingerprints across early and late lists, respectively, for each condition
 (Fig.~\ref{fig:fingerprint-trajectories}B). We found that participants'
@@
 clearer story emerged. Within each order manipulation condition, participants
-exhibited a range of feature clustering scores, on both early and late lists
+exhibited a range of feature clustering scores on both early and late lists
 (Fig.~\ref{fig:clustering-scatterplots}A, B). Across every order manipulation
@@
 all $p$s $\leq 0.010$) lists. We found no evidence of a condition-level trend;
-for example the conditions where participants tended to show stronger
+for example, the conditions where participants tended to show stronger
 clustering scores were not correlated with the conditions where participants
 remembered more words (early: $r(4) = 0.526, p = 0.284$; late: $r(4) = -0.257,
-p = 0.623$; see insets of panels A and B). We observed carryover associations
-between feature clustering and recall performance
-(Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who showed stronger
-feature clustering on early lists tended to recall more items on late lists
-(across conditions: $r(179) = 0.492, p < 0.001$; all conditions individually:
-$r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who recalled more items
-on early lists also tended to show stronger feature clustering on late lists
-(across conditions: $r(179) = 0.280, p < 0.001$; all non-visual conditions:
-$r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29) = 0.298, p = 0.103$;
-location: $r(28) = 0.354, p = 0.055$). Neither of these effects showed
-condition-level trends (early feature clustering versus late recall
-probability: $r(4) = -0.299, p = 0.565$; early recall probability versus late
-feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for associations
-between feature clustering and temporal clustering. Across every order
-manipulation condition, participants who exhibited stronger feature clustering
-also exhibited stronger temporal clustering. For early lists
-(Fig.~\ref{fig:clustering-scatterplots}E), this trend held
-overall ($r(179) = 0.924, p < 0.001$), for each condition individually (all
-$r$s $\geq 0.822$, all $p$s $< 0.001$), and across conditions ($r(4) = 0.964, p
-= 0.002$). For late lists
-(Fig.~\ref{fig:clustering-scatterplots}F), the results were more
-variable (overall: $r(179) = 0.348, p < 0.001$; all non-visual conditions: $r$s
-$\geq 0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453, p = 0.011$;
-location: $ r(28) = 0.190, p = 0.314$; across-conditions: $r(4) = -0.036, p =
-0.945$). While less robust than the carryover associations between feature
-clustering and recall performance, we also observed some carryover associations
-between feature clustering and temporal clustering
-(Fig.~\ref{fig:clustering-scatterplots}G, H). Participants who
-showed stronger feature clustering on early lists trended towards showing
-stronger temporal clustering on later lists (overall: $r(179) = 0.301, p <
-0.001$; for individual conditions: all $r$s $\geq 0.297$, all $p$s $\leq
-0.111$; across conditions: $ r(4) = 0.107, p = 0.840$). And participants who
-showed stronger temporal clustering on early lists trended towards showing
-stronger feature clustering on later lists (overall: $r(179) = 0.579, p <
-0.001$; all non-visual conditions: $r$s $\geq 0.323$, all $p$s $\leq 0.082$;
-visual conditions: $r$s $\geq 0.089$, all $p$s $\leq 0.632$; across conditions:
-$r(4) = 0.916, p = 0.010$). Taken together, the results displayed in
+p = 0.623$; see insets of Fig.~\ref{fig:clustering-scatterplots}A and B). We
+observed carryover associations between feature clustering and recall
+performance (Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who
+showed stronger feature clustering on early lists tended to recall more items
+on late lists (across conditions: $r(179) = 0.492, p < 0.001$; all conditions
+individually: $r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who
+recalled more items on early lists also tended to show stronger feature
+clustering on late lists (across conditions: $r(179) = 0.280, p < 0.001$; all
+non-visual conditions: $r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29)
+= 0.298, p = 0.103$; location: $r(28) = 0.354, p = 0.055$). Neither of these
+effects showed condition-level trends (early feature clustering versus late
+recall probability: $r(4) = -0.299, p = 0.565$; early recall probability versus
+late feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for
+associations between feature clustering and temporal clustering. Across every
+order manipulation condition, participants who exhibited stronger feature
+clustering also exhibited stronger temporal clustering. For early lists
+(Fig.~\ref{fig:clustering-scatterplots}E), this trend held overall ($r(179) =
+0.924, p < 0.001$), for each condition individually (all $r$s $\geq 0.822$, all
+$p$s $< 0.001$), and across conditions ($r(4) = 0.964, p = 0.002$). For late
+lists (Fig.~\ref{fig:clustering-scatterplots}F), the results were more variable
+(overall: $r(179) = 0.348, p < 0.001$; all non-visual conditions: $r$s $\geq
+0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453, p = 0.011$; location: $
+r(28) = 0.190, p = 0.314$; across-conditions: $r(4) = -0.036, p = 0.945$).
+While less robust than the carryover associations between feature clustering
+and recall performance, we also observed some carryover associations between
+feature clustering and temporal clustering
+(Fig.~\ref{fig:clustering-scatterplots}G, H). Participants who showed stronger
+feature clustering on early lists trended towards showing stronger temporal
+clustering on later lists (overall: $r(179) = 0.301, p < 0.001$; for individual
+conditions: all $r$s $\geq 0.297$, all $p$s $\leq 0.111$; across conditions: $
+r(4) = 0.107, p = 0.840$). And participants who showed stronger temporal
+clustering on early lists trended towards showing stronger feature clustering
+on later lists (overall: $r(179) = 0.579, p < 0.001$; all non-visual
+conditions: $r$s $\geq 0.323$, all $p$s $\leq 0.082$; visual conditions: $r$s
+$\geq 0.089$, all $p$s $\leq 0.632$; across conditions: $r(4) = 0.916, p =
+0.010$). Taken together, the results displayed in
 Figure~\ref{fig:clustering-scatterplots} show that participants who were more
@@
 of the participant's memory fingerprint (see \textit{Online “fingerprint”
-analysis}). Third, on \textit{destabilize} lists, the presentation is adjusted
-to be \textit{minimally} similar to the current estimate of the participant's
-memory fingerprint (see \textit{Ordering ``stabilize'' and ``destabilize''
-lists by an estimated fingerprint}). The orders in which participants
-experienced each type of list were counterbalanced across participants to help
-reduce the influence of potential list order effects. Because the presentation
-orders on stabilize and destabilize lists are adjusted to best match each
-participant's (potentially unique) memory fingerprint, the adaptive condition
-removes uncertainty about whether participants' assigned conditions might just
-``happen'' to match their preferred ways of organizing their memories.
+analysis}). Third, on \textit{destabilize} lists, the presentation order is
+adjusted to be \textit{minimally} similar to the current estimate of the
+participant's memory fingerprint (see \textit{Ordering ``stabilize'' and
+``destabilize'' lists by an estimated fingerprint}). The orders in which
+participants experienced each type of list were counterbalanced across
+participants to help reduce the influence of potential list-order effects.
+Because the presentation orders on stabilize and destabilize lists are adjusted
+to best match each participant's (potentially unique) memory fingerprint, the
+adaptive condition removes uncertainty about whether participants' assigned
+conditions might just ``happen'' to match their preferred ways of organizing
+their memories.
 
@@
 Participants' fingerprints on stabilize and random lists tended to become
-(numerically) slightly more similar to their average fingerprints computed from the
-previous lists they had experienced, and their fingerprints on destabilize
+(numerically) slightly more similar to their average fingerprints computed from
+the previous lists they had experienced, and their fingerprints on destabilize
 lists tended to become numerically less similar (Fig.~\ref{fig:adaptive}D).
 Overall, we found that participants tended to be better at remembering words on
-stabilize lists relative to words on random ($t(59) = 1.740, p = 0.087$) or
-destabilize ($t(59) = 1.714, p = 0.092$) lists (Fig.~\ref{fig:adaptive}A).
+stabilize lists relative to words on both random ($t(59) = 1.740, p = 0.087$)
+and destabilize ($t(59) = 1.714, p = 0.092$) lists (Fig.~\ref{fig:adaptive}A).
 Participants showed no reliable differences in their memory performance on
@@
 two lexicographic features (word length and first letter), which reflected
-aspects of the words' \textit{letters}. These semantic and lexicographic
-features are intrinsic to each word. We also considered and manipulated two
-additional visual features (color and location) that affected the
-\textit{appearance} of each studied item, but could be varied independently of
-the words' identities. Across different experimental conditions, we manipulated
-how the visual features varied across words (within each list), along with the
-orders of each list's words. Although the participants' task (verbally
-recalling as many words as possible, in any order, within one minute) remained
-constant across all of these conditions, and although the set of words they
-studied on each list remained constant, our manipulations substantially
+characteristics of the words' \textit{letters}. These semantic and
+lexicographic features are intrinsic to each word. We also considered and
+manipulated two additional visual features (color and location) that affected
+the \textit{appearance} of each studied item, but could be varied independently
+of the words' identities. Across different experimental conditions, we
+manipulated how the visual features varied across words (within each list),
+along with the orders of each list's words. Although the participants' task
+(verbally recalling as many words as possible, in any order, within one minute)
+remained constant across all of these conditions, and although the set of words
+they studied from each list remained constant, our manipulations substantially
 affected participants' memories. The impact of some of the manipulations also
@@
 fixed across lists (in the feature rich and reduced conditions), they
-remembered more words on early versus late lists. On feature rich lists, they
-also showed stronger clustering on early versus late lists. However, when we
-\textit{varied} participants' visual experiences across lists (in the ``reduced
-(early)'' and ``reduced (late)'' conditions), these early versus late accuracy
-and clustering differences disappeared. Abruptly changing how irrelevant visual
-features varied across words seemed to act as a sort of ``event boundary'' that
-partially reset how participants processed and remembered post-boundary lists.
-Within-list clustering also increased in these manipulations, suggesting that
-the "within-event" words were being more tightly associated with each other.
-
-When we held the visual features constant on early lists, but then varied
-words' appearances on later lists (i.e., the reduced (early) condition), this improved
-participants' overall memory performance. However, this impact was directional:
-when we \textit{removed} visual features on late lists that had been present on
-early lists (i.e., the reduced (late) condition), we saw no memory improvement.
+remembered more words from early lists than from late lists. For feature rich
+lists, they also showed stronger clustering for early versus late lists.
+However, when we \textit{varied} participants' visual experiences across lists
+(in the ``reduced (early)'' and ``reduced (late)'' conditions), these early
+versus late accuracy and clustering differences disappeared. Abruptly changing
+how irrelevant visual features varied across words seemed to act as a sort of
+``event boundary'' that partially reset how participants processed and
+remembered post-boundary lists. Within-list clustering also increased in these
+manipulations, suggesting that the "within-event" words were being more tightly
+associated with each other.
+
+When we held the visual features constant during early lists, but then varied
+words' appearances in later lists (i.e., the reduced (early) condition),
+participants' overall memory performance improved. However, this impact was
+directional: when we \textit{removed} visual features from words in late lists
+that had been present in early lists (i.e., the reduced (late) condition), we
+saw no memory improvement.
 
@@
 late lists. At the group level, we found little evidence for lingering
-``carryover'' effects of these manipulations; participants in the order
+``carryover'' effects of these manipulations: participants in the order
 manipulation conditions showed similar memory performance and clustering on
@@
 
-Participants who showed stronger feature clustering on early (order
-manipulated) lists tended to better remember late (randomly ordered) lists.
-Participants who remembered early lists better also tended to show stronger
-feature clustering (along their condition's feature dimension) on late lists
-(even though the words on those late lists were presented in a random order).
-We also observed some (weaker) carryover effects of temporal clustering.
-Participants who showed stronger feature clustering (along their condition's
-feature dimension) on early lists tended to show stronger temporal clustering
-on late lists. And participants who showed stronger temporal clustering on
-early lists also tended to show stronger feature clustering on late lists.
-Essentially, these order manipulations appeared to affect each participant
-differently. Some participants were sensitive to our manipulations, and those
-participants showed stronger impacts on their memory performance for the ordered
-lists as well as future (random) lists. Other participants appeared relatively
-insensitive to our manipulations, and those participants showed little
-carryover effects on late lists.
+Participants who showed stronger feature clustering on early
+(order-manipulated) lists tended to better remember late (randomly ordered)
+lists. Participants who remembered early lists better also tended to show
+stronger feature clustering (along their condition's feature dimension) on late
+lists (even though the words on those late lists were presented in a random
+order). We also observed some (weaker) carryover effects of temporal
+clustering. Participants who showed stronger feature clustering (along their
+condition's feature dimension) on early lists tended to show stronger temporal
+clustering on late lists. And participants who showed stronger temporal
+clustering on early lists also tended to show stronger feature clustering on
+late lists. Essentially, these order manipulations appeared to affect each
+participant differently. Some participants were sensitive to our manipulations,
+and those participants' memory performance was impacted more strongly, both for
+the ordered lists and for future (random) lists. Other participants appeared
+relatively insensitive to our manipulations, and those participants showed
+little carryover effects on late lists.
 
@@
 along \textit{particular} (e.g., preferred) feature dimensions. To help
-distinguish between these possibilities, we designed an adaptive manipulation
+distinguish between these possibilities, we designed an adaptive condition
 whereby we attempted to manipulate whether participants studied words in an
-order that matched (or mismatched) our estimate of how they would cluster or
-organize the studied words in memory (i.e., their idiosyncratic memory
+order that either matched or mismatched our estimate of how they would cluster
+or organize the studied words in memory (i.e., their idiosyncratic memory
 fingerprint). We found that when we presented words in orders that were
@@
 bring into the experiment their own idiosyncratic preferences and strategies
-for organizing the words in their memories. When we presented the words in an
+for organizing the words in their memory. When we presented the words in an
 order consistent with each participant's idiosyncratic fingerprint, their
@@
 sounds~\citep{BeamJone10, SahaSmit14}, secondary tasks~\citep{MasiSaha14,
-PolyEtal09}, and more can all impact how participants remember, and organize in
-memory, lists of studied items.
+PolyEtal09, OberLewa08}, and more can all impact how participants remember, and
+organize in memory, lists of studied items.
 
@@
 that it is not only the contextual features themselves that affect memory, but
-also the \textit{dynamics} of context-- i.e., how the contextual features
+also the \textit{dynamics} of context---i.e., how the contextual features
 associated with each item change over time.
@@
 experiences, expectations, and other real, perceived, or inferred cues to help
-resolve the ambiguities. We may also be overtly or covertly ``primed'' to
+resolve these ambiguities. We may also be overtly or covertly ``primed'' to
 influence how we are likely to resolve ambiguities. For example, before
@@
 
-In more simplified scenarios, like list learning paradigms, the stimuli and
+In more simplified scenarios, like list-learning paradigms, the stimuli and
 tasks participants encounter before studying a given list can influence what
@@
 our memory for pre-boundary information and enhance our memory for
-post-boundary information~\citep{RadvCope06, SahaKell02, MannEtal16}. Event
-boundaries are also tightly associated with the notion of \textit{situation
-models} and \textit{schemas}-- mental frameworks for organizing our
-understanding about the rules of how we and others are likely to behave, how
-events are likely to unfold over time, how different elements are likely to
-interact, and so on. For example, a situation model pertaining to a particular
-restaurant might set our expectations about what we are likely to experience
-when we visit that restaurant (e.g., what the building will look like, how it
-will smell when we enter, how crowded the restaurant is likely to be, the
-sounds we are likely to hear, etc.). Similarly, as mentioned in the
+post-boundary information~\citep{RadvCope06, SahaKell02, MannEtal16,
+DuBrDava13}. Event boundaries are also tightly associated with the notion of
+\textit{situation models} and \textit{schemas}---mental frameworks for
+organizing our understanding about the rules of how we and others are likely to
+behave, how events are likely to unfold over time, how different elements are
+likely to interact, and so on. For example, a situation model pertaining to a
+particular restaurant might set our expectations about what we are likely to
+experience when we visit that restaurant (e.g., what the building will look
+like, how it will smell when we enter, how crowded the restaurant is likely to
+be, the sounds we are likely to hear, etc.). Similarly, as mentioned in the
 \textit{Introduction}, we might learn a schema describing how events are likely
-to unfold \textit{across} any sit-down restaurant-- e.g., open the door, wait
+to unfold \textit{across} any sit-down restaurant---e.g., open the door, wait
 to be seated, receive a menu, decide what to order, place the order, and so on.
@@
 are presented, can lead participants to behave similarly to what one might
-expect upon crossing an event boundary. Adding in variability in font color and
-presentation locations for words on late lists, after those visual features had
+expect upon crossing an event boundary. Adding variability in font color and
+presentation location for words on late lists, after those visual features had
 been held constant on early lists, led participants to remember more words on
@@
 list. According to contextual change accounts of memory across event
-boundaries~\citep[e.g.,][]{SahaKell02}, this could help to explain why
-participants in the reduced (early) and reduced (late) conditions exhibited
-better overall memory performance. Specifically, their memory for late list
-items could benefit from less interference from early list items, and the
-contextual features associated with late list items (after the ``event
-boundary'') might serve as more specific recall cues for those late items
-(relative to if the boundary had not occurred).
+boundaries~\citep[e.g.,][]{SahaKell02, PettEtal16, GoldEtal17, FlorEtal17},
+this could help to explain why participants in the reduced (early) and reduced
+(late) conditions exhibited better overall memory performance. Specifically,
+their memory for late list items could benefit from less interference from
+early list items, and the contextual features associated with late list items
+(after the ``event boundary'') might serve as more specific recall cues for
+those late items (relative to if the boundary had not occurred).
 
@@
 Although most modern formal theories of episodic memory have been developed and
-tested to explain memory for list learning tasks~\citep{Kaha20}, a number of
+tested to explain memory for list-learning tasks~\citep{Kaha20}, a number of
 recent studies suggest some substantial differences between memory for lists
@@
 naturalistic stimuli are often much more engaging than the highly simplified
-list learning tasks typically employed in the psychological laboratory, perhaps
+list-learning tasks typically employed in the psychological laboratory, perhaps
 leading participants to pay more attention, exert more effort, and stay more
@@
 The experiment we report in this paper was designed to help bridge some of this
-gap between naturalistic tasks and more traditional list learning tasks. We had
+gap between naturalistic tasks and more traditional list-learning tasks. We had
 people study word lists similar to those used in classic memory studies, but we
@@
 the manipulations led to changes that were common across people (e.g., more
-temporal clustering when words' appearances were varied; enhanced memory for
-lists following an ``event boundary;'' more feature clustering on
-order-manipulated lists; etc.). Other manipulations led to changes that were
+temporal clustering when words' appearances were varied, enhanced memory for
+lists following an ``event boundary,'' more feature clustering on
+order-manipulated lists, etc.). Other manipulations led to changes that were
 idiosyncratic (especially carryover effects from order manipulations; e.g.,
@@
 to show stronger feature clustering for their condition's feature dimension on
-late randomly ordered lists; etc.). We also found that participants remembered
+late randomly ordered lists, etc.). We also found that participants remembered
 more words from lists that were sorted to align with their idiosyncratic
@@
 which they recalled the words, and the effects of early list manipulations on
-later lists, varied according to how we presented the to-be-remembered words.
+later lists all varied according to how we presented the to-be-remembered
+words.
 
@@
 these manipulations be extended? In other words, might there be more
-sophisticated or clever feature or order manipulations that one might implement
+sophisticated or clever feature or order manipulations that one could implement
 to have stronger impacts on memory? Are there limits to how much impact (on
@@
 classroom content, perhaps they could be utilized manually by teachers, or even
-by automated content presentation systems, to optimize how and what students
+by automated content-presentation systems, to optimize how and what students
 remember.
@@
 content? We know that our expectations, strategies, situation models learned
-through prior experiences, and more, collectively shape how our experiences are
+through prior experiences, and more collectively shape how our experiences are
 remembered. But those aspects of our memory are not fixed: when we are exposed

### 2023-01-24 paper/main.tex
quotation fix (thanks PCF!)

@@
 remembered post-boundary lists. Within-list clustering also increased in these
-manipulations, suggesting that the "within-event" words were being more tightly
+manipulations, suggesting that the ``within-event'' words were being more tightly
 associated with each other.

### 2023-01-24 paper/main.tex

@@
 yet-to-be-experienced future~\citep{XuEtal23, SchaTurk15, RangRitc12,
-BoweEtal79, MomeEtal17}. On the other hand, the rich structure of real-world
+BoweEtal79, MomeEtal17}. On the other hand, the rich structures of real-world
 experiences and other naturalistic stimuli that enable people to form deep and
-meaningful situation models and schemas have no obvious analog in simple word
+meaningful situation models and schemas have no obvious analogs in simple word
 lists. Often, lists in free recall studies are explicitly \textit{designed} to
@@
 items (words) were designed to vary along three general dimensions: semantic
-(word \textit{category}, and physical \textit{size} of the referent),
+(word \textit{category} and physical \textit{size} of the referent),
 lexicographic (word \textit{length} and \textit{first letter}), and visual
@@
 visual features. In two conditions, we manipulated whether the words'
-appearances were fixed or variable within each list. In six manipulation
-conditions, we asked participants to first study and recall eight lists whose
-items were sorted by a target feature (e.g., word category), and then study and
-recall an additional eight lists whose items had the same features, but that
-were sorted in a random temporal order. We were interested in how these
-manipulations affected participants' recall behaviors on early (manipulated)
-lists, as well as how order manipulations on early lists affected recall
-behaviors on later (randomly ordered) lists. Finally, in an \textit{adaptive}
-experimental condition we used participants' recall behaviors on early lists to
-manipulate, in real-time, the presentation orders of subsequent lists. In this
-adaptive condition, we varied the agreement between how participants preferred
-to organize their memories of the studied items versus the orders in which the
+appearances were fixed or variable within each list. In six conditions, we
+asked participants to first study and recall eight lists whose items were
+sorted by a target feature (e.g., word category), and then study and recall an
+additional eight lists whose items had the same features, but that were sorted
+in a random temporal order. We were interested in how these manipulations
+affected participants' recall behaviors on early (manipulated) lists, as well
+as how order manipulations on early lists affected recall behaviors on later
+(randomly ordered) lists. Finally, in an \textit{adaptive} experimental
+condition we used participants' recall behaviors on early lists to manipulate,
+in real-time, the presentation orders of subsequent lists. In this adaptive
+condition, we varied the agreement between how participants preferred to
+organize their memories of the studied items versus the orders in which the
 items were presented.
@@
 We enrolled a total of 491 members of the Dartmouth College community across 11
-experimental conditions. The conditions included two controls (feature rich,
+experimental conditions. The conditions included two controls (feature rich and
 reduced), two visual manipulation conditions [reduced (early) and reduced
@@
 experimental condition as follows: feature rich (67 participants), reduced (61
-participants), reduced (early), (42 participants), reduced (late) (41
+participants), reduced (early) (42 participants), reduced (late) (41
 participants), category (30 participants), size (30 participants), length (30
@@
 studied words along each dimension. To obtain stable estimates of these
-fingerprints for each participant, we averaged clustering scores across lists.
-We also tracked and characterized how participants' fingerprints changed across
-lists (e.g.,
+fingerprints for each participant, we averaged their clustering scores across
+lists. We also tracked and characterized how participants' fingerprints changed
+across lists (e.g.,
 Figs.~\ref{fig:fingerprint-trajectories},~\fingerprintTrajectoryRandom).
@@
 package~\citep{HeusEtal17} to apply speech-to-text algorithms to the
-just-collected data, aggregate the data for the given participant, and estimate
+just-collected audio data, aggregate the data for the given participant, and estimate
 the participant's memory fingerprint using all of their available data up to
@@
 it were a potential ``perfect'' recall sequence. (We did not include temporal
-clustering scores in these fingerprints.) This yielded a ``simulated
+clustering scores in these fingerprints, since the temporal clustering score
+for every permutation is always equal to 1.) This yielded a ``simulated
 fingerprint'' vector, $\hat{f}_p$ for each permutation $p$. We used these
@@
 \textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
-(early)} condition, we held the irrelevant visual features constant on early
-lists, but allowed them to vary randomly on late lists. In a \textit{reduced
-(late)} condition, we allowed the irrelevant visual features to vary randomly
-on early lists, but held them constant on late lists. Given our above findings
-that (a) participants tended to remember more words and exhibit stronger
-clustering effects on feature rich (versus reduced) lists, and (b) participants
-tended to remember more words and exhibit stronger clustering effects on early
-(versus late) lists, we expected these early versus late differences to be
-enhanced in the reduced (early) condition and diminished in the reduced (late)
-condition. However, to our surprise, participants in \textit{neither} condition
-exhibited reliable early versus late differences in accuracy (reduced (early):
-$t(41) = 1.499, p = 0.141$; reduced (late): $t(40) = 1.462, p = 0.152$),
-temporal clustering (reduced (early): $t(41) = 0.998, p = 0.324$; reduced
-(late): $t(40) = 1.099, p = 0.278$), nor feature-based clustering (reduced
-(early), category: $t(41) = 0.753, p = 0.456$; reduced (early), size: $t(41) =
-0.721, p = 0.475$; reduced (early), length: $t(41) = 0.493, p = 0.625$; reduced
-(early), first letter: $t(41) = 0.780, p = 0.440$; reduced (late), category:
-$t(40) = -0.086, p = 0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$;
-reduced (late), length: $t(40) = 1.476, p = 0.148$; reduced (late), first
-letter: $t(40) = 0.966, p = 0.340$). We hypothesized that adding or removing
-the irrelevant features was acting as a sort of ``event boundary'' between
-early and late lists. In prior work, we (and others) have found that memories
-formed just after event boundaries can be enhanced~\citep[e.g., due to less
-contextual interference between pre- and post-boundary items;][]{MannEtal16,
-PettEtal16, GoldEtal17, FlorEtal17}.
-
-We found that \textit{adding} irrelevant visual features on later lists that
+(early)} condition, we held the visual features constant on early lists, but
+allowed them to vary randomly on late lists. In a \textit{reduced (late)}
+condition, we allowed the visual features to vary randomly on early lists, but
+held them constant on late lists. Given our above findings that (a)
+participants tended to remember more words and exhibit stronger clustering
+effects on feature rich (versus reduced) lists, and (b) participants tended to
+remember more words and exhibit stronger clustering effects on early (versus
+late) lists, we expected these early versus late differences to be enhanced in
+the reduced (early) condition and diminished in the reduced (late) condition.
+However, to our surprise, participants in \textit{neither} condition exhibited
+reliable early versus late differences in accuracy (reduced (early): $t(41) =
+1.499, p = 0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal
+clustering (reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40)
+= 1.099, p = 0.278$), nor feature-based clustering (reduced (early), category:
+$t(41) = 0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$;
+reduced (early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first
+letter: $t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086,
+p = 0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
+length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
+0.966, p = 0.340$). We hypothesized that adding or removing the variability in
+the visual features was acting as a sort of ``event boundary'' between early
+and late lists. In prior work, we (and others) have found that memories formed
+just after event boundaries can be enhanced~\citep[e.g., due to less contextual
+interference between pre- and post-boundary items;][]{MannEtal16, PettEtal16,
+GoldEtal17, FlorEtal17}.
+
+We found that \textit{adding} incidental visual features on later lists that
 had not been present on early lists (as in the reduced (early) condition)
@@
 \textit{not} been present earlier. In contrast, we found that participants
-exhibited more temporal and feature-based clustering when we added irrelevant
+exhibited more temporal and feature-based clustering when we added incidental
 visual features to \textit{any} lists (comparisons of clustering on feature
@@
 memories. Next, we turn to examine the memory effects of varying the temporal
-ordering of different stimulus features while holding the features themselves
-constant. We hypothesized that changing the orders in which participants were
-exposed to the words on a given list might enhance (or diminish) the relative
-influence of different features. For example, presenting a set of words
-alphabetically might enhance participants' attention to the studied items'
-first letters, whereas sorting the same list of words by semantic category
-might instead enhance participants' attention to the words' semantic
-attributes. Importantly, we expected these order manipulations to hold even
-when the variation in the total set of features (across words) was held
-constant across lists (e.g., unlike in the reduced (early) and reduced (late)
-conditions, where variations in visual features were added or removed from a
-subset of the lists participants studied).
+ordering of different stimulus features. We hypothesized that changing the
+orders in which participants were exposed to the words on a given list might
+enhance (or diminish) the relative influence of different features. For
+example, presenting a set of words alphabetically might enhance participants'
+attention to the studied items' first letters, whereas sorting the same list of
+words by semantic category might instead enhance participants' attention to the
+words' semantic attributes. Importantly, we expected these order manipulations
+to hold even when the variation in the total set of features (across words) was
+held constant across lists (e.g., unlike in the reduced (early) and reduced
+(late) conditions, where variations in visual features were added or removed
+from a subset of the lists participants studied).
 
@@
 ``event boundaries'' (to use the term loosely) can be induced across lists by
-adding or removing irrelevant visual features. At shorter timescales, ``event
+adding or removing incidental visual features. At shorter timescales, ``event
 boundaries'' can be induced across items (within a single list) by adjusting
@@
 lexicographic clustering on lexicographically sorted lists than semantically
-(length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) = 3.686, p =
-0.000$) or visually (length: $t(119) = 3.024, p = 0.003$; first letter: $t(119)
+(length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) = 3.686, p <
+0.001$) or visually (length: $t(119) = 3.024, p = 0.003$; first letter: $t(119)
 = 2.644, p = 0.009$) sorted lists; and more visual clustering on visually
@@
 participant's memory fingerprint describes how they tend to retrieve memories
-of the studied items, perhaps searching through several feature spaces (or
-along several representational dimensions). To gain insights into the dynamics
-of how participants' clustering scores tended to change over time, we computed
-the average (across participants) fingerprint from each list, from each order
-manipulation condition (Fig.~\ref{fig:fingerprint-trajectories}). We projected
-these fingerprints into a two-dimensional space to help visualize the dynamics
-(top panels; see \textit{Computing low-dimensional embeddings of memory
-fingerprints}). We found that participants' average fingerprints tended to
-remain relatively stable on early lists, and exhibited a ``jump'' to another
+of the studied items, perhaps searching in parallel through several feature
+spaces (or along several representational dimensions). To gain insights into
+the dynamics of how participants' clustering scores tended to change over time,
+we computed the average (across participants) fingerprint from each list, from
+each order manipulation condition (Fig.~\ref{fig:fingerprint-trajectories}). We
+projected these fingerprints into a two-dimensional space to help visualize the
+dynamics (top panels; see \textit{Computing low-dimensional embeddings of
+memory fingerprints}). We found that participants' average fingerprints tended
+to remain relatively stable on early lists, and exhibited a ``jump'' to another
 stable state on later lists. The sizes of these jumps varied somewhat across
@@
 versus late accuracy and clustering differences disappeared. Abruptly changing
-how irrelevant visual features varied across words seemed to act as a sort of
+how incidental visual features varied across words seemed to act as a sort of
 ``event boundary'' that partially reset how participants processed and
@@
 boundaries~\citep[e.g.,][]{SahaKell02, PettEtal16, GoldEtal17, FlorEtal17},
-this could help to explain why participants in the reduced (early) and reduced
-(late) conditions exhibited better overall memory performance. Specifically,
-their memory for late list items could benefit from less interference from
-early list items, and the contextual features associated with late list items
-(after the ``event boundary'') might serve as more specific recall cues for
-those late items (relative to if the boundary had not occurred).
+this could help to explain why participants in the reduced (early) condition
+exhibited better overall memory performance. Specifically, their memory for
+late list items could benefit from less interference from early list items, and
+the contextual features associated with late list items (after the ``event
+boundary'') might serve as more specific recall cues for those late items
+(relative to if the boundary had not occurred).
 

### 2023-01-24 paper/supplement.tex

@@
 
-    \caption{\textbf{Recall accuracy by study list number.}  Each panel displays the average recall accuracy (across participants) as a function of the number of studied lists for the
-    random conditions (\textbf{A.}), order manipulation conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}).  The conditions are denoted by color.  Note that the first four
-    lists in all of the ``adaptive'' conditions were ordered randomly to compute a baseline fingerprint for each participant prior to initiating the adaptive ordering procedure.  All error
-    ribbons denote bootstrap-estimated 95\% confidence intervals, calculated across participants.}
+\caption{\textbf{Recall accuracy by study list number.} Each panel displays the
+average recall accuracy (across participants) as a function of the number of
+studied lists for the random conditions (\textbf{A.}), order manipulation
+conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}). The conditions
+are denoted by color. Note that the first four lists in all of the ``adaptive''
+conditions were ordered randomly to compute a baseline fingerprint for each
+participant prior to initiating the adaptive ordering procedure. \textbf{All
+panels.} All error ribbons denote bootstrap-estimated 95\% confidence
+intervals, calculated across participants.} 
 \label{fig:accuracy-by-list}
+
 \end{figure}
@@
     The across-participant distributions of clustering scores for each feature
-    type ($x$-coordinate) are displayed for each experimental condition
+    type ($x$-axis) are displayed for each experimental condition
     (color), separately for early (top) and late (bottom) lists.}
@@
     The across-participant distributions of clustering scores for each feature
-    type ($x$-coordinate) are displayed for each experimental condition
+    type ($x$-axis) are displayed for each experimental condition
     (color).}
@@
 
-    \caption{\textbf{Probability of $n$\textsuperscript{th} recall matrices.}  Each sub-panel displays
-    the average probability of recalling the given word (Presentation position, $x$-coordinate)
-    at the given output position ($y$-coordinate); color denotes the probability.  \textbf{A. Random conditions.}  The top rows
-    display data from early (order manipulation) lists, the middle rows display data from late (randomly ordered)
-    lists, and the bottom rows display the differences between the matrices in the top and middle rows.  Panel columns denote experimental conditions.
-    \textbf{B. Order manipulation conditions.}  The matrices are displayed in the same format as those in Panel A.
-    Panel columns are organized by feature type (semantic, lexicographic, or visual).  \textbf{C. Adaptive conditions.}
-    The sub-panels are displayed in the same formats as Panels A and B, but here the matrices
-    and contrasts (indicated in $y$-axis labels) reflect different list manipulation conditions.}
+    \caption{\textbf{Probability of $n$\textsuperscript{th} recall matrices.}
+    Each sub-panel displays the average probability of recalling the given word
+    (Presentation position, matrix column) at the given output position (matrix
+    row); color denotes the probability. \textbf{A. Random conditions.} The top
+    rows of matrices display data from early (order manipulation) lists, the
+    middle rows of matrices display data from late (randomly ordered) lists,
+    and the bottom rows of matrices display the differences between the
+    matrices in the top and middle rows. Panel columns denote experimental
+    conditions. \textbf{B. Order manipulation conditions.} The matrices are
+    displayed in the same format as those in Panel A. Panel columns are
+    organized by feature type (semantic, lexicographic, or visual). \textbf{C.
+    Adaptive conditions.} The sub-panels are displayed in the same formats as
+    Panels A and B, but here the matrices and contrasts (indicated in $y$-axis
+    labels) reflect different list manipulation conditions.}
 

### 2023-01-30 paper/supplement.tex
re-organized figures notebook, updated readme file

@@
 
-    \caption{\textbf{Recall dynamics in feature rich free recall (adaptive conditions).} \textbf{Left panels.} The probabilities of
+    \caption{\textbf{Recall dynamics in feature rich free recall (adaptive conditions).} \textbf{Left panel.} The probabilities of
     initiating recall with each word are plotted as a function of presentation
-    position. \textbf{Middle panels.} The conditional probabilities of recalling
+    position. \textbf{Middle panel.} The conditional probabilities of recalling
     each word are plotted as a function of the relative position (Lag) to the words
-    recalled just-prior. \textbf{Right panels.} The overall probabilities of
+    recalled just-prior. \textbf{Right panel.} The overall probabilities of
     recalling each word are plotted as a function of presentation position.

### 2023-02-03 paper/admin/cover_letter.tex
added draft of cover letter

\title{cover letter}
%
% See http://texblog.org/2013/11/11/latexs-alternative-letter-class-newlfm/
% and http://www.ctan.org/tex-archive/macros/latex/contrib/newlfm
% for more information.
%
\documentclass[11pt,stdletter,orderfromtodate,sigleft]{newlfm}
\usepackage{blindtext, xfrac, animate, hyperref, pxfonts, geometry}

  \setlength{\voffset}{0in}

\newlfmP{dateskipbefore=0pt}
\newlfmP{sigsize=20pt}
\newlfmP{sigskipbefore=10pt}

\newlfmP{Headlinewd=0pt,Footlinewd=0pt}

\newcommand{\journal}{Psychological Review}
\newcommand{\articletype}{Article}
\newcommand{\myTitle}{Feature and order manipulations in a free recall task affect memory for current and future lists}
\newcommand{\corresponding}{Jeremy R. Manning}

\namefrom{\vspace{-0.3in}\corresponding}
\addrfrom{
  Dartmouth College\\
  Department of Psychological and Brain Sciences\\
  HB 6207 Moore Hall\\
  Hanover, NH, 03755}
\addrto{}
\dateset{\today}

\greetto{To the editors of \textit{\journal}:}

\closeline{Sincerely,}
%\usepackage{setspace}
%\linespread{0.85}
% How will your work make others in the field think differently and move the field forward?
% How does your work relate to the current literature on the topic?
% Who do you consider to be the most relevant audience for this work?
% Have you made clear in the letter what the work has and has not achieved?

\begin{document}
\begin{newlfm}

  We have enclosed our manuscript entitled \textit{\myTitle} to be considered
  for publication as an \textit{\articletype}. The experiment reported on in
  our manuscript shows how manipulating the appearances and presentation orders
  of items in a free recall task affects .

  All of the 452 participants in our study encountered the same 16 lists of 16
  words each. The items on each list varied along two semantic dimensions (word
  category, size of referent), two lexicographic dimensions (word length, first
  letter), and two visual dimensions (text color and onscreen presentation
  location). Across different conditions in the experiment, we manipulated
  whether the visual features varied across items (i.e., each word is presented
  in a different font color and at a different location) or whether the visual
  features were held constant across items (i.e., each word is presented in
  black at the center of the screen). We also manipulated the presentation
  orders such that, across different conditions, participants studied some
  lists in a randomized order, and other lists in an order that sorted items
  along one or more feature dimensions. These manipulations changed
  systematically across lists, enabling us to study the consequences of these
  manipulations on the manuplated lists as well as on later lists (subjected to
  potentially different manipulations).

  We observed several particularly exciting findings. First, even though the
  visual features were incidental to the participants' task (i.e., remembering
  as many words as possible), and even though every participant saw exactly the
  same words and lists overall, some of the visual manipulations let
  participants to remember more words. Second, participants showed substantial
  variability in how ``susceptible'' they were to the order manipulations.
  Participants whose behaviors were affected most by the order manipulations
  also showed lingering impacts on later lists that were \textit{not} subjected
  to the same order manipulations. Third, we used a real-time behavioral
  manipulation to change how lists were presented on future lists, according to
  how participants recalled earlier lists. We show that when we organize
  to-be-memorized lists in ways that are compatable with how participants
  ``naturally'' recalled earlier lists, the participants remember more words
  overall. Taken together, our work shows that the \textit{way}
  to-be-remembered content is presented can affect what is remembered, and how.

  We draw several important links between our work and other recent findings
  related to naturalistic memory; context effects on memory performance and
  organization; priming effects on memory; and expectation, event boundaries,
  and situation models. Essentially, we see our manipulations (and findings) as
  helping to bridge key gaps between a large literature describing memory in
  ``traditional'' list-learning tasks, and a growing literature describing
  memory in more ``naturalistic'' settings such as real-world experiences,
  written narratives, movies, and so on. For example, we designed several of
  our experimental manipulations to mimic several aspects of the temporal and
  conceptual structure that are fundamental to naturalistic scenarios. However,
  whereas it can be very difficult to explicitly model or formalize the
  temporal progression of naturalistic experiences, our paradigm is more
  amenable to modeling and incorporating stimulus manipulations.

  We expect that this article will be of interest to readership interested in
  learning and memory, real-time experimental design, priming effects on
  memory, and more.

Thank you for considering our manuscript, and I hope you will find it suitable
for publication in \textit{\journal}.

\end{newlfm}
\end{document}

### 2023-02-06 paper/admin/cover_letter.tex
updated cover letter

@@
   We have enclosed our manuscript entitled \textit{\myTitle} to be considered
-  for publication as an \textit{\articletype}. The experiment reported on in
-  our manuscript shows how manipulating the appearances and presentation orders
-  of items in a free recall task affects .
+  for publication as an \textit{\articletype}. The main question we explore in
+  our manuscript is: can changing the way we present the same lists of words
+  affect how people remember those lists and/or future lists?
 
@@
   category, size of referent), two lexicographic dimensions (word length, first
-  letter), and two visual dimensions (text color and onscreen presentation
+  letter), and two visual dimensions (text color, onscreen presentation
   location). Across different conditions in the experiment, we manipulated
@@
   orders such that, across different conditions, participants studied some
-  lists in a randomized order, and other lists in an order that sorted items
+  lists in a randomized order and other lists in an order that sorted items
   along one or more feature dimensions. These manipulations changed
-  systematically across lists, enabling us to study the consequences of these
-  manipulations on the manuplated lists as well as on later lists (subjected to
-  potentially different manipulations).
+  systematically across lists, enabling us to study the memory consequences of
+  these manipulations on the manuplated lists as well as on later lists
+  (subjected to potentially different manipulations).
 
-  We observed several particularly exciting findings. First, even though the
-  visual features were incidental to the participants' task (i.e., remembering
-  as many words as possible), and even though every participant saw exactly the
-  same words and lists overall, some of the visual manipulations let
-  participants to remember more words. Second, participants showed substantial
-  variability in how ``susceptible'' they were to the order manipulations.
-  Participants whose behaviors were affected most by the order manipulations
-  also showed lingering impacts on later lists that were \textit{not} subjected
-  to the same order manipulations. Third, we used a real-time behavioral
-  manipulation to change how lists were presented on future lists, according to
-  how participants recalled earlier lists. We show that when we organize
-  to-be-memorized lists in ways that are compatable with how participants
-  ``naturally'' recalled earlier lists, the participants remember more words
-  overall. Taken together, our work shows that the \textit{way}
-  to-be-remembered content is presented can affect what is remembered, and how.
+  We observed several particularly exciting demonstrations that how and what
+  people remember can be influenced by the way they experience the
+  to-be-remembered content and on their prior experiences. First, even though
+  the visual features were incidental to the participants' task (i.e.,
+  remembering as many words as possible), and even though every participant saw
+  exactly the same words and lists overall, some of the visual manipulations
+  led participants to remember more words. Second, participants showed
+  substantial variability in how ``susceptible'' they were to the order
+  manipulations. Participants whose behaviors were affected most by the order
+  manipulations also showed lingering impacts on later lists that were
+  \textit{not} subjected to the same order manipulations. Third, we used a
+  real-time behavioral manipulation to change how new lists were presented,
+  according to how individual participants had recalled earlier lists. We show
+  that when we organize to-be-memorized lists in ways that are compatable with
+  how participants ``naturally'' recalled earlier lists, the participants
+  remember more words overall.
 
-  We draw several important links between our work and other recent findings
-  related to naturalistic memory; context effects on memory performance and
-  organization; priming effects on memory; and expectation, event boundaries,
-  and situation models. Essentially, we see our manipulations (and findings) as
-  helping to bridge key gaps between a large literature describing memory in
-  ``traditional'' list-learning tasks, and a growing literature describing
-  memory in more ``naturalistic'' settings such as real-world experiences,
-  written narratives, movies, and so on. For example, we designed several of
-  our experimental manipulations to mimic several aspects of the temporal and
-  conceptual structure that are fundamental to naturalistic scenarios. However,
-  whereas it can be very difficult to explicitly model or formalize the
-  temporal progression of naturalistic experiences, our paradigm is more
-  amenable to modeling and incorporating stimulus manipulations.
+  Our study touches on deep questions about how we learn and remember. At the
+  core, our findings show how the same to-be-remembered content can be encoded,
+  organized, and retrieved differently depending on how it was experienced and
+  what other experiences happenend nearby in time. We also show that different
+  people have different ``preferences'' regarding how they will encode and
+  organize the words they study. Conforming to or violating those preferences
+  enabled us to manipulate memory performance. If memory performance depends in
+  part on how the to-be-remembered content is presented to us, what are the
+  limits of our memory systems? Might \textit{other} manipulations in some
+  future study or application show even stronger impacts on memory? How far can
+  the capacities of our memory systems be pushed?  Are there implications
+  of these effects on other domains, such as personalized instruction or
+  automated tutoring systems?
+
+  We also draw several important links between our work and other recent
+  findings related to naturalistic memory; context effects on memory
+  performance and organization; priming effects on memory; and expectation,
+  event boundaries, and situation models. Essentially, we see our manipulations
+  (and findings) as helping to bridge key gaps between a large literature
+  describing memory in ``traditional'' list-learning tasks, and a growing
+  literature describing memory in more ``naturalistic'' settings such as
+  real-world experiences, written narratives, movies, classrooms, and so on.
+  For example, we designed several of our experimental manipulations to mimic
+  several aspects of the temporal and conceptual structure that are fundamental
+  to naturalistic scenarios. However, whereas it can be very difficult to
+  explicitly model or formalize the temporal progression of naturalistic
+  experiences, our paradigm is more amenable to modeling and incorporating
+  stimulus manipulations.
 

### 2023-07-04 paper/main.tex
re-ran all notebooks, made some minor format changes per editor request
starting to address reviewer comments

@@
 unrelated one. In general, people organize memories for words along a wide
-variety of stimulus dimensions. As formalized by models like the
+variety of stimulus dimensions. According to models like the
 \textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the
@@
 between two words as the Euclidean distances between their $(r, g, b)$ color or
-$(x, y)$ location vectors, respectively. Therefore, the color and location
-distance measures always take on non-negative real values (upper-bounded at
-441.67 for color, or 27~in for location, reflecting the distances between the
-corresponding maximally different vectors).
+$(x, y)$ location vectors (specified in inches), respectively. Therefore, the
+color and location distance measures always take on non-negative real values
+(upper-bounded at 441.67 for color, or 27~in for location, reflecting the
+distances between the corresponding maximally different vectors).
 
@@
 transitions. In carrying out this analysis, we excluded all incorrect recalls
-and successive repetitions (i.e., recalling the same word twice in a row). This
-yielded, for each list, a 1 by number-of-lags ($-15$ to $+15$; 30 lags in
-total, excluding lags of 0) array of conditional probabilities. We averaged
-these probabilities across lists to obtain a single lag-CRP for each
-participant. Because transitions at large absolute lags are rare, these curves
-are typically displayed using range restrictions~\citep{Kaha12}.
+and repetitions (i.e., recalling a word that had already appeared previously in
+the current recall sequence). This yielded, for each list, a 1 by
+number-of-lags ($-15$ to $+15$; 30 lags in total, excluding lags of 0) array of
+conditional probabilities. We averaged these probabilities across lists to
+obtain a single lag-CRP for each participant. Because transitions at large
+absolute lags are rare, these curves are typically displayed using range
+restrictions~\citep{Kaha12}.
 
@@
 feature dimension of interest).
+
+\subsection*{Data and code availability}
+
+All of the data analyzed in this manuscript, along with all of the code for
+carrying out the analyses may be found at https://github.com/ContextLab/FRFR-analyses.
 
@@
 held them constant on late lists. Given our above findings that (a)
-participants tended to remember more words and exhibit stronger clustering
-effects on feature rich (versus reduced) lists, and (b) participants tended to
-remember more words and exhibit stronger clustering effects on early (versus
-late) lists, we expected these early versus late differences to be enhanced in
-the reduced (early) condition and diminished in the reduced (late) condition.
-However, to our surprise, participants in \textit{neither} condition exhibited
-reliable early versus late differences in accuracy (reduced (early): $t(41) =
-1.499, p = 0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal
-clustering (reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40)
-= 1.099, p = 0.278$), nor feature-based clustering (reduced (early), category:
-$t(41) = 0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$;
-reduced (early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first
-letter: $t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086,
-p = 0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
+participants tended to exhibit stronger clustering effects on feature rich
+(versus reduced) lists, and (b) participants tended to remember more words and
+exhibit stronger clustering effects on early (versus late) lists, we expected
+these early versus late differences to be enhanced in the reduced (early)
+condition and diminished in the reduced (late) condition. However, to our
+surprise, participants in \textit{neither} condition exhibited reliable early
+versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
+0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
+(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
+= 0.278$), nor feature-based clustering (reduced (early), category: $t(41) =
+0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
+(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
+$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
+0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
 length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
@@
 late lists) for the semantic order manipulations (but not other manipulations)
-also tended to show a larger improvement in recall
+also tended to show a smaller decrease in recall on early versus late lists
 (Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.378, p < 0.001$;
@@
 
-\subsection*{Data and code availability}
+\subsection*{Author note}
 
@@
 Python toolbox for analyzing free recall data, which may be found at
-https://cdl-quail.read\-the\-docs.io/\-en/\-latest/.
+https://cdl-quail.read\-the\-docs.io/\-en/\-latest/.  Note that this study was not
+preregistered.  Some of the ideas and data presented in this manuscript
+were also presented at the Annual Meeting of the Society for Neuroscience (2017).
 

### 2023-07-04 code/analyze.py
figure format changes to address reviewer requests:   - use 1-indexing instead of 0-indexing   - use "vlag" colormap for difference plots (instead of mako)   - increase font size in figure S8 legend   - cut all violin plots off outside of observed ranges

@@ -106,7 +106,21 @@ def analyze_data(analyses=['fingerprint', 'pfr', 'lagcrp', 'spc', 'accuracy'], d
     return results, analyses, listgroups
 
 
+# update PFR and SPC curves to use 1-indexing instead of 0-indexing, per reviewer request
+def increment_presentation_positions(r):
+    if type(r) is dict:
+        return {k + 1 if type(k) is int else k: increment_presentation_positions(v) for k, v in r.items()}
+    
+    x = r.get_data().copy()
+    if type(x.columns) is pd.RangeIndex:
+        x.columns = pd.RangeIndex(start=x.columns.start + 1, stop=x.columns.stop + 1, step=x.columns.step)
+    
+    r.data = x
+    return r
+
+
 results, analyses, listgroups = analyze_data(savefile=results_file)
+results = {k: increment_presentation_positions(v) if k in ['pfr', 'spc', 'pnr'] else v for k, v in results.items()}
 
 
 # hack for recovering fingerprint features (feature tags are lost in the pickling process)
@@ -155,7 +169,6 @@ def organize_by_listgroup(x, groups):
 
 results_by_list = results  # per-list results
 results = {a: organize_by_listgroup(results_by_list[a], listgroups) for a in results_by_list.keys()}  # averaged within each listgroup
-
 
 def select_conds(results, conds='all'):
     return {k: v for k, v in results.items() if conds == 'all' or k in conds}
@@ -265,7 +278,7 @@ def stack_diffs(diffs, include_conds='all'):
 
 
 def pnr_matrix(pnr_results, include_conds='all', include_lists='all'):
-    positions = range(16)
+    positions = range(1, 17)
 
     if type(include_conds) is str:
         include_conds = [include_conds]

### 2023-07-04 code/plot.py

@@ -126,7 +126,7 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
                    x='Dimension', y='Clustering score',
                    hue='Condition',
                    order=['Category', 'Size', 'Length', 'First letter', 'Color', 'Location'],
-                   palette=palette, linewidth=1, inner='quartile', scale='width')
+                   palette=palette, linewidth=1, inner='quartile', scale='width', cut=0)
     
     if ylim is not None:
         plt.ylim(ylim)
@@ -199,7 +199,7 @@ def plot_heatmaps(results, include_conds='all', include_lists='all', contrasts=N
                 ax = axes[r, column]
             except IndexError:
                 ax = axes[max([r, column])]
-            heatmap(results[cond][k] - results[cond][v], '', dvmin, dvmax, ax, showx=(len(include_lists) + i + 1) == n_rows, showy=column == 0, cmap='mako', yprepend=f'{k.capitalize()} $-$ {v.lower()}\n')
+            heatmap(results[cond][k] - results[cond][v], '', dvmin, dvmax, ax, showx=(len(include_lists) + i + 1) == n_rows, showy=column == 0, cmap='vlag', center=0, yprepend=f'{k.capitalize()} $-$ {v.lower()}\n')
     
     plt.tight_layout()
 
@@ -208,12 +208,16 @@ def plot_heatmaps(results, include_conds='all', include_lists='all', contrasts=N
     return fig
 
 
-def accuracy_by_list(x, xlim=[0, 15], ylim=[0, 1], fname=None):
+def accuracy_by_list(x, xlim=[1, 16], ylim=[0, 1], fname=None):
     _, inds = np.unique(x['Condition'], return_index=True)
     conds = [x['Condition'][i] for i in sorted(inds)]
     
     palette = [colors[c] for c in conds]
     fig = plt.figure(figsize=(5.5, 2.5))
+
+    # convert lists to 1-indexed
+    x['List'] += 1
+    
     sns.lineplot(data=x, x='List', y='Accuracy', hue='Condition', palette=palette, legend=False)
     plt.xlabel('List number', fontsize=12)
     plt.ylabel('Recall probability', fontsize=12)
@@ -358,7 +362,7 @@ def plot_trajectories(fingerprints, include_conds='all', fname=None, xlim=None, 
 
     if xlim is not None:
         if type(xlim[0]) is not list:
[…]

### 2023-10-30 paper/main.tex
checking in edits (workiing on integrating R1 responses into paper)

@@
 organized their memories of the studied words. Because random word lists are
-unstructured by design, it is not clear if, or how, non-trivial situation models
-might apply to these stimuli. Nevertheless, there are \textit{some}
-commonalities between memory for word lists and memory for real-world
-experiences.
+unstructured by design, it is not clear if, or how, non-trivial situation
+models might apply to these stimuli. As we unpack below, this provides an
+important motivation for our current study, which uses free recall of
+\textit{structured} lists to help bridge the gap between these two lines of
+research.
 
@@
 items were presented.
+
+% define the theoretical stakes of our study
+
+From a theoretical perspective, we are interested in several core questions
+organized around the central theme of how structure in our experiences affect
+how we remember \textit{those} experiences, and also how we remember
+\textit{future} experiences (which may or may not exhibit similar structure).
+For example, when we distill participants' experiences down to simple word
+lists that vary (meaningfully) along just a few feature dimensions, are there
+important differences in which dimenesions influence participants' memories? Or
+are all features essentially ``equally'' influential? Further, are there
+differences in how specific features influence participants' memories for
+ongoing versus future experiences? Are there interaction effects between
+different features, or do people appear to treat each feature independently?
+Finally, are there individual differences in how people organize their
+memories, or in how people are influenced by our experimental manipulations? If
+so, what are those differences and which aspects of memory do they affect?
 

### 2023-11-01 paper/main.tex
recompiled all figures, changed fig 5 format, added stats annotations, p-value correction, and started to input edits to address R1 concerns

@@
 different features, or do people appear to treat each feature independently?
-Finally, are there individual differences in how people organize their
+And are there individual differences in how people organize their
 memories, or in how people are influenced by our experimental manipulations? If
@@
 the meaning or semantic content of the words (e.g., word category, size of the
-referent, etc.) and of the lexicographic properties of the words (e.g., word length,
-first letter, etc.). We wondered whether this additional word-independent information
-might facilitate recall (e.g., by providing new potential ways of organizing or
-retrieving memories of the studied words) or impair recall (e.g., by
-distracting participants with irrelevant information). Second, we manipulated
-the orders in which words were studied (and how those orderings changed over
-time). We wondered whether presenting the same list of words with different
-appearances (e.g., by manipulating font size and onscreen location) or in
-different orders (e.g., sorted along one feature dimension versus another)
-might serve to influence how participants organized their memories of the
-words. We also wondered whether some order manipulations might be temporally
-``sticky'' by influencing how \textit{future} lists were remembered.
+referent, etc.) and of the lexicographic properties of the words (e.g., word
+length, first letter, etc.). We wondered whether this additional
+word-independent information might facilitate recall~\citep[e.g., by providing
+new or richer potential ways of organizing or retrieving memories of the
+studied words;][]{HargEtal12, Mada21, MeinEtal20, DavaEtal03, SlamBarl79,
+DrewMurd80, SochEtal09} or impair recall~\citep[e.g., by distracting or
+confusing participants with irrelevant information][]{MarsEtal15, MarsEtal12,
+Lang05, ReinEtal92}. Second, we manipulated the orders in which words were
+studied (and how those orderings changed over time). We wondered whether
+presenting the same list of words with different appearances (e.g., by
+manipulating font size and onscreen location) or in different orders (e.g.,
+sorted along one feature dimension versus another) might serve to influence how
+participants organized their memories of the words~\citep[e.g., ][]{PolyKaha08,
+MannEtal15}. We also wondered whether some order manipulations might be
+temporally ``sticky'' by influencing how \textit{future} lists were
+remembered~\citep[e.g., ][]{SiroEtal05, LohnEtal10, Whit27, Badd68,
+DarlMurd71}.
 
@@
 the visual features was acting as a sort of ``event boundary'' between early
-and late lists. In prior work, we (and others) have found that memories formed
-just after event boundaries can be enhanced~\citep[e.g., due to less contextual
-interference between pre- and post-boundary items;][]{MannEtal16, PettEtal16,
-GoldEtal17, FlorEtal17}.
+and late lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior
+work, we (and others) have found that memories formed just after event
+boundaries can be enhanced~\citep[e.g., due to less contextual interference
+between pre- and post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17,
+FlorEtal17}.
 
@@
 \begin{figure}[tp] \centering
-    \includegraphics[width=0.7\textwidth]{figures/memory_perf_barchart_compare}
+    \includegraphics[width=\textwidth]{figures/memory_perf_barchart_compare}
 

### 2023-11-01 code/plot.py

@@ -13,7 +13,7 @@ from tqdm import tqdm
 from matplotlib import pyplot as plt
 from matplotlib.ticker import MaxNLocator
 from sklearn.decomposition import IncrementalPCA as PCA
-import statannot
+from statannotations.Annotator import Annotator 
 
 
 figdir = os.path.join(os.path.split(os.getcwd())[0], 'paper', 'figures', 'source')
@@ -123,11 +123,11 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
             palette.append(colors[c])
             order.append(c)
     
-    sns.violinplot(data=reorder_df(pd.concat(fingerprints, axis=0), 'Condition', order),
-                   x='Dimension', y='Clustering score',
-                   hue='Condition',
-                   order=['Category', 'Size', 'Length', 'First letter', 'Color', 'Location'],
-                   palette=palette, linewidth=1, inner='quartile', scale='width', cut=0)
+    sns.barplot(data=reorder_df(pd.concat(fingerprints, axis=0), 'Condition', order),
+                x='Dimension', y='Clustering score',
+                hue='Condition',
+                order=['Category', 'Size', 'Length', 'First letter', 'Color', 'Location'],
+                palette=palette, linewidth=1)
     
     if ylim is not None:
         plt.ylim(ylim)
@@ -580,21 +580,14 @@ def barplot_helper(clustering_results, x='Condition', y=None, hue=None, palette=
         y_ref = clustering_results.query('Condition == @ref')[y].mean()
         plt.plot(xlim, [y_ref, y_ref], '--', color='black', linewidth=1)
         plt.xlim(xlim)
-
-        pairs = [((ref, 'Early'), (c, 'Early')) for c in clustering_results['Condition'].unique() if c != ref]
+        
+        pairs = [((c, 'Early'), (c, 'Late')) for c in clustering_results['Condition'].unique()]
+        pairs.extend([((ref, 'Early'), (c, 'Early')) for c in clustering_results['Condition'].unique() if c != ref])
         pairs.extend([((ref, 'Late'), (c, 'Late')) for c in clustering_results['Condition'].unique() if c != ref])
-        pairs.extend([((c, 'Early'), (c, 'Late')) for c in clustering_results['Condition'].unique()])
 
-        statannot.add_stat_annotation(
-            ax,
-            data=clustering_results,
-            x=x,
-            y=y,
-            box_pairs=pairs,
-            test='t-test_paired',
[…]

### 2023-11-02 paper/main.tex
recomputing fingerprints after (hopefully) fixing quail "bug"
note: the "bug" was an edge case related to computing permutation-corrected clustering scores.  in the previous implementation, if all list items had the same feature value for a given feature, the corrected clustering score for that feature would be 1, since the comparisons for permuted vs. observed scores were made using a strictly "less than" comparison (and none of the permuted scores would be greater than the observed score).

to fix this edge case, quail now considers 3 cases for each permuted score (and computed the average a

@@
 \caption{\textbf{Memory ``fingerprints'' (order manipulation conditions).} The
-across-participant distributions of clustering scores for each feature type
-($x$-coordinate) are displayed for each experimental condition (color),
-separately for order manipulation (early, top) and randomly ordered (late,
-bottom) lists. See Figures~\fingerprintsRandom~and~\fingerprintsAdaptive~for
-analogous plots for the random and adaptive conditions.}
-\label{fig:fingerprints}
+across-participant clustering scores for each feature type ($x$-coordinate) are
+displayed for each experimental condition (color), separately for order
+manipulation (early, top) and randomly ordered (late, bottom) lists. Error bars
+denote bootstrap-estimated 95\% confidence intervals. See
+Figures~\fingerprintsRandom~and~\fingerprintsAdaptive~for analogous plots for
+the random and adaptive conditions.} \label{fig:fingerprints}
 
@@
 lists.} The bar heights display the average (across participants) recall
-probabilities (\textbf{A.}), temporal clustering scores (\textbf{B.}), and
-feature clustering scores (\textbf{C.}) for early (gray) and late (gold) lists.
-For the feature rich bars (left), the feature clustering scores are averaged
-across features. For the order manipulation conditions, feature clustering
-scores are displayed for the focused-on feature for each condition (e.g.,
-category clustering scores are displayed for the category condition, and so
-on). All panels: error bars denote bootstrap-estimated 95\% confidence
-intervals. The horizontal dotted lines denote the average values (across all
-lists and participants) for the feature rich condition.
-\textbf} \label{fig:barplots}
+probabilities (\textbf{A.}), temporal clustering scores (\textbf{B.}), feature
+clustering scores (\textbf{C.}), and residual feature clustering scores (after
+factoring out temporal clustering effects; \textbf{D.}) for early (gray) and
+late (gold) lists. For the feature rich bars (left), the feature clustering
+scores are averaged across features. For the order manipulation conditions,
+feature clustering scores are displayed for the focused-on feature for each
+condition (e.g., category clustering scores are displayed for the category
+condition, and so on). All panels: error bars denote bootstrap-estimated 95\%
+confidence intervals. The horizontal dotted lines denote the average values
+(across all lists and participants) for the feature rich condition. The bars
+denote $t$-tests between the corresponding bars, and the asterisks denote the
+Benjamini-Hochberg-corrected $p$-values. Comparisons for which corrected $p
+\geq 0.05$ are not shown.} \label{fig:barplots}
 

### 2023-11-02 paper/supplement.tex
recomputing fingerprints after (hopefully) fixing quail "bug"
note: the "bug" was an edge case related to computing permutation-corrected clustering scores.  in the previous implementation, if all list items had the same feature value for a given feature, the corrected clustering score for that feature would be 1, since the comparisons for permuted vs. observed scores were made using a strictly "less than" comparison (and none of the permuted scores would be greater than the observed score).

to fix this edge case, quail now considers 3 cases for each permuted score (and computed the average a

@@
 
-    \caption{\textbf{Memory ``fingerprints'' (random conditions).}
-    The across-participant distributions of clustering scores for each feature
-    type ($x$-axis) are displayed for each experimental condition
-    (color), separately for early (top) and late (bottom) lists.}
-        \label{fig:fingerprints-random}
-\end{figure}
+\caption{\textbf{Memory ``fingerprints'' (random conditions).} The
+across-participant clustering scores for each feature type ($x$-axis) are
+displayed for each experimental condition (color), separately for early (top)
+and late (bottom) lists. Error bars denote bootstrap-estimated 95\% confidence
+intervals.}
+
+\label{fig:fingerprints-random} \end{figure}
 
@@
 
-    \caption{\textbf{Memory ``fingerprints'' (adaptive conditions).}
-    The across-participant distributions of clustering scores for each feature
-    type ($x$-axis) are displayed for each experimental condition
-    (color).}
-        \label{fig:fingerprints-adaptive}
-\end{figure}
+\caption{\textbf{Memory ``fingerprints'' (adaptive conditions).} The
+across-participant clustering scores for each feature type ($x$-axis) are
+displayed for each experimental condition (color). Error bars denote
+bootstrap-estimated 95\% confidence intervals.}
+
+\label{fig:fingerprints-adaptive} \end{figure}
 

### 2023-11-03 paper/main.tex
backing up progress:   - switched to updated version of quail to fix "all features have the same value" corrected clustering score edge case   - working on re-computing all fingerprints (this will take a few days to complete; only the first batch are included in this update)   - working on addressing reviewer 1 comments in the main text (not yet complete)

@@
 % connections to prior work: 
+%     - richness of encoding
 %     - context effects: role of irrelevant features, temporal context and order effects
@@
 %     - situation models:
+
+\subsection*{Memory consequences of feature variability}
+
+Several prior studies have examined how varying the richness or experiences, or
+the extensive of encoding, can affect memory. Although specific details
+differ~\citep{BoniEtal22}, in general these studies have found that richer and
+more deeply or extensively encoded experiences are remembered
+better~\citep{HargEtal12, Mada21, MeinEtal20}. Our findings help to elucidate
+an additional factor that may contribute to these phenomenon. For example, our
+finding that participants better remember ``feature rich'' lists (where words'
+appearances are varied) than ``reduced'' lists (where words' appearances are
+held constant) only when those feature rich lists are presented \textit{after}
+reduced lists suggests that some factors that influence the richness or depth
+of encoding may be relative, rather than absolute. In other words,
+\textit{changes} in richness may be more important than the overall complexity
+or numbers of features.
+
+We did \textit{not} find any evidence that changing words' appearances
+\textit{harmed} memory performance, e.g., by distracting them with irrelevant
+information~\citep{MarsEtal15, Lang05, MarsEtal12, ReinEtal92}. Nor did we find
+any evidence that \textit{changes} in the presence of potentially
+``distracting'' features adversely affected memory. For example, when we
+increased or decreased the variability in words' appearances on late versus
+early lists (as in the reduced (early) and reduced (late) conditions), we found
+no evidence that this harmed participants' memories.
+
+% add something about multiple pathways to recall (for different feature dimensions)
 

### 2023-11-03 code/analyze.py

@@ -81,11 +81,13 @@ def analyze_data(analyses=['fingerprint', 'pfr', 'lagcrp', 'spc', 'accuracy'], d
             tmpfile = f'{a}.pkl'
 
         # save out temp files
-        if a not in ['pfr', 'pnr']:
-            with Pool(min([cpu_count(), len(data)])) as p:
-                p.map(apply_wrapper, [[a, x, d, tmpfile, kwargs] for x, d in data.items()])
-        else:
-            [apply_wrapper([a, x, d, tmpfile, kwargs]) for x, d in data.items()]
+        # if a not in ['pfr', 'pnr']:
+        #     for x, d in data.items():
+        #         apply_wrapper()
+        #     with Pool(min([cpu_count(), len(data)])) as p:
+        #         p.map(apply_wrapper, [[a, x, d, tmpfile, kwargs] for x, d in data.items()])
+        # else:
+        [apply_wrapper([a, x, d, tmpfile, kwargs]) for x, d in data.items()]
         
         # load in temp files and update results
         for x in data.keys():

### 2023-11-07 paper/main.tex
working on addressing reviewer comments:   - fixed a "bug" in quail (https://github.com/ContextLab/quail/releases/tag/v0.2.2) related to computing permutation-corrected clustering scores    - regenerated all figures using a quick run of the updated clustering score metric, using N = 10 iterations    - I'm now rerunning the "full" analysis using N = 500 iterations (as we used in the previous implementation) to generate the production-ready figures and stats     - added a new discusion section on theoretical implications related to feature variability     - added description to methods on tempor

@@
 (seven-dimensional) feature spaces.
+
+\subsubsection*{Factoring out the effects of temporal clustering}
+
+For a given list of words, if the values along two feature dimensions (e.g.,
+category and size) are correlated, then the clustering scores for those two
+dimensions will also be correlated. When lists are sorted along a given feature
+dimension, the sorted feature values will also tend to be correlated with the
+serial positions of the words in the list. This means that the temporal
+clustering score will \textit{also} tend to be correlated with the clustering
+scores for the sorted feature dimension. These correlations mean that it can be
+difficult to specifically identify when participants are using one feature
+versus another (or a manipulated feature versus temporal information) to
+organize or search their memories.
+
+We developed a permutation-based procedure to factor out the effects of
+temporal clustering from the clustering scores for each feature dimension. For
+a given set of recalled items (whose presentation positions are given by $x_1,
+x_2, x_3, …, x_N$), we circularly shift the presentation positions by a
+randomly chosen amount (between 1 and the list length) to obtain a new set of
+items. Since the new set of items will have the same (average) temporal
+distances between successive recalls, the temporal clustering score for the new
+set of items is equal (on average) to the temporal clustering score for the
+original recalls. However, we can then re-compute the feature clustering score
+for those new items. Finally, we can compute a ``temporally corrected'' feature
+clustering score by computing the average percentile rank of the observed (raw)
+feature clustering score within the distribution of circularly shifted feature
+clustering scores. This new temporally corrected score provides an estimate of
+the observed degree of feature clustering over and above what could be
+accounted for by temporal clustering alone.
+
+While these temporally corrected clustering scores are useful for identifying
+when feature clustering cannot be accounted for by temporal clustering alone,
+they are \textit{not} necessarily valid estimates of the ``true'' degree to
+which participants are organizing their memories along a given feature
+dimension. For example, on a list where the presentation order and feature
+values (along the given feature dimension) are perfectly correlated, the
+temporally corrected score will have an expected value of 0.5 no matter which
+words (or in what order) are recalled. Therefore these temporally corrected
+clustering scores are interpretable only to the extent that presentation order
+and feature values are decoupled.
 
@@
 of encoding may be relative, rather than absolute. In other words,
-\textit{changes} in richness may be more important than the overall complexity
-or numbers of features.
+\textit{increases} in richness (e.g., relative to a recency-weighted baseline)
+may be more important than the overall complexity or numbers of features.
+
+Some prior studies have suggested that people can ``cue'' their memories using
+different ``strategies'' or ``pathways'' for searching for the target
+information. For example, modern accounts of free recall typically posit that
+memory search typically begins by matching the current state of mental context
+with the contexts associated with other items in memory~\citep{Kaha20}. Since
+context is the defining hallmark of episodic memory~\citep{Tulv83},
+context-based search can be described as an ``episodic'' pathway to recall.
+When episodic cueing fails to elicit a match, participants may then search for
+items that are similar to the current mental context or mental state along
+other dimensions, such as semantic similarity~\citep{DavaEtal03, SochEtal09}.
+These multiple pathways accounts of memory search also provide a potential
+explanation of why participants might have an easier time remembering richer
+stimuli (or experiences): richer stimuli and experiences might have more
+features that could be used to cue memory search. Our work suggests that there
+may be some additional factors at play with respect to the \textit{dynamics} of
+these processes. In particular, we only observed memory benefits for ``richer''
+stimuli when they were encountered after more ``impoverished'' stimuli (in the
+reduced (early) condition). This suggests that the pathways available to recall
+a given item may also depend on recent prior experiences.
 
@@
 early lists (as in the reduced (early) and reduced (late) conditions), we found
-no evidence that this harmed participants' memories.
-
-% add something about multiple pathways to recall (for different feature dimensions)
+no evidence that this harmed participants' memories. One potential
+interpretation under the ``multiple pathways to recall'' framework is that the
+availability of multiple pathways to recall do not appear to specifically
+interfere with each other.
 

### 2023-11-07 code/analyze.py

@@ -17,7 +17,9 @@ from copy import deepcopy
 
 from dataloader import fetch_data, get_listgroups, datadir, feature_groupings
 
-results_file = 'analyzed_500_iter.pkl'
+N_ITER = 500
+
+results_file = f'analyzed_{N_ITER}_iter.pkl'
 
 random = ['feature rich', 'reduced (early)', 'reduced (late)', 'reduced']
 adaptive = ['adaptive']
@@ -73,7 +75,7 @@ def analyze_data(analyses=['fingerprint', 'pfr', 'lagcrp', 'spc', 'accuracy'], d
 
         if a == 'fingerprint':
             kwargs['permute'] = True
-            kwargs['n_perms'] = 500
+            kwargs['n_perms'] = N_ITER
             tmpfile = f'{a}-{kwargs["permute"]}-{kwargs["n_perms"]}.pkl'
         else:
             if a == 'pfr':
@@ -544,7 +546,7 @@ def filter_egg(data, g, listgroups):
     return p, r
 
 
-def recall_accuracy_near_boundaries(data, bounds, listgroups, maxlag=10):    
+def recall_accuracy_near_boundaries(data, bounds, listgroups, maxlag=10):
     results = {}
     for g in np.unique(listgroups):
         p, r = filter_egg(data, g, listgroups)
@@ -725,7 +727,7 @@ def compute_temporally_corrected_fingerprints(results, results_by_list, listgrou
         elif type(positions) is pd.DataFrame:
             return pd.DataFrame([get_recalls(positions.iloc[i], presentations.iloc[i]) for i in range(positions.shape[0])], index=positions.index, columns=positions.columns)
     
-    def temporally_corrected_fingerprints(x, n=500, savefile=None):
+    def temporally_corrected_fingerprints(x, n=N_ITER, savefile=None):
         if savefile is not None and os.path.exists(savefile):
             with open(savefile, 'rb') as f:
                 observed, shuffled = pickle.load(f)

### 2023-11-07 paper/main.tex
added text related to temporal correction, incorporated refs to barplot figure, converted barplot figure to a sidewaysfigure so that it's readable

@@
 \usepackage{amsopn}
+\usepackage{rotating}
 
@@
 clustering score by computing the average percentile rank of the observed (raw)
-feature clustering score within the distribution of circularly shifted feature
-clustering scores. This new temporally corrected score provides an estimate of
-the observed degree of feature clustering over and above what could be
-accounted for by temporal clustering alone.
+feature clustering score within the distributions of circularly shifted feature
+clustering scores, across $N = 500$ repetitions of this procedure. This new
+temporally corrected score provides an estimate of the observed degree of
+feature clustering over and above what could be accounted for by temporal
+clustering alone.
 
@@
 on early lists (again, relative to early feature rich lists: $t(96) = 1.850, p
-= 0.067$). Participants' performances on early lists in all of the other order
-manipulation conditions were indistinguishable from performance on the early
-feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$). Participants in both of
-the semantically ordered conditions exhibited stronger temporal clustering on
-early lists (versus early feature rich lists; category: $t(95) = 8.508, p <
-0.001$; size: $t(95) = 2.429, p = 0.017$). Participants in the length-ordered
-condition tended to exhibit \textit{less} temporal clustering on early lists
-relative to early feature rich lists ($t(95) = -1.666, p = 0.099$), whereas
-participants in the first letter-ordered condition exhibited stronger temporal
-clustering on early lists ($t(95) = 2.587, p = 0.011$). Participants in the
-visually ordered conditions exhibited more similar performance on early lists,
-relative to early feature rich lists (color: $t(96) = -1.064, p = 0.290$; we
-found a trending enhancement for participants in the location-ordered
-condition: $t(95) = 1.682, p = 0.096$). We also compared feature-based
-clustering on early lists across the order manipulation and feature rich
-conditions. Since these results were similar across both semantic conditions
-(category and size), both lexicographic conditions (length and first letter),
-and both visual conditions (color and location), here we aggregate data from
-conditions that manipulated each of these three feature groupings in our
-comparisons, to simplify the presentation. On early lists, participants in the
-semantically ordered conditions exhibited stronger semantic clustering relative
-to participants in the feature rich condition (category: $t(125) = 2.524, p =
-0.013$; size: $t(125) = 3.510, p = 0.001$), but showed no reliable differences
-in lexicographic (length: $t(125) = 0.539, p = 0.591$; first letter: $t(125) =
--0.587, p = 0.558$) or visual (color: $t(125) = -0.579, p = 0.564$; location:
-$t(125) = -0.346, p = 0.730$) clustering. Similarly, participants in the
-lexicographically ordered conditions exhibited stronger (relative to feature
-rich participants) lexicographic clustering (length: $t(125) = 3.426, p =
-0.001$; first letter: $t(125) = 3.236, p = 0.002$) on early lists, but showed
-no reliable differences in semantic (category: $t(125) = -1.078, p = 0.283$;
-size: $t(125) = -0.310, p = 0.757$) or visual (color: $t(125) = -0.209, p =
-0.835$; location: $t(125) = -0.004, p = 0.997$) clustering. And participants in
-the visually ordered conditions exhibited stronger visual clustering (again,
-relative to feature rich participants, and on early lists; color: $t(126) =
-2.099, p = 0.038$; location: $t(126) = 4.392, p < 0.001$), but showed no
-reliable differences in semantic (category: $t(126) = 0.204, p = 0.839$; size:
-$t(126) = -0.093, p = 0.926$) or lexicographic (length: $t(126) = 0.714, p =
-0.476$; first letter: $t(126) = 0.820, p = 0.414$) clustering. Taken together,
-these order manipulation results suggest several broad patterns
-(Figs.~\ref{fig:recall-dynamics}A, \ref{fig:fingerprints}). First, most of the
-order manipulations we carried out did \textit{not} reliably affect overall
-recall performance. Second, most of the order manipulations increased
-participants' tendencies to temporally cluster their recalls. Third, all of the
-order manipulations enhanced participants' clustering of each condition's
-target feature (i.e., semantic manipulations enhanced semantic clustering,
-lexicographic manipulations enhanced lexicographic clustering, and visual
-manipulations enhanced visual clustering) while leaving clustering along other
+= 0.067$; Fig.~\ref{fig:barplots}A). Participants' performances on early lists
+in all of the other order manipulation conditions were indistinguishable from
+performance on the early feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$).
+Participants in both of the semantically ordered conditions exhibited stronger
+temporal clustering on early lists (versus early feature rich lists; category:
+$t(95) = 8.508, p < 0.001$; size: $t(95) = 2.429, p = 0.017$;
+Fig.~\ref{fig:barplots}B). Participants in the length-ordered condition tended
+to exhibit \textit{less} temporal clustering on early lists relative to early
+feature rich lists ($t(95) = -1.666, p = 0.099$), whereas participants in the
+first letter-ordered condition exhibited stronger temporal clustering on early
+lists ($t(95) = 2.587, p = 0.011$). Participants in the visually ordered
+conditions exhibited more similar performance on early lists, relative to early
+feature rich lists (color: $t(96) = -1.064, p = 0.290$; we found a trending
+enhancement for participants in the location-ordered condition: $t(95) = 1.682,
+p = 0.096$). We also compared feature-based clustering on early lists across
+the order manipulation and feature rich conditions. Since these results were
+similar across both semantic conditions (category and size), both lexicographic
+conditions (length and first letter), and both visual conditions (color and
+location), here we aggregate data from conditions that manipulated each of
+these three feature groupings in our comparisons, to simplify the presentation.
+On early lists, participants in the semantically ordered conditions exhibited
+stronger semantic clustering relative to participants in the feature rich
+condition (category: $t(125) = 2.524, p = 0.013$; size: $t(125) = 3.510, p =
+0.001$), but showed no reliable differences in lexicographic (length: $t(125) =
+0.539, p = 0.591$; first letter: $t(125) = -0.587, p = 0.558$) or visual
+(color: $t(125) = -0.579, p = 0.564$; location: $t(125) = -0.346, p = 0.730$)
+clustering. Similarly, participants in the lexicographically ordered conditions
+exhibited stronger (relative to feature rich participants) lexicographic
+clustering (length: $t(125) = 3.426, p = 0.001$; first letter: $t(125) = 3.236,
+p = 0.002$) on early lists, but showed no reliable differences in semantic
+(category: $t(125) = -1.078, p = 0.283$; size: $t(125) = -0.310, p = 0.757$) or
+visual (color: $t(125) = -0.209, p = 0.835$; location: $t(125) = -0.004, p =
+0.997$) clustering. And participants in the visually ordered conditions
+exhibited stronger visual clustering (again, relative to feature rich
+participants, and on early lists; color: $t(126) = 2.099, p = 0.038$; location:
+$t(126) = 4.392, p < 0.001$), but showed no reliable differences in semantic
+(category: $t(126) = 0.204, p = 0.839$; size: $t(126) = -0.093, p = 0.926$) or
+lexicographic (length: $t(126) = 0.714, p = 0.476$; first letter: $t(126) =
+0.820, p = 0.414$) clustering. Taken together, these order manipulation results
+suggest several broad patterns (Figs.~\ref{fig:recall-dynamics}A,
+\ref{fig:fingerprints}). First, most of the order manipulations we carried out
+did \textit{not} reliably affect overall recall performance. Second, most of
+the order manipulations increased participants' tendencies to temporally
+cluster their recalls. Third, all of the order manipulations enhanced
+participants' clustering of each condition's target feature (i.e., semantic
+manipulations enhanced semantic clustering, lexicographic manipulations
+enhanced lexicographic clustering, and visual manipulations enhanced visual
+clustering; Fig.~\ref{fig:barplots}C) while leaving clustering along other
 feature dimensions roughly unchanged (i.e., semantic manipulations did not
-affect lexicographic or visual clustering, and so on).
+affect lexicographic or visual clustering, and so on). Although it is not
+possible to fully separate feature versus temporal clustering when considering
+sorted lists, we used a permutation-based procedure to identify the degree of
+feature clustering over and above what could be accounted for by temporal
+clustering alone (see \textit{Factoring out the effects of temporal
+clustering}). When we carried out this analysis (Fig.~\ref{fig:barplots}D), we
+found that participants exhibited more semantic clustering on semantically
+sorted lists than on randomly ordered lists (category: $t(XXX) = XXX, p = XXX$;
+size: $t(XXX) = XXX, p = XXX$), but the effects of the other order
+manipulations could not reliably be separated from temporal clustering alone
+($t$s $> XXX, p$s $> XXX$).
 
@@
 
-\begin{figure}[tp] \centering
+\begin{sidewaysfigure}[p]
+
+    \centering
     \includegraphics[width=\textwidth]{figures/memory_perf_barchart_compare}
@@
 late (gold) lists. For the feature rich bars (left), the feature clustering
-scores are averaged across features. For the order manipulation conditions,
-feature clustering scores are displayed for the focused-on feature for each
-condition (e.g., category clustering scores are displayed for the category
-condition, and so on). All panels: error bars denote bootstrap-estimated 95\%
-confidence intervals. The horizontal dotted lines denote the average values
-(across all lists and participants) for the feature rich condition. The bars
-denote $t$-tests between the corresponding bars, and the asterisks denote the
-Benjamini-Hochberg-corrected $p$-values. Comparisons for which corrected $p
-\geq 0.05$ are not shown.} \label{fig:barplots}
-
-\end{figure}
+scores are averaged across all feature dimensions. For the order manipulation
+conditions, feature clustering scores are displayed for the focused-on feature
+for each condition (e.g., category clustering scores are displayed for the
+category condition, and so on). All panels: error bars denote
+bootstrap-estimated 95\% confidence intervals. The horizontal dotted lines
+denote the average values (across all lists and participants) for the feature
+rich condition. The bars denote $t$-tests between the corresponding bars, and
+the asterisks denote the Benjamini-Hochberg-corrected $p$-values. Comparisons
+for which corrected $p \geq 0.05$ are not shown.} \label{fig:barplots}
+
+\end{sidewaysfigure}
 

### 2023-11-09 paper/main.tex
added new section on "Free recall of blocked versus random categorized word lists" along with a bunch of new refs

@@
 
+\subsection*{Free recall of blocked versus random categorized word lists}
+
+A large number of prior studies have compared participants' memories for
+categorized word lists that are presented in blocked versus random orders. In
+``blocked'' lists, all of the words from a given semantic category (e.g.,
+animals) are presented consecutively, whereas in ``random'' lists, the words
+from different categories are intermixed. Most of these studies report that
+participants tend to better remember blocked (versus random)
+lists~\citep{CofeEtal66, DAgo69, Dall64, Shap70, ShapPalm70, Kint70,
+BoweEtal68, BoweEtal69, LuekEtal71}. Other studies suggest that these order
+effects may also be modulated by factors like list length and the numbers of
+examplars in each cateogry~\citep[e.g.,][]{BorgMang72}.
+
+Although we did not directly manipulate ``blocking'' in our order manipulation
+conditions, our sorting procedures in those conditions (see
+\textit{Constructing feature-sorted lists}) have \textit{indirect} effects on
+the lists' blockiness. For example, lists that are stochastically sorted by
+semantic category will tend to contain runs of several same-category words in
+succession. Consistent with the above work on blocked versus random categorized
+lists, we found that participants tended to better remember lists that were
+sorted semantically (Fig.~\ref{fig:barplots}B). However, this memory
+improvement did not appear to extend to the other order manipulation conditions
+we considered (e.g., to lexicographically or visually sorted lists). One
+possibility is that the memory benefits of blocked versus random lists are
+specific to semantic categories, and do not generalize to other feature
+dimensions. Another possibility is that the memory benefits are due to the
+presence of infrequent ``jumps'' between successive items (e.g., from different
+categories). Because the features we manipulated in the lexicographic and
+visual conditions were less categorical than the semantic features, feature
+values across words in those conditions tended to vary more gradually.
+Relatively stable features that are punctuated by infrequent large changes
+(e.g., as words transition from a same-category sequence to a new category) may
+also relate to perceived ``event boundaries,'' which can have important
+consequences for memory~\citep{RadvZack17, DuBrEtal17, DuBrDava16, DuBrDava13}.
+
 \subsection*{Expectation, event boundaries, and situation models}

### 2023-11-10 paper/main.tex
backing up next batch of shuffled fingerprints [WIP]
also working on adding additional discussion requested by reviewers

@@
 lists~\citep{CofeEtal66, DAgo69, Dall64, Shap70, ShapPalm70, Kint70,
-BoweEtal68, BoweEtal69, LuekEtal71}. Other studies suggest that these order
+BoweEtal68, BoweEtal69, LuekEtal71, Puff74}. Other studies suggest that these order
 effects may also be modulated by factors like list length and the numbers of

### 2023-11-10 code/analyze.py
added effect sizes and 95% confidence intervals to stats

@@ -584,7 +584,7 @@ def get_boundaries(n_stddev):
     return boundaries, accuracy_near_boundaries
 
 
-def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_sample=True):
+def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_sample=True, n_iter=1000, alpha=0.05):
     x = x.data
     y = y.data
 
@@ -607,17 +607,53 @@ def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_
 
     if independent_sample: 
         tfun = stats.ttest_ind
-        df = len(x) + len(y) - 2
+        df = len(x) + len(y) - 2        
     else:
         tfun = stats.ttest_rel 
         df = len(x) - 1
-    result = tfun(x, y) 
+    result = tfun(x, y)
+
+    # cohen's d
+    x_mean = np.mean(x, axis=0)
+    y_mean = np.mean(y, axis=0)
+    x_var = np.var(x, ddof=1, axis=0)
+    y_var = np.var(y, ddof=1, axis=0)
+    pooled_sd = np.sqrt(((len(x) - 1) * x_var + (len(y) - 1) * y_var) / df)
+    d = (x_mean - y_mean) / pooled_sd
 
     try:
-        print(f't({df}) = {result.statistic[0]:.3f}, p = {result.pvalue[0]:.3f}')
-    except IndexError:
-        print(f't({df}) = {result.statistic:.3f}, p = {result.pvalue:.3f}')
+        d = d[0]
+    except:
+        pass
+    
+    # bootstrap-estimated 95% confidence interval for t-statistic
+    t_dist = np.zeros([n_iter])
+    for i in range(n_iter):
+        if independent_sample:
+            x_sample = x.sample(n=len(x), replace=True)
+            y_sample = y.sample(n=len(y), replace=True)
+            t_dist[i] = stats.ttest_ind(x_sample, y_sample)[0]
+        else:
+            inds = np.random.choice(len(x), size=len(x), replace=True)
+            x_sample = x.iloc[inds]
+            y_sample = y.iloc[inds]
+            t_dist[i] = stats.ttest_rel(x_sample, y_sample)[0]
+    low_conf, high_conf = np.percentile(t_dist, [alpha * 50, 100 - (alpha * 50)])
 
+    try:
+        t = result.statistic[0]
[…]

### 2023-11-10 code/analyze.py
working through remaining reviewer comments, new/modified analysis
i've added/modified several analyses:

- new fingerprint "stability" analysis comparing average similarity of held-out fingerprints to the average fingerprint from other lists from the same subject, or the average fingerprint from other subjects.
- fingerprint stability analysis: changed the bottom panels of figs 6 and S8 to compare each new list's fingerprint to the *average up to the prior list* rather than only to the list 1 fingerprint
- updated Fig 9D to use 1-indexed list numbers
- updated aspect ratios in Fig S7 to be sq

@@ -426,7 +426,7 @@ def trajectorize(x, n_dims=2, model=PCA, average=False):
     return {c: pd.DataFrame(m.transform(x), index=x.index) for c, x in data.items()}
 
 
-def get_dists(fingerprints):
+def get_dists(fingerprints, ref=0):
     if type(fingerprints) is dict:
         return {c: get_dists(x.data) for c, x in fingerprints.items()}
     elif type(fingerprints) is pd.DataFrame:
@@ -434,7 +434,7 @@ def get_dists(fingerprints):
         subjs = fingerprints.index.get_level_values('Subject').unique().tolist()
 
         for s in subjs:
-            by_subj.append(get_dists(fingerprints.query('Subject == @s').values))
+            by_subj.append(get_dists(fingerprints.query('Subject == @s').values, ref=ref))
         
         df = pd.DataFrame(pd.concat(by_subj, axis=0, ignore_index=True))
         df['Subject'] = subjs
@@ -442,7 +442,10 @@ def get_dists(fingerprints):
 
     dists = []
     for i in range(fingerprints.shape[0]):
-        dists.append(np.linalg.norm(fingerprints[0, :] - fingerprints[i, :]))
+        if i == 0 or str(ref).isnumeric():
+            dists.append(np.linalg.norm(fingerprints[0, :] - fingerprints[i, :]))
+        elif ref == 'mean':
+            dists.append(np.linalg.norm(fingerprints[:i, :].mean(axis=0) - fingerprints[i, :]))
     return pd.DataFrame(dists).T
 
 

### 2023-11-10 code/plot.py

@@ -145,12 +145,14 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
     return fig
 
 
-def plot_heatmaps(results, include_conds='all', include_lists='all', contrasts=None, fname=None, vmin=0, vmax=1.0, dvmin=-1.0, dvmax=1.0, fontsize=12, width=2.5, height=2, xlabel='', ylabel=''):
+def plot_heatmaps(results, include_conds='all', include_lists='all', contrasts=None, fname=None, vmin=0, vmax=1.0, dvmin=-1.0, dvmax=1.0, fontsize=12, width=2.5, height=2, xlabel='', ylabel='', rotx=None):
     def heatmap(m, title, vmn, vmx, ax, showx=False, showy=False, show_title=False, yprepend='', **kwargs):
         sns.heatmap(m, vmin=vmn, vmax=vmx, ax=ax, cbar=False, **kwargs)
 
         if showx:
             ax.set_xlabel(xlabel, fontsize=fontsize)
+            if rotx is not None:
+                ax.set_xticklabels(ax.get_xticklabels(), rotation=rotx)
         if showy:
             ax.set_ylabel(yprepend + ylabel, fontsize=fontsize)
         if show_title:
@@ -397,14 +399,14 @@ def plot_trajectories(fingerprints, include_conds='all', fname=None, xlim=None, 
             ax[0, i].set_ylim(ylim[0])
         
         # plot distances
-        dists = get_dists(fingerprints[c].data).reset_index().melt(id_vars='Subject', var_name='List', value_name='Distance')
+        dists = get_dists(fingerprints[c].data, ref='mean').reset_index().melt(id_vars='Subject', var_name='List', value_name='Distance')
         dists['List'] += 1 # convert from 0-indexed to 1-indexed
         sns.barplot(data=dists, x='List', y='Distance', ax=ax[1, i], color=colors[c])
         ax[1, i].plot([7.5, 7.5], [0, 1.2], 'k--', linewidth=2)
 
         ax[1, i].set_xlabel('List', fontsize=14)
         if i == 0:
-            ax[1, i].set_ylabel('Distance from\nlist 1 fingerprint', fontsize=14)
+            ax[1, i].set_ylabel('Distance from previous\naverage fingerprint', fontsize=14)
         else:
             ax[1, i].set_yticklabels([])
             ax[1, i].set_ylabel('')

### 2023-11-11 code/analyze.py
fixed mislabling of "word length" as a discrete (categorical) feature instead of a "continuous" feature-- this affects the "boundary probability" figure (3B)

@@ -449,40 +449,42 @@ def get_dists(fingerprints, ref=0):
     return pd.DataFrame(dists).T
 
 
-def get_event_boundaries(data, focus=None, n_stddev=2):
-    discrete_features = ['category', 'size', 'length']
-    continuous_features = ['color', 'location', 'first letter']
-    features = discrete_features + continuous_features
+discrete_features = ['category', 'size']
+continuous_features = ['color', 'location', 'length', 'first letter']
+features = discrete_features + continuous_features
 
+
+def field2feature(x):
     rename = {'firstLetter': 'first letter', 'first_letter': 'first letter', 'wordLength': 'length', 'pos': 'location'}
-
-    def field2feature(x):
-        if type(x) is list:
-            return [field2feature(y) for y in x]
-        
-        if x in rename:
-            return rename[x]
-        else:
-            return x
-
-    def rename_dict(d):
-        return {field2feature(k): v for k, v in d.items()}
+    if type(x) is list:
+        return [field2feature(y) for y in x]
     
-    def get_dists(y1, y2, focus=None):    
-        if focus is None or focus in continuous_features:
-            return [np.linalg.norm(np.array(a) - np.array(b)) for a, b in zip(y1.values, y2.values)]
-        else:
-            return [int(a != b) for a, b in zip(y1.values, y2.values)]
+    if x in rename:
+        return rename[x]
+    else:
+        return x
 
+
+def rename_dict(d):
+    return {field2feature(k): v for k, v in d.items()}
+
+
+def get_dists(y1, y2, focus=None):    
+    if focus is None or focus in continuous_features:
+        return [np.linalg.norm(np.array(a) - np.array(b)) for a, b in zip(y1.values, y2.values)]
+    else:
+        return [int(a != b) for a, b in zip(y1.values, y2.values)]
+
+
+def get_event_boundaries(data, focus=None, n_stddev=2):    
     if type(data) is dict:
         x = {}
         for k in data.keys():
             try:
                 x[k] = get_event_boundaries(data[k])
             except:
-                print('problem with', k, 'in get_event_boundaries')
+                print(f'problem with {k} in get_event_boundaries')
         return x
-        #return {k: get_event_boundaries(data[c]) for c in data.keys()}
     elif type(data) is quail.Egg:
         x = data.get_pres_features().applymap(lambda v: rename_dict(v))
         return {f: get_event_boundaries(x, focus=f) for f in features}

### 2023-11-12 code/analyze.py
added new analyses:   - correlations between features (aggregated across all lists/conditions)   - z-scored distances to previous items (for order manipulation conditions)
also corrected two bugs in the event boundary estimation:

1.) the prior version was using a standard deviation threshold of 2 (it's default) even though a threshold of 1 was specified in the notebook.  i've updated the code to actually use the specified threshold, and i've updated the threshold in the notebook to specify a threshold of 2 (for consistency with prior version)

2.) the prior version treated list length as a "d

@@ -469,7 +469,7 @@ def rename_dict(d):
     return {field2feature(k): v for k, v in d.items()}
 
 
-def get_dists(y1, y2, focus=None):    
+def feature_dists(y1, y2, focus=None):    
     if focus is None or focus in continuous_features:
         return [np.linalg.norm(np.array(a) - np.array(b)) for a, b in zip(y1.values, y2.values)]
     else:
@@ -481,13 +481,13 @@ def get_event_boundaries(data, focus=None, n_stddev=2):
         x = {}
         for k in data.keys():
             try:
-                x[k] = get_event_boundaries(data[k])
+                x[k] = get_event_boundaries(data[k], n_stddev=n_stddev)
             except:
                 print(f'problem with {k} in get_event_boundaries')
         return x
     elif type(data) is quail.Egg:
         x = data.get_pres_features().applymap(lambda v: rename_dict(v))
-        return {f: get_event_boundaries(x, focus=f) for f in features}
+        return {f: get_event_boundaries(x, focus=f, n_stddev=n_stddev) for f in features}
     elif type(data) is pd.DataFrame:
         if focus != 'first letter':
             x = data.applymap(lambda v: v[focus])
@@ -499,15 +499,19 @@ def get_event_boundaries(data, focus=None, n_stddev=2):
             y1 = x[i]
             y2 = x[i + 1]
 
-            dists[i + 1] = get_dists(y1, y2, focus=focus)
+            dists[i + 1] = feature_dists(y1, y2, focus=focus)
         
-        if focus in discrete_features:
-            return dists.astype(int)
+        if np.char.isnumeric(str(n_stddev)):
+            if focus in discrete_features:
+                return dists.astype(int)
+            else:
+                # compute a threshold (add an n_stddev param to function definition)
+                # binarize the distances as being above vs. below the threshold
+            
+                thresh = np.mean(dists.values) + n_stddev * np.std(dists.values)
+                return (dists > thresh).astype(int)
         else:
-            # compute a threshold (add an n_stddev param to function definition)
-            # binarize the distances as being above vs. below the threshold
-            thresh = np.mean(dists.values) + n_stddev * np.std(dists.values)
-            return (dists > thresh).astype(int)
+            return dists
     
 
 def shift(x, n):
@@ -578,10 +582,11 @@ def get_boundaries(n_stddev):
[…]

### 2023-11-12 code/plot.py

@@ -147,8 +147,8 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
 
 def plot_heatmaps(results, include_conds='all', include_lists='all', contrasts=None, fname=None, vmin=0, vmax=1.0, dvmin=-1.0, dvmax=1.0, fontsize=12, width=2.5, height=2, xlabel='', ylabel='', rotx=None):
     def heatmap(m, title, vmn, vmx, ax, showx=False, showy=False, show_title=False, yprepend='', **kwargs):
-        sns.heatmap(m, vmin=vmn, vmax=vmx, ax=ax, cbar=False, **kwargs)
-
+        sns.heatmap(m, vmin=vmn, vmax=vmx, ax=ax, cbar=False, square=True, **kwargs)
+        
         if showx:
             ax.set_xlabel(xlabel, fontsize=fontsize)
             if rotx is not None:

### 2023-11-13 code/plot.py
finished implementing reviewer-requested new analyses and figure modifications (still need to lay out figures for the paper, since some figures will now have new panels)

@@ -11,6 +11,7 @@ import string
 
 from tqdm import tqdm
 from matplotlib import pyplot as plt
+import matplotlib.colors as mcolors
 from matplotlib.ticker import MaxNLocator
 from sklearn.decomposition import IncrementalPCA as PCA
 from statannotations.Annotator import Annotator 
@@ -83,6 +84,16 @@ def combo_lineplot(x, include_conds='all', include_lists='all', fname=None, xlab
     return fig
 
 
+def is_light_color(color):
+    # Convert color to RGB and then to HSV
+    rgb = mcolors.to_rgb(color)
+    hsv = mcolors.rgb_to_hsv(rgb)
+    # Use the Value component to determine brightness
+    brightness = hsv[2]
+    # If brightness is high, use black color, else white
+    return brightness > 0.5
+
+
 def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=None, ylim=None, palette=None, xlabel='Dimension', ylabel='Clustering score', figsize=(15, 2.5)):
     def melt_by_list(x, cond='-'):
         x = rename_features(x).reset_index()
@@ -123,11 +134,21 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
             palette.append(colors[c])
             order.append(c)
     
-    sns.barplot(data=reorder_df(pd.concat(fingerprints, axis=0), 'Condition', order),
-                x='Dimension', y='Clustering score',
-                hue='Condition',
-                order=['Category', 'Size', 'Length', 'First letter', 'Color', 'Location'],
-                palette=palette, linewidth=1)
+    bplot = sns.barplot(data=reorder_df(pd.concat(fingerprints, axis=0), 'Condition', order),
+                        x='Dimension', y='Clustering score',
+                        hue='Condition',
+                        order=['Category', 'Size', 'Length', 'First letter', 'Color', 'Location'],
+                        palette=palette, linewidth=1)
+
+    ylim = ax.get_ylim()
+
+    # Annotate each bar
+    for p in bplot.patches:
+        height = p.get_height()
+        # Decide text color based on bar color
+        text_color = 'black' if is_light_color(p.get_facecolor()) else 'white'
+        ax.text(p.get_x() + p.get_width() / 2. + 0.01, ylim[0] + 0.05, f'{height:.2f}',
+                ha='center', va='bottom', color=text_color, fontsize=8, rotation=90)
     
     if ylim is not None:
         plt.ylim(ylim)

### 2023-11-13 code/plot.py
updated compiled figures

@@ -140,7 +140,10 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
                         order=['Category', 'Size', 'Length', 'First letter', 'Color', 'Location'],
                         palette=palette, linewidth=1)
 
-    ylim = ax.get_ylim()
+    if ylim is not None:
+        plt.ylim(ylim)
+    else:
+        ylim = ax.get_ylim()
 
     # Annotate each bar
     for p in bplot.patches:
@@ -150,8 +153,8 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
         ax.text(p.get_x() + p.get_width() / 2. + 0.01, ylim[0] + 0.05, f'{height:.2f}',
                 ha='center', va='bottom', color=text_color, fontsize=8, rotation=90)
     
-    if ylim is not None:
-        plt.ylim(ylim)
+    plt.ylim(ylim)
+    
     plt.xlabel(xlabel, fontsize=14)
     plt.ylabel(ylabel, fontsize=14)
 

### 2023-11-13 paper/main.tex
added note on pnr curves

@@
 reflect the probability that an item will be recalled first, as a function of
-its serial position during encoding. To carry out this analysis, we initialized
-(for each participant) a number-of-lists (16) by number-of-words-per-list (16)
-matrix of 0s. Then, for each list, we found the index of the word that was
-recalled first, and we filled in that position in the matrix with a 1. Finally,
-we averaged over the rows of the matrix to obtain a 1 by 16 array of
-probabilities, for each participant. We used an analogous procedure to compute
-probability of $n$\textsuperscript{th} recall curves for each participant.
-Specifically, we filled in the corresponding matrices according to the
-$n$\textsuperscript{th} recall on each list that each participant made. When a
-given participant had made fewer than $n$ recalls for a given list, we simply
-excluded that list from our analysis when computing that participant's
-curve(s). The probability of first recall curve corresponds to a special case
-where $n = 1$.
+its serial position during encoding. We used an analogous approach to compute
+the proportion of trials on which each item (as a function of its presentation
+position) was recalled at output position $n$~\citep{HowaKaha99, PolyEtal09,
+ZhanEtal23}. To carry out this analysis, we initialized (for each participant)
+a number-of-lists (16) by number-of-words-per-list (16) matrix of 0s. Then, for
+each list, we found the index of the word that was recalled first, and we
+filled in that position in the matrix with a 1. Finally, we averaged over the
+rows of the matrix to obtain a 1 by 16 array of probabilities, for each
+participant. We used an analogous procedure to compute probability of
+$n$\textsuperscript{th} recall curves for each participant. Specifically, we
+filled in the corresponding matrices according to the $n$\textsuperscript{th}
+recall on each list that each participant made. When a given participant had
+made fewer than $n$ recalls for a given list, we simply excluded that list from
+our analysis when computing that participant's curve(s). The probability of
+first recall curve corresponds to a special case where $n = 1$.
+
+We note that several other studies have used a slightly different approach to
+compute these curves, by correcting for the ``availability'' of a given word to
+be recalled. For example, if a participant recalls item 1, then item 2 on a
+given list, our approach places a 0 into the item 1 column for that list when
+computing the ``probability of second recall'' curve. However, accounting for
+the fact that the participant had already recalled item 1, an alternative
+approach~\citep[e.g.,][]{Farr10} would be to count the item 1 column as
+``unobserved'' (i.e., missing data). Ultimately we chose to use the simpler
+variant of this approach in our work, but we direct the reader to further
+discussion of this issue in other work~\citep{MoraGosh14, Farr14}.
 

### 2023-11-14 paper/main.tex
updated stats in main text

@@
 difference in recall accuracy (i.e., the proportions of correctly recalled
-words) for feature rich versus reduced lists ($t(126) = -0.290, p = 0.772$).
-However, participants in the feature rich condition clustered their recalls
-substantially more along every dimension we examined (temporal clustering:
-$t(126) = 10.624, p < 0.001$; semantic category clustering: $t(126) = 10.077, p
-< 0.001$; size clustering: $t(126) = 11.829, p < 0.001$; word length
-clustering: $t(126) = 10.639, p < 0.001$; first letter clustering: $t(126) =
-7.775, p < 0.001$; see \textit{Permutation-corrected feature clustering scores}
-for more information about how we quantified each participant's clustering
-tendencies.) Taken together, these comparisons suggest that adding new features
-changes how participants organize their memories of studied words, even when
-those new features are independent of the words themselves and even when the
-new features vary randomly across words. We found no evidence that those
-additional uninformative features were distracting (in terms of their impact on
-memory performance), but they did affect participants' recall dynamics
-(measured via their clustering scores).
+words) for feature rich versus reduced lists ($t(126) = -0.290, p = 0.772,
+\mathrm{Cohen's}~d~(d) = -0.051,
+\mathrm{bootstrap~estimated~95\%~confidence~interval}~(\mathrm{CI}) = [-2.387,
+1.768]$). However, participants in the feature rich condition clustered their
+recalls substantially more along every dimension we examined (temporal
+clustering: $t(126) = 10.632, p < 0.001, d = 1.882, \mathrm{CI} = [7.786,
+14.386]$; semantic category clustering: $t(126) = 10.148, p < 0.001, d = 1.796,
+\mathrm{CI} = [7.324, 13.778] $; size clustering: $t(126) = 12.033, p < 0.001,
+d = 2.129, \mathrm{CI} = [9.030, 15.918]$; word length clustering: $t(126) =
+10.720, p < 0.001, d = 1.897, \mathrm{CI} = [7.442, 15.174]$; first letter
+clustering: $t(126) = 6.679, p < 0.001, d = 1.182, \mathrm{CI} = [4.490,
+9.611]$; see \textit{Permutation-corrected feature clustering scores} for more
+information about how we quantified each participant's clustering tendencies.)
+Taken together, these comparisons suggest that adding new features changes how
+participants organize their memories of studied words, even when those new
+features are independent of the words themselves and even when the new features
+vary randomly across words. We found no evidence that those additional
+uninformative features were distracting (in terms of their impact on memory
+performance), but they did affect participants' recall dynamics (measured via
+their clustering scores).
 
@@
 both conditions remembered more words on early versus late lists (feature rich:
-$t(66) = 4.553, p < 0.001$; reduced: $t(60) = 2.434, p = 0.018$). Participants
-in the feature rich (but not reduced) conditions exhibited more temporal
-clustering on early versus late lists (feature rich: $t(66) = 2.318, p =
-0.024$; reduced: $t(60) = 0.929, p = 0.357$). And participants in both
-conditions exhibited more semantic (category and size) clustering on early
-versus late lists (feature rich, category: $t(66) = 3.805, p < 0.001$; feature
-rich, size: $t(66) = 2.190, p = 0.032$; reduced, category: $t(60) = 2.856, p =
-0.006$; reduced, size: $t(60) = 2.947, p = 0.005$). Participants in the reduced
-(but not feature rich) conditions exhibited more lexicographic clustering on
-early versus late lists (feature rich, word length: $t(66) = 0.161, p = 0.872$;
-feature rich, first letter: $t(66) = 0.410, p = 0.683$; reduced, word length:
-$t(60) = 3.528, p = 0.001$; reduced, first letter: $t(60) = 2.275, p = 0.026$).
-Taken together, these comparisons suggest that even when the presence or
-absence of incidental visual features is stable across lists, participants
-still exhibit some differences in their performance and memory organization
-tendencies for early versus late lists.
+$t(66) = 4.553, p < 0.001, d = 0.233, \mathrm{CI} = [2.427, 7.262]$; reduced:
+$t(60) = 2.434, p = 0.018, d = 0.134, \mathrm{CI} = [0.493, 4.910]$).
+Participants in the feature rich (but not reduced) conditions exhibited more
+temporal clustering on early versus late lists (feature rich: $t(66) = 2.268, p
+= 0.027, d = 0.181, \mathrm{CI} = [0.437, 4.425]$; reduced: $t(60) = 0.986, p =
+0.328, d = 0.061, \mathrm{CI} = [-0.897, 3.348]$). And participants in both
+conditions tended to exhibit more semantic clustering on early versus late
+lists (feature rich, category: $t(66) = 3.684, p < 0.001, d = 0.220,
+\mathrm{CI} = [1.733, 5.732]$; feature rich, size: $t(66) = 1.629, p = 0.108, d
+= 0.100, \mathrm{CI} = [-0.207, 3.905] $; reduced, category: $t(60) = 2.755, p
+= 0.008, d = 0.177, \mathrm{CI} = [0.761, 5.189]$; reduced, size: $t(60) =
+3.081, p = 0.003, d = 0.201, \mathrm{CI} = [1.210, 5.326]$). Participants in
+the reduced (but not feature rich) conditions tended to exhibit more
+lexicographic clustering on early versus late lists (feature rich, word length:
+$t(66) = -0.100, p = 0.921, d = -0.010, \mathrm{CI} = [-2.217, 1.899]$; feature
+rich, first letter: $t(66) = -0.412, p = 0.681, d = -0.045, \mathrm{CI} =
+[-2.461, 1.645]$; reduced, word length: $t(60) = 3.762, p < 0.001, d = 0.261,
+\mathrm{CI} = [1.604, 6.821]$; reduced, first letter: $t(60) = 1.721, p =
+0.090, d = 0.175, \mathrm{CI} = [-0.138, 4.098]$). Taken together, these
+comparisons suggest that even when the presence or absence of incidental visual
+features is stable across lists, participants still exhibit some differences in
+their performance and memory organization tendencies for early versus late
+lists.
 
@@
 versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
-0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
-(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
-= 0.278$), nor feature-based clustering (reduced (early), category: $t(41) =
-0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
-(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
-$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
-0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
-length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
-0.966, p = 0.340$). We hypothesized that adding or removing the variability in
-the visual features was acting as a sort of ``event boundary'' between early
-and late lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior
-work, we (and others) have found that memories formed just after event
-boundaries can be enhanced~\citep[e.g., due to less contextual interference
-between pre- and post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17,
-FlorEtal17}.
+0.141, d = 0.098, \mathrm{CI} = [-0.345, 3.579]$; reduced (late): $t(40) =
+1.462, p = 0.152, d = 0.121, \mathrm{CI} = [-0.376, 2.993]$), temporal
+clustering (reduced (early): $t(41) = 0.857, p = 0.396, d = 0.068, \mathrm{CI}
+= [-1.012, 2.896]$; reduced (late): $t(40) = 1.244, p = 0.221, d = 0.128,
+\mathrm{CI} = [-0.894, 3.088]$), nor feature-based clustering (reduced (early),
+category: $t(41) = 0.707, p = 0.484, d = 0.068, \mathrm{CI} = [-1.314, 2.830]$;
+reduced (early), size: $t(41) = 0.803, p = 0.427, d = 0.079, \mathrm{CI} =
+[-1.142, 2.953]$; reduced (early), length: $t(41) = 0.461, p = 0.648, d =
+0.060, \mathrm{CI} = [-1.545, 2.462]$; reduced (early), first letter: $t(41) =
+0.781, p = 0.439, d = 0.101, \mathrm{CI} = [-1.039, 2.881]$; reduced (late),
+category: $t(40) = -0.101, p = 0.920, d = -0.009, \mathrm{CI} = [-2.307,
+1.776]$; reduced (late), size: $t(40) = 0.555, p = 0.582, d = 0.058,
+\mathrm{CI} = [-1.444, 2.274] $; reduced (late), length: $t(40) = 1.482, p =
+0.146, d = 0.126, \mathrm{CI} = [-0.444, 3.743]$; reduced (late), first letter:
+$t(40) = -0.143, p = 0.887, d = -0.017, \mathrm{CI} = [-2.204, 1.830] $). We
+hypothesized that adding or removing the variability in the visual features was
+acting as a sort of ``event boundary'' between early and late
+lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we
+(and others) have found that memories formed just after event boundaries can be
+enhanced~\citep[e.g., due to less contextual interference between pre- and
+post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17, FlorEtal17}.
 
@@
 the same blends of features (accuracy for feature rich versus reduced (early):
-$t(107) = -2.230, p = 0.028$; reduced versus reduced (early): $t(101) = -2.045,
-p = 0.043$; also see Fig.~\accuracyByList A). However, \textit{subtracting}
-irrelevant visual features on later lists that \textit{had} been present on
-early lists (as in the reduced (late) condition) did not appear to impact
-recall performance (accuracy for feature rich versus reduced (late): $t(106) =
--0.638, p = 0.525$; reduced versus reduced (late): $t(100) = -0.407, p =
-0.685$). These comparisons suggest that recall accuracy has a directional
-component: accuracy is affected differently by removing features later
-that had been present earlier versus adding features later that had
-\textit{not} been present earlier. In contrast, we found that participants
-exhibited more temporal and feature-based clustering when we added incidental
-visual features to \textit{any} lists (comparisons of clustering on feature
-rich versus reduced lists are reported above; temporal clustering in reduced
-versus reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
--9.780$, $p$s $< 0.001$; feature-based clustering in reduced versus reduced
-(early) and reduced versus reduced (late) conditions: $t$s $\leq -5.443$, $p$s
-$< 0.001$). Temporal and feature-based clustering were not reliably different
-in the feature rich, reduced (early), and reduced (late) conditions (temporal
-clustering in feature rich versus reduced (early) and feature rich versus
-reduced (late) conditions: $t$s $\geq -1.434$, $p$s $\geq 0.154$; feature-based
-clustering in feature rich versus reduced (early) and feature rich versus
-reduced (late) conditions: $t$s $\geq -1.359$, $p$s $> 0.177$).
+$t(107) = -2.230, p = 0.028, d = -0.439, \mathrm{CI} = [-4.252, -0.229]$;
+reduced versus reduced (early): $t(101) = -2.045, p = 0.043, d = -0.410,
+\mathrm{CI} = [-3.826, 0.112]$; also see Fig.~\accuracyByList A). However,
+\textit{subtracting} irrelevant visual features on later lists that
+\textit{had} been present on early lists (as in the reduced (late) condition)
+did not appear to impact recall performance (accuracy for feature rich versus
+reduced (late): $t(106) = -0.638, p = 0.525, d = -0.126, \mathrm{CI} = [-2.720,
+1.362]$; reduced versus reduced (late): $t(100) = -0.407, p = 0.685, d =
+-0.082, \mathrm{CI} = [-2.477, 1.626]$). These comparisons suggest that recall
+accuracy has a directional component: accuracy is affected differently by
+removing features later that had been present earlier versus adding features
+later that had \textit{not} been present earlier. In contrast, we found that
+participants exhibited more temporal and feature-based clustering when we added
+incidental visual features to \textit{any} lists (comparisons of clustering on
+feature rich versus reduced lists are reported above; temporal clustering in
+reduced versus reduced (early) and reduced versus reduced (late) conditions:
+$t$s $\leq -9.885$, $p$s $< 0.001$; feature-based clustering in reduced versus
+reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
+-4.555$, $p$s $< 0.001$). Temporal and feature-based clustering were not
+reliably different in the feature rich, reduced (early), and reduced (late)
+conditions (temporal clustering in feature rich versus reduced (early) and
+feature rich versus reduced (late) conditions: $t$s $\geq -1.379$, $p$s $\geq
+0.171$; feature-based clustering in feature rich versus reduced (early) and
+feature rich versus reduced (late) conditions: $|t|$s $\leq 1.441$, $p$s $\geq
+0.153$).
 
@@
 showed an increase in memory performance on early lists (accuracy, relative to
-early feature rich lists; $t(95) = 3.034, p = 0.003$). Participants in the
-color-ordered condition also showed a trending increase in memory performance
-on early lists (again, relative to early feature rich lists: $t(96) = 1.850, p
-= 0.067$; Fig.~\ref{fig:barplots}A). Participants' performances on early lists
-in all of the other order manipulation conditions were indistinguishable from
-performance on the early feature rich lists ($\|t\|$s $< 1.013, p$s $> 0.314$).
-Participants in both of the semantically ordered conditions exhibited stronger
-temporal clustering on early lists (versus early feature rich lists; category:
-$t(95) = 8.508, p < 0.001$; size: $t(95) = 2.429, p = 0.017$;
-Fig.~\ref{fig:barplots}B). Participants in the length-ordered condition tended
-to exhibit \textit{less} temporal clustering on early lists relative to early
-feature rich lists ($t(95) = -1.666, p = 0.099$), whereas participants in the
-first letter-ordered condition exhibited stronger temporal clustering on early
-lists ($t(95) = 2.587, p = 0.011$). Participants in the visually ordered
-conditions exhibited more similar performance on early lists, relative to early
-feature rich lists (color: $t(96) = -1.064, p = 0.290$; we found a trending
-enhancement for participants in the location-ordered condition: $t(95) = 1.682,
-p = 0.096$). We also compared feature-based clustering on early lists across
-the order manipulation and feature rich conditions. Since these results were
+early feature rich lists; $t(95) = 3.034, p = 0.003, d = 0.667, \mathrm{CI} =
+[1.048, 5.113]$). Participants in the color-ordered condition also showed a
+trending increase in memory performance on early lists (again, relative to
+early feature rich lists: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
+[-0.010, 3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
+early lists in all of the other order manipulation conditions were
+indistinguishable from performance on the early feature rich lists ($\|t\|$s
+$\leq 1.013, p$s $\geq 0.314$). Participants in both of the semantically
+ordered conditions exhibited stronger temporal clustering on early lists
+(versus early feature rich lists; category: $t(95) = 8.813, p < 0.001, d =
+1.936, \mathrm{CI} = [6.793, 11.751] $; size: $t(95) = 2.630, p = 0.010, d =
+0.578, \mathrm{CI} = [0.831, 4.866]$; Fig.~\ref{fig:barplots}B). Participants
+in the length-ordered condition tended to exhibit \textit{less} temporal
+clustering on early lists relative to early feature rich lists ($t(95) =
+-1.547, p = 0.125, d = -0.340, \mathrm{CI} = [-3.693, 0.341]$), whereas
+participants in the first letter-ordered condition exhibited stronger temporal
+clustering on early lists ($t(95) = 2.858, p = 0.005, d = 0.628, \mathrm{CI} =
+[1.031, 4.886]$). Participants in the visually ordered conditions exhibited
+more similar performance (accuracy) on early lists, relative to early feature
+rich lists (we found a trending enhancement for participants in the
+color-ordered condition: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
+[-0.010, 3.712]$; location: $t(95) = 0.043, p = 0.966, d = 0.010, \mathrm{CI} =
+[-1.598, 1.729]$). Participants in the visually ordered conditions also showed
+similar temporal clustering on early lists, relative to early feature rich
+lists (color: $t(96) = -1.339, p = 0.184, d = -0.291, \mathrm{CI} = [-3.238,
+0.394]$, we found a trending increase for participants in the location-ordered
+condition: $t(95) = 1.705, p = 0.092, d = 0.374, \mathrm{CI} = [-0.155,
+3.521]$). We also compared feature-based clustering on early lists across the
+order manipulation and feature rich conditions. Since these results were
 similar across both semantic conditions (category and size), both lexicographic
@@
 stronger semantic clustering relative to participants in the feature rich
-condition (category: $t(125) = 2.524, p = 0.013$; size: $t(125) = 3.510, p =
-0.001$), but showed no reliable differences in lexicographic (length: $t(125) =
-0.539, p = 0.591$; first letter: $t(125) = -0.587, p = 0.558$) or visual
-(color: $t(125) = -0.579, p = 0.564$; location: $t(125) = -0.346, p = 0.730$)
-clustering. Similarly, participants in the lexicographically ordered conditions
-exhibited stronger (relative to feature rich participants) lexicographic
-clustering (length: $t(125) = 3.426, p = 0.001$; first letter: $t(125) = 3.236,
-p = 0.002$) on early lists, but showed no reliable differences in semantic
-(category: $t(125) = -1.078, p = 0.283$; size: $t(125) = -0.310, p = 0.757$) or
-visual (color: $t(125) = -0.209, p = 0.835$; location: $t(125) = -0.004, p =
-0.997$) clustering. And participants in the visually ordered conditions
-exhibited stronger visual clustering (again, relative to feature rich
-participants, and on early lists; color: $t(126) = 2.099, p = 0.038$; location:
-$t(126) = 4.392, p < 0.001$), but showed no reliable differences in semantic
-(category: $t(126) = 0.204, p = 0.839$; size: $t(126) = -0.093, p = 0.926$) or
-lexicographic (length: $t(126) = 0.714, p = 0.476$; first letter: $t(126) =
-0.820, p = 0.414$) clustering. Taken together, these order manipulation results
-suggest several broad patterns (Figs.~\ref{fig:recall-dynamics}A,
-\ref{fig:fingerprints}). First, most of the order manipulations we carried out
-did \textit{not} reliably affect overall recall performance. Second, most of
-the order manipulations increased participants' tendencies to temporally
-cluster their recalls. Third, all of the order manipulations enhanced
-participants' clustering of each condition's target feature (i.e., semantic
-manipulations enhanced semantic clustering, lexicographic manipulations
-enhanced lexicographic clustering, and visual manipulations enhanced visual
-clustering; Fig.~\ref{fig:barplots}C) while leaving clustering along other
-feature dimensions roughly unchanged (i.e., semantic manipulations did not
-affect lexicographic or visual clustering, and so on). Although it is not
-possible to fully separate feature versus temporal clustering when considering
-sorted lists, we used a permutation-based procedure to identify the degree of
-feature clustering over and above what could be accounted for by temporal
-clustering alone (see \textit{Factoring out the effects of temporal
-clustering}). When we carried out this analysis (Fig.~\ref{fig:barplots}D), we
-found that participants exhibited more semantic clustering on semantically
-sorted lists than on randomly ordered lists (category: $t(XXX) = XXX, p = XXX$;
-size: $t(XXX) = XXX, p = XXX$), but the effects of the other order
-manipulations could not reliably be separated from temporal clustering alone
-($t$s $> XXX, p$s $> XXX$).
+condition (category: $t(125) = 2.722, p = 0.007, d = 0.484, \mathrm{CI} =
+[0.827, 4.932]$; size: $t(125) = 3.866, p < 0.001, d = 0.687, \mathrm{CI} =
+[2.020, 5.983]$), but showed no reliable differences in lexicographic (length:
+$t(125) = 0.521, p = 0.603, d = 0.093, \mathrm{CI} = [-1.311, 2.333]$; first
+letter: $t(125) = -0.842, p = 0.401, d = -0.150, \mathrm{CI} = [-2.825,
+1.095]$) or visual (color: $t(125) = -0.650, p = 0.517, d = -0.116, \mathrm{CI}
+= [-2.680, 1.249]$; location: $t(125) = -0.251, p = 0.802, d = -0.045,
+\mathrm{CI} = [-2.257, 1.524]$) clustering. Similarly, participants in the
+lexicographically ordered conditions exhibited stronger (relative to feature
+rich participants) lexicographic clustering (length: $t(125) = 3.682, p <
+0.001, d = 0.655, \mathrm{CI} = [1.890, 5.569]$; first letter: $t(125) = 5.134,
+p < 0.001, d = 0.912, \mathrm{CI} = [3.251, 7.258]$) on early lists, but showed
+no reliable differences in semantic (category: $t(125) = -1.040, p = 0.301, d =
+-0.185, \mathrm{CI} = [-3.095, 1.092]$; size: $t(125) = 0.006, p = 0.995, d =
+0.001, \mathrm{CI} = [-1.933, 1.952]$) or visual (color: $t(125) = 0.092, p =
+0.927, d = 0.016, \mathrm{CI} = [-1.834, 1.867]$; location: $t(125) = 0.407, p
+= 0.685, d = 0.072, \mathrm{CI} = [-1.655, 2.463]$) clustering. And
+participants in the visually ordered conditions exhibited stronger visual
+clustering (again, relative to feature rich participants, and on early lists;
+color: $t(126) = 2.022, p = 0.045, d = 0.358, \mathrm{CI} = [0.056, 3.965]$;
+location: $t(126) = 4.390, p < 0.001, d = 0.777, \mathrm{CI} = [2.730, 6.199]
+$), but showed no reliable differences in semantic (category: $t(126) = 0.012,
+p = 0.991, d = 0.002, \mathrm{CI} = [-1.988, 1.871]$; size: $t(126) = -0.104, p
+= 0.917, d = -0.018, \mathrm{CI} = [-2.166, 1.847] $) or lexicographic (length:
+$t(126) = 0.592, p = 0.555, d = 0.105, \mathrm{CI} = [-1.361, 2.420]$; first
+letter: $t(126) = 0.040, p = 0.968, d = 0.007, \mathrm{CI} = [-1.791, 1.863]$)
+clustering. Taken together, these order manipulation results suggest several
+broad patterns (Figs.~\ref{fig:recall-dynamics}A, \ref{fig:fingerprints}).
+First, most of the order manipulations we carried out did \textit{not} reliably
+affect overall recall performance. Second, most of the order manipulations
+increased participants' tendencies to temporally cluster their recalls. Third,
+all of the order manipulations enhanced participants' clustering of each
+condition's target feature (i.e., semantic manipulations enhanced semantic
+clustering, lexicographic manipulations enhanced lexicographic clustering, and
+visual manipulations enhanced visual clustering; Fig.~\ref{fig:barplots}C)
+while leaving clustering along other feature dimensions roughly unchanged
+(i.e., semantic manipulations did not affect lexicographic or visual
+clustering, and so on). Although it is not possible to fully separate feature
+versus temporal clustering when considering sorted lists, we used a
+permutation-based procedure to identify the degree of feature clustering over
+and above what could be accounted for by temporal clustering alone (see
+\textit{Factoring out the effects of temporal clustering}). When we carried out
+this analysis (Fig.~\ref{fig:barplots}D), we found that participants exhibited
+more semantic clustering on semantically sorted lists than on randomly ordered
+lists, but the effects of the other order manipulations could not reliably be
+separated from temporal clustering alone (reliable comparisons are reported in
+the figure).
 
@@
 lists that were sorted semantically better than lexicographically sorted lists
-($t(118) = 1.936, p = 0.055$). Participants also remembered visually sorted
-lists better than lexicographically sorted lists ($t(119) = 2.145, p = 0.034$).
-However, participants showed no reliable differences in recall for semantically
-versus visually sorted lists ($t(119) = 0.113, p = 0.910$). Participants
+($t(118) = 1.936, p = 0.055, d = 0.353, \mathrm{CI} = [0.057, 3.916]$).
+Participants also remembered visually sorted lists better than
+lexicographically sorted lists ($t(119) = 2.145, p = 0.034, d = 0.390,
+\mathrm{CI} = [0.208, 4.254]$). However, participants showed no reliable
+differences in recall for semantically versus visually sorted lists ($t(119) =
+0.113, p = 0.910, d = 0.021, \mathrm{CI} = [-1.987, 2.097]$). Participants
 temporally clustered semantically sorted lists more strongly than either
-lexicographically ($t(118) = 5.572, p < 0.001$) or visually ($t(119) = 6.215, p
-< 0.001$) sorted lists, but did not show reliable differences in temporal
-clustering on lexicographically versus visually sorted lists ($t(119) = 0.189,
-p = 0.850$). Participants also showed reliably more semantic clustering on
-semantically sorted lists than lexicographically (category: $t(118) = 3.492, p
-= 0.001$, size: $t(118) = 3.972, p < 0.001$) or visually (category: $t(119) =
-2.702, p = 0.008$, size: $t(119) = 4.230, p < 0.001$) sorted lists; more
-lexicographic clustering on lexicographically sorted lists than semantically
-(length: $t(118) = 3.112, p = 0.002$; first letter: $t(118) = 3.686, p <
-0.001$) or visually (length: $t(119) = 3.024, p = 0.003$; first letter: $t(119)
-= 2.644, p = 0.009$) sorted lists; and more visual clustering on visually
-sorted lists than semantically (color: $t(119) = -2.659, p = 0.009$; location:
-$t(119) = -4.604, p < 0.001$) or lexicographically (color: $t(119) = -2.366, p
-= 0.020$; location: $t(119) = -4.265, p < 0.001$) sorted lists. In summary,
-sorting lists by different features appeared to have slightly different effects
-on overall memory performance and temporal clustering.  Participants also tended to
-cluster their recalls along a given feature dimension more when the studied
-lists were (versus were not) sorted along that dimension.
+lexicographically ($t(118) = 5.620, p < 0.001, d = 1.026, \mathrm{CI} = [3.486,
+8.010]$) or visually ($t(119) = 6.613, p < 0.001, d = 1.202, \mathrm{CI} =
+[4.481, 9.464]$) sorted lists, but did not show reliable differences in
+temporal clustering on lexicographically versus visually sorted lists ($t(119)
+= 0.589, p = 0.557, d = 0.107, \mathrm{CI} = [-1.336, 2.539]$). Participants
+also showed reliably more semantic clustering on semantically sorted lists than
+lexicographically (category: $t(118) = 3.667, p < 0.001, d = 0.670, \mathrm{CI}
+= [1.822, 5.942]$, size: $t(118) = 3.972, p < 0.001$) or visually (category:
+$t(119) = 2.702, p = 0.008$, size: $t(118) = 4.043, p < 0.001, d = 0.738,
+\mathrm{CI} = [2.145, 6.296]$) sorted lists; more lexicographic clustering on
+lexicographically sorted lists than semantically (length: $t(118) = 3.390, p <
+0.001, d = 0.619, \mathrm{CI} = [1.499, 5.661]$; first letter: $t(118) = 5.705,
+p < 0.001, d = 1.042, \mathrm{CI} = [3.841, 7.790]$) or visually (length:
+$t(119) = 3.399, p < 0.001, d = 0.618, \mathrm{CI} = [1.500, 5.527]$; first
+letter: $t(119) = 4.859, p < 0.001, d = 0.883, \mathrm{CI} = [2.860, 6.849]$)
+sorted lists; and more visual clustering on visually sorted lists than
+semantically (color: $t(119) = 2.673, p = 0.009, d = 0.486, \mathrm{CI} =
+[0.848, 4.567]$; location: $t(119) = 4.499, p < 0.001, d = 0.818, \mathrm{CI} =
+[2.721, 6.399]$) or lexicographically (color: $t(119) = 1.988, p = 0.049, d =
+0.361, \mathrm{CI} = [0.102, 3.894]$; location: $t(119) = 3.966, p < 0.001, d =
+0.721, \mathrm{CI} = [2.099, 5.862]$) sorted lists. In summary, sorting lists
+by different features appeared to have slightly different effects on overall
+memory performance and temporal clustering. Participants also tended to cluster
+their recalls along a given feature dimension more when the studied lists were
+(versus were not) sorted along that dimension.
 
@@
 in all six order manipulation conditions showed no reliable differences
-(semantic: $t(125) = 0.487, p = 0.627$; lexicographic: $t(125) = 0.878, p =
-0.382$; visual: $t(126) = 1.437, p = 0.153$). Nor did we observe any reliable
-differences in temporal clustering on late lists (relative to late feature rich
-lists; semantic: $t(125) = 0.146, p = 0.884$; lexicographic: $t(125) = 0.923, p
-= 0.358$; visual: $t(126) = 0.525, p = 0.601$). Aside from a slightly increased
-tendency for participants to cluster words by their length on late visual order
-manipulation lists (more than late feature rich lists; $t(126) = 2.199, p =
-0.030$), we observed no reliable differences in any type of feature clustering
-on late order manipulation condition lists versus late feature rich lists
-($\|t\|$s $\leq 1.234, p$s $\geq 0.220$).
+(semantic: $t(125) = 0.487, p = 0.627, d = 0.087, \mathrm{CI} = [-1.661,
+2.323]$; lexicographic: $t(125) = 0.878, p = 0.382, d = 0.156, \mathrm{CI} =
+[-1.226, 3.044]$; visual: $t(126) = 1.437, p = 0.153, d = 0.254, \mathrm{CI} =
+[-0.447, 3.519]$). Nor did we observe any reliable differences in temporal
+clustering on late lists (relative to late feature rich lists; semantic:
+$t(125) = 0.157, p = 0.875, d = 0.028, \mathrm{CI} = [-1.859, 1.974]$;
+lexicographic: $t(125) = 0.998, p = 0.320, d = 0.177, \mathrm{CI} = [-0.902,
+2.920]$; visual: $t(126) = 0.548, p = 0.585, d = 0.097, \mathrm{CI} = [-1.450,
+2.365]$). Aside from a slightly increased tendency for participants to cluster
+words by their length on late visual order manipulation lists (more than late
+feature rich lists; $t(126) = 2.005, p = 0.047, d = 0.355, \mathrm{CI} =
+[0.211, 3.722]$), we observed no reliable differences in any type of feature
+clustering on late order manipulation condition lists versus late feature rich
+lists ($\|t\|$s $\leq 1.124, p$s $\geq 0.263$).
 
@@
 condition's manipulated feature) recalled more words. This trend held overall
-across conditions and participants (early: $r(179) = 0.537, p < 0.001$; late: $
-r(179) = 0.492, p < 0.001$) as well as for each condition individually for
-early ($r$s $\geq 0.386$, all $p$s $\leq 0.035$) and late ($r$s $\geq 0.462$,
-all $p$s $\leq 0.010$) lists. We found no evidence of a condition-level trend;
-for example, the conditions where participants tended to show stronger
-clustering scores were not correlated with the conditions where participants
-remembered more words (early: $r(4) = 0.526, p = 0.284$; late: $r(4) = -0.257,
-p = 0.623$; see insets of Fig.~\ref{fig:clustering-scatterplots}A and B). We
-observed carryover associations between feature clustering and recall
-performance (Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who
-showed stronger feature clustering on early lists tended to recall more items
-on late lists (across conditions: $r(179) = 0.492, p < 0.001$; all conditions
-individually: $r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who
-recalled more items on early lists also tended to show stronger feature
-clustering on late lists (across conditions: $r(179) = 0.280, p < 0.001$; all
-non-visual conditions: $r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29)
-= 0.298, p = 0.103$; location: $r(28) = 0.354, p = 0.055$). Neither of these
-effects showed condition-level trends (early feature clustering versus late
-recall probability: $r(4) = -0.299, p = 0.565$; early recall probability versus
-late feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for
-associations between feature clustering and temporal clustering. Across every
-order manipulation condition, participants who exhibited stronger feature
-clustering also exhibited stronger temporal clustering. For early lists
+across conditions and participants (early: $r(179) = 0.492, p < 0.001,
+\mathrm{CI} = [0.352, 0.606]$; late: $r(179) = 0.403, p < 0.001, \mathrm{CI} =
+[0.271, 0.517]$) as well as for each condition individually for early ($r$s
+$\geq 0.331$, all $p$s $\leq 0.069$) and late ($r$s $\geq 0.404$, all $p$s
+$\leq 0.027$) lists. We found no evidence of a condition-level trend; for
+example, the conditions where participants tended to show stronger clustering
+scores were not correlated with the conditions where participants remembered
+more words (early: $r(4) = 0.511, p = 0.300, \mathrm{CI} = [-0.999, 0.996]$;
+late: $r(4) = -0.304, p = 0.559, \mathrm{CI} = [-0.833, 0.748]$; see insets of
+Fig.~\ref{fig:clustering-scatterplots}A and B). We observed carryover
+associations between feature clustering and recall performance
+(Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who showed stronger
+feature clustering on early lists tended to recall more items on late lists
+(across conditions: $r(179) = 0.492, p < 0.001$; all conditions individually:
+$r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who recalled more items
+on early lists also tended to show stronger feature clustering on late lists
+(across conditions: $r(179) = 0.280, p < 0.001$; all non-visual conditions:
+$r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29) = 0.298, p = 0.103$;
+location: $r(28) = 0.354, p = 0.055$). Neither of these effects showed
+condition-level trends (early feature clustering versus late recall
+probability: $r(4) = -0.299, p = 0.565$; early recall probability versus late
+feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for associations
+between feature clustering and temporal clustering. Across every order
+manipulation condition, participants who exhibited stronger feature clustering
+also exhibited stronger temporal clustering. For early lists
 (Fig.~\ref{fig:clustering-scatterplots}E), this trend held overall ($r(179) =
@@
 feature clustering on early lists trended towards showing stronger temporal
-clustering on later lists (overall: $r(179) = 0.301, p < 0.001$; for individual
-conditions: all $r$s $\geq 0.297$, all $p$s $\leq 0.111$; across conditions: $
-r(4) = 0.107, p = 0.840$). And participants who showed stronger temporal
-clustering on early lists trended towards showing stronger feature clustering
-on later lists (overall: $r(179) = 0.579, p < 0.001$; all non-visual
-conditions: $r$s $\geq 0.323$, all $p$s $\leq 0.082$; visual conditions: $r$s
-$\geq 0.089$, all $p$s $\leq 0.632$; across conditions: $r(4) = 0.916, p =
-0.010$). Taken together, the results displayed in
-Figure~\ref{fig:clustering-scatterplots} show that participants who were more
-sensitive to the order manipulations (i.e., participants who showed stronger
-feature clustering for their condition's feature on early lists) remembered
-more words and showed stronger temporal clustering. These associations also
-appeared to carry over across lists, even when the items on later lists were
-presented in a random order.
+clustering on later lists (overall: $r(179) = 0.464, p < 0.001, \mathrm{CI} =
+[0.321, 0.582]$; for individual conditions: all $r$s $\geq 0.377$, all $p$s
+$\leq 0.040$; across conditions: $ r(4) = 0.451, p = 0.369, \mathrm{CI} =
+[-0.986, 0.998]$). And participants who showed stronger temporal clustering on
+early lists trended towards showing stronger feature clustering on later lists
+(overall: $r(179) = 0.266, p < 0.001, \mathrm{CI} = [0.129, 0.396]$; for
+individual conditions: all $r$s $\geq 0.298$, all $p$s $\leq 0.110$; across
+conditions: $r(4) = 0.064, p = 0.903, \mathrm{CI} = [-0.972,$). Taken together,
+the results displayed in Figure~\ref{fig:clustering-scatterplots} show that
+participants who were more sensitive to the order manipulations (i.e.,
+participants who showed stronger feature clustering for their condition's
+feature on early lists) remembered more words and showed stronger temporal
+clustering. These associations also appeared to carry over across lists, even
+when the items on later lists were presented in a random order.
 
@@
 feature clustering on late lists (Fig.~\ref{fig:clustering-carryover}A; overall
-across participants and conditions: $r(179) = 0.592, p < 0.001$; non-visual
-feature conditions: all $r$s $\geq 0.350$, all $p$s $\leq 0.058$; color: $r(29)
-= -0.071, p = 0.704$; location: $r(28) = 0.032, p = 0.868$; across conditions:
-$r(4) = 0.934, p = 0.006$). Although participants tended to show weaker feature
-clustering on late lists (Fig.~\ref{fig:fingerprint-trajectories}) on
-\textit{average}, the associations between early and late lists for individual
-participants suggests that some influence of early order manipulations may
-linger on late lists. We found that participants who exhibited larger carryover
-in feature clustering (i.e., continued to show strong feature clustering on
-late lists) for the semantic order manipulations (but not other manipulations)
-also tended to show a smaller decrease in recall on early versus late lists
-(Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.378, p < 0.001$;
-category: $r(28) = 0.419, p = 0.021$; size: $r(28) = 0.737, p < 0.001$;
-non-semantic conditions: all $r$s $\leq 0.252$, all $p$s $\geq 0.179$; across
-conditions: $r(4) = 0.773, p = 0.072$) on late lists, relative to early lists.
-Participants who exhibited larger carryover in feature clustering also tended
-to show stronger temporal clustering on late lists (relative to early lists)
-for all but the category condition (Fig.~\ref{fig:clustering-carryover}C;
-overall: $r(179) = 0.434, p < 0.001$; category: $r(28) = 0.229, p = 0.223$; all
-non-category conditions: all $r$s $\geq 0.448$, all $p$s $\leq 0.012$; across
-conditions: $r(4) = 0.598, p = 0.210$).
+across participants and conditions: $r(179) = 0.591, p < 0.001, \mathrm{CI} =
+[0.472, 0.682]$; category: $r(28) = 0.590, p < 0.001, \mathrm{CI} = [0.354,
+0.756]$; size: $r(28) = 0.488, p = 0.006, \mathrm{CI} = [0.134, 0.732]$;
+length: $r(28) = 0.384, p = 0.036, \mathrm{CI} = [0.040, 0.681]$; first letter:
+$r(28) = 0.202, p = 0.284, \mathrm{CI} = [-0.273, 0.620]$; color: $r(29) =
+-0.183, p = 0.325, \mathrm{CI} = [-0.562, 0.258]$; location: $r(28) = 0.031, p
+= 0.870, \mathrm{CI} = [-0.240, 0.296]$; across conditions: $r(4) = 0.942, p =
+0.005, \mathrm{CI} = [0.442, 1.000]$). Although participants tended to show
+weaker feature clustering on late lists
+(Fig.~\ref{fig:fingerprint-trajectories}) on \textit{average}, the associations
+between early and late lists for individual participants suggests that some
+influence of early order manipulations may linger on late lists. We found that
+participants who exhibited larger carryover in feature clustering (i.e.,
+continued to show strong feature clustering on late lists) for the semantic
+order manipulations (but not other manipulations) also tended to show a smaller
+decrease in recall on early versus late lists
+(Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.307, p < 0.001,
+\mathrm{CI} = [0.148, 0.469]$; category: $r(28) = 0.350, p = 0.058, \mathrm{CI}
+= [0.050, 0.642]$; size: $r(28) = 0.708, p < 0.001, \mathrm{CI} = [0.472,
+0.862]$; length: $r(28) = 0.205, p = 0.276, \mathrm{CI} = [-0.109, 0.492]$;
+first letter: $r(28) = 0.081, p = 0.672, \mathrm{CI} = [-0.433, 0.597]$; color:
+$r(29) = 0.155, p = 0.406, \mathrm{CI} = [-0.174, 0.541]$; location: $r(28) =
+0.052, p = 0.787, \mathrm{CI} = [-0.307, 0.360]$; across conditions: $r(4) =
+0.635, p = 0.176, \mathrm{CI} = [-0.924, 0.981]$. Participants who exhibited
+larger carryover in feature clustering also tended to show stronger temporal
+clustering on late lists (relative to early lists) for all but the category
+condition (Fig.~\ref{fig:clustering-carryover}C; overall: $r(179) = 0.434, p <
+0.001$; category: $r(28) = 0.229, p = 0.223$; all non-category conditions: all
+$r$s $\geq 0.448$, all $p$s $\leq 0.012$; across conditions: $r(4) = 0.598, p =
+0.210$).
 
@@
 Overall, we found that participants tended to be better at remembering words on
-stabilize lists relative to words on both random ($t(59) = 1.740, p = 0.087$)
-and destabilize ($t(59) = 1.714, p = 0.092$) lists (Fig.~\ref{fig:adaptive}A).
-Participants showed no reliable differences in their memory performance on
-destabilize versus random lists ($t(59) = -0.249, p = 0.804$). Participants
-also exhibited stronger temporal clustering on stabilize lists, relative to
-random ($t(59) = 3.554, p = 0.001$) and destabilize ($t(59) = 4.045, p <
-0.001$) lists (Fig.~\ref{fig:adaptive}B). We found no reliable differences in
+stabilize lists relative to words on both random ($t(59) = 1.740, p = 0.087, d
+= 0.095, \mathrm{CI} = [-0.187, 3.761]$) and destabilize ($t(59) = 1.714, p =
+0.092, d = 0.114, \mathrm{CI} = [-0.351, 4.108]$) lists
+(Fig.~\ref{fig:adaptive}A). Participants showed no reliable differences in
+their memory performance on destabilize versus random lists ($t(59) = -0.249, p
+= 0.804, d = -0.017, \mathrm{CI} = [-2.327, 1.578]$). Participants also
+exhibited stronger temporal clustering on stabilize lists, relative to random
+($t(59) = 3.428, p = 0.001, d = 0.306, \mathrm{CI} = [1.635, 5.460]$) and
+destabilize ($t(59) = 4.174, p < 0.001, d = 0.374, \mathrm{CI} = [1.964,
+6.968]$) lists (Fig.~\ref{fig:adaptive}B). We found no reliable differences in
 temporal clustering for items on random versus destabilize lists ($t(59) =
--0.781, p = 0.438$).
+-0.880, p = 0.382, d = -0.081, \mathrm{CI} = [-3.165, 1.127]$).
 
@@
 scores also tended to recall more items. This held across subjects, aggregating
-across all list types ($r(178) = 0.721, p < 0.001$), and for each list type
-individually (all $r$s $\geq 0.683$, all $p$s $\leq 0.001$). Taken together,
-the results from the adaptive condition suggest that each participant comes
-into the experiment with their own unique memory organization tendencies, as
-characterized by their memory fingerprint. When participants study lists whose
-items come pre-sorted according to their unique preferences, they tend to remember
-more and show stronger temporal clustering.
+across all list types ($r(178) = 0.701, p < 0.001, \mathrm{CI} = [0.590,
+0.789]$), and for each list type individually (all $r$s $\geq 0.651$, all $p$s
+$< 0.001$). Taken together, the results from the adaptive condition suggest
+that each participant comes into the experiment with their own unique memory
+organization tendencies, as characterized by their memory fingerprint. When
+participants study lists whose items come pre-sorted according to their unique
+preferences, they tend to remember more and show stronger temporal clustering.
 

### 2023-11-14 code/analyze.py

@@ -662,7 +662,7 @@ def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_
     else:
         p_string = f'p = {p:.3f}'
 
-    print(f't({df}) = {t:.3f}, {p_string}, d = {d:.3f}, CI = [{low_conf:.3f}, {high_conf:.3f}]')
+    print(f't({df}) = {t:.3f}, {p_string}, d = {d:.3f}, ' + '\mathrm{CI}' + f' = [{low_conf:.3f}, {high_conf:.3f}]')
     
 
 def stack_fried_eggs(*args):

### 2023-11-14 paper/main.tex
added discussion of "tension" between different types of clustering

@@
 the proportion of trials on which each item (as a function of its presentation
-position) was recalled at output position $n$~\citep{HowaKaha99, PolyEtal09,
-ZhanEtal23}. To carry out this analysis, we initialized (for each participant)
-a number-of-lists (16) by number-of-words-per-list (16) matrix of 0s. Then, for
-each list, we found the index of the word that was recalled first, and we
-filled in that position in the matrix with a 1. Finally, we averaged over the
-rows of the matrix to obtain a 1 by 16 array of probabilities, for each
-participant. We used an analogous procedure to compute probability of
-$n$\textsuperscript{th} recall curves for each participant. Specifically, we
-filled in the corresponding matrices according to the $n$\textsuperscript{th}
-recall on each list that each participant made. When a given participant had
-made fewer than $n$ recalls for a given list, we simply excluded that list from
-our analysis when computing that participant's curve(s). The probability of
-first recall curve corresponds to a special case where $n = 1$.
+position) was recalled at output position $n$~\citep{Hoga75, HowaKaha99,
+PolyEtal09, ZhanEtal23}. To carry out this analysis, we initialized (for each
+participant) a number-of-lists (16) by number-of-words-per-list (16) matrix of
+0s. Then, for each list, we found the index of the word that was recalled
+first, and we filled in that position in the matrix with a 1. Finally, we
+averaged over the rows of the matrix to obtain a 1 by 16 array of
+probabilities, for each participant. We used an analogous procedure to compute
+probability of $n$\textsuperscript{th} recall curves for each participant.
+Specifically, we filled in the corresponding matrices according to the
+$n$\textsuperscript{th} recall on each list that each participant made. When a
+given participant had made fewer than $n$ recalls for a given list, we simply
+excluded that list from our analysis when computing that participant's
+curve(s). The probability of first recall curve corresponds to a special case
+where $n = 1$.
 
@@
 
+\subsection*{How do different types of clustering relate to each other, and to memory performance?}
+
+When the words on a studied list are presented in a random order, different types
+of clustering in participants' recalls often tend to be negatively correlated.
+For example, words that occur nearby on the list will not (on average)
+tend to be semantically related, and vice versa. Therefore a participant who
+shows a strong tendency to temporally cluster their recalls will tend to show weaker
+semantic clustering, and so on~\citep{HealUitv19, HowaKaha02b, SedeEtal10}.  Further,
+there is some evidence that temporal clustering is positively correlated with
+memory performance, whereas semantic clustering is negatively correlated with
+memory performance~\citep{SedeEtal10}.
+
+The notion of ``multiple pathways to recall'' discussed above (see
+\textit{Memory consequences of feature variability}) suggests one potential
+explanation for these patterns. For example, temporal clustering has been
+proposed to reflect reliance on contextual cues in an ``episodic'' pathway to
+search memory, whereas semantic clustering reflects a relies on specific item
+features. These two pathways may ``compete'' with each other during
+recall~\citep{SochEtal09}. Meanwhile, extra-list intrusion errors (i.e., false
+``recalls'' of items that were never encountered on the list) often tend to
+share semantic features with recently recalled items~\citep{ZaroEtal06} and
+also often lead the participant to stop recalling additional
+items~\citep{MillEtal12a}. Speculatively, over-reliance on semantic cues may lead to
+more intrusion errors, which in turn may lead to fewer recalls overall.
+
+Our findings extend these prior results to consider lists that are \textit{not}
+ordered randomly. Because ordering the words on a list along a particular
+feature dimension removes the ``conflict'' between temporal and feature
+clustering, the order manipulation conditions in our study represent an ``edge
+case'' whereby different pathways to recall are not necessarily in conflict
+with each other. For example, the same participants who exhibit strong feature
+clustering \textit{also} show strong temporal clustering on ordered lists
+(Fig.~\ref{fig:clustering-scatterplots}E). This is presumably at least partly
+due to an inability to separate temporal and feature clustering on ordered
+lists (also see \textit{Factoring out the effects of temporal clustering}).
+However, features that change gradually with time (i.e., presentation position)
+could also serve to strengthen the episodic (contextual) cues associated with
+each item. In other words, participants might essentially combine multiple
+noisy measures of change to form a more stable internal representation of
+temporal context.
+
 % implications for:

### 2023-11-15 paper/main.tex
complete draft incorporating edits to address all reviewer concerns (still needs final round of proofreading)

@@
 \newcommand{\fingerprintTrajectoryRandom}{S8}
+\newcommand{\featureCorrs}{S9}
 
@@
 lists that vary (meaningfully) along just a few feature dimensions, are there
-important differences in which dimenesions influence participants' memories? Or
+important differences in which dimensions influence participants' memories? Or
 are all features essentially ``equally'' influential? Further, are there
@@
 words referred to objects that tended to be a particular size, which meant that
-category and size were not fully independent. On late lists, the words were
-always presented in a randomized order (chosen anew for each participant). In
-all of the order manipulation conditions, we varied words' font colors and
-onscreen locations, as in the feature rich condition.
+category and size were not fully independent (Fig.~\featureCorrs). On late
+lists, the words were always presented in a randomized order (chosen anew for
+each participant). In all of the order manipulation conditions, we varied
+words' font colors and onscreen locations, as in the feature rich condition.
 
@@
 fingerprints (for all lists and experimental conditions) to create a large
-matrix with number-of-lists (16) $\times$ number-of-conditions (10, encluding
+matrix with number-of-lists (16) $\times$ number-of-conditions (10, including
 the adaptive condition) rows and seven columns (one for each feature clustering
@@
 their clustering scores).
+
+A core assumption of our approach is that each participant organizes
+their memories in a unique way. We defined each participant's \textit{memory
+fingerprint} as the set of their permutation-corrected clustering scores across
+all dimensions we tracked in our study, including their six feature-based
+clustering scores (category, size, length, first letter, color, and location)
+and their temporal clustering score. Conceptually, a participant's memory
+fingerprint describes their tendency to order in their recall sequences (and,
+presumably, organize in memory) the studied words along each dimension. If
+these memory fingerprints are truly unique to each participant, then we would
+expect that the estimated fingerprints computed for a given participant, on
+different lists, should be more similar than the estimated fingerprints computed
+for different participants. We reasoned that the feature rich condition would
+provide the best opportunity to test this assumption, since the clustering
+scores would not be potentially confounded by order manipulations. To test our
+``unique memory fingerprint'' assumption, we compared the similarity
+(correlation) between the fingerprint from a single list (from one participant)
+and (a) the average fingerprint from all other lists from the same participant
+versus (b) the average fingerprint from each other participant (across all of
+their lists). We found that participants' fingerprints on a held-out list are
+reliably more similar to the same participant's fingerprints on other lists
+than to other participants' fingerprints ($t(70280) = 5.077, p < 0.001, d =
+0.162, \mathrm{CI} = [3.086, 6.895]$). This suggests that participants'
+fingerprints are stable across lists, and that each participant's fingerprint
+is unique to them.
 
@@
 See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
-random and adaptive conditions. \textbf{B.} Proportion of event boundaries (see
-\textit{Identifying event boundaries}) for each condition's feature of focus,
-plotted as a function of presentation position.}
+random and adaptive conditions. \textbf{B.} Feature distances ($z$-scored
+within condition) between the features of successively presented words (see
+\textit{Defining feature-based distances}), for each condition's feature of
+focus, plotted as a function of presentation position.\textbf{C.} Proportion of
+event boundaries (see \textit{Identifying event boundaries}) for each
+condition's feature of focus, plotted as a function of presentation position.}
 
@@
 sorting early lists by one feature dimension could affect how participants
-cluster \textit{other} features (on early and/or late lists) as well. We
-defined participants' \textit{memory fingerprints} as the set of their temporal
-and feature clustering scores (see \textit{Memory fingerprints}). A
-participant's memory fingerprint describes how they tend to retrieve memories
-of the studied items, perhaps searching in parallel through several feature
-spaces (or along several representational dimensions). To gain insights into
-the dynamics of how participants' clustering scores tended to change over time,
-we computed the average (across participants) fingerprint from each list, from
-each order manipulation condition (Fig.~\ref{fig:fingerprint-trajectories}). We
-projected these fingerprints into a two-dimensional space to help visualize the
-dynamics (top panels; see \textit{Computing low-dimensional embeddings of
-memory fingerprints}). We found that participants' average fingerprints tended
-to remain relatively stable on early lists, and exhibited a ``jump'' to another
-stable state on later lists. The sizes of these jumps varied somewhat across
-conditions (the Euclidean distances between fingerprints in their original high
-dimensional spaces are displayed in the bottom panels). We also averaged the
-fingerprints across early and late lists, respectively, for each condition
+cluster \textit{other} features (on early and/or late lists) as well. As
+described above, a participant's memory fingerprint describes how they tend to
+retrieve memories of the studied items, perhaps searching in parallel through
+several feature spaces (or along several representational dimensions). To gain
+insights into the dynamics of how participants' clustering scores tended to
+change over time, we computed the average (across participants) fingerprint
+from each list, from each order manipulation condition
+(Fig.~\ref{fig:fingerprint-trajectories}). We projected these fingerprints into
+a two-dimensional space to help visualize the dynamics (top panels; see
+\textit{Computing low-dimensional embeddings of memory fingerprints}). We found
+that participants' average fingerprints tended to remain relatively stable on
+early lists, and exhibited a ``jump'' to another stable state on later lists.
+The sizes of these jumps varied somewhat across conditions (the Euclidean
+distances between fingerprints in their original high dimensional spaces are
+displayed in the bottom panels). We also averaged the fingerprints across early
+and late lists, respectively, for each condition
 (Fig.~\ref{fig:fingerprint-trajectories}B). We found that participants'
@@
 
+We note that the multivariate aspect of the adaptive condition (i.e., sorting
+lists simultaneously along multiple feature dimensions) provides an important
+contrast with the order order manipulation conditions, where we sort lists
+along only a single feature dimension in each condition. We found that
+participants ``naturally'' clustered their recalls along multiple feature
+dimensions, even when the lists they studied were not sorted along those
+dimensions (as in the feature rich condition). A caveat is that the
+\textit{specific} feature dimensions participants tended to cluster along
+varied across participants. One way to quantify the multidimensional nature of
+participants' clustering tendencies is to sort each partipant's clustering
+scores (for each of the six feature dimensions, along with a seventh dimension
+to capture temporal clustering). We can then ask whether the distribution of
+clustering scores at each ``rank'' within the sorted set of scores for each
+participant has a mean that is reliably different from a chance value of 0.5.
+We carried out these tests for each set of ranked scores, and found that
+participants in the feature rich condition reliably cluster their recalls along
+at least three dimensions, including temporal clustering (which was often
+ranked highest); Rank 1: $t(66) = 12.751, p < 0.001, d = 0.162, \mathrm{CI} =
+[8.702, 20.013]$; Rank 2: $t(66) = 8.196, p < 0.001, d = 0.162, \mathrm{CI} =
+[4.794, 12.978]$; Rank 3: $t(66) = 3.243, p = 0.002, d = 0.162, \mathrm{CI} =
+[1.028, 7.051]$; Rank 4: $t(66) = -3.112, p = 0.003, d = 0.162, \mathrm{CI} =
+[-5.282, -1.920]$; Rank 5: $t(66) = -7.154, p < 0.001, d = 0.162, \mathrm{CI} =
+[-12.649, -5.568]$; Rank 6: $t(66) = -12.608, p < 0.001, d = 0.162, \mathrm{CI}
+= [-22.114, -9.347]$; Rank 7: $t(66) = -18.397, p < 0.001, d = 0.162,
+\mathrm{CI} = [-27.238, -14.073]$.
+
 \section*{Discussion}
@@
 effects may also be modulated by factors like list length and the numbers of
-examplars in each cateogry~\citep[e.g.,][]{BorgMang72}.
+exemplars in each category~\citep[e.g.,][]{BorgMang72}.
 

### 2023-11-15 paper/supplement.tex
complete draft incorporating edits to address all reviewer concerns (still needs final round of proofreading)

@@
 
+    \begin{figure}[tp] \centering
+        \includegraphics[width=0.6\textwidth]{figures/feature_correlations}
+
+        \caption{\textbf{Correlations between features.} Within each list, for each participant, we
+        computed the set of distances between each pair of words, along each feature dimension.  We
+        then combined these pairwise distances across all lists and participants, and computed the
+        Spearman correlation between the distances for each pair of feature dimensions.  The correlation
+        coefficients are displayed in the heatmap, annotated with each cell's value.} \label{fig:feature-correlations}
+
+        \end{figure}
+
 % accuracy by list (list order effects)

### 2023-11-15 paper/main.tex
another pass through main text + supplement, recompiled changes doc
-- sending this version to co-authors for comments...

@@
 this correction procedure to all of the clustering scores (feature and
-temporal) reported in this paper.
+temporal) reported in this paper. In Figure~\clusterCorrs~we report how
+participants' clustering scores along different feature dimensions (in the
+order manipulation conditions) are correlated, and how clustering scores change
+across lists.
 
@@
 randomly chosen amount (between 1 and the list length) to obtain a new set of
-items. Since the new set of items will have the same (average) temporal
-distances between successive recalls, the temporal clustering score for the new
-set of items is equal (on average) to the temporal clustering score for the
-original recalls. However, we can then re-compute the feature clustering score
-for those new items. Finally, we can compute a ``temporally corrected'' feature
-clustering score by computing the average percentile rank of the observed (raw)
-feature clustering score within the distributions of circularly shifted feature
-clustering scores, across $N = 500$ repetitions of this procedure. This new
-temporally corrected score provides an estimate of the observed degree of
-feature clustering over and above what could be accounted for by temporal
-clustering alone.
+items at the (now altered) positions of the original recalls. Since the new
+set of items will have the same (average) temporal distances between successive
+recalls, the temporal clustering score for the new set of items is equal (on
+average) to the temporal clustering score for the original recalls. However, we
+can then re-compute the feature clustering score for those new items. Finally,
+we can compute a ``temporally corrected'' feature clustering score by computing
+the average percentile rank of the observed (raw) feature clustering score
+within the distributions of circularly shifted feature clustering scores,
+across $N = 500$ repetitions of this procedure. This new temporally corrected
+score provides an estimate of the observed degree of feature clustering over
+and above what could be accounted for by temporal clustering alone.
 
@@
 
-A core assumption of our approach is that each participant organizes
-their memories in a unique way. We defined each participant's \textit{memory
+A core assumption of our approach is that each participant organizes their
+memories in a unique way. We defined each participant's \textit{memory
 fingerprint} as the set of their permutation-corrected clustering scores across
@@
 and their temporal clustering score. Conceptually, a participant's memory
-fingerprint describes their tendency to order in their recall sequences (and,
-presumably, organize in memory) the studied words along each dimension. If
+fingerprint describes their tendency to order, in their recall sequences (and,
+presumably, organize in memory), the studied words along each dimension. If
 these memory fingerprints are truly unique to each participant, then we would
 expect that the estimated fingerprints computed for a given participant, on
-different lists, should be more similar than the estimated fingerprints computed
-for different participants. We reasoned that the feature rich condition would
-provide the best opportunity to test this assumption, since the clustering
-scores would not be potentially confounded by order manipulations. To test our
-``unique memory fingerprint'' assumption, we compared the similarity
-(correlation) between the fingerprint from a single list (from one participant)
-and (a) the average fingerprint from all other lists from the same participant
-versus (b) the average fingerprint from each other participant (across all of
-their lists). We found that participants' fingerprints on a held-out list are
+different lists, should be more similar than the estimated fingerprints
+computed for different participants. We reasoned that the feature rich
+condition would provide the best opportunity to test this assumption, since the
+clustering scores would not be potentially confounded by order manipulations.
+To test our ``unique memory fingerprint'' assumption, we compared the
+similarity (correlation) between the fingerprint from a single list (from one
+participant) and (a) the average fingerprint from all other lists from the same
+participant versus (b) the average fingerprint from each other participant
+(across all of their lists). Repeating this procedure for all lists and
+participants, we found that participants' fingerprints on a held-out list are
 reliably more similar to the same participant's fingerprints on other lists
@@
 
-We also wondered whether adding these incidental visual features to later lists
+We next asked whether adding these incidental visual features to later lists
 (after the participants had already studied impoverished lists), or removing
@@
 $t(66) = -0.100, p = 0.921, d = -0.010, \mathrm{CI} = [-2.217, 1.899]$; feature
-rich, first letter: $t(66) = -0.412, p = 0.681, d = -0.045, \mathrm{CI} =
-[-2.461, 1.645]$; reduced, word length: $t(60) = 3.762, p < 0.001, d = 0.261,
+rich, first letter: $t(66) = 0.412, p = 0.681, d = 0.045, \mathrm{CI} =
+[-1.645, 2.461]$; reduced, word length: $t(60) = 3.762, p < 0.001, d = 0.261,
 \mathrm{CI} = [1.604, 6.821]$; reduced, first letter: $t(60) = 1.721, p =
@@
 0.781, p = 0.439, d = 0.101, \mathrm{CI} = [-1.039, 2.881]$; reduced (late),
-category: $t(40) = -0.101, p = 0.920, d = -0.009, \mathrm{CI} = [-2.307,
-1.776]$; reduced (late), size: $t(40) = 0.555, p = 0.582, d = 0.058,
-\mathrm{CI} = [-1.444, 2.274] $; reduced (late), length: $t(40) = 1.482, p =
-0.146, d = 0.126, \mathrm{CI} = [-0.444, 3.743]$; reduced (late), first letter:
-$t(40) = -0.143, p = 0.887, d = -0.017, \mathrm{CI} = [-2.204, 1.830] $). We
+category: $t(40) = 0.101, p = 0.920, d = 0.009, \mathrm{CI} = [-1.776, 2.307]$;
+reduced (late), size: $t(40) = 0.555, p = 0.582, d = 0.058, \mathrm{CI} =
+[-1.444, 2.274] $; reduced (late), length: $t(40) = 1.482, p = 0.146, d =
+0.126, \mathrm{CI} = [-0.444, 3.743]$; reduced (late), first letter: $t(40) =
+-0.143, p = 0.887, d = -0.017, \mathrm{CI} = [-2.204, 1.830] $). We
 hypothesized that adding or removing the variability in the visual features was
@@
 early lists in all of the other order manipulation conditions were
-indistinguishable from performance on the early feature rich lists ($\|t\|$s
+indistinguishable from performance on the early feature rich lists ($|t|$s
 $\leq 1.013, p$s $\geq 0.314$). Participants in both of the semantically
@@
 positions of feature changes in the study sequence
-(Fig.~\ref{fig:recall-dynamics}B; see \textit{Identifying event boundaries})
+(Fig.~\ref{fig:recall-dynamics}C; see \textit{Identifying event boundaries})
 with the positions of items participants recalled first, we noticed a striking
@@
 lists, the category labels changed every four items on average (dark purple
-peaks in Fig.~\ref{fig:recall-dynamics}B), and participants also seemed to
+peaks in Figs.~\ref{fig:recall-dynamics}B, C), and participants also seemed to
 display an increased tendency (relative to other order manipulation and random
@@
 the size labels changed every eight items on average (light purple peaks in
-Fig.~\ref{fig:recall-dynamics}B), and participants also seemed to display an
+Figs.~\ref{fig:recall-dynamics}B, C), and participants also seemed to display an
 increased tendency to initiate recall of size-ordered lists with items whose
@@
 clustering on late order manipulation condition lists versus late feature rich
-lists ($\|t\|$s $\leq 1.124, p$s $\geq 0.263$).
+lists ($|t|$s $\leq 1.124, p$s $\geq 0.263$).
 
@@
 cluster \textit{other} features (on early and/or late lists) as well. As
-described above, a participant's memory fingerprint describes how they tend to
-retrieve memories of the studied items, perhaps searching in parallel through
-several feature spaces (or along several representational dimensions). To gain
-insights into the dynamics of how participants' clustering scores tended to
-change over time, we computed the average (across participants) fingerprint
-from each list, from each order manipulation condition
+described above, a participant's memory fingerprint characterizes how they tend
+to retrieve memories of the studied items, perhaps searching in parallel
+through several feature spaces (or along several representational dimensions).
+To gain insights into the dynamics of how participants' clustering scores
+tended to change over time, we computed the average (across participants)
+fingerprint from each list, from each order manipulation condition
 (Fig.~\ref{fig:fingerprint-trajectories}). We projected these fingerprints into
@@
     the bottom panels display the Euclidean distances of the per-list memory
-    fingerprints to the list 0 fingerprint, for each condition. Error bars
-    denote bootstrap-estimated 95\% confidence intervals. The dotted vertical
-    lines denote the boundaries between early and late lists. \textbf{B.} In
-    this panel, the fingerprints for early (circle) and late (star) lists are
-    averaged across lists and participants before projecting the fingerprints
-    into a (new) 2D space. See Figure~\fingerprintTrajectoryRandom~for
-    analogous plots for the random conditions. }
-    \label{fig:fingerprint-trajectories}
+    fingerprints to the average fingerprint across all prior lists, for each
+    condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
+    The dotted vertical lines denote the boundaries between early and late
+    lists. \textbf{B.} In this panel, the fingerprints for early (circle) and
+    late (star) lists are averaged across lists and participants before
+    projecting the fingerprints into a (new) 2D space. See
+    Figure~\fingerprintTrajectoryRandom~for analogous plots for the random
+    conditions. } \label{fig:fingerprint-trajectories}
 
@@
 (Fig.~\ref{fig:clustering-scatterplots}C, D). Participants who showed stronger
-feature clustering on early lists tended to recall more items on late lists
-(across conditions: $r(179) = 0.492, p < 0.001$; all conditions individually:
-$r$s $\geq 0.462$, all $p$s $\leq 0.010$). Participants who recalled more items
-on early lists also tended to show stronger feature clustering on late lists
-(across conditions: $r(179) = 0.280, p < 0.001$; all non-visual conditions:
-$r$s $\geq 0.445$, all $p$s $\leq 0.014$; color: $r(29) = 0.298, p = 0.103$;
-location: $r(28) = 0.354, p = 0.055$). Neither of these effects showed
-condition-level trends (early feature clustering versus late recall
-probability: $r(4) = -0.299, p = 0.565$; early recall probability versus late
-feature clustering: $r(4) = 0.400, p = 0.432$). We also looked for associations
-between feature clustering and temporal clustering. Across every order
-manipulation condition, participants who exhibited stronger feature clustering
-also exhibited stronger temporal clustering. For early lists
+feature clustering on early lists in the non-visual conditions tended to recall
+more items on late lists (across conditions: $r(179) = 0.230, p = 0.002,
+\mathrm{CI} = [0.072, 0.372]$; all non-visual conditions individually: $r$s
+$\geq 0.405$, all $p$s $\leq 0.027$; color: $r(29) = 0.212, p = 0.251,
+\mathrm{CI} = [-0.164, 0.532]$; location: $r(28) = 0.320, p = 0.085,
+\mathrm{CI} = [0.011, 0.584]$). Participants who recalled more items on early
+lists also tended to show stronger feature clustering on late lists (across
+conditions: $r(179) = 0.464, p < 0.001, \mathrm{CI} = [0.321, 0.582]$;
+individual conditions: all $r$s $\geq 0.377$, all $p$s $\leq 0.040$). Neither
+of these effects showed condition-level trends (early feature clustering versus
+late recall probability: $r(4) = -0.338, p = 0.512, \mathrm{CI} = [-0.971,
+0.634]$; early recall probability versus late feature clustering: $r(4) =
+0.451, p = 0.369, \mathrm{CI} = [-0.986, 0.998]$). We also looked for
+associations between feature clustering and temporal clustering. Across every
+order manipulation condition, participants who exhibited stronger feature
+clustering also exhibited stronger temporal clustering. For early lists
 (Fig.~\ref{fig:clustering-scatterplots}E), this trend held overall ($r(179) =
-0.924, p < 0.001$), for each condition individually (all $r$s $\geq 0.822$, all
-$p$s $< 0.001$), and across conditions ($r(4) = 0.964, p = 0.002$). For late
-lists (Fig.~\ref{fig:clustering-scatterplots}F), the results were more variable
+0.916, p < 0.001, \mathrm{CI} = [0.893, 0.936]$), for each condition
+individually (all $r$s $\geq 0.822$, all $p$s $< 0.001$), and across conditions
+($r(4) = 0.964, p = 0.002$). For late lists
+(Fig.~\ref{fig:clustering-scatterplots}F), the results were more variable
 (overall: $r(179) = 0.348, p < 0.001$; all non-visual conditions: $r$s $\geq
@@
 (Fig.~\ref{fig:clustering-scatterplots}G, H). Participants who showed stronger
-feature clustering on early lists trended towards showing stronger temporal
+feature clustering on early lists showed stronger temporal
 clustering on later lists (overall: $r(179) = 0.464, p < 0.001, \mathrm{CI} =
@@
 clustering on late lists (relative to early lists) for all but the category
-condition (Fig.~\ref{fig:clustering-carryover}C; overall: $r(179) = 0.434, p <
-0.001$; category: $r(28) = 0.229, p = 0.223$; all non-category conditions: all
-$r$s $\geq 0.448$, all $p$s $\leq 0.012$; across conditions: $r(4) = 0.598, p =
-0.210$).
+condition (Fig.~\ref{fig:clustering-carryover}C; overall: $r(179) = 0.426, p <
+0.001, \mathrm{CI} = [0.285, 0.544]$; category: $r(28) = 0.110, p = 0.564,
+\mathrm{CI} = [-0.284, 0.442]$; all non-category conditions: all $r$s $\geq
+0.406$, all $p$s $\leq 0.023$; across conditions: $r(4) = 0.649, p = 0.163,
+\mathrm{CI} = [-0.856, 0.988]$).
 

### 2023-11-15 paper/supplement.tex
another pass through main text + supplement, recompiled changes doc
-- sending this version to co-authors for comments...

@@
     space. The bar plots in the bottom panels display the Euclidean distances
-    of the per-list memory fingerprints to the list 0 fingerprint, for each
-    condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
-    The dotted vertical lines denote the boundaries between early and late
-    lists. \textbf{B.} In this panel, the fingerprints for early (circle) and
-    late (star) lists are averaged across lists and participants before
-    projecting the fingerprints into a (new) 2D space.} \label{fig:fingerprint-trajectories-random}
+    of the per-list memory fingerprints to the average fingerprint across all
+    prior lists, for each condition. Error bars denote bootstrap-estimated 95\%
+    confidence intervals. The dotted vertical lines denote the boundaries
+    between early and late lists. \textbf{B.} In this panel, the fingerprints
+    for early (circle) and late (star) lists are averaged across lists and
+    participants before projecting the fingerprints into a (new) 2D space.}
+    \label{fig:fingerprint-trajectories-random}
 

### 2023-11-17 paper/main.tex
edits through introduction

@@
 2}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili
-Interactive\\\textsuperscript{*}Corresponding author:
+Interactive Labs\\\textsuperscript{*}Corresponding author:
 jeremy.r.manning@dartmouth.edu}
@@
 
-\begin{abstract} \footnotesize{We perceive, interpret, and remember ongoing
+\begin{abstract} \footnotesize{We perceive, interpret, and remember our ongoing
 experiences through the lens of our prior experiences. Inferring that we are in
@@
 Experience is subjective: different people who encounter identical physical
-experiences can take away very different meanings and memories. One reason is
+experiences can take away very different meanings and memories. One reason for this is
 that our moment-by-moment subjective experiences are shaped in part by the
@@
 by more traditional paradigms like free recall of random word
-lists~\citep{Kaha12, Kaha20}. In free recall, participants study lists of items
+lists~\citep{Kaha12, Kaha20}. In free recall paradigms, participants study lists of items
 and are instructed to recall the items in any order they choose. The orders in
@@
 organized their memories of the studied words. Because random word lists are
-unstructured by design, it is not clear if, or how, non-trivial situation
+unstructured by design, it is not clear if or how non-trivial situation
 models might apply to these stimuli. As we unpack below, this provides an
@@
 special context representation that is associated with each list. According to
-early theories~\citep[e.g.][]{Este55a,AndeBowe72} context representations are
+early theories~\citep[e.g.,][]{Este55a,AndeBowe72} context representations are
 composed of many features which fluctuate from moment to moment, slowly
@@
 phenomena include the well known recency and primacy effects (superior recall
-of items from the end and, to a lesser extent, from the beginning of the study
+of items from the end and, to a lesser extent, from the beginning of the studied
 list), as well as semantic and temporal clustering effects~\citep{KahaEtal08b,
@@
 variety of stimulus dimensions. According to models like the
-\textit{Context Maintenance and Retrieval Model}~\citep{PolyEtal09}, the
-stimulus features associated with each word (e.g.\ the word's meaning, size of
-the object the word represents, the letters that make up the word, font size,
-font color, location on the screen, etc.) are incorporated into the
+\textit{Context Maintenance and Retrieval} model~\citep{PolyEtal09}, the
+stimulus features associated with each word (e.g.,\ the word's meaning, size of
+the object the word represents, letters that make up the word, font size,
+font color, location on the screen, etc.)~are incorporated into the
 participant's mental context representation~\citep{SmitVela01, MannEtal11,
 MannEtal12, MannEtal15, Mann20}. During a memory test, any of these features
-may serve as a memory cue, which in turn leads the participant to recall in
-succession words that share stimulus features.
+may serve as a memory cue, which in turn leads the participant to successively recall
+words that share stimulus features.
 
@@
 lists. Often, lists in free recall studies are explicitly \textit{designed} to
-be devoid of exploitable temporal structure, for example, by sorting the words
+be devoid of exploitable temporal structure, for example by sorting the words
 in a random order~\citep{Kaha12}.
@@
 across different ``situations,'' analogous to how the content of real-world
-experiences change across different real-world situations. We asked
+experiences changes across different real-world situations. We asked
 participants to study and freely recall a series of word lists
@@
 (font \textit{color} and the onscreen \textit{location} of each word). We used
-two control conditions as a baseline; in these control conditions all of the
+two control conditions as a baseline; in these control conditions, all of the
 lists were sorted randomly, but we manipulated the presence or absence of the
@@
 sorted by a target feature (e.g., word category), and then study and recall an
-additional eight lists whose items had the same features, but that were sorted
+additional eight lists whose items had the same features but were sorted
 in a random temporal order. We were interested in how these manipulations
-affected participants' recall behaviors on early (manipulated) lists, as well
-as how order manipulations on early lists affected recall behaviors on later
+might affect participants' recall behaviors on early (manipulated) lists, as well
+as how order manipulations on early lists would affect recall behaviors on later
 (randomly ordered) lists. Finally, in an \textit{adaptive} experimental
-condition we used participants' recall behaviors on early lists to manipulate,
-in real-time, the presentation orders of subsequent lists. In this adaptive
-condition, we varied the agreement between how participants preferred to
-organize their memories of the studied items versus the orders in which the
-items were presented.
+condition, we used participants' recall behaviors on prior lists to manipulate,
+in real time, the presentation orders of subsequent lists. In this adaptive
+condition, we varied whether the order in which items were presented agreed 
+or disagreed with how each participant preferred to organize their memories of 
+the studied items.
 
@@
 From a theoretical perspective, we are interested in several core questions
-organized around the central theme of how structure in our experiences affect
-how we remember \textit{those} experiences, and also how we remember
+organized around the central theme of how structure in our experiences affects
+how we remember those experiences, as well as how we remember
 \textit{future} experiences (which may or may not exhibit similar structure).
@@
 lists that vary (meaningfully) along just a few feature dimensions, are there
-important differences in which dimensions influence participants' memories? Or
+important differences in these dimensions' influence on participants' memories? Or
 are all features essentially ``equally'' influential? Further, are there
@@
 ongoing versus future experiences? Are there interaction effects between
-different features, or do people appear to treat each feature independently?
+different features, or is the influence of each feature independent of all others'?
 And are there individual differences in how people organize their
-memories, or in how people are influenced by our experimental manipulations? If
+memories, or in how participants are influenced by our experimental manipulations? If
 so, what are those differences and which aspects of memory do they affect?

### 2023-11-17 paper/main.tex
hyphenate feature-rich throughout main.tex

@@
 We enrolled a total of 491 members of the Dartmouth College community across 11
-experimental conditions. The conditions included two controls (feature rich and
+experimental conditions. The conditions included two controls (feature-rich and
 reduced), two visual manipulation conditions [reduced (early) and reduced
@@
 merges. In total, this left usable data from 452 participants, broken down by
-experimental condition as follows: feature rich (67 participants), reduced (61
+experimental condition as follows: feature-rich (67 participants), reduced (61
 participants), reduced (early) (42 participants), reduced (late) (41
@@
 help interpret performance in other ``manipulation'' conditions. In the first
-control condition, which we call the \textit{feature rich} condition, we
+control condition, which we call the \textit{feature-rich} condition, we
 randomly shuffled the presentation order (independently for each participant)
@@
 ``reduced'' procedure (presenting each word in black at the center of the
-screen) for early lists, and followed the ``feature rich'' procedure
+screen) for early lists, and followed the ``feature-rich'' procedure
 (presenting each word in a random color and location) for late lists. Finally,
-in the \textit{reduced (late)} condition, we followed the feature rich
+in the \textit{reduced (late)} condition, we followed the feature-rich
 procedure for early lists and the reduced procedure for late lists.
@@
 each participant). In all of the order manipulation conditions, we varied
-words' font colors and onscreen locations, as in the feature rich condition.
+words' font colors and onscreen locations, as in the feature-rich condition.
 
@@
 randomized order. We varied the words' colors and locations for every word
-presentation, as in the feature rich and order manipulation conditions.
+presentation, as in the feature-rich and order manipulation conditions.
 
@@
 Lists in the random batches were sorted randomly (as on the initialization
-lists and in the feature rich condition). Lists in the stabilize and
+lists and in the feature-rich condition). Lists in the stabilize and
 destabilize batches were sorted in ways that either matched or mismatched each
@@
 at the center of the display) we compared participants' performance on the
-\textit{feature rich} and \textit{reduced} experimental conditions (see
-\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature rich
+\textit{feature-rich} and \textit{reduced} experimental conditions (see
+\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature-rich
 condition the words' colors and locations varied randomly across words, and in
@@
 difference in recall accuracy (i.e., the proportions of correctly recalled
-words) for feature rich versus reduced lists ($t(126) = -0.290, p = 0.772,
+words) for feature-rich versus reduced lists ($t(126) = -0.290, p = 0.772,
 \mathrm{Cohen's}~d~(d) = -0.051,
 \mathrm{bootstrap~estimated~95\%~confidence~interval}~(\mathrm{CI}) = [-2.387,
-1.768]$). However, participants in the feature rich condition clustered their
+1.768]$). However, participants in the feature-rich condition clustered their
 recalls substantially more along every dimension we examined (temporal
@@
 different lists, should be more similar than the estimated fingerprints
-computed for different participants. We reasoned that the feature rich
+computed for different participants. We reasoned that the feature-rich
 condition would provide the best opportunity to test this assumption, since the
@@
 help interpret our results, we compared participants' memories on early versus
-late lists in the above feature rich and reduced conditions. Participants in
-both conditions remembered more words on early versus late lists (feature rich:
+late lists in the above feature-rich and reduced conditions. Participants in
+both conditions remembered more words on early versus late lists (feature-rich:
 $t(66) = 4.553, p < 0.001, d = 0.233, \mathrm{CI} = [2.427, 7.262]$; reduced:
 $t(60) = 2.434, p = 0.018, d = 0.134, \mathrm{CI} = [0.493, 4.910]$).
-Participants in the feature rich (but not reduced) conditions exhibited more
-temporal clustering on early versus late lists (feature rich: $t(66) = 2.268, p
+Participants in the feature-rich (but not reduced) conditions exhibited more
+temporal clustering on early versus late lists (feature-rich: $t(66) = 2.268, p
 = 0.027, d = 0.181, \mathrm{CI} = [0.437, 4.425]$; reduced: $t(60) = 0.986, p =
@@
 conditions tended to exhibit more semantic clustering on early versus late
-lists (feature rich, category: $t(66) = 3.684, p < 0.001, d = 0.220,
-\mathrm{CI} = [1.733, 5.732]$; feature rich, size: $t(66) = 1.629, p = 0.108, d
+lists (feature-rich, category: $t(66) = 3.684, p < 0.001, d = 0.220,
+\mathrm{CI} = [1.733, 5.732]$; feature-rich, size: $t(66) = 1.629, p = 0.108, d
 = 0.100, \mathrm{CI} = [-0.207, 3.905] $; reduced, category: $t(60) = 2.755, p
@@
 3.081, p = 0.003, d = 0.201, \mathrm{CI} = [1.210, 5.326]$). Participants in
-the reduced (but not feature rich) conditions tended to exhibit more
-lexicographic clustering on early versus late lists (feature rich, word length:
+the reduced (but not feature-rich) conditions tended to exhibit more
+lexicographic clustering on early versus late lists (feature-rich, word length:
 $t(66) = -0.100, p = 0.921, d = -0.010, \mathrm{CI} = [-2.217, 1.899]$; feature
@@
 held them constant on late lists. Given our above findings that (a)
-participants tended to exhibit stronger clustering effects on feature rich
+participants tended to exhibit stronger clustering effects on feature-rich
 (versus reduced) lists, and (b) participants tended to remember more words and
@@
 served to enhance recall performance relative to conditions where all lists had
-the same blends of features (accuracy for feature rich versus reduced (early):
+the same blends of features (accuracy for feature-rich versus reduced (early):
 $t(107) = -2.230, p = 0.028, d = -0.439, \mathrm{CI} = [-4.252, -0.229]$;
@@
 \textit{had} been present on early lists (as in the reduced (late) condition)
-did not appear to impact recall performance (accuracy for feature rich versus
+did not appear to impact recall performance (accuracy for feature-rich versus
 reduced (late): $t(106) = -0.638, p = 0.525, d = -0.126, \mathrm{CI} = [-2.720,
@@
 incidental visual features to \textit{any} lists (comparisons of clustering on
-feature rich versus reduced lists are reported above; temporal clustering in
+feature-rich versus reduced lists are reported above; temporal clustering in
 reduced versus reduced (early) and reduced versus reduced (late) conditions:
@@
 -4.555$, $p$s $< 0.001$). Temporal and feature-based clustering were not
-reliably different in the feature rich, reduced (early), and reduced (late)
-conditions (temporal clustering in feature rich versus reduced (early) and
-feature rich versus reduced (late) conditions: $t$s $\geq -1.379$, $p$s $\geq
-0.171$; feature-based clustering in feature rich versus reduced (early) and
-feature rich versus reduced (late) conditions: $|t|$s $\leq 1.441$, $p$s $\geq
+reliably different in the feature-rich, reduced (early), and reduced (late)
+conditions (temporal clustering in feature-rich versus reduced (early) and
+feature-rich versus reduced (late) conditions: $t$s $\geq -1.379$, $p$s $\geq
+0.171$; feature-based clustering in feature-rich versus reduced (early) and
+feature-rich versus reduced (late) conditions: $|t|$s $\leq 1.441$, $p$s $\geq
 0.153$).
@@
 
-\caption{\textbf{Recall dynamics in feature rich free recall (order
+\caption{\textbf{Recall dynamics in feature-rich free recall (order
 manipulation conditions).} \textbf{A.} Behavioral plots. \textbf{Left panels.}
@@
 showed an increase in memory performance on early lists (accuracy, relative to
-early feature rich lists; $t(95) = 3.034, p = 0.003, d = 0.667, \mathrm{CI} =
+early feature-rich lists; $t(95) = 3.034, p = 0.003, d = 0.667, \mathrm{CI} =
 [1.048, 5.113]$). Participants in the color-ordered condition also showed a
 trending increase in memory performance on early lists (again, relative to
-early feature rich lists: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
+early feature-rich lists: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
 [-0.010, 3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
 early lists in all of the other order manipulation conditions were
-indistinguishable from performance on the early feature rich lists ($|t|$s
+indistinguishable from performance on the early feature-rich lists ($|t|$s
 $\leq 1.013, p$s $\geq 0.314$). Participants in both of the semantically
 ordered conditions exhibited stronger temporal clustering on early lists
-(versus early feature rich lists; category: $t(95) = 8.813, p < 0.001, d =
+(versus early feature-rich lists; category: $t(95) = 8.813, p < 0.001, d =
 1.936, \mathrm{CI} = [6.793, 11.751] $; size: $t(95) = 2.630, p = 0.010, d =
@@
 in the length-ordered condition tended to exhibit \textit{less} temporal
-clustering on early lists relative to early feature rich lists ($t(95) =
+clustering on early lists relative to early feature-rich lists ($t(95) =
 -1.547, p = 0.125, d = -0.340, \mathrm{CI} = [-3.693, 0.341]$), whereas
@@
 [-1.598, 1.729]$). Participants in the visually ordered conditions also showed
-similar temporal clustering on early lists, relative to early feature rich
+similar temporal clustering on early lists, relative to early feature-rich
 lists (color: $t(96) = -1.339, p = 0.184, d = -0.291, \mathrm{CI} = [-3.238,
@@
 3.521]$). We also compared feature-based clustering on early lists across the
-order manipulation and feature rich conditions. Since these results were
+order manipulation and feature-rich conditions. Since these results were
 similar across both semantic conditions (category and size), both lexicographic
@@
 On early lists, participants in the semantically ordered conditions exhibited
-stronger semantic clustering relative to participants in the feature rich
+stronger semantic clustering relative to participants in the feature-rich
 condition (category: $t(125) = 2.722, p = 0.007, d = 0.484, \mathrm{CI} =
@@
 participants in the visually ordered conditions exhibited stronger visual
-clustering (again, relative to feature rich participants, and on early lists;
+clustering (again, relative to feature-rich participants, and on early lists;
 color: $t(126) = 2.022, p = 0.045, d = 0.358, \mathrm{CI} = [0.056, 3.965]$;
@@
 The above comparisons between memory performance on early lists in the order
-manipulation versus feature rich conditions highlight how sorted lists are
+manipulation versus feature-rich conditions highlight how sorted lists are
 remembered differently from random lists. We also wondered how sorting lists
@@
 factoring out temporal clustering effects; \textbf{D.}) for early (gray) and
-late (gold) lists. For the feature rich bars (left), the feature clustering
+late (gold) lists. For the feature-rich bars (left), the feature clustering
 scores are averaged across all feature dimensions. For the order manipulation
@@
 these null results in aggregate across the three feature groupings. Relative to
-memory performance on late feature rich lists, participants' memory performance
+memory performance on late feature-rich lists, participants' memory performance
 in all six order manipulation conditions showed no reliable differences
@@
 [-0.447, 3.519]$). Nor did we observe any reliable differences in temporal
-clustering on late lists (relative to late feature rich lists; semantic:
+clustering on late lists (relative to late feature-rich lists; semantic:
 $t(125) = 0.157, p = 0.875, d = 0.028, \mathrm{CI} = [-1.859, 1.974]$;
@@
 words by their length on late visual order manipulation lists (more than late
-feature rich lists; $t(126) = 2.005, p = 0.047, d = 0.355, \mathrm{CI} =
+feature-rich lists; $t(126) = 2.005, p = 0.047, d = 0.355, \mathrm{CI} =
 [0.211, 3.722]$), we observed no reliable differences in any type of feature
-clustering on late order manipulation condition lists versus late feature rich
+clustering on late order manipulation condition lists versus late feature-rich
 lists ($|t|$s $\leq 1.124, p$s $\geq 0.263$).
@@
 three key types of lists. On \textit{random} lists, words are ordered randomly
-(as in the feature rich condition). On \textit{stabilize} lists, the
+(as in the feature-rich condition). On \textit{stabilize} lists, the
 presentation order is adjusted to be maximally similar to the current estimate
@@
 dimensions, even when the lists they studied were not sorted along those
-dimensions (as in the feature rich condition). A caveat is that the
+dimensions (as in the feature-rich condition). A caveat is that the
 \textit{specific} feature dimensions participants tended to cluster along
@@
 We carried out these tests for each set of ranked scores, and found that
-participants in the feature rich condition reliably cluster their recalls along
+participants in the feature-rich condition reliably cluster their recalls along
 at least three dimensions, including temporal clustering (which was often
@@
 % specific findings:
-%   reduced (early/late) vs. feature rich and reduced: event boundaries
+%   reduced (early/late) vs. feature-rich and reduced: event boundaries
 %   order manipulations: lingering effects
@@
 
-We found that participants in our feature rich condition (where we varied
+We found that participants in our feature-rich condition (where we varied
 words' appearances) recalled similar proportions of words to participants in a
@@
 When we held the within-list variability in participants' visual experiences
-fixed across lists (in the feature rich and reduced conditions), they
-remembered more words from early lists than from late lists. For feature rich
+fixed across lists (in the feature-rich and reduced conditions), they
+remembered more words from early lists than from late lists. For feature-rich
 lists, they also showed stronger clustering for early versus late lists.
@@
 manipulation conditions showed similar memory performance and clustering on
-late lists to participants in the corresponding control (feature rich)
+late lists to participants in the corresponding control (feature-rich)
 condition. At the level of individual participants, however, we found several
@@
 an additional factor that may contribute to these phenomenon. For example, our
-finding that participants better remember ``feature rich'' lists (where words'
+finding that participants better remember ``feature-rich'' lists (where words'
 appearances are varied) than ``reduced'' lists (where words' appearances are
-held constant) only when those feature rich lists are presented \textit{after}
+held constant) only when those feature-rich lists are presented \textit{after}
 reduced lists suggests that some factors that influence the richness or depth

### 2023-11-17 paper/supplement.tex
hyphenate feature-rich throughout supplement.tex

@@
 
-\caption{\textbf{Recall dynamics in feature rich free recall (random conditions).} \textbf{Left panels.} The probabilities of
+\caption{\textbf{Recall dynamics in feature-rich free recall (random conditions).} \textbf{Left panels.} The probabilities of
 initiating recall with each word are plotted as a function of presentation
@@
 
-    \caption{\textbf{Recall dynamics in feature rich free recall (adaptive conditions).} \textbf{Left panel.} The probabilities of
+    \caption{\textbf{Recall dynamics in feature-rich free recall (adaptive conditions).} \textbf{Left panel.} The probabilities of
     initiating recall with each word are plotted as a function of presentation

### 2023-11-18 paper/main.tex
edits up to Analyses methods subsection

@@
+%! suppress = LineBreak
 \documentclass[11pt]{article}
@@
 SiroEtal05, KimbEtal07, PolyKaha08, SedeEtal08, PolyEtal09, ShanHowa12}. These
-phenomena include the well known recency and primacy effects (superior recall
+phenomena include the well-known recency and primacy effects (superior recall
 of items from the end and, to a lesser extent, from the beginning of the studied
@@
 
-Participants either received course credit or a one-time \$10 payment for
+Participants received either course credit or a one-time \$10 cash payment for
 enrolling in our study. We asked each participant to fill out a demographic
 survey that included questions about their age, gender, ethnicity, race,
-education, vision, reading impairments, medications or recent injuries, coffee
+education, vision, reading impairments, medications and recent injuries, coffee
 consumption on the day of testing, and level of alertness at the time of
@@
 ranged from 17 to 31 years (mean: 19.1 years; standard deviation: 1.356 years).
-A total of 318 participants reported their gender as female, 170 as male, and
+A total of 318 participants reported their gender as female, 170 reported their gender as male, and
 two participants declined to report their gender. A total of 442 participants
-reported their ethnicity as ``not Hispanic or Latino,'' 39 as ``Hispanic or
+reported their ethnicity as ``not Hispanic or Latino,'' 39 reported their ethnicity as ``Hispanic or
 Latino,'' and nine declined to report their ethnicity. Participants reported
@@
 participants), ``Doctorate'' (one participant), and ``Master's degree'' (one
-participant). A total of 482 participants reported no reading impairments, and
+participant). A total of 482 participants reported no reading impairments;
 eight reported having mild reading impairments. A total of 489 participants
-reported having normal color vision and one participant reported that they were
-red-green color blind. A total of 482 participants reported taking no
+reported having normal color vision and one participant reported having
+impaired color vision. A total of 482 participants reported taking no
 prescription medications and having no recent injuries; four participants
@@
 allergies, one reported a recently torn ACL/MCL, and one reported a concussion
-from several months prior. The participants reported consuming 0--3 cups of
-coffee prior to the testing session (mean: 0.32 cups; standard deviation: 0.58
+from several months prior. The participants reported having consumed 0--3 cups of
+coffee on the day of the testing session (mean: 0.32 cups; standard deviation: 0.58
 cups). Participants reported their current level of alertness, and we converted
@@
 letters, identity of the word's first letter), and two visual features that are
-independent of the words themselves (text color, presentation location). Each
+independent of the words themselves (font color, presentation location). Each
 list contains four words from each of four different semantic categories, with
-two object sizes reflected across all of the words. After studying each list,
+two referent object sizes reflected across all of the words. After studying each list,
 the participant attempts to recall as many words as they can from that list, in
 any order they choose. Because each individual word is associated with several
-well defined (and quantifiable) features, and because each list incorporates a
+well-defined (and quantifiable) features, and because each list incorporates a
 diverse mix of feature values along each dimension, this allows us to estimate
@@
 of blank screen. The red microphone icons during the ``Recall'' phase denote
-the one minute verbal recall interval. The labels on the right (and
+the one-minute verbal recall interval. The labels on the right (and
 corresponding groupings on the left) denote experimental condition labels.}
@@
 The stimuli in our paradigm were 256 English words selected in a previous
-study~\citep{ZimaEtal18}. The words all referred to concrete nouns, and were
+study~\citep{ZimaEtal18}. All words referred to concrete nouns and were
 chosen from 15 unique semantic categories: body parts, building-related,
@@
 kitchen-related, mammals, (US) states, tools, trees, and vegetables. We also
-tagged each word according to the approximate size of the object the word
-referred to. Words were labeled as ``small'' if the corresponding object was
+tagged each word according to the approximate size of the object it
+referred to. Words were labeled as ``small'' if the referent object was
 likely able to ``fit in a standard shoebox'' or ``large'' if the object was
@@
 mean number of different sizes per category: 1.33; standard deviation: 0.49).
-The numbers of words in each semantic category also varied from 12--28 (mean
-number of words per category: 17.07; standard deviation number of words: 4.65).
-We also identified lexicographic features for each word, including the words'
-first letters and lengths (i.e., number of letters). Across all categories, all
+The number of words in each semantic category also varied from 12--28 (mean
+number of words per category: 17.07; standard deviation: 4.65).
+We also identified lexicographic features for each word, including its
+first letter and length (i.e., number of letters). Across all categories, all
 possible first letters were represented except for `Q' (average number of
@@
 We assigned the categorized words into a total of 16 lists with several
-constraints. First, we required that each list contained words from exactly
-four unique categories, each with exactly four exemplars from each category.
-Second, we required that (across all words on the list) at least one instance
-of both object sizes were represented. On average, each category was
+constraints. First, we required that each list contain exactly four unique words
+from each of four unique categories.
+Second, we required that each list contain at least one word representing each
+of the two object sizes (``small'' and ``large''). On average, each category was
 represented in 4.27 lists (standard deviation: 1.16 lists). Aside from these
-two constraints, we assigned each word to a unique list. After random
+two constraints, we randomly assigned each word to a single list (i.e., such that
+no words appeared in multiple lists or were omitted entirely). After random
 assignment, each list contained words with an average of 11.13 unique starting
-letters (standard deviation: 1.15 letters) and an average word length of 6.17
+letters (standard deviation: 1.15 letters) and an average length of 6.17
 letters (standard deviation: 0.34 letters).
@@
 participants, such that every participant studied the same set of 16 lists. In
-every condition we randomized the study order of these lists across
+every condition, we randomized the study order of these lists across
 participants. For participants in most conditions, on some or all of the lists,
 we also randomly varied two additional visual features associated with each
-word: the presentation font color, and the word's onscreen location. These
+word: the presentation font color and the word's onscreen location. These
 attributes were assigned independently for each word (and for every
@@
 lists of the ``reduced (early)'' condition, and the last eight lists of the
-``reduced (late)'' condition. In these latter cases, words were all presented
+``reduced (late)'' condition. In these latter cases, all words were presented
 in black at the center of the experimental computer's display.
@@
 To select a random font color for each word, we drew three integers uniformly
-and at random from the interval $\left[0, 255\right]$, corresponding to the red
+and at random from the interval $\left[0,~254\right]$, corresponding to the red
 (r), green (g), and blue (b) color channels for that word. To assign random
@@
 sorting the presented words differently on the first eight lists participants
-studied (which we call \textit{early} lists) versus on the final eight lists
-they studied (\textit{late} lists). Since every participant studied exactly 16
+studied (which we call ``\textit{early}'' lists) versus on the final eight lists
+they studied (``\textit{late}'' lists). Since every participant studied exactly 16
 lists, every list was either ``early'' or ``late'' depending on its order in
@@
 We used two ``control'' conditions to evaluate and explore participants'
-baseline behaviors. We also used performance on these control conditions to
+baseline behaviors. We also used performance in these control conditions to
 help interpret performance in other ``manipulation'' conditions. In the first
@@
 of the words on each list. In the second control condition, which we call the
-\textit{reduced} condition, we randomized word presentations as in the feature
-rich condition. However, rather than assigning each word a random color and
+\textit{reduced} condition, we randomized word presentations as in the feature-rich
+condition. However, rather than assigning each word a random color and
 location, we instead displayed all of the words in black and at the center of
@@
 
-We also designed two conditions where we varied the words' visual appearances
+We also designed two conditions in which we varied the words' visual appearances
 across lists. In the \textit{reduced (early)} condition, we followed the
@@
 each participant). In all of the order manipulation conditions, we varied
-words' font colors and onscreen locations, as in the feature-rich condition.
+words' font colors and onscreen locations as in the feature-rich condition.
 
@@
 given relevant feature requires first defining a distance function for
-quantifying the dissimilarity between each pair of features. This function
+quantifying the dissimilarity between the values of that feature for each pair of words. This function
 varied according to the type of feature under consideration. Semantic features
@@
 category or both of the same size). If two words' labels were different for a
-given feature, we defined the words to have a distance of 1 for that feature.
+given feature, we defined the words to have a distance of 1.
 Lexicographic features (length and first letter) are \textit{discrete}. For
-these features we defined a discrete distance function. Specifically, we
+these features, we defined a discrete distance function. Specifically, we
 defined the distance between two words as either the absolute difference
@@
 (positive) real values. We defined the ``color'' and ``location'' distances
-between two words as the Euclidean distances between their $(r, g, b)$ color or
-$(x, y)$ location vectors (specified in inches), respectively. Therefore, the
+between two words as the Euclidean distances between their $(r, g, b)$ color vectors and
+$(x, y)$ location vectors (specified as percentages of screen width and height), respectively. Therefore, the
 color and location distance measures always take on non-negative real values
-(upper-bounded at 441.67 for color, or 27~in for location, reflecting the
+(upper-bounded at 439.94 for color, or 124.52 for location, reflecting the
 distances between the corresponding maximally different vectors).
@@
 value of $\tau$ would diminish the influence of similarity on order. Also note
-that this approach requires $\tau > 0$. Finally, we computed a set of
+that this approach requires $\tau > 0$. Finally, we compute a set of
 normalized similarity values by dividing the similarities by their sum:
@@
 \end{equation}
-where in the denominator, $i$ takes on each of the $n$ feature values of the
+where in the denominator, $i$ takes on the feature value of each of the $n$
 to-be-presented words. The resulting set of normalized similarity scores sums
@@
 normalized similarity scores. To select the next to-be-presented item, we draw
-a random number, $r$, from the uniform distribution bounded between 0 and 1
+a random number $r$ from the uniform distribution bounded between 0 and 1
 (inclusive). The identity of the next item is given by the stick adjacent to an
@@
 been presented. The result is an ordered list that tends to change gradually
-along the selected feature dimension (for example ``sorted'' lists, see
+along the selected feature dimension (for examples of ``sorted'' lists, see
 Fig.~\ref{fig:exp}, \textit{Order manipulation} lists).
@@
 ``naturally'' organized their memories. Like the other conditions, all
-participants in the adaptive condition studied a total of 16 lists, in a
+participants in the adaptive condition studied a total of 16 word lists in a
 randomized order. We varied the words' colors and locations for every word
@@
 All participants in the adaptive condition began the experiment by studying a
-set of four \textit{initialization} lists. Words and features on these lists
+set of four \textit{initialization} lists. Words on these lists
 were presented in a randomized order (computed independently for each
 participant). These initialization lists were used to estimate each
-participant's ``memory fingerprint,'' defined below. At a high level, a
+participant's ``memory fingerprint,'' which we define below. At a high level, a
 participant's memory fingerprint describes how they prioritize or consider
@@
 presented items together in their recall sequences, where ``similarity''
-considers one given feature dimension (e.g., category, color, etc.). We base
+considers one given feature dimension (e.g., category, color, etc.). We based
 our main approach to computing clustering scores on analogous temporal and
@@
 0.5. For each recall transition (and separately for each participant), we
-sorted all not-yet-recalled words according to their absolute lag (that is,
-distance away in the list). We then computed the percentile rank of the next
+sorted all not-yet-recalled words according to their absolute lag (i.e.,
+their distance from the just-recalled word in the presented list). We then computed the percentile rank of the next
 word the participant recalled. We took an average of these percentile ranks
@@
 we were most interested in. After computing the uncorrected clustering score
-(for the given list and observed recall sequence), we compute a ``null''
-distribution of $n$ additional clustering scores after randomly shuffling the
-order of the recalled words (we use $n = 500$ in the present study). This null
+(for the given list and observed recall sequence), we constructed a ``null''
+distribution of $n$ additional clustering scores by repeatedly randomly
+shuffling the order of the recalled words and recomputing the clustering score
+for these shuffled recall sequences (we use $n = 500$ in the present study). This null
 distribution represents an approximation of the range of clustering scores one
@@
 this correction procedure to all of the clustering scores (feature and
-temporal) reported in this paper. In Figure~\clusterCorrs~we report how
+temporal) reported in this paper. In Figure~\clusterCorrs, we report how
 participants' clustering scores along different feature dimensions (in the
@@
 experiment (see \textit{Adaptive condition}) were sorted according to
-participants' \textit{current} memory fingerprint, estimated using all of the
+each individual participant's \textit{current} memory fingerprint, estimated using all of the
 lists they had studied up to that point in the experiment. Because our
@@
 that point in the experiment. Two aspects of our implementation are worth
-noting. First, because memory fingerprints are computed independently for each
+noting: First, because memory fingerprints are computed independently for each
 list and then averaged across lists, the already-computed memory fingerprints
-for earlier lists could be cached and loaded as needed in future computations.
-This meant that our computations pertaining to updating our estimate of a
-participant's memory fingerprint only needed to consider data from the most
-recent list. Second, each element of the null distributions of uncorrected
-fingerprint scores (see \textit{Permutation-corrected feature clustering
-scores}) could be estimated independently from the others. This enabled us to
-make use of the testing computers' multi-core CPU architectures by considering
-(in parallel) elements of the null distributions in batches of eight (i.e., the
-number of CPU cores on each testing computer). Taken together, we were able to
+for earlier lists could be cached and retrieved as needed in future computations.
+This meant that updating our estimate of a participant's memory fingerprint required
+computing feature and temporal clustering scores only for the single most
+recent list. Second, the clustering scores for each dimension of a participant's
+memory fingerprint could be estimated independently from the others, as could
+each element of the null distributions of uncorrected clustering scores computed
+for each dimension (see \textit{Permutation-corrected feature clustering scores}).
+This enabled us to aggressively parallelize the fingerprint-updating procedure and
 compress the relevant computations into just a few seconds of computing time.
@@
 score. For the remaining feature dimensions, we arranged the clustering scores
-in the fingerprint into a template vector, $f$. Second, we computed $n = 2500$
+in the fingerprint into a template vector $f$. Second, we computed $n = 2500$
 random permutations of the to-be-presented items. These permutations served as
@@
 for every permutation is always equal to 1.) This yielded a ``simulated
-fingerprint'' vector, $\hat{f}_p$ for each permutation $p$. We used these
-simulated fingerprints to select a specific permutation, $i$, that either
+fingerprint'' vector $\hat{f}_p$ for each permutation $p$. We used these
+simulated fingerprints to select a specific permutation $i$ that either
 maximized (for stabilize lists) or minimized (for destabilize lists) the
@@
 
-Following some of our prior work~\citep{HeusEtal18a, HeusEtal21, MannEtal22},
-we use low-dimensional embeddings to help visualize how participants' memory
+Following some of our prior work~\citep{FitzEtal23, HeusEtal18a, HeusEtal21, MannEtal22},
+we used low-dimensional embeddings to help visualize how participants' memory
 fingerprints change across lists
@@
 most variance in the data). For two visualizations
-(Figs.~\ref{fig:fingerprint-trajectories}B,~and~\fingerprintTrajectoryRandom B),
+(Figs.~\ref{fig:fingerprint-trajectories}B,~\fingerprintTrajectoryRandom B),
 we computed an additional set of two-dimensional embeddings for the
@@
 rather than for each individual list. We used these embeddings solely for
-visualization. All statistical tests were carried out in the original
+visualization; all statistical tests were carried out in the original
 (seven-dimensional) feature spaces.
@@
 a given set of recalled items (whose presentation positions are given by $x_1,
-x_2, x_3, …, x_N$), we circularly shift the presentation positions by a
+x_2, x_3, …, x_N$), we circularly shifted the presentation positions by a
 randomly chosen amount (between 1 and the list length) to obtain a new set of
@@
 set of items will have the same (average) temporal distances between successive
-recalls, the temporal clustering score for the new set of items is equal (on
+recalls, the temporal clustering score for the new set of items will be equal (on
 average) to the temporal clustering score for the original recalls. However, we
@@
 temporally corrected score will have an expected value of 0.5 no matter which
-words (or in what order) are recalled. Therefore these temporally corrected
+words a participant recalls, or the order in which they recall them. Therefore these temporally corrected
 clustering scores are interpretable only to the extent that presentation order
-and feature values are decoupled.
+and feature value are decoupled.
 

### 2023-11-18 paper/main.tex
edits through end of methods section

@@
 the proportion of trials on which each item (as a function of its presentation
-position) was recalled at output position $n$~\citep{Hoga75, HowaKaha99,
+position) was recalled at each output position $n$~\citep{Hoga75, HowaKaha99,
 PolyEtal09, ZhanEtal23}. To carry out this analysis, we initialized (for each
 participant) a number-of-lists (16) by number-of-words-per-list (16) matrix of
-0s. Then, for each list, we found the index of the word that was recalled
-first, and we filled in that position in the matrix with a 1. Finally, we
+0s. Then, for each list, we found the presentation index of the word that was recalled
+first, and filled in that position in the matrix with a 1. Finally, we
 averaged over the rows of the matrix to obtain a 1 by 16 array of
@@
 reflects the probability of recalling a given item after the just-recalled
-item, as a function of their relative encoding positions (lag). In other words,
+item, as a function of the items' relative encoding positions (lag). In other words,
 a lag of 1 indicates that a recalled item was presented immediately after the
@@
 three items before the previously recalled item. For each recall transition
-(following the first recall), we computed the lag between the just-recalled
-word's presentation position and the next-recalled word's presentation
-position. We computed the proportions of transitions (between successively
+(following the first recall), we computed the lag between the presentation positions
+of the just-recalled word and the next-recalled word. 
+We then computed the proportions of transitions (between successively
 recalled words) for each lag, normalizing for the total numbers of possible
 transitions. In carrying out this analysis, we excluded all incorrect recalls
-and repetitions (i.e., recalling a word that had already appeared previously in
+and repetitions (i.e., recalling a word that had already appeared in
 the current recall sequence). This yielded, for each list, a 1 by
@@
 All of the data analyzed in this manuscript, along with all of the code for
-carrying out the analyses may be found at https://github.com/ContextLab/FRFR-analyses.
+carrying out the analyses, may be found at https://github.com/ContextLab/FRFR-analyses.
+Code for running the non-adaptive
+experimental conditions may be found at
+https://github.com/Con\-text\-Lab/efficient-learning-code. Code for running the
+adaptive experimental condition may be found at
+https://github.com/ContextLab/adaptiveFR. We have also released an associated
+Python toolbox for analyzing free recall data, which may be found at
+https://cdl-quail.read\-the\-docs.io/\-en/\-latest.
 
@@
 Python toolbox for analyzing free recall data, which may be found at
-https://cdl-quail.read\-the\-docs.io/\-en/\-latest/.  Note that this study was not
+https://cdl-quail.read\-the\-docs.io/\-en/\-latest.  Note that this study was not
 preregistered.  Some of the ideas and data presented in this manuscript
-were also presented at the Annual Meeting of the Society for Neuroscience (2017).
+were also presented at the Annual Meeting of the Society for Neuroscience (2017)
+and the Context and Episodic Memory Symposium (2017).
 

### 2023-11-18 paper/supplement.tex
edits to supplement

@@
 position. \textbf{Middle panels.} The conditional probabilities of recalling
-each word are plotted as a function of the relative position (Lag) to the words
+each word are plotted as a function of the relative position (lag) to the word
 recalled just-prior. \textbf{Right panels.} The overall probabilities of
@@
 intervals (calculated across participants). Top panels display the recall
-dynamics for early (order manipulation) lists in each condition (color). Bottom
-panels display the recall dynamics for late (randomly ordered) lists.}
+dynamics for early lists in each condition, and bottom
+panels display the recall dynamics for late lists.}
 
@@
 
-    \caption{\textbf{Recall dynamics in feature-rich free recall (adaptive conditions).} \textbf{Left panel.} The probabilities of
+    \caption{\textbf{Recall dynamics in feature-rich free recall (adaptive condition).} \textbf{Left panel.} The probabilities of
     initiating recall with each word are plotted as a function of presentation
     position. \textbf{Middle panel.} The conditional probabilities of recalling
-    each word are plotted as a function of the relative position (Lag) to the words
+    each word are plotted as a function of the relative position (lag) to the word
     recalled just-prior. \textbf{Right panel.} The overall probabilities of
@@
     \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
-    intervals (calculated across participants). Condition is denoted by color.}
+    intervals (calculated across participants). Word-sorting policy (batch) is denoted by color.}
 
@@
 studied lists for the random conditions (\textbf{A.}), order manipulation
-conditions (\textbf{B.}), and adaptive conditions (\textbf{C.}). The conditions
-are denoted by color. Note that the first four lists in all of the ``adaptive''
-conditions were ordered randomly to compute a baseline fingerprint for each
+conditions (\textbf{B.}), and word-sorting policy (batch) in the adaptive condition (\textbf{C.}). The conditions (or batches)
+are denoted by color. Note that words in the first four lists of the ``adaptive''
+condition were ordered randomly to compute a baseline fingerprint for each
 participant prior to initiating the adaptive ordering procedure. \textbf{All
-panels.} All error ribbons denote bootstrap-estimated 95\% confidence
-intervals, calculated across participants.} 
+panels.} Error ribbons denote bootstrap-estimated 95\% confidence
+intervals (calculated across participants).} 
 \label{fig:accuracy-by-list}
@@
     rows display across-participant correlations between clustering scores for each feature dimension (top: order manipulation lists; middle: randomly ordered lists).  The matrices
-    in the bottom rows display the differences between the top and middle rows.}
+    in the bottom row display the differences between the top and middle rows.}
         \label{fig:clustering-correlations}
@@
 \caption{\textbf{Memory ``fingerprints'' (random conditions).} The
-across-participant clustering scores for each feature type ($x$-axis) are
+across-participant average clustering scores for each feature type ($x$-axis) are
 displayed for each experimental condition (color), separately for early (top)
@@
 
-\caption{\textbf{Memory ``fingerprints'' (adaptive conditions).} The
-across-participant clustering scores for each feature type ($x$-axis) are
-displayed for each experimental condition (color). Error bars denote
+\caption{\textbf{Memory ``fingerprints'' (adaptive condition).} The
+across-participant average clustering scores for each feature type ($x$-axis) are
+displayed for each batch of lists in the adaptive condition (color). Error bars denote
 bootstrap-estimated 95\% confidence intervals.}
@@
     \caption{\textbf{Probability of $n$\textsuperscript{th} recall matrices.}
-    Each sub-panel displays the average probability of recalling the given word
-    (Presentation position, matrix column) at the given output position (matrix
+    Each sub-panel displays the average probability of recalling the word with the given 
+    presentation position (matrix column) at the given output position (matrix
     row); color denotes the probability. \textbf{A. Random conditions.} The top
-    rows of matrices display data from early (order manipulation) lists, the
-    middle rows of matrices display data from late (randomly ordered) lists,
-    and the bottom rows of matrices display the differences between the
+    row of matrices displays data from early lists, the
+    middle row of matrices displays data from late lists,
+    and the bottom row of matrices displays the differences between the
     matrices in the top and middle rows. Panel columns denote experimental
     conditions. \textbf{B. Order manipulation conditions.} The matrices are
-    displayed in the same format as those in Panel A. Panel columns are
-    organized by feature type (semantic, lexicographic, or visual). \textbf{C.
-    Adaptive conditions.} The sub-panels are displayed in the same formats as
-    Panels A and B, but here the matrices and contrasts (indicated in $y$-axis
-    labels) reflect different list manipulation conditions.}
+    displayed in the same format as those in Panel A. In these conditions, word order 
+    was manipulated in early lists (top row) and randomized in late lists (middle row). Panel columns are
+    grouped by feature type (semantic, lexicographic, or visual). \textbf{C.
+    Adaptive condition.} The sub-panels are displayed in the same formats as
+    Panels A and B, but here the matrices and contrasts (indicated by $y$-axis
+    labels) reflect different word-sorting policies (batches).}
 
@@
     the top panels, each marker displays a 2D projection of the
-    (across-participant) average memory fingerprint for one list. Early lists
+    (across-participant) average memory fingerprint for a single list. Early lists
     are denoted by circles and late lists are denoted by stars. All of the
@@
     space. The bar plots in the bottom panels display the Euclidean distances
-    of the per-list memory fingerprints to the average fingerprint across all
+    between each per-list memory fingerprint and the average fingerprint across all
     prior lists, for each condition. Error bars denote bootstrap-estimated 95\%
@@
     \begin{figure}[tp] \centering
-        \includegraphics[width=0.6\textwidth]{figures/feature_correlations}
+        \includegraphics[width=0.7\textwidth]{figures/feature_correlations}
 
@@
         Spearman correlation between the distances for each pair of feature dimensions.  The correlation
-        coefficients are displayed in the heatmap, annotated with each cell's value.} \label{fig:feature-correlations}
+        coefficients are displayed in their corresponding cells of the heatmap.} \label{fig:feature-correlations}
 

### 2023-11-19 code/analyze.py
change "feature rich" to "feature-rich" (hyphenated) in all figures

@@ -21,7 +21,7 @@ N_ITER = 500
 
 results_file = f'analyzed_{N_ITER}_iter.pkl'
 
-random = ['feature rich', 'reduced (early)', 'reduced (late)', 'reduced']
+random = ['feature-rich', 'reduced (early)', 'reduced (late)', 'reduced']
 adaptive = ['adaptive']
 non_adaptive_exclude_random = ['category', 'size', 'length', 'first letter', 'color', 'location']
 
@@ -704,7 +704,7 @@ def merge_results(results, groups):
 
 
 def create_clustering_df(results):
-    conds = ['feature rich', 'category', 'size', 'length', 'first letter', 'color', 'location']
+    conds = ['feature-rich', 'category', 'size', 'length', 'first letter', 'color', 'location']
     decoder = {'category': 'category', 'size': 'size', 'length': 'wordLength', 'first letter': 'firstLetter', 'color': 'color', 'location': 'location'}
 
     dfs = []
@@ -718,7 +718,7 @@ def create_clustering_df(results):
         x = pd.DataFrame(index=pd.MultiIndex.from_tuples(idx_vals, names=results['fingerprint'][c].data.index.names))
         x['Condition'] = c.capitalize()
 
-        if c == 'feature rich':
+        if c == 'feature-rich':
             x['Feature clustering score'] = results['fingerprint'][c].data[list(decoder.values())].mean(axis=1).values
             x['Corrected feature clustering score'] = results['corrected fingerprint'][c].data[list(decoder.values())].mean(axis=1).values
         else:

### 2023-11-19 code/dataloader.py

@@ -34,7 +34,7 @@ urls = {
 
 
 descriptions = {
-    'exp1': 'feature rich',
+    'exp1': 'feature-rich',
     'exp2': 'category',
     'exp3': 'color',
     'exp4': 'length',
@@ -49,7 +49,7 @@ descriptions = {
 
 
 feature_groupings = {
-    'random': ['feature rich', 'reduced (early)', 'reduced (late)', 'reduced'],
+    'random': ['feature-rich', 'reduced (early)', 'reduced (late)', 'reduced'],
     'semantic': ['category', 'size'],
     'lexicographic': ['length', 'first letter'],
     'visual': ['color', 'location'],

### 2023-11-19 code/plot.py

@@ -26,7 +26,7 @@ from analyze import analyze_data, filter, reorder_df, rename_features, mini_filt
 
 
 colors = {
-    'feature rich': '#c0673c',
+    'feature-rich': '#c0673c',
     'reduced (early)': '#f15a22',
     'reduced (late)': '#f7996c',
     'reduced': '#fddac5',
@@ -116,7 +116,7 @@ def combo_fingerprint_plot(x, include_conds='all', include_lists='all', fname=No
     fig = plt.figure(figsize=figsize)
     ax = plt.gca()
     
-    force_order = ['feature rich', 'reduced (early)', 'reduced (late)', 'reduced',
+    force_order = ['feature-rich', 'reduced (early)', 'reduced (late)', 'reduced',
                    'category', 'size', 'length', 'first letter', 'color', 'location',
                    'adaptive', 'random', 'stabilize', 'destabilize']        
 

### 2023-11-19 code/analyze.py
change "feature rich" to "feature-rich" in stats notebook; also include "~" (forced space) in t-test and correlation outputs

@@ -662,7 +662,7 @@ def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_
     else:
         p_string = f'p = {p:.3f}'
 
-    print(f't({df}) = {t:.3f}, {p_string}, d = {d:.3f}, ' + '\mathrm{CI}' + f' = [{low_conf:.3f}, {high_conf:.3f}]')
+    print(f't({df}) = {t:.3f},~{p_string},~d = {d:.3f},~' + '\mathrm{CI}' + f' = [{low_conf:.3f},~{high_conf:.3f}]')
     
 
 def stack_fried_eggs(*args):

### 2023-11-19 paper/main.tex

@@
 randomly varying the font colors and locations of presented words (versus
-holding the font color fixed at black, and holding the display locations fixed
-at the center of the display) we compared participants' performance on the
+holding the font color fixed at black, and the words' locations fixed
+at the center of the screen), we compared participants' performance on the
 \textit{feature-rich} and \textit{reduced} experimental conditions (see
 \textit{Random conditions}, Fig.~\dynamicsRandom). In the feature-rich
-condition the words' colors and locations varied randomly across words, and in
-the reduced condition words were always presented in black, at the center of
+condition, the words' colors and locations varied randomly, and in
+the reduced condition, words were always presented in black, at the center of
 the display. Aggregating across all lists for each participant, we found no
-difference in recall accuracy (i.e., the proportions of correctly recalled
-words) for feature-rich versus reduced lists ($t(126) = -0.290, p = 0.772,
-\mathrm{Cohen's}~d~(d) = -0.051,
-\mathrm{bootstrap~estimated~95\%~confidence~interval}~(\mathrm{CI}) = [-2.387,
+difference in recall accuracy (i.e., the proportions of words successfully recalled)
+for feature-rich versus reduced lists ($t(126) = -0.290,~p = 0.772$,
+Cohen's~$d = -0.051$, bootstrap-estimated 95\% confidence interval (CI)~$= [-2.387,~
 1.768]$). However, participants in the feature-rich condition clustered their
 recalls substantially more along every dimension we examined (temporal
-clustering: $t(126) = 10.632, p < 0.001, d = 1.882, \mathrm{CI} = [7.786,
-14.386]$; semantic category clustering: $t(126) = 10.148, p < 0.001, d = 1.796,
-\mathrm{CI} = [7.324, 13.778] $; size clustering: $t(126) = 12.033, p < 0.001,
-d = 2.129, \mathrm{CI} = [9.030, 15.918]$; word length clustering: $t(126) =
-10.720, p < 0.001, d = 1.897, \mathrm{CI} = [7.442, 15.174]$; first letter
-clustering: $t(126) = 6.679, p < 0.001, d = 1.182, \mathrm{CI} = [4.490,
+clustering: $t(126) = 10.632,~p < 0.001,~d = 1.882,~\mathrm{CI} = [7.786,~
+14.386]$; semantic category clustering: $t(126) = 10.148,~p < 0.001,~d = 1.796,~
+\mathrm{CI} = [7.324,~13.778]$; size clustering: $t(126) = 12.033,~p < 0.001,~
+d = 2.129,~\mathrm{CI} = [9.030,~15.918]$; word length clustering: $t(126) =
+10.720,~p < 0.001,~d = 1.897,~\mathrm{CI} = [7.442,~15.174]$; first letter
+clustering: $t(126) = 6.679,~p < 0.001,~d = 1.182,~\mathrm{CI} = [4.490,~
 9.611]$; see \textit{Permutation-corrected feature clustering scores} for more
@@
 participants organize their memories of studied words, even when those new
-features are independent of the words themselves and even when the new features
+features are independent of the words themselves and
 vary randomly across words. We found no evidence that those additional
@@
 and their temporal clustering score. Conceptually, a participant's memory
-fingerprint describes their tendency to order, in their recall sequences (and,
+fingerprint describes their tendency to order, in their recall sequences (and
 presumably, organize in memory), the studied words along each dimension. If
@@
 (across all of their lists). Repeating this procedure for all lists and
-participants, we found that participants' fingerprints on a held-out list are
-reliably more similar to the same participant's fingerprints on other lists
-than to other participants' fingerprints ($t(70280) = 5.077, p < 0.001, d =
-0.162, \mathrm{CI} = [3.086, 6.895]$). This suggests that participants'
-fingerprints are stable across lists, and that each participant's fingerprint
+participants, we found that participants' fingerprints for a held-out list were
+reliably more similar to their fingerprints for other lists than they were to
+ other participants' fingerprints ($t(70280) = 5.077,~p < 0.001,~d =
+0.162,~\mathrm{CI} = [3.086,~6.895]$). This suggests that participants'
+fingerprints are relatively stable across lists, and that each participant's fingerprint
 is unique to them.
@@
 both conditions remembered more words on early versus late lists (feature-rich:
-$t(66) = 4.553, p < 0.001, d = 0.233, \mathrm{CI} = [2.427, 7.262]$; reduced:
-$t(60) = 2.434, p = 0.018, d = 0.134, \mathrm{CI} = [0.493, 4.910]$).
+$t(66) = 4.553,~p < 0.001,~d = 0.233,~\mathrm{CI} = [2.427,~7.262]$; reduced:
+$t(60) = 2.434,~p = 0.018,~d = 0.134,~\mathrm{CI} = [0.493,~4.910]$).
 Participants in the feature-rich (but not reduced) conditions exhibited more
-temporal clustering on early versus late lists (feature-rich: $t(66) = 2.268, p
-= 0.027, d = 0.181, \mathrm{CI} = [0.437, 4.425]$; reduced: $t(60) = 0.986, p =
-0.328, d = 0.061, \mathrm{CI} = [-0.897, 3.348]$). And participants in both
+temporal clustering on early versus late lists (feature-rich: $t(66) = 2.268,~p
+= 0.027,~d = 0.181,~\mathrm{CI} = [0.437,~4.425]$; reduced: $t(60) = 0.986,~p =
+0.328,~d = 0.061,~\mathrm{CI} = [-0.897,~3.348]$). And participants in both
 conditions tended to exhibit more semantic clustering on early versus late
-lists (feature-rich, category: $t(66) = 3.684, p < 0.001, d = 0.220,
-\mathrm{CI} = [1.733, 5.732]$; feature-rich, size: $t(66) = 1.629, p = 0.108, d
-= 0.100, \mathrm{CI} = [-0.207, 3.905] $; reduced, category: $t(60) = 2.755, p
-= 0.008, d = 0.177, \mathrm{CI} = [0.761, 5.189]$; reduced, size: $t(60) =
-3.081, p = 0.003, d = 0.201, \mathrm{CI} = [1.210, 5.326]$). Participants in
+lists (feature-rich, category: $t(66) = 3.684,~p < 0.001,~d = 0.220,~
+\mathrm{CI} = [1.733,~5.732]$; feature-rich, size: $t(66) = 1.629,~p = 0.108,~d
+= 0.100,~\mathrm{CI} = [-0.207,~3.905] $; reduced, category: $t(60) = 2.755,~p
+= 0.008,~d = 0.177,~\mathrm{CI} = [0.761,~5.189]$; reduced, size: $t(60) =
+3.081,~p = 0.003,~d = 0.201,~\mathrm{CI} = [1.210,~5.326]$). Participants in
 the reduced (but not feature-rich) conditions tended to exhibit more
 lexicographic clustering on early versus late lists (feature-rich, word length:
-$t(66) = -0.100, p = 0.921, d = -0.010, \mathrm{CI} = [-2.217, 1.899]$; feature
-rich, first letter: $t(66) = 0.412, p = 0.681, d = 0.045, \mathrm{CI} =
-[-1.645, 2.461]$; reduced, word length: $t(60) = 3.762, p < 0.001, d = 0.261,
-\mathrm{CI} = [1.604, 6.821]$; reduced, first letter: $t(60) = 1.721, p =
-0.090, d = 0.175, \mathrm{CI} = [-0.138, 4.098]$). Taken together, these
+$t(66) = -0.100,~p = 0.921,~d = -0.010,~\mathrm{CI} = [-2.217,~1.899]$; feature
+rich, first letter: $t(66) = 0.412,~p = 0.681,~d = 0.045,~\mathrm{CI} =
+[-1.645,~2.461]$; reduced, word length: $t(60) = 3.762,~p < 0.001,~d = 0.261,~
+\mathrm{CI} = [1.604,~6.821]$; reduced, first letter: $t(60) = 1.721,~p =
+0.090,~d = 0.175,~\mathrm{CI} = [-0.138,~4.098]$). Taken together, these
 comparisons suggest that even when the presence or absence of incidental visual
@@
 condition and diminished in the reduced (late) condition. However, to our
-surprise, participants in \textit{neither} condition exhibited reliable early
-versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
-0.141, d = 0.098, \mathrm{CI} = [-0.345, 3.579]$; reduced (late): $t(40) =
-1.462, p = 0.152, d = 0.121, \mathrm{CI} = [-0.376, 2.993]$), temporal
-clustering (reduced (early): $t(41) = 0.857, p = 0.396, d = 0.068, \mathrm{CI}
-= [-1.012, 2.896]$; reduced (late): $t(40) = 1.244, p = 0.221, d = 0.128,
-\mathrm{CI} = [-0.894, 3.088]$), nor feature-based clustering (reduced (early),
-category: $t(41) = 0.707, p = 0.484, d = 0.068, \mathrm{CI} = [-1.314, 2.830]$;
-reduced (early), size: $t(41) = 0.803, p = 0.427, d = 0.079, \mathrm{CI} =
-[-1.142, 2.953]$; reduced (early), length: $t(41) = 0.461, p = 0.648, d =
-0.060, \mathrm{CI} = [-1.545, 2.462]$; reduced (early), first letter: $t(41) =
-0.781, p = 0.439, d = 0.101, \mathrm{CI} = [-1.039, 2.881]$; reduced (late),
-category: $t(40) = 0.101, p = 0.920, d = 0.009, \mathrm{CI} = [-1.776, 2.307]$;
-reduced (late), size: $t(40) = 0.555, p = 0.582, d = 0.058, \mathrm{CI} =
-[-1.444, 2.274] $; reduced (late), length: $t(40) = 1.482, p = 0.146, d =
-0.126, \mathrm{CI} = [-0.444, 3.743]$; reduced (late), first letter: $t(40) =
--0.143, p = 0.887, d = -0.017, \mathrm{CI} = [-2.204, 1.830] $). We
+surprise, participants in \textit{neither} condition exhibited reliable early-versus-late
+differences in accuracy (reduced (early): $t(41) = 1.499, ~p =
+0.141,~d = 0.098,~\mathrm{CI} = [-0.345,~3.579]$; reduced (late): $t(40) =
+1.462,~p = 0.152,~d = 0.121,~\mathrm{CI} = [-0.376,~2.993]$), temporal
+clustering (reduced (early): $t(41) = 0.857,~p = 0.396,~d = 0.068,~\mathrm{CI}
+= [-1.012,~2.896]$; reduced (late): $t(40) = 1.244,~p = 0.221,~d = 0.128,~
+\mathrm{CI} = [-0.894,~3.088]$), nor feature-based clustering (reduced (early),
+category: $t(41) = 0.707,~p = 0.484,~d = 0.068,~\mathrm{CI} = [-1.314,~2.830]$;
+reduced (early), size: $t(41) = 0.803,~p = 0.427,~d = 0.079,~\mathrm{CI} =
+[-1.142,~2.953]$; reduced (early), length: $t(41) = 0.461,~p = 0.648,~d =
+0.060,~\mathrm{CI} = [-1.545,~2.462]$; reduced (early), first letter: $t(41) =
+0.781,~p = 0.439,~d = 0.101,~\mathrm{CI} = [-1.039,~2.881]$; reduced (late),
+category: $t(40) = 0.101,~p = 0.920,~d = 0.009,~\mathrm{CI} = [-1.776,~2.307]$;
+reduced (late), size: $t(40) = 0.555,~p = 0.582,~d = 0.058,~\mathrm{CI} =
+[-1.444,~2.274]$; reduced (late), length: $t(40) = 1.482,~p = 0.146,~d =
+0.126,~\mathrm{CI} = [-0.444,~3.743]$; reduced (late), first letter: $t(40) =
+-0.143,~p = 0.887,~d = -0.017,~\mathrm{CI} = [-2.204,~1.830] $). We
 hypothesized that adding or removing the variability in the visual features was
 acting as a sort of ``event boundary'' between early and late
-lists~\citep[e.g., ][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we
+lists~\citep[e.g.,][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we
 (and others) have found that memories formed just after event boundaries can be
@@
 served to enhance recall performance relative to conditions where all lists had
-the same blends of features (accuracy for feature-rich versus reduced (early):
-$t(107) = -2.230, p = 0.028, d = -0.439, \mathrm{CI} = [-4.252, -0.229]$;
-reduced versus reduced (early): $t(101) = -2.045, p = 0.043, d = -0.410,
-\mathrm{CI} = [-3.826, 0.112]$; also see Fig.~\accuracyByList A). However,
+the same blends of features (accuracy for feature-rich vs.~reduced (early):
+$t(107) = -2.230,~p = 0.028,~d = -0.439,~\mathrm{CI} = [-4.252,~-0.229]$;
+reduced vs.~reduced (early): $t(101) = -2.045,~p = 0.043,~d = -0.410,~
+\mathrm{CI} = [-3.826,~0.112]$; also see Fig.~\accuracyByList A). However,
 \textit{subtracting} irrelevant visual features on later lists that
 \textit{had} been present on early lists (as in the reduced (late) condition)
-did not appear to impact recall performance (accuracy for feature-rich versus
-reduced (late): $t(106) = -0.638, p = 0.525, d = -0.126, \mathrm{CI} = [-2.720,
-1.362]$; reduced versus reduced (late): $t(100) = -0.407, p = 0.685, d =
--0.082, \mathrm{CI} = [-2.477, 1.626]$). These comparisons suggest that recall
+did not appear to impact recall performance (accuracy for feature-rich vs.~
+reduced (late): $t(106) = -0.638,~p = 0.525,~d = -0.126,~\mathrm{CI} = [-2.720,~
+1.362]$; reduced vs.~reduced (late): $t(100) = -0.407,~p = 0.685,~d =
+-0.082,~\mathrm{CI} = [-2.477,~1.626]$). These comparisons suggest that recall
 accuracy has a directional component: accuracy is affected differently by
-removing features later that had been present earlier versus adding features
-later that had \textit{not} been present earlier. In contrast, we found that
+removing features that had initially been present versus adding features
+that had initially been absent. In contrast, we found that
 participants exhibited more temporal and feature-based clustering when we added
 incidental visual features to \textit{any} lists (comparisons of clustering on
-feature-rich versus reduced lists are reported above; temporal clustering in
-reduced versus reduced (early) and reduced versus reduced (late) conditions:
-$t$s $\leq -9.885$, $p$s $< 0.001$; feature-based clustering in reduced versus
+feature-rich vs.~reduced lists are reported above; temporal clustering in
+reduced vs.~reduced (early) and reduced vs.~reduced (late) conditions:
+$t$s $\leq -9.885$,~$p$s $< 0.001$; feature-based clustering in reduced versus
 reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
--4.555$, $p$s $< 0.001$). Temporal and feature-based clustering were not
+-4.555$,~$p$s $< 0.001$). Temporal and feature-based clustering were not
 reliably different in the feature-rich, reduced (early), and reduced (late)
-conditions (temporal clustering in feature-rich versus reduced (early) and
-feature-rich versus reduced (late) conditions: $t$s $\geq -1.379$, $p$s $\geq
-0.171$; feature-based clustering in feature-rich versus reduced (early) and
-feature-rich versus reduced (late) conditions: $|t|$s $\leq 1.441$, $p$s $\geq
+conditions (temporal clustering in feature-rich vs.~reduced (early) and
+feature-rich vs.~reduced (late) conditions: $t$s $\geq -1.379$,~$p$s $\geq
+0.171$; feature-based clustering in feature-rich vs.~reduced (early) and
+feature-rich vs.~reduced (late) conditions: $|t|$s $\leq 1.441$,~$p$s $\geq
 0.153$).
@@
 of presentation position. \textbf{Middle panels.} The conditional probabilities
-of recalling each word are plotted as a function of the relative position (Lag)
-to the words recalled just-prior. \textbf{Right panels.} The overall
+of recalling each word are plotted as a function of the relative position (lag)
+to the word recalled just-prior. \textbf{Right panels.} The overall
 probabilities of recalling each word are plotted as a function of presentation
@@
 See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
-random and adaptive conditions. \textbf{B.} Feature distances ($z$-scored
-within condition) between the features of successively presented words (see
-\textit{Defining feature-based distances}), for each condition's feature of
-focus, plotted as a function of presentation position.\textbf{C.} Proportion of
+random and adaptive conditions. \textbf{B.} Distances between successively
+presented words ($z$-scored within condition) computed based on each condition's
+feature of focus, and plotted as a function of presentation position. See
+\textit{Defining feature-based distances} for additional information. \textbf{C.} Proportion of
 event boundaries (see \textit{Identifying event boundaries}) for each
@@
 letter, color, and location). Participants in the category-ordered condition
-showed an increase in memory performance on early lists (accuracy, relative to
-early feature-rich lists; $t(95) = 3.034, p = 0.003, d = 0.667, \mathrm{CI} =
-[1.048, 5.113]$). Participants in the color-ordered condition also showed a
+showed an increase in memory performance on early lists (accuracy relative to
+early feature-rich lists; $t(95) = 3.034,~p = 0.003,~d = 0.667,~\mathrm{CI} =
+[1.048,~5.113]$). Participants in the color-ordered condition also showed a
 trending increase in memory performance on early lists (again, relative to
-early feature-rich lists: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
-[-0.010, 3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
+early feature-rich lists: $t(96) = 1.850,~p = 0.067,~d = 0.402,~\mathrm{CI} =
+[-0.010,~3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
 early lists in all of the other order manipulation conditions were
 indistinguishable from performance on the early feature-rich lists ($|t|$s
-$\leq 1.013, p$s $\geq 0.314$). Participants in both of the semantically
+$\leq 1.013,~p$s $\geq 0.314$). Participants in both of the semantically
 ordered conditions exhibited stronger temporal clustering on early lists
-(versus early feature-rich lists; category: $t(95) = 8.813, p < 0.001, d =
-1.936, \mathrm{CI} = [6.793, 11.751] $; size: $t(95) = 2.630, p = 0.010, d =
-0.578, \mathrm{CI} = [0.831, 4.866]$; Fig.~\ref{fig:barplots}B). Participants
+(vs.~early feature-rich lists; category: $t(95) = 8.813,~p < 0.001,~d =
+1.936,~\mathrm{CI} = [6.793,~11.751] $; size: $t(95) = 2.630,~p = 0.010,~d =
+0.578,~\mathrm{CI} = [0.831,~4.866]$; Fig.~\ref{fig:barplots}B). Participants
 in the length-ordered condition tended to exhibit \textit{less} temporal
 clustering on early lists relative to early feature-rich lists ($t(95) =
--1.547, p = 0.125, d = -0.340, \mathrm{CI} = [-3.693, 0.341]$), whereas
+-1.547,~p = 0.125,~d = -0.340,~\mathrm{CI} = [-3.693,~0.341]$), whereas
 participants in the first letter-ordered condition exhibited stronger temporal
-clustering on early lists ($t(95) = 2.858, p = 0.005, d = 0.628, \mathrm{CI} =
-[1.031, 4.886]$). Participants in the visually ordered conditions exhibited
-more similar performance (accuracy) on early lists, relative to early feature
+clustering on early lists ($t(95) = 2.858,~p = 0.005,~d = 0.628,~\mathrm{CI} =
+[1.031,~4.886]$). Participants in the visually ordered conditions exhibited
+more similar performance (accuracy) on early lists relative to early feature
 rich lists (we found a trending enhancement for participants in the
-color-ordered condition: $t(96) = 1.850, p = 0.067, d = 0.402, \mathrm{CI} =
-[-0.010, 3.712]$; location: $t(95) = 0.043, p = 0.966, d = 0.010, \mathrm{CI} =
-[-1.598, 1.729]$). Participants in the visually ordered conditions also showed
-similar temporal clustering on early lists, relative to early feature-rich
-lists (color: $t(96) = -1.339, p = 0.184, d = -0.291, \mathrm{CI} = [-3.238,
+color-ordered condition: $t(96) = 1.850,~p = 0.067,~d = 0.402,~\mathrm{CI} =
+[-0.010,~3.712]$; location: $t(95) = 0.043,~p = 0.966,~d = 0.010,~\mathrm{CI} =
+[-1.598,~1.729]$). Participants in the visually ordered conditions also showed
+similar temporal clustering on early lists relative to early feature-rich
+lists (color: $t(96) = -1.339,~p = 0.184,~d = -0.291,~\mathrm{CI} = [-3.238,~
 0.394]$, we found a trending increase for participants in the location-ordered
-condition: $t(95) = 1.705, p = 0.092, d = 0.374, \mathrm{CI} = [-0.155,
+condition: $t(95) = 1.705,~p = 0.092,~d = 0.374,~\mathrm{CI} = [-0.155,~
 3.521]$). We also compared feature-based clustering on early lists across the
@@
 stronger semantic clustering relative to participants in the feature-rich
-condition (category: $t(125) = 2.722, p = 0.007, d = 0.484, \mathrm{CI} =
-[0.827, 4.932]$; size: $t(125) = 3.866, p < 0.001, d = 0.687, \mathrm{CI} =
-[2.020, 5.983]$), but showed no reliable differences in lexicographic (length:
-$t(125) = 0.521, p = 0.603, d = 0.093, \mathrm{CI} = [-1.311, 2.333]$; first
-letter: $t(125) = -0.842, p = 0.401, d = -0.150, \mathrm{CI} = [-2.825,
-1.095]$) or visual (color: $t(125) = -0.650, p = 0.517, d = -0.116, \mathrm{CI}
-= [-2.680, 1.249]$; location: $t(125) = -0.251, p = 0.802, d = -0.045,
-\mathrm{CI} = [-2.257, 1.524]$) clustering. Similarly, participants in the
+condition (category: $t(125) = 2.722,~p = 0.007,~d = 0.484,~\mathrm{CI} =
+[0.827,~4.932]$; size: $t(125) = 3.866,~p < 0.001,~d = 0.687,~\mathrm{CI} =
+[2.020,~5.983]$), but showed no reliable differences in lexicographic (length:
+$t(125) = 0.521,~p = 0.603,~d = 0.093,~\mathrm{CI} = [-1.311,~2.333]$; first
+letter: $t(125) = -0.842,~p = 0.401,~d = -0.150,~\mathrm{CI} = [-2.825,~
+1.095]$) or visual (color: $t(125) = -0.650,~p = 0.517,~d = -0.116,~\mathrm{CI}
+= [-2.680,~1.249]$; location: $t(125) = -0.251,~p = 0.802,~d = -0.045,
+\mathrm{CI} = [-2.257,~1.524]$) clustering. Similarly, participants in the
 lexicographically ordered conditions exhibited stronger (relative to feature
-rich participants) lexicographic clustering (length: $t(125) = 3.682, p <
-0.001, d = 0.655, \mathrm{CI} = [1.890, 5.569]$; first letter: $t(125) = 5.134,
-p < 0.001, d = 0.912, \mathrm{CI} = [3.251, 7.258]$) on early lists, but showed
-no reliable differences in semantic (category: $t(125) = -1.040, p = 0.301, d =
--0.185, \mathrm{CI} = [-3.095, 1.092]$; size: $t(125) = 0.006, p = 0.995, d =
-0.001, \mathrm{CI} = [-1.933, 1.952]$) or visual (color: $t(125) = 0.092, p =
-0.927, d = 0.016, \mathrm{CI} = [-1.834, 1.867]$; location: $t(125) = 0.407, p
-= 0.685, d = 0.072, \mathrm{CI} = [-1.655, 2.463]$) clustering. And
+rich participants) lexicographic clustering (length: $t(125) = 3.682,~p <
+0.001,~d = 0.655,~\mathrm{CI} = [1.890,~5.569]$; first letter: $t(125) = 5.134,
+p < 0.001,~d = 0.912,~\mathrm{CI} = [3.251,~7.258]$) on early lists, but showed
+no reliable differences in semantic (category: $t(125) = -1.040,~p = 0.301,~d =
+-0.185,~\mathrm{CI} = [-3.095,~1.092]$; size: $t(125) = 0.006,~p = 0.995,~d =
+0.001,~\mathrm{CI} = [-1.933,~1.952]$) or visual (color: $t(125) = 0.092,~p =
+0.927,~d = 0.016,~\mathrm{CI} = [-1.834,~1.867]$; location: $t(125) = 0.407, p
+= 0.685,~d = 0.072,~\mathrm{CI} = [-1.655,~2.463]$) clustering. And
 participants in the visually ordered conditions exhibited stronger visual
 clustering (again, relative to feature-rich participants, and on early lists;
-color: $t(126) = 2.022, p = 0.045, d = 0.358, \mathrm{CI} = [0.056, 3.965]$;
-location: $t(126) = 4.390, p < 0.001, d = 0.777, \mathrm{CI} = [2.730, 6.199]
+color: $t(126) = 2.022,~p = 0.045,~d = 0.358,~\mathrm{CI} = [0.056,~3.965]$;
+location: $t(126) = 4.390,~p < 0.001,~d = 0.777,~\mathrm{CI} = [2.730,~6.199]
 $), but showed no reliable differences in semantic (category: $t(126) = 0.012,
-p = 0.991, d = 0.002, \mathrm{CI} = [-1.988, 1.871]$; size: $t(126) = -0.104, p
-= 0.917, d = -0.018, \mathrm{CI} = [-2.166, 1.847] $) or lexicographic (length:
-$t(126) = 0.592, p = 0.555, d = 0.105, \mathrm{CI} = [-1.361, 2.420]$; first
-letter: $t(126) = 0.040, p = 0.968, d = 0.007, \mathrm{CI} = [-1.791, 1.863]$)
+p = 0.991,~d = 0.002,~\mathrm{CI} = [-1.988,~1.871]$; size: $t(126) = -0.104, p
+= 0.917,~d = -0.018,~\mathrm{CI} = [-2.166,~1.847] $) or lexicographic (length:
+$t(126) = 0.592,~p = 0.555,~d = 0.105,~\mathrm{CI} = [-1.361,~2.420]$; first
+letter: $t(126) = 0.040,~p = 0.968,~d = 0.007,~\mathrm{CI} = [-1.791,~1.863]$)
 clustering. Taken together, these order manipulation results suggest several
-broad patterns (Figs.~\ref{fig:recall-dynamics}A, \ref{fig:fingerprints}).
+broad patterns (Figs.~\ref{fig:recall-dynamics}A,~\ref{fig:fingerprints}).
 First, most of the order manipulations we carried out did \textit{not} reliably
@@
 (i.e., semantic manipulations did not affect lexicographic or visual
-clustering, and so on). Although it is not possible to fully separate feature
+clustering, and so on). Although it is not possible to fully separate feature-based
 versus temporal clustering when considering sorted lists, we used a
@@
 \caption{\textbf{Memory ``fingerprints'' (order manipulation conditions).} The
-across-participant clustering scores for each feature type ($x$-coordinate) are
-displayed for each experimental condition (color), separately for order
-manipulation (early, top) and randomly ordered (late, bottom) lists. Error bars
+across-participant average clustering scores for each feature type ($x$-axis) are
+displayed for each experimental condition (color), separately for order-manipulated
+(early, top) and randomly ordered (late, bottom) lists. Error bars
 denote bootstrap-estimated 95\% confidence intervals. See
@@
 randomly ordered lists (Figs.~\dynamicsRandom,~\recallInit). One difference is
-that participants in the category condition (dark purple curves,
-Fig.~\ref{fig:recall-dynamics}) most often initiated recall with the
+that participants in the category condition (Fig.~\ref{fig:recall-dynamics},
+dark purple curves) most often initiated recall with the
 fourth-from-last item (\textit{Recall initiation}, top left panel), whereas
@@
 The above comparisons between memory performance on early lists in the order
-manipulation versus feature-rich conditions highlight how sorted lists are
+manipulation and feature-rich conditions highlight how sorted lists are
 remembered differently from random lists. We also wondered how sorting lists
@@
 the other feature dimensions. Participants trended towards remembering early
-lists that were sorted semantically better than lexicographically sorted lists
-($t(118) = 1.936, p = 0.055, d = 0.353, \mathrm{CI} = [0.057, 3.916]$).
+lists that were sorted semantically better than those sorted lexicographically
+($t(118) = 1.936,~p = 0.055,~d = 0.353,~\mathrm{CI} = [0.057,~3.916]$).
 Participants also remembered visually sorted lists better than
-lexicographically sorted lists ($t(119) = 2.145, p = 0.034, d = 0.390,
-\mathrm{CI} = [0.208, 4.254]$). However, participants showed no reliable
+lexicographically sorted lists ($t(119) = 2.145,~p = 0.034,~d = 0.390,~
+\mathrm{CI} = [0.208,~4.254]$). However, participants showed no reliable
 differences in recall for semantically versus visually sorted lists ($t(119) =
-0.113, p = 0.910, d = 0.021, \mathrm{CI} = [-1.987, 2.097]$). Participants
-temporally clustered semantically sorted lists more strongly than either
-lexicographically ($t(118) = 5.620, p < 0.001, d = 1.026, \mathrm{CI} = [3.486,
-8.010]$) or visually ($t(119) = 6.613, p < 0.001, d = 1.202, \mathrm{CI} =
-[4.481, 9.464]$) sorted lists, but did not show reliable differences in
+0.113,~p = 0.910,~d = 0.021,~\mathrm{CI} = [-1.987,~2.097]$). Participants
+temporally clustered semantically sorted lists more strongly than lists sorted either
+lexicographically ($t(118) = 5.620,~p < 0.001,~d = 1.026,~\mathrm{CI} = [3.486,~
+8.010]$) or visually ($t(119) = 6.613,~p < 0.001,~d = 1.202,~\mathrm{CI} =
+[4.481,~9.464]$), but did not show reliable differences in
 temporal clustering on lexicographically versus visually sorted lists ($t(119)
-= 0.589, p = 0.557, d = 0.107, \mathrm{CI} = [-1.336, 2.539]$). Participants
+= 0.589,~p = 0.557,~d = 0.107,~\mathrm{CI} = [-1.336,~2.539]$). Participants
 also showed reliably more semantic clustering on semantically sorted lists than
-lexicographically (category: $t(118) = 3.667, p < 0.001, d = 0.670, \mathrm{CI}
-= [1.822, 5.942]$, size: $t(118) = 3.972, p < 0.001$) or visually (category:
-$t(119) = 2.702, p = 0.008$, size: $t(118) = 4.043, p < 0.001, d = 0.738,
-\mathrm{CI} = [2.145, 6.296]$) sorted lists; more lexicographic clustering on
-lexicographically sorted lists than semantically (length: $t(118) = 3.390, p <
-0.001, d = 0.619, \mathrm{CI} = [1.499, 5.661]$; first letter: $t(118) = 5.705,
-p < 0.001, d = 1.042, \mathrm{CI} = [3.841, 7.790]$) or visually (length:
-$t(119) = 3.399, p < 0.001, d = 0.618, \mathrm{CI} = [1.500, 5.527]$; first
-letter: $t(119) = 4.859, p < 0.001, d = 0.883, \mathrm{CI} = [2.860, 6.849]$)
+lexicographically (category: $t(118) = 3.667,~p < 0.001,~d = 0.670,~\mathrm{CI}
+= [1.822,~5.942]$; size: $t(118) = 3.972,~p < 0.001$) or visually (category:
+$t(119) = 2.702,~p = 0.008$; size: $t(118) = 4.043,~p < 0.001,~d = 0.738,~
+\mathrm{CI} = [2.145,~6.296]$) sorted lists; more lexicographic clustering on
+lexicographically sorted lists than semantically (length: $t(118) = 3.390,~p <
+0.001,~d = 0.619,~\mathrm{CI} = [1.499,~5.661]$; first letter: $t(118) = 5.705,~
+p < 0.001,~d = 1.042,~\mathrm{CI} = [3.841,~7.790]$) or visually (length:
+$t(119) = 3.399,~p < 0.001,~d = 0.618,~\mathrm{CI} = [1.500,~5.527]$; first
+letter: $t(119) = 4.859,~p < 0.001,~d = 0.883,~\mathrm{CI} = [2.860,~6.849]$)
 sorted lists; and more visual clustering on visually sorted lists than
-semantically (color: $t(119) = 2.673, p = 0.009, d = 0.486, \mathrm{CI} =
-[0.848, 4.567]$; location: $t(119) = 4.499, p < 0.001, d = 0.818, \mathrm{CI} =
-[2.721, 6.399]$) or lexicographically (color: $t(119) = 1.988, p = 0.049, d =
-0.361, \mathrm{CI} = [0.102, 3.894]$; location: $t(119) = 3.966, p < 0.001, d =
-0.721, \mathrm{CI} = [2.099, 5.862]$) sorted lists. In summary, sorting lists
+semantically (color: $t(119) = 2.673,~p = 0.009,~d = 0.486,~\mathrm{CI} =
+[0.848,~4.567]$; location: $t(119) = 4.499,~p < 0.001,~d = 0.818,~\mathrm{CI} =
+[2.721,~6.399]$) or lexicographically (color: $t(119) = 1.988,~p = 0.049,~d =
+0.361,~\mathrm{CI} = [0.102,~3.894]$; location: $t(119) = 3.966,~p < 0.001,~d =
+0.721,~\mathrm{CI} = [2.099,~5.862]$) sorted lists. In summary, sorting lists
 by different features appeared to have slightly different effects on overall
@@
 in all six order manipulation conditions showed no reliable differences
-(semantic: $t(125) = 0.487, p = 0.627, d = 0.087, \mathrm{CI} = [-1.661,
-2.323]$; lexicographic: $t(125) = 0.878, p = 0.382, d = 0.156, \mathrm{CI} =
-[-1.226, 3.044]$; visual: $t(126) = 1.437, p = 0.153, d = 0.254, \mathrm{CI} =
-[-0.447, 3.519]$). Nor did we observe any reliable differences in temporal
+(semantic: $t(125) = 0.487,~p = 0.627,~d = 0.087,~\mathrm{CI} = [-1.661,~
+2.323]$; lexicographic: $t(125) = 0.878,~p = 0.382,~d = 0.156,~\mathrm{CI} =
+[-1.226,~3.044]$; visual: $t(126) = 1.437,~p = 0.153,~d = 0.254,~\mathrm{CI} =
+[-0.447,~3.519]$). Nor did we observe any reliable differences in temporal
 clustering on late lists (relative to late feature-rich lists; semantic:
-$t(125) = 0.157, p = 0.875, d = 0.028, \mathrm{CI} = [-1.859, 1.974]$;
-lexicographic: $t(125) = 0.998, p = 0.320, d = 0.177, \mathrm{CI} = [-0.902,
-2.920]$; visual: $t(126) = 0.548, p = 0.585, d = 0.097, \mathrm{CI} = [-1.450,
+$t(125) = 0.157,~p = 0.875,~d = 0.028,~\mathrm{CI} = [-1.859,~1.974]$;
+lexicographic: $t(125) = 0.998,~p = 0.320,~d = 0.177,~\mathrm{CI} = [-0.902,~
+2.920]$; visual: $t(126) = 0.548,~p = 0.585,~d = 0.097,~\mathrm{CI} = [-1.450,~
 2.365]$). Aside from a slightly increased tendency for participants to cluster
 words by their length on late visual order manipulation lists (more than late
-feature-rich lists; $t(126) = 2.005, p = 0.047, d = 0.355, \mathrm{CI} =
-[0.211, 3.722]$), we observed no reliable differences in any type of feature
+feature-rich lists; $t(126) = 2.005,~p = 0.047,~d = 0.355,~\mathrm{CI} =
+[0.211,~3.722]$), we observed no reliable differences in any type of feature
 clustering on late order manipulation condition lists versus late feature-rich
-lists ($|t|$s $\leq 1.124, p$s $\geq 0.263$).
+lists ($|t|$s $\leq 1.124,~p$s $\geq 0.263$).
 
@@
     condition. In the top panels, each marker displays a 2D projection of the
-    (across-participant) average memory fingerprint for one list. Order
+    (across-participant) average memory fingerprint for a single list. Order
     manipulation (early) lists are denoted by circles and randomly ordered
@@
     conditions and lists) are projected into a common space. The bar plots in
-    the bottom panels display the Euclidean distances of the per-list memory
-    fingerprints to the average fingerprint across all prior lists, for each
+    the bottom panels display the Euclidean distances between each per-list memory
+    fingerprint and the average fingerprint across all prior lists, for each
     condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
@@
 condition's manipulated feature) recalled more words. This trend held overall
-across conditions and participants (early: $r(179) = 0.492, p < 0.001,
-\mathrm{CI} = [0.352, 0.606]$; late: $r(179) = 0.403, p < 0.001, \mathrm{CI} =
-[0.271, 0.517]$) as well as for each condition individually for early ($r$s
+across conditions and participants (early: $r(179) = 0.492,~p < 0.001,
+\mathrm{CI} = [0.352, 0.606]$; late: $r(179) = 0.403,~p < 0.001,~\mathrm{CI} =
+[0.271,~0.517]$) as well as for each condition individually for early ($r$s
 $\geq 0.331$, all $p$s $\leq 0.069$) and late ($r$s $\geq 0.404$, all $p$s
@@
 scores were not correlated with the conditions where participants remembered
-more words (early: $r(4) = 0.511, p = 0.300, \mathrm{CI} = [-0.999, 0.996]$;
-late: $r(4) = -0.304, p = 0.559, \mathrm{CI} = [-0.833, 0.748]$; see insets of
+more words (early: $r(4) = 0.511,~p = 0.300,~\mathrm{CI} = [-0.999,~0.996]$;
+late: $r(4) = -0.304,~p = 0.559,~\mathrm{CI} = [-0.833,~0.748]$; see insets of
 Fig.~\ref{fig:clustering-scatterplots}A and B). We observed carryover
@@
 feature clustering on early lists in the non-visual conditions tended to recall
-more items on late lists (across conditions: $r(179) = 0.230, p = 0.002,
-\mathrm{CI} = [0.072, 0.372]$; all non-visual conditions individually: $r$s
-$\geq 0.405$, all $p$s $\leq 0.027$; color: $r(29) = 0.212, p = 0.251,
-\mathrm{CI} = [-0.164, 0.532]$; location: $r(28) = 0.320, p = 0.085,
-\mathrm{CI} = [0.011, 0.584]$). Participants who recalled more items on early
+more items on late lists (across conditions: $r(179) = 0.230,~p = 0.002,
+\mathrm{CI} = [0.072,~0.372]$; all non-visual conditions individually: $r$s
+$\geq 0.405$, all $p$s $\leq 0.027$; color: $r(29) = 0.212,~p = 0.251,
+\mathrm{CI} = [-0.164,~0.532]$; location: $r(28) = 0.320,~p = 0.085,
+\mathrm{CI} = [0.011,~0.584]$). Participants who recalled more items on early
 lists also tended to show stronger feature clustering on late lists (across
-conditions: $r(179) = 0.464, p < 0.001, \mathrm{CI} = [0.321, 0.582]$;
+conditions: $r(179) = 0.464,~p < 0.001,~\mathrm{CI} = [0.321,~0.582]$;
 individual conditions: all $r$s $\geq 0.377$, all $p$s $\leq 0.040$). Neither
 of these effects showed condition-level trends (early feature clustering versus
-late recall probability: $r(4) = -0.338, p = 0.512, \mathrm{CI} = [-0.971,
+late recall probability: $r(4) = -0.338,~p = 0.512,~\mathrm{CI} = [-0.971,~
 0.634]$; early recall probability versus late feature clustering: $r(4) =
-0.451, p = 0.369, \mathrm{CI} = [-0.986, 0.998]$). We also looked for
+0.451,~p = 0.369,~\mathrm{CI} = [-0.986,~0.998]$). We also looked for
 associations between feature clustering and temporal clustering. Across every
@@
 (Fig.~\ref{fig:clustering-scatterplots}E), this trend held overall ($r(179) =
-0.916, p < 0.001, \mathrm{CI} = [0.893, 0.936]$), for each condition
+0.916,~p < 0.001,~\mathrm{CI} = [0.893,~0.936]$), for each condition
 individually (all $r$s $\geq 0.822$, all $p$s $< 0.001$), and across conditions
-($r(4) = 0.964, p = 0.002$). For late lists
+($r(4) = 0.964,~p = 0.002$). For late lists
 (Fig.~\ref{fig:clustering-scatterplots}F), the results were more variable
-(overall: $r(179) = 0.348, p < 0.001$; all non-visual conditions: $r$s $\geq
-0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453, p = 0.011$; location: $
-r(28) = 0.190, p = 0.314$; across-conditions: $r(4) = -0.036, p = 0.945$).
+(overall: $r(179) = 0.348,~p < 0.001$; all non-visual conditions: $r$s $\geq
+0.382$, all $p$s $\leq 0.037$; color: $r(29) = 0.453,~p = 0.011$; location: $
+r(28) = 0.190,~p = 0.314$; across-conditions: $r(4) = -0.036,~p = 0.945$).
 While less robust than the carryover associations between feature clustering
@@
 feature clustering on early lists showed stronger temporal
-clustering on later lists (overall: $r(179) = 0.464, p < 0.001, \mathrm{CI} =
+clustering on later lists (overall: $r(179) = 0.464,~p < 0.001,~\mathrm{CI} =
 [0.321, 0.582]$; for individual conditions: all $r$s $\geq 0.377$, all $p$s
-$\leq 0.040$; across conditions: $ r(4) = 0.451, p = 0.369, \mathrm{CI} =
+$\leq 0.040$; across conditions: $ r(4) = 0.451,~p = 0.369,~\mathrm{CI} =
 [-0.986, 0.998]$). And participants who showed stronger temporal clustering on
 early lists trended towards showing stronger feature clustering on later lists
-(overall: $r(179) = 0.266, p < 0.001, \mathrm{CI} = [0.129, 0.396]$; for
+(overall: $r(179) = 0.266,~p < 0.001,~\mathrm{CI} = [0.129, 0.396]$; for
 individual conditions: all $r$s $\geq 0.298$, all $p$s $\leq 0.110$; across
-conditions: $r(4) = 0.064, p = 0.903, \mathrm{CI} = [-0.972,$). Taken together,
+conditions: $r(4) = 0.064,~p = 0.903,~\mathrm{CI} = [-0.972,$). Taken together,
 the results displayed in Figure~\ref{fig:clustering-scatterplots} show that
@@
 feature clustering on late lists (Fig.~\ref{fig:clustering-carryover}A; overall
-across participants and conditions: $r(179) = 0.591, p < 0.001, \mathrm{CI} =
-[0.472, 0.682]$; category: $r(28) = 0.590, p < 0.001, \mathrm{CI} = [0.354,
-0.756]$; size: $r(28) = 0.488, p = 0.006, \mathrm{CI} = [0.134, 0.732]$;
-length: $r(28) = 0.384, p = 0.036, \mathrm{CI} = [0.040, 0.681]$; first letter:
-$r(28) = 0.202, p = 0.284, \mathrm{CI} = [-0.273, 0.620]$; color: $r(29) =
--0.183, p = 0.325, \mathrm{CI} = [-0.562, 0.258]$; location: $r(28) = 0.031, p
-= 0.870, \mathrm{CI} = [-0.240, 0.296]$; across conditions: $r(4) = 0.942, p =
-0.005, \mathrm{CI} = [0.442, 1.000]$). Although participants tended to show
+across participants and conditions: $r(179) = 0.591,~p < 0.001,~\mathrm{CI} =
+[0.472,~0.682]$; category: $r(28) = 0.590,~p < 0.001,~\mathrm{CI} = [0.354,~
+0.756]$; size: $r(28) = 0.488,~p = 0.006,~\mathrm{CI} = [0.134,~0.732]$;
+length: $r(28) = 0.384,~p = 0.036,~\mathrm{CI} = [0.040,~0.681]$; first letter:
+$r(28) = 0.202,~p = 0.284,~\mathrm{CI} = [-0.273,~0.620]$; color: $r(29) =
+-0.183,~p = 0.325,~\mathrm{CI} = [-0.562,~0.258]$; location: $r(28) = 0.031, p
+= 0.870,~\mathrm{CI} = [-0.240,~0.296]$; across conditions: $r(4) = 0.942,~p =
+0.005,~\mathrm{CI} = [0.442,~1.000]$). Although participants tended to show
 weaker feature clustering on late lists
@@
 decrease in recall on early versus late lists
-(Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.307, p < 0.001,
-\mathrm{CI} = [0.148, 0.469]$; category: $r(28) = 0.350, p = 0.058, \mathrm{CI}
-= [0.050, 0.642]$; size: $r(28) = 0.708, p < 0.001, \mathrm{CI} = [0.472,
-0.862]$; length: $r(28) = 0.205, p = 0.276, \mathrm{CI} = [-0.109, 0.492]$;
-first letter: $r(28) = 0.081, p = 0.672, \mathrm{CI} = [-0.433, 0.597]$; color:
-$r(29) = 0.155, p = 0.406, \mathrm{CI} = [-0.174, 0.541]$; location: $r(28) =
-0.052, p = 0.787, \mathrm{CI} = [-0.307, 0.360]$; across conditions: $r(4) =
-0.635, p = 0.176, \mathrm{CI} = [-0.924, 0.981]$. Participants who exhibited
+(Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.307,~p < 0.001,
+\mathrm{CI} = [0.148,~0.469]$; category: $r(28) = 0.350,~p = 0.058,~\mathrm{CI}
+= [0.050,~0.642]$; size: $r(28) = 0.708,~p < 0.001,~\mathrm{CI} = [0.472,~
+0.862]$; length: $r(28) = 0.205,~p = 0.276,~\mathrm{CI} = [-0.109,~0.492]$;
+first letter: $r(28) = 0.081,~p = 0.672,~\mathrm{CI} = [-0.433,~0.597]$; color:
+$r(29) = 0.155,~p = 0.406,~\mathrm{CI} = [-0.174,~0.541]$; location: $r(28) =
+0.052,~p = 0.787,~\mathrm{CI} = [-0.307,~0.360]$; across conditions: $r(4) =
+0.635,~p = 0.176,~\mathrm{CI} = [-0.924,~0.981]$. Participants who exhibited
 larger carryover in feature clustering also tended to show stronger temporal
 clustering on late lists (relative to early lists) for all but the category
-condition (Fig.~\ref{fig:clustering-carryover}C; overall: $r(179) = 0.426, p <
-0.001, \mathrm{CI} = [0.285, 0.544]$; category: $r(28) = 0.110, p = 0.564,
-\mathrm{CI} = [-0.284, 0.442]$; all non-category conditions: all $r$s $\geq
-0.406$, all $p$s $\leq 0.023$; across conditions: $r(4) = 0.649, p = 0.163,
+condition (Fig.~\ref{fig:clustering-carryover}C; overall: $r(179) = 0.426,~p <
+0.001,~\mathrm{CI} = [0.285,~0.544]$; category: $r(28) = 0.110,~p = 0.564,~
+\mathrm{CI} = [-0.284,~0.442]$; all non-category conditions: all $r$s $\geq
+0.406$, all $p$s $\leq 0.023$; across conditions: $r(4) = 0.649,~p = 0.163,
 \mathrm{CI} = [-0.856, 0.988]$).
@@
         \caption{\textbf{Adaptive free recall.} \textbf{A.} Average probability
-        of recall (taken across words, lists, and participants) for lists from
-        each adaptive condition. \textbf{B.} Average temporal clustering scores
-        for lists from each adaptive condition. \textbf{C.} Recall probability
+        of recall (taken across words, lists, and participants) for each batch of
+        four lists in the adaptive condition. \textbf{B.} Average temporal clustering scores
+        for each batch of lists. \textbf{C.} Recall probability
         versus temporal clustering scores by participant (main panel; each
-        participant contributes one dot per condition) and averaged within
-        condition (inset; each dot represents a single condition). \textbf{D.}
+        participant contributes one dot per batch) and averaged within
+        batch (inset; each dot represents a single batch). \textbf{D.}
         Per-list correlations between the current list's fingerprint and the
@@
         that the participant had experienced at the time of the current list.
-        All panels: Colors denote the sorting type (condition) for each list.
+        All panels: Colors denote the sorting policy (batch) for each list.
         Error bars and ribbons denote bootstrap-estimated 95\% confidence
         intervals. For additional details about participants' behavior and
-        performance during the adaptive conditions, see
+        performance during the adaptive condition, see
         Figure~\dynamicsAdaptive.}
@@
 Overall, we found that participants tended to be better at remembering words on
-stabilize lists relative to words on both random ($t(59) = 1.740, p = 0.087, d
-= 0.095, \mathrm{CI} = [-0.187, 3.761]$) and destabilize ($t(59) = 1.714, p =
-0.092, d = 0.114, \mathrm{CI} = [-0.351, 4.108]$) lists
+stabilize lists relative to words on both random ($t(59) = 1.740,~p = 0.087, d
+= 0.095,~\mathrm{CI} = [-0.187,~3.761]$) and destabilize ($t(59) = 1.714,~p =
+0.092,~d = 0.114,~\mathrm{CI} = [-0.351,~4.108]$) lists
 (Fig.~\ref{fig:adaptive}A). Participants showed no reliable differences in
 their memory performance on destabilize versus random lists ($t(59) = -0.249, p
-= 0.804, d = -0.017, \mathrm{CI} = [-2.327, 1.578]$). Participants also
+= 0.804,~d = -0.017,~\mathrm{CI} = [-2.327,~1.578]$). Participants also
 exhibited stronger temporal clustering on stabilize lists, relative to random
-($t(59) = 3.428, p = 0.001, d = 0.306, \mathrm{CI} = [1.635, 5.460]$) and
-destabilize ($t(59) = 4.174, p < 0.001, d = 0.374, \mathrm{CI} = [1.964,
+($t(59) = 3.428,~p = 0.001,~d = 0.306,~\mathrm{CI} = [1.635,~5.460]$) and
+destabilize ($t(59) = 4.174,~p < 0.001,~d = 0.374,~\mathrm{CI} = [1.964,~
 6.968]$) lists (Fig.~\ref{fig:adaptive}B). We found no reliable differences in
 temporal clustering for items on random versus destabilize lists ($t(59) =
--0.880, p = 0.382, d = -0.081, \mathrm{CI} = [-3.165, 1.127]$).
+-0.880,~p = 0.382,~d = -0.081,~\mathrm{CI} = [-3.165,~1.127]$).
 
@@
 scores also tended to recall more items. This held across subjects, aggregating
-across all list types ($r(178) = 0.701, p < 0.001, \mathrm{CI} = [0.590,
+across all list types ($r(178) = 0.701,~p < 0.001,~\mathrm{CI} = [0.590,~
 0.789]$), and for each list type individually (all $r$s $\geq 0.651$, all $p$s
@@
 at least three dimensions, including temporal clustering (which was often
-ranked highest); Rank 1: $t(66) = 12.751, p < 0.001, d = 0.162, \mathrm{CI} =
-[8.702, 20.013]$; Rank 2: $t(66) = 8.196, p < 0.001, d = 0.162, \mathrm{CI} =
-[4.794, 12.978]$; Rank 3: $t(66) = 3.243, p = 0.002, d = 0.162, \mathrm{CI} =
-[1.028, 7.051]$; Rank 4: $t(66) = -3.112, p = 0.003, d = 0.162, \mathrm{CI} =
-[-5.282, -1.920]$; Rank 5: $t(66) = -7.154, p < 0.001, d = 0.162, \mathrm{CI} =
-[-12.649, -5.568]$; Rank 6: $t(66) = -12.608, p < 0.001, d = 0.162, \mathrm{CI}
-= [-22.114, -9.347]$; Rank 7: $t(66) = -18.397, p < 0.001, d = 0.162,
-\mathrm{CI} = [-27.238, -14.073]$.
+ranked highest); Rank 1: $t(66) = 12.751,~p < 0.001,~d = 0.162,~\mathrm{CI} =
+[8.702,~20.013]$; Rank 2: $t(66) = 8.196,~p < 0.001,~d = 0.162,~\mathrm{CI} =
+[4.794,~12.978]$; Rank 3: $t(66) = 3.243,~p = 0.002,~d = 0.162,~\mathrm{CI} =
+[1.028,~7.051]$; Rank 4: $t(66) = -3.112,~p = 0.003,~d = 0.162,~\mathrm{CI} =
+[-5.282,~-1.920]$; Rank 5: $t(66) = -7.154,~p < 0.001,~d = 0.162,~\mathrm{CI} =
+[-12.649,~-5.568]$; Rank 6: $t(66) = -12.608,~p < 0.001,~d = 0.162,~\mathrm{CI}
+= [-22.114,~-9.347]$; Rank 7: $t(66) = -18.397,~p < 0.001,~d = 0.162,
+\mathrm{CI} = [-27.238,~-14.073]$.
 

### 2023-11-20 paper/main.tex
addressed PFC comments
in addition to (mostly) text changes, i've also changed figure S9 to show spearman correlation coefficients in the lower triangle and p-values in the upper triangle, using two different colormaps.

@@
 \maketitle
+\thispagestyle{empty}
 
@@
 ranged from 17 to 31 years (mean: 19.1 years; standard deviation: 1.356 years).
-A total of 318 participants reported their gender as female, 170 reported their gender as male, and
-two participants declined to report their gender. A total of 442 participants
-reported their ethnicity as ``not Hispanic or Latino,'' 39 reported their ethnicity as ``Hispanic or
-Latino,'' and nine declined to report their ethnicity. Participants reported
-their races as White (345 participants), Asian (120 participants), Black or
-African American (31 participants), American Indian or Alaska Native (11
-participants), Native Hawaiian or Other Pacific Islander (four participants),
-Mixed race (three participants), Middle Eastern (one participant), and Arab
-(one participant). A total of five participants declined to report their race.
-We note that several participants reported more than one of the above racial
-categories. Participants reported their highest degrees achieved as ``Some
-college'' (359 participants), ``High school graduate'' (117 participants),
-``College graduate'' (seven participants), ``Some high school'' (five
-participants), ``Doctorate'' (one participant), and ``Master's degree'' (one
-participant). A total of 482 participants reported no reading impairments;
-eight reported having mild reading impairments. A total of 489 participants
-reported having normal color vision and one participant reported having
-impaired color vision. A total of 482 participants reported taking no
-prescription medications and having no recent injuries; four participants
-reported having ADHD, one reported having dyslexia, one reported having
-allergies, one reported a recently torn ACL/MCL, and one reported a concussion
-from several months prior. The participants reported having consumed 0--3 cups of
-coffee on the day of the testing session (mean: 0.32 cups; standard deviation: 0.58
-cups). Participants reported their current level of alertness, and we converted
-their responses to numerical scores as follows: ``very sluggish'' (-2), ``a
-little sluggish'' (-1), ``neutral'' (0), ``a little alert'' (1), and ``very
-alert'' (2). Across all participants, the full range of alertness levels were
-reported (range: -2--2; mean: 0.35; standard deviation: 0.89).
+A total of 318 participants reported their gender as female, 170 reported their
+gender as male, and two participants declined to report their gender. A total
+of 442 participants reported their ethnicity as ``not Hispanic or Latino,'' 39
+reported their ethnicity as ``Hispanic or Latino,'' and nine declined to report
+their ethnicity. Participants reported their races as White (345 participants),
+Asian (120 participants), Black or African American (31 participants), American
+Indian or Alaska Native (11 participants), Native Hawaiian or Other Pacific
+Islander (four participants), Mixed race (three participants), Middle Eastern
+(one participant), and Arab (one participant). A total of five participants
+declined to report their race. We note that several participants reported more
+than one of the above racial categories. Participants reported their highest
+degrees achieved as ``Some college'' (359 participants), ``High school
+graduate'' (117 participants), ``College graduate'' (seven participants),
+``Some high school'' (five participants), ``Doctorate'' (one participant), and
+``Master's degree'' (one participant). A total of 482 participants reported no
+reading impairments; eight reported having mild reading impairments. A total of
+489 participants reported having normal color vision and one participant
+reported having impaired color vision. A total of 482 participants reported
+taking no prescription medications and having no recent injuries; four
+participants reported having ADHD, one reported having dyslexia, one reported
+having allergies, one reported a recently torn ACL/MCL, and one reported a
+concussion from several months prior. The participants reported having consumed
+0--3 cups of coffee on the day of the testing session (mean: 0.32 cups;
+standard deviation: 0.58 cups). Participants reported their current level of
+alertness, and we converted their responses to numerical scores as follows:
+``very sluggish'' (-2), ``a little sluggish'' (-1), ``neutral'' (0), ``a little
+alert'' (1), and ``very alert'' (2). Across all participants, the full range of
+alertness levels were reported (range: -2--2; mean: 0.35; standard deviation:
+0.89).
 
@@
 list contains four words from each of four different semantic categories, with
-two referent object sizes reflected across all of the words. After studying each list,
-the participant attempts to recall as many words as they can from that list, in
-any order they choose. Because each individual word is associated with several
-well-defined (and quantifiable) features, and because each list incorporates a
-diverse mix of feature values along each dimension, this allows us to estimate
-which features participants are considering or leveraging in organizing their
-memories.
+two referent object sizes reflected across all of the words. After studying
+each list, the participant attempts to recall as many words as they can from
+that list, in any order they choose. Because each individual word is associated
+with several well-defined (and quantifiable) features, and because each list
+incorporates a diverse mix of feature values along each dimension, this allows
+us to estimate which features participants are considering or leveraging in
+organizing their memories.
 
@@
 The stimuli in our paradigm were 256 English words selected in a previous
-study~\citep{ZimaEtal18}. All words referred to concrete nouns and were
-chosen from 15 unique semantic categories: body parts, building-related,
-cities, clothing, countries, flowers, fruits, insects, instruments,
-kitchen-related, mammals, (US) states, tools, trees, and vegetables. We also
-tagged each word according to the approximate size of the object it
-referred to. Words were labeled as ``small'' if the referent object was
-likely able to ``fit in a standard shoebox'' or ``large'' if the object was
-larger than a shoebox. Most semantic categories comprised words that reflected
-both ``small'' and ``large'' object sizes, but several included only one or the
-other (e.g., all countries, US states, and cities are larger than a shoebox;
-mean number of different sizes per category: 1.33; standard deviation: 0.49).
-The number of words in each semantic category also varied from 12--28 (mean
-number of words per category: 17.07; standard deviation: 4.65).
-We also identified lexicographic features for each word, including its
-first letter and length (i.e., number of letters). Across all categories, all
-possible first letters were represented except for `Q' (average number of
-unique first letters per category: 11; standard deviation: 2 letters). Word
-lengths ranged from 3--12 letters (average: 6.17 letters; standard deviation:
-2.06 letters).
+study~\citep{ZimaEtal18}. All words referred to concrete nouns and were chosen
+from 15 unique semantic categories: body parts, building-related, cities,
+clothing, countries, flowers, fruits, insects, instruments, kitchen-related,
+mammals, (US) states, tools, trees, and vegetables. We also tagged each word
+according to the approximate size of the object it referred to. Words were
+labeled as ``small'' if the referent object was likely able to ``fit in a
+standard shoebox'' or ``large'' if the object was larger than a shoebox. Most
+semantic categories comprised words that reflected both ``small'' and ``large''
+object sizes, but several included only one or the other (e.g., all countries,
+US states, and cities are larger than a shoebox; mean number of different sizes
+per category: 1.33; standard deviation: 0.49). The number of words in each
+semantic category also varied from 12--28 (mean number of words per category:
+17.07; standard deviation: 4.65). We also identified lexicographic features for
+each word, including its first letter and length (i.e., number of letters).
+Across all categories, all possible first letters were represented except for
+`Q' (average number of unique first letters per category: 11.00; standard
+deviation: 2.00 letters). Word lengths ranged from 3--12 letters (average: 6.17
+letters; standard deviation: 2.06 letters).
 
 We assigned the categorized words into a total of 16 lists with several
-constraints. First, we required that each list contain exactly four unique words
-from each of four unique categories.
-Second, we required that each list contain at least one word representing each
-of the two object sizes (``small'' and ``large''). On average, each category was
-represented in 4.27 lists (standard deviation: 1.16 lists). Aside from these
-two constraints, we randomly assigned each word to a single list (i.e., such that
-no words appeared in multiple lists or were omitted entirely). After random
-assignment, each list contained words with an average of 11.13 unique starting
-letters (standard deviation: 1.15 letters) and an average length of 6.17
-letters (standard deviation: 0.34 letters).
+constraints. First, we required that each list contain exactly four unique
+words from each of four unique categories. Second, we required that each list
+contain at least one word representing each of the two object sizes (``small''
+and ``large''). On average, each category was represented in 4.27 lists
+(standard deviation: 1.16 lists). Aside from these two constraints, we randomly
+assigned each word to a single list (i.e., such that no words appeared in
+multiple lists or were omitted entirely). After random assignment, each list
+contained words with an average of 11.13 unique starting letters (standard
+deviation: 1.15 letters) and an average length of 6.17 letters (standard
+deviation: 0.34 letters).
 
@@
 any part of the words. The words were shown on 27-in (diagonal) Retina 5K iMac
-displays (resolution: 5120 $\times$ 2880 pixels).
+displays (resolution: 5120 by 2880 pixels).
 
@@
 sorting the presented words differently on the first eight lists participants
-studied (which we call ``\textit{early}'' lists) versus on the final eight lists
-they studied (``\textit{late}'' lists). Since every participant studied exactly 16
-lists, every list was either ``early'' or ``late'' depending on its order in
-the list study sequence.
+studied (which we call ``\textit{early}'' lists) versus on the final eight
+lists they studied (``\textit{late}'' lists). Since every participant studied
+exactly 16 lists, every list was either ``early'' or ``late'' depending on its
+order in the list study sequence. (In other words, the ``early'' and ``late''
+labels capture all of the lists participants studied.)
 
@@
 
-Our experimental paradigm incorporates the Google Cloud Speech API
-speech-to-text engine~\citep{HalpEtal16} to automatically transcribe
-participants' verbal recalls into text. This allows recalls to be transcribed
-in real time---a distinguishing feature of the experiment; in typical verbal
-recall experiments, the audio data must be parsed and transcribed manually. In
-prior work, we used a similar experimental setup (equivalent to the ``reduced''
-condition in the present study) to verify that the automatically transcribed
-recalls were sufficiently close to human-transcribed recalls to yield reliable
+Our experimental paradigm incorporates the Google Cloud Speech-to-Text
+engine~\citep{HalpEtal16} to automatically transcribe participants' verbal
+recalls into text. This allows recalls to be transcribed in real time---a
+distinguishing feature of the experiment; in typical verbal recall experiments,
+the audio data must be parsed and transcribed manually. In prior work, we used
+a similar experimental setup (equivalent to the ``reduced'' condition in the
+present study) to verify that the automatically transcribed recalls were
+sufficiently close to human-transcribed recalls to yield reliable
 data~\citep{ZimaEtal18}. This real-time speech processing component of the
@@
 of the words on each list. In the second control condition, which we call the
-\textit{reduced} condition, we randomized word presentations as in the feature-rich
-condition. However, rather than assigning each word a random color and
-location, we instead displayed all of the words in black and at the center of
-the screen.
-
-We also designed two conditions in which we varied the words' visual appearances
-across lists. In the \textit{reduced (early)} condition, we followed the
-``reduced'' procedure (presenting each word in black at the center of the
-screen) for early lists, and followed the ``feature-rich'' procedure
+\textit{reduced} condition, we randomized word presentations as in the
+feature-rich condition. However, rather than assigning each word a random color
+and location, we instead displayed all of the words in black and at the center
+of the screen.
+
+We also designed two conditions in which we varied the words' visual
+appearances across lists. In the \textit{reduced (early)} condition, we
+followed the ``reduced'' procedure (presenting each word in black at the center
+of the screen) for early lists, and followed the ``feature-rich'' procedure
 (presenting each word in a random color and location) for late lists. Finally,
@@
 given relevant feature requires first defining a distance function for
-quantifying the dissimilarity between the values of that feature for each pair of words. This function
-varied according to the type of feature under consideration. Semantic features
-(category and size) are \textit{categorical}. For these features, we defined a
-binary distance function: two words were considered to ``match'' (i.e., have a
-distance of 0) if their labels were the same (i.e., both from the same semantic
-category or both of the same size). If two words' labels were different for a
-given feature, we defined the words to have a distance of 1.
-Lexicographic features (length and first letter) are \textit{discrete}. For
-these features, we defined a discrete distance function. Specifically, we
-defined the distance between two words as either the absolute difference
-between their lengths, or the absolute distance between their starting letters
-in the English alphabet, respectively. For example, two words that started with
-the same letter would have a ``first letter'' distance of 0, and a pair of
-words starting with `J' and `A' would have a first letter distance of 9.
-Because words' lengths and letters' positions in the alphabet are always
-integers, these discrete distances always take on integer values. Finally, the
-visual features (color and location) are \textit{continuous} and
+quantifying the dissimilarity between the values of that feature for each pair
+of words. This function varied according to the type of feature under
+consideration. Semantic features (category and size) are \textit{categorical}.
+For these features, we defined a binary distance function: two words were
+considered to ``match'' (i.e., have a distance of 0) if their labels were the
+same (i.e., both from the same semantic category or both of the same size). If
+two words' labels were different for a given feature, we defined the words to
+have a distance of 1. Lexicographic features (length and first letter) are
+\textit{discrete}. For these features, we defined a discrete distance function.
+Specifically, we defined the distance between two words as either the absolute
+difference between their lengths, or the absolute distance between their
+starting letters in the English alphabet, respectively. For example, two words
+that started with the same letter would have a ``first letter'' distance of 0,
+and a pair of words starting with `J' and `A' would have a first letter
+distance of 9. Because words' lengths and letters' positions in the alphabet
+are always integers, these discrete distances always take on integer values.
+Finally, the visual features (color and location) are \textit{continuous} and
 \textit{multivariate}, in that each ``feature'' is defined by multiple
 (positive) real values. We defined the ``color'' and ``location'' distances
-between two words as the Euclidean distances between their $(r, g, b)$ color vectors and
-$(x, y)$ location vectors (specified as percentages of screen width and height), respectively. Therefore, the
-color and location distance measures always take on non-negative real values
-(upper-bounded at 439.94 for color, or 124.52 for location, reflecting the
-distances between the corresponding maximally different vectors).
+between two words as the Euclidean distances between their $(r, g, b)$ color
+vectors and $(x, y)$ location vectors (specified as percentages of screen width
+and height), respectively. Therefore, the color and location distance measures
+always take on non-negative real values (upper-bounded at 439.94 for color, or
+124.52 for location, reflecting the distances between the corresponding
+maximally different vectors).
 
@@
 words on the to-be-presented list. Second, we compute the distances between the
-chosen word's feature(s) and the corresponding feature(s) of all
-yet-to-be-presented words. Third, we convert these distances (between the
-previously presented word's feature values, $a$, and the candidate word's
-feature values, $b$) to similarity scores:
+chosen word's feature(s) value(s) and the corresponding feature(s) value(s) of
+all yet-to-be-presented words. Third, we convert these distances (between the
+previously presented word's feature values, $a$, and each of the $W$
+yet-to-be-presented candidate words' feature values, $b_{i \in 1...W}$) to
+similarity scores:
 
 \begin{equation}
- \mathrm{similarity}(a, b) = \exp\{-\tau \cdot
-\mathrm{distance}(a, b)\}, \label{eqn:similarity}
+ \mathrm{similarity}(a, b_i) = \exp\{-\tau \cdot
+\mathrm{distance}(a, b_i)\}, \label{eqn:similarity}
 \end{equation}
@@
 \begin{equation}
-\mathrm{similarity}_{\mathrm{normalized}}(a, b) = \frac{\mathrm{similarity}(a,
-b)}{\sum_{i=1}^n \mathrm{similarity}(a, i)},
+\mathrm{similarity}_{\mathrm{normalized}}(a, b_i) = \frac{\mathrm{similarity}(a,
+b)}{\sum_{j=1}^W \mathrm{similarity}(a, b_j)},
 \end{equation}
-where in the denominator, $i$ takes on the feature value of each of the $n$
+where in the denominator, $b_j$ takes on the feature value of each of the $W$
 to-be-presented words. The resulting set of normalized similarity scores sums
@@
 0.5. For each recall transition (and separately for each participant), we
-sorted all not-yet-recalled words according to their absolute lag (i.e.,
-their distance from the just-recalled word in the presented list). We then computed the percentile rank of the next
-word the participant recalled. We took an average of these percentile ranks
-across all of the participant’s recalls to obtain a single (uncorrected)
-temporal clustering score for the participant.
+sorted all not-yet-recalled words according to their absolute lag (i.e., their
+distance from the just-recalled word in the presented list). We then computed
+the percentile rank of the next word the participant recalled. We took an
+average of these percentile ranks across all of the participant’s recalls to
+obtain a single (uncorrected) temporal clustering score for the participant.
 
@@
 $B$ have been recalled, the best possible next matching recall will be a large
-item. All subsequent correct recalls must also be large items---so for
-those later recalls it becomes difficult to determine whether the participant
-is successively recalling large items because they are organizing their
-memories according to size, or (alternatively), whether they are simply
-recalling the yet-to-be-recalled items in a random order. In general, the
-precise order and blend of feature values expressed in a given list, the order
-and number of correct recalls a participant makes, the number of intervening
-presentation positions between successive recalls, and so on, can all affect
-the range of clustering scores that are possible to observe for a given list.
-An uncorrected clustering score therefore conflates participants' actual memory
-organization with other ``nuisance'' factors.
+item. All subsequent correct recalls must also be large items---so for those
+later recalls it becomes difficult to determine whether the participant is
+successively recalling large items because they are organizing their memories
+according to size, or (alternatively), whether they are simply recalling the
+yet-to-be-recalled items in a random order. In general, the precise order and
+blend of feature values expressed in a given list, the order and number of
+correct recalls a participant makes, the number of intervening presentation
+positions between successive recalls, and so on, can all affect the range of
+clustering scores that are possible to observe for a given list. An uncorrected
+clustering score therefore conflates participants' actual memory organization
+with other ``nuisance'' factors.
 
@@
 shuffling the order of the recalled words and recomputing the clustering score
-for these shuffled recall sequences (we use $n = 500$ in the present study). This null
-distribution represents an approximation of the range of clustering scores one
-might expect to observe by ``chance,'' given that a hypothetical participant
-was \textit{not} truly clustering their recalls, but where the hypothetical
-participant still studied and recalled exactly the same items (with the same
-features) as the true participant. We define the \textit{permutation-corrected
-clustering score} as the percentile rank of the observed uncorrected clustering
-score in this estimated null distribution. In this way, a corrected score of 1
-indicates that the observed score was greater than any clustering score one
-might expect by chance---in other words, good evidence that the participant was
-truly clustering their recalls along the given feature dimension. We applied
-this correction procedure to all of the clustering scores (feature and
-temporal) reported in this paper. In Figure~\clusterCorrs, we report how
-participants' clustering scores along different feature dimensions (in the
-order manipulation conditions) are correlated, and how clustering scores change
-across lists.
+for these shuffled recall sequences (we use $n = 500$ in the present study).
+This null distribution represents an approximation of the range of clustering
+scores one might expect to observe by ``chance,'' given that a hypothetical
+participant was \textit{not} truly clustering their recalls, but where the
+hypothetical participant still studied and recalled exactly the same items
+(with the same features) as the true participant. We define the
+\textit{permutation-corrected clustering score} as the percentile rank of the
+observed uncorrected clustering score in this estimated null distribution. In
+this way, a corrected score of 1 indicates that the observed score was greater
+than any clustering score one might expect by chance---in other words, good
+evidence that the participant was truly clustering their recalls along the
+given feature dimension. We applied this correction procedure to all of the
+clustering scores (feature and temporal) reported in this paper. In
+Figure~\clusterCorrs, we report how participants' clustering scores along
+different feature dimensions (in the order manipulation conditions) are
+correlated, and how clustering scores change across lists.
 
@@
 The presentation orders of some lists in the adaptive condition of our
-experiment (see \textit{Adaptive condition}) were sorted according to
-each individual participant's \textit{current} memory fingerprint, estimated using all of the
-lists they had studied up to that point in the experiment. Because our
-experiment incorporated a speech-to-text component, all of the behavioral data
-for each participant could be analyzed just a few seconds after the conclusion
-of the recall intervals for each list. We used the \texttt{Quail} Python
-package~\citep{HeusEtal17} to apply speech-to-text algorithms to the
-just-collected audio data, aggregate the data for the given participant, and estimate
-the participant's memory fingerprint using all of their available data up to
-that point in the experiment. Two aspects of our implementation are worth
+experiment (see \textit{Adaptive condition}) were sorted according to each
+individual participant's \textit{current} memory fingerprint, estimated using
+all of the lists they had studied up to that point in the experiment. Because
+our experiment incorporated a speech-to-text component, all of the behavioral
+data for each participant could be analyzed just a few seconds after the
+conclusion of the recall intervals for each list. We used the \texttt{Quail}
+Python package~\citep{HeusEtal17} to apply speech-to-text algorithms to the
+just-collected audio data, aggregate the data for the given participant, and
+estimate the participant's memory fingerprint using all of their available data
+up to that point in the experiment. Two aspects of our implementation are worth
 noting: First, because memory fingerprints are computed independently for each
 list and then averaged across lists, the already-computed memory fingerprints
-for earlier lists could be cached and retrieved as needed in future computations.
-This meant that updating our estimate of a participant's memory fingerprint required
-computing feature and temporal clustering scores only for the single most
-recent list. Second, the clustering scores for each dimension of a participant's
-memory fingerprint could be estimated independently from the others, as could
-each element of the null distributions of uncorrected clustering scores computed
-for each dimension (see \textit{Permutation-corrected feature clustering scores}).
-This enabled us to aggressively parallelize the fingerprint-updating procedure and
-compress the relevant computations into just a few seconds of computing time.
-The combined processing time for the speech-to-text algorithm, fingerprint
-computations, and permutation-based ordering procedure (described next) easily
-fit within the inter-list intervals, where participants paused for a self-paced
-break before moving on to study and recall the next list.
+for earlier lists could be cached and retrieved as needed in future
+computations. This meant that updating our estimate of a participant's memory
+fingerprint required computing feature and temporal clustering scores only for
+the single most recent list. Second, the clustering scores for each dimension
+of a participant's memory fingerprint could be estimated independently from the
+others, as could each element of the null distributions of uncorrected
+clustering scores computed for each dimension (see
+\textit{Permutation-corrected feature clustering scores}). This enabled us to
+aggressively parallelize the fingerprint-updating procedure and compress the
+relevant computations into just a few seconds of computing time. The combined
+processing time for the speech-to-text algorithm, fingerprint computations, and
+permutation-based ordering procedure (described next) easily fit within the
+inter-list intervals, where participants paused for a self-paced break before
+moving on to study and recall the next list.
 
@@
 
-Following some of our prior work~\citep{FitzEtal23, HeusEtal18a, HeusEtal21, MannEtal22},
-we used low-dimensional embeddings to help visualize how participants' memory
-fingerprints change across lists
+Following some of our prior work~\citep{FitzEtal23, HeusEtal18a, HeusEtal21,
+MannEtal22}, we used low-dimensional embeddings to help visualize how
+participants' memory fingerprints change across lists
 (Figs.~\ref{fig:fingerprint-trajectories}A,~\fingerprintTrajectoryRandom A). To
@@
 fingerprints (for all lists and experimental conditions) to create a large
-matrix with number-of-lists (16) $\times$ number-of-conditions (10, including
-the adaptive condition) rows and seven columns (one for each feature clustering
+matrix with number-of-lists (16) by number-of-conditions (10, including the
+adaptive condition) rows and seven columns (one for each feature clustering
 score, plus an additional temporal clustering score column). We used principal
@@
 most variance in the data). For two visualizations
-(Figs.~\ref{fig:fingerprint-trajectories}B,~\fingerprintTrajectoryRandom B),
-we computed an additional set of two-dimensional embeddings for the
+(Figs.~\ref{fig:fingerprint-trajectories}B,~\fingerprintTrajectoryRandom B), we
+computed an additional set of two-dimensional embeddings for the
 \textit{average} fingerprints across lists within a given list grouping (i.e.,
@@
 randomly chosen amount (between 1 and the list length) to obtain a new set of
-items at the (now altered) positions of the original recalls. Since the new
-set of items will have the same (average) temporal distances between successive
-recalls, the temporal clustering score for the new set of items will be equal (on
-average) to the temporal clustering score for the original recalls. However, we
-can then re-compute the feature clustering score for those new items. Finally,
-we can compute a ``temporally corrected'' feature clustering score by computing
-the average percentile rank of the observed (raw) feature clustering score
-within the distributions of circularly shifted feature clustering scores,
-across $N = 500$ repetitions of this procedure. This new temporally corrected
-score provides an estimate of the observed degree of feature clustering over
-and above what could be accounted for by temporal clustering alone.
+items at the (now altered) positions of the original recalls. Since the new set
+of items will have the same (average) temporal distances between successive
+recalls, the temporal clustering score for the new set of items will be equal
+(on average) to the temporal clustering score for the original recalls.
+However, we can then re-compute the feature clustering score for those new
+items. Finally, we can compute a ``temporally corrected'' feature clustering
+score by computing the average percentile rank of the observed (raw) feature
+clustering score within the distributions of circularly shifted feature
+clustering scores, across $N = 500$ repetitions of this procedure. This new
+temporally corrected score provides an estimate of the observed degree of
+feature clustering over and above what could be accounted for by temporal
+clustering alone.
 
@@
 temporally corrected score will have an expected value of 0.5 no matter which
-words a participant recalls, or the order in which they recall them. Therefore these temporally corrected
-clustering scores are interpretable only to the extent that presentation order
-and feature value are decoupled.
+words a participant recalls, or the order in which they recall them. Therefore
+these temporally corrected clustering scores are interpretable only to the
+extent that presentation order and feature value are decoupled.
 
@@
 participant) a number-of-lists (16) by number-of-words-per-list (16) matrix of
-0s. Then, for each list, we found the presentation index of the word that was recalled
-first, and filled in that position in the matrix with a 1. Finally, we
+0s. Then, for each list, we found the presentation index of the word that was
+recalled first, and filled in that position in the matrix with a 1. Finally, we
 averaged over the rows of the matrix to obtain a 1 by 16 array of
@@
 reflects the probability of recalling a given item after the just-recalled
-item, as a function of the items' relative encoding positions (lag). In other words,
-a lag of 1 indicates that a recalled item was presented immediately after the
-previously recalled item, and a lag of $-3$ indicates that a recalled item came
-three items before the previously recalled item. For each recall transition
-(following the first recall), we computed the lag between the presentation positions
-of the just-recalled word and the next-recalled word.
-We then computed the proportions of transitions (between successively
-recalled words) for each lag, normalizing for the total numbers of possible
-transitions. In carrying out this analysis, we excluded all incorrect recalls
-and repetitions (i.e., recalling a word that had already appeared in
-the current recall sequence). This yielded, for each list, a 1 by
-number-of-lags ($-15$ to $+15$; 30 lags in total, excluding lags of 0) array of
-conditional probabilities. We averaged these probabilities across lists to
-obtain a single lag-CRP for each participant. Because transitions at large
-absolute lags are rare, these curves are typically displayed using range
+item, as a function of the items' relative encoding positions (lag). In other
+words, a lag of 1 indicates that a recalled item was presented immediately
+after the previously recalled item, and a lag of $-3$ indicates that a recalled
+item came three items before the previously recalled item. For each recall
+transition (following the first recall), we computed the lag between the
+presentation positions of the just-recalled word and the next-recalled word. We
+then computed the proportions of transitions (between successively recalled
+words) for each lag, normalizing for the total numbers of possible transitions.
+In carrying out this analysis, we excluded all incorrect recalls and
+repetitions (i.e., recalling a word that had already appeared in the current
+recall sequence). This yielded, for each list, a 1 by number-of-lags ($-15$ to
+$+15$; 30 lags in total, excluding lags of 0) array of conditional
+probabilities. We averaged these probabilities across lists to obtain a single
+lag-CRP for each participant. Because transitions at large absolute lags are
+rare, these curves are typically displayed using range
 restrictions~\citep{Kaha12}.
@@
 over the rows of the matrix to yield a 1 by 16 array representing the
-proportion of words at each position that the participant remembered.
+proportion of words at each position that the participants remembered.
 
@@
 
-\subsection*{Data and code availability}
+\subsection*{Transparency and openness}
 
@@
 randomly varying the font colors and locations of presented words (versus
-holding the font color fixed at black, and the words' locations fixed
-at the center of the screen), we compared participants' performance on the
+holding the font color fixed at black, and the words' locations fixed at the
+center of the screen), we compared participants' performance on the
 \textit{feature-rich} and \textit{reduced} experimental conditions (see
 \textit{Random conditions}, Fig.~\dynamicsRandom). In the feature-rich
-condition, the words' colors and locations varied randomly, and in
-the reduced condition, words were always presented in black, at the center of
-the display. Aggregating across all lists for each participant, we found no
-difference in recall accuracy (i.e., the proportions of words successfully recalled)
-for feature-rich versus reduced lists ($t(126) = -0.290,~p = 0.772$,
-Cohen's~$d = -0.051$, bootstrap-estimated 95\% confidence interval (CI)~$= [-2.387,~
+condition, the words' colors and locations varied randomly, and in the reduced
+condition, words were always presented in black, at the center of the display.
+Aggregating across all lists for each participant, we found no difference in
+recall accuracy (i.e., the proportions of words successfully recalled) for
+feature-rich versus reduced lists ($t(126) = -0.290,~p = 0.772$, Cohen's~$d =
+-0.051$, bootstrap-estimated 95\% confidence interval (CI)~$= [-2.387,~
 1.768]$). However, participants in the feature-rich condition clustered their
@@
 clustering: $t(126) = 10.632,~p < 0.001,~d = 1.882,~\mathrm{CI} = [7.786,~
-14.386]$; semantic category clustering: $t(126) = 10.148,~p < 0.001,~d = 1.796,~
-\mathrm{CI} = [7.324,~13.778]$; size clustering: $t(126) = 12.033,~p < 0.001,~
-d = 2.129,~\mathrm{CI} = [9.030,~15.918]$; word length clustering: $t(126) =
-10.720,~p < 0.001,~d = 1.897,~\mathrm{CI} = [7.442,~15.174]$; first letter
-clustering: $t(126) = 6.679,~p < 0.001,~d = 1.182,~\mathrm{CI} = [4.490,~
-9.611]$; see \textit{Permutation-corrected feature clustering scores} for more
-information about how we quantified each participant's clustering tendencies.)
-Taken together, these comparisons suggest that adding new features changes how
-participants organize their memories of studied words, even when those new
-features are independent of the words themselves and
-vary randomly across words. We found no evidence that those additional
-uninformative features were distracting (in terms of their impact on memory
-performance), but they did affect participants' recall dynamics (measured via
-their clustering scores).
+14.386]$; semantic category clustering: $t(126) = 10.148,~p < 0.001,~d =
+1.796,~ \mathrm{CI} = [7.324,~13.778]$; size clustering: $t(126) = 12.033,~p <
+0.001,~ d = 2.129,~\mathrm{CI} = [9.030,~15.918]$; word length clustering:
+$t(126) = 10.720,~p < 0.001,~d = 1.897,~\mathrm{CI} = [7.442,~15.174]$; first
+letter clustering: $t(126) = 6.679,~p < 0.001,~d = 1.182,~\mathrm{CI} =
+[4.490,~ 9.611]$; see \textit{Permutation-corrected feature clustering scores}
+for more information about how we quantified each participant's clustering
+tendencies.) Taken together, these comparisons suggest that adding new features
+changes how participants organize their memories of studied words, even when
+those new features are independent of the words themselves and vary randomly
+across words. We found no evidence that those additional uninformative features
+were distracting (in terms of their impact on memory performance), but they did
+affect participants' recall dynamics (measured via their clustering scores).
 
@@
 other participants' fingerprints ($t(70280) = 5.077,~p < 0.001,~d =
-0.162,~\mathrm{CI} = [3.086,~6.895]$). This suggests that participants'
-fingerprints are relatively stable across lists, and that each participant's fingerprint
-is unique to them.
+0.162,~\mathrm{CI} = [3.086,~6.895]$). That within-participant fingerprint
+similarity was greater than between-participant fingerprint similarity suggests
+that participants' memory fingerprints are relatively stable across lists, and
+that each participant's fingerprint is unique to them.
 
@@
 probabilities of recalling each word are plotted as a function of presentation
-position. \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\%
-confidence intervals (calculated across participants). Top panels display the
-recall dynamics for early (order manipulation) lists in each condition (color).
-Bottom panels display the recall dynamics for late (randomly ordered) lists.
-See Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
+position. Top panels display the recall dynamics for early (order manipulation)
+lists in each condition (color). Bottom panels display the recall dynamics for
+late (randomly ordered) lists. See
+Figures~\dynamicsRandom~and~\dynamicsAdaptive~for analogous plots for the
 random and adaptive conditions. \textbf{B.} Distances between successively
-presented words ($z$-scored within condition) computed based on each condition's
-feature of focus, and plotted as a function of presentation position. See
-\textit{Defining feature-based distances} for additional information. \textbf{C.} Proportion of
-event boundaries (see \textit{Identifying event boundaries}) for each
-condition's feature of focus, plotted as a function of presentation position.}
+presented words ($z$-scored within condition) computed based on each
+condition's feature of focus, and plotted as a function of presentation
+position. See \textit{Defining feature-based distances} for additional
+information. \textbf{C.} Proportion of event boundaries (see
+\textit{Identifying event boundaries}) for each condition's feature of focus,
+plotted as a function of presentation position. \textbf{All panels.} Error
+ribbons denote bootstrap-estimated 95\% confidence intervals (calculated across
+participants in Panel A, and across lists in Panels B and C).}
 
@@
     condition. In the top panels, each marker displays a 2D projection of the
-    (across-participant) average memory fingerprint for a single list. Order
-    manipulation (early) lists are denoted by circles and randomly ordered
-    (late) lists are denoted by stars. All of the fingerprints (across all
-    conditions and lists) are projected into a common space. The bar plots in
-    the bottom panels display the Euclidean distances between each per-list memory
-    fingerprint and the average fingerprint across all prior lists, for each
-    condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
-    The dotted vertical lines denote the boundaries between early and late
-    lists. \textbf{B.} In this panel, the fingerprints for early (circle) and
-    late (star) lists are averaged across lists and participants before
-    projecting the fingerprints into a (new) 2D space. See
+    (across-participant) average memory fingerprint for a single list. Lines
+    connect successive lists. Order manipulation (early) lists are denoted by
+    circles and randomly ordered (late) lists are denoted by stars. All of the
+    fingerprints (across all conditions and lists) are projected into a common
+    space. The bar plots in the bottom panels display the Euclidean distances
+    between each per-list memory fingerprint and the average fingerprint across
+    all prior lists, for each condition. Error bars denote bootstrap-estimated
+    95\% confidence intervals. The dotted vertical lines denote the boundaries
+    between early and late lists. \textbf{B.} In this panel, the fingerprints
+    for early (circle) and late (star) lists are averaged across lists and
+    participants before projecting the fingerprints into a (new) 2D space. See
     Figure~\fingerprintTrajectoryRandom~for analogous plots for the random
@@
 participants tend to better remember blocked (versus random)
-lists~\citep{CofeEtal66, DAgo69, Dall64, Shap70, ShapPalm70, Kint70,
-BoweEtal68, BoweEtal69, LuekEtal71, Puff74}. Other studies suggest that these order
+lists~\citep{CofeEtal66, DAgo69, Dall64, Shap70, ShapPale70, Kint70,
+BoweEtal69, LuekEtal71, Puff74}. Other studies suggest that these order
 effects may also be modulated by factors like list length and the numbers of
@@
 
-Conceptualization: JRM and ACH. Methodology: JRM and ACH. Software: JRM, PCF,
-CEF, and ACH. Analysis: JRM, PCF, and ACH. Data collection: ECW, PCF, MRL, AMF,
-BJB, DR, and CEF. Data curation and management: ECW, PCF, MRL, and ACH. Writing
-(original draft): JRM. Writing (review and editing): ECW, PCF, MRL, AMF, BJB,
-DR, CEF, and ACH. Supervision: JRM and ACH. Project administration: ECW and
-PCF. Funding acquisition: JRM.
+Conceptualization: JRM and ACH. Data curation: JRM, PCF, ACH. Formal analysis:
+JRM, PCF, and ACH. Funding acquisition: JRM. Investigation: ECW, PCF, MRL, AMF,
+BJB, DR, CEF, and ACH. Methodology: JRM and ACH. Project administration: ECW
+and PCF. Resources: JRM. Software: JRM, PCF, CEF, and ACH. Supervision: JRM and
+ACH. Validation: JRM, PCF, and ACH. Writing (original draft): JRM. Writing
+(review and editing): ECW, PCF, MRL, AMF, BJB, DR, CEF, and ACH.
 

### 2023-11-20 paper/supplement.tex
addressed PFC comments
in addition to (mostly) text changes, i've also changed figure S9 to show spearman correlation coefficients in the lower triangle and p-values in the upper triangle, using two different colormaps.

@@
 
-\renewcommand{\figurename}{Supplementary Figure}
+%\renewcommand{\figurename}{Supplementary Figure}
 
@@
 
-\caption{\textbf{Recall dynamics in feature-rich free recall (random conditions).} \textbf{Left panels.} The probabilities of
-initiating recall with each word are plotted as a function of presentation
-position. \textbf{Middle panels.} The conditional probabilities of recalling
-each word are plotted as a function of the relative position (lag) to the word
-recalled just-prior. \textbf{Right panels.} The overall probabilities of
-recalling each word are plotted as a function of presentation position.
-\textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
-intervals (calculated across participants). Top panels display the recall
-dynamics for early lists in each condition, and bottom
-panels display the recall dynamics for late lists.}
+\caption{\textbf{Recall dynamics in feature-rich free recall (random
+conditions).} \textbf{Left panels.} The probabilities of initiating recall with
+each word are plotted as a function of presentation position. \textbf{Middle
+panels.} The conditional probabilities of recalling each word are plotted as a
+function of the relative position (lag) to the word recalled just-prior.
+\textbf{Right panels.} The overall probabilities of recalling each word are
+plotted as a function of presentation position. \textbf{All panels.} Error
+ribbons denote bootstrap-estimated 95\% confidence intervals (calculated across
+participants). Top panels display the recall dynamics for early lists in each
+condition, and bottom panels display the recall dynamics for late lists.}
 
@@
     the top panels, each marker displays a 2D projection of the
-    (across-participant) average memory fingerprint for a single list. Early lists
-    are denoted by circles and late lists are denoted by stars. All of the
-    fingerprints (across all conditions and lists) are projected into a common
-    space. The bar plots in the bottom panels display the Euclidean distances
-    between each per-list memory fingerprint and the average fingerprint across all
-    prior lists, for each condition. Error bars denote bootstrap-estimated 95\%
-    confidence intervals. The dotted vertical lines denote the boundaries
-    between early and late lists. \textbf{B.} In this panel, the fingerprints
-    for early (circle) and late (star) lists are averaged across lists and
-    participants before projecting the fingerprints into a (new) 2D space.}
+    (across-participant) average memory fingerprint for a single list. Early
+    lists are denoted by circles and late lists are denoted by stars. Lines
+    connect successive lists. All of the fingerprints (across all conditions
+    and lists) are projected into a common space. The bar plots in the bottom
+    panels display the Euclidean distances between each per-list memory
+    fingerprint and the average fingerprint across all prior lists, for each
+    condition. Error bars denote bootstrap-estimated 95\% confidence intervals.
+    The dotted vertical lines denote the boundaries between early and late
+    lists. \textbf{B.} In this panel, the fingerprints for early (circle) and
+    late (star) lists are averaged across lists and participants before
+    projecting the fingerprints into a (new) 2D space.}
     \label{fig:fingerprint-trajectories-random}
@@
 
-        \caption{\textbf{Correlations between features.} Within each list, for each participant, we
-        computed the set of distances between each pair of words, along each feature dimension.  We
-        then combined these pairwise distances across all lists and participants, and computed the
-        Spearman correlation between the distances for each pair of feature dimensions.  The correlation
-        coefficients are displayed in their corresponding cells of the heatmap.} \label{fig:feature-correlations}
+        \caption{\textbf{Correlations between features.} Within each list, for
+        each participant, we computed the set of distances between each pair of
+        words, along each feature dimension. We then combined these pairwise
+        distances across all lists and participants, and computed the Spearman
+        correlation coefficient ($\rho$) between the distances for each pair of
+        feature dimensions. The correlation coefficients are displayed in their
+        corresponding cells of the heatmap in the lower triangle, and the
+        corresponding $p$-values are displayed in the upper triangle.}
+        \label{fig:feature-correlations}
 

### 2023-11-20 paper/main.tex
moved most stats to tables

@@
 \usepackage{rotating}
+\usepackage{booktabs}
 
@@
 
+% figures
 \newcommand{\dynamicsRandom}{S1}
@@
 \newcommand{\featureCorrs}{S9}
+
+% tables
+\newcommand{\abbreviations}{S1}
 
@@
 phenomena include the well-known recency and primacy effects (superior recall
-of items from the end and, to a lesser extent, from the beginning of the studied
-list), as well as semantic and temporal clustering effects~\citep{KahaEtal08b,
-HowaKaha02b}. The contiguity effect is an example of temporal clustering, which
-is perhaps the dominant form of organization in free recall. This effect can be
-seen in people's tendencies to successively recall items that occupied
-neighboring positions in the studied list~\citep{Kaha96}. There are also
-striking effects of semantic clustering~\citep{RomnEtal93, Bous53, BousEtal54,
-JenkRuss52, MannKaha12}, whereby the recall of a given item is more likely to
-be followed by recall of a similar or related item than a dissimilar or
-unrelated one. In general, people organize memories for words along a wide
-variety of stimulus dimensions. According to models like the
-\textit{Context Maintenance and Retrieval} model~\citep{PolyEtal09}, the
-stimulus features associated with each word (e.g.,\ the word's meaning, size of
-the object the word represents, letters that make up the word, font size,
-font color, location on the screen, etc.)~are incorporated into the
-participant's mental context representation~\citep{SmitVela01, MannEtal11,
-MannEtal12, MannEtal15, Mann20}. During a memory test, any of these features
-may serve as a memory cue, which in turn leads the participant to successively recall
-words that share stimulus features.
+of items from the end and, to a lesser extent, from the beginning of the
+studied list), as well as semantic and temporal clustering
+effects~\citep{KahaEtal08b, HowaKaha02b}. The contiguity effect is an example
+of temporal clustering, which is perhaps the dominant form of organization in
+free recall. This effect can be seen in people's tendencies to successively
+recall items that occupied neighboring positions in the studied
+list~\citep{Kaha96}. There are also striking effects of semantic
+clustering~\citep{RomnEtal93, Bous53, BousEtal54, JenkRuss52, MannKaha12},
+whereby the recall of a given item is more likely to be followed by recall of a
+similar or related item than a dissimilar or unrelated one. In general, people
+organize memories for words along a wide variety of stimulus dimensions.
+According to models like the \textit{Context Maintenance and Retrieval}
+model~\citep{PolyEtal09}, the stimulus features associated with each word
+(e.g.,\ the word's meaning, size of the object the word represents, letters
+that make up the word, font size, font color, location on the screen, etc.)~are
+incorporated into the participant's mental context
+representation~\citep{SmitVela01, MannEtal11, MannEtal12, MannEtal15, Mann20}.
+During a memory test, any of these features may serve as a memory cue, which in
+turn leads the participant to successively recall words that share stimulus
+features.
 
@@
 
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/feature_rich_vs_reduced.tex}
+
+\caption{\textbf{Comparing memory in the feature-rich versus reduced conditions
+(all lists).} The $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all lists from each
+participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:feature-rich-vs-reduced}
+\end{table}
+
 To obtain a clean preliminary estimate of the consequences on memory of
@@
 \textit{feature-rich} and \textit{reduced} experimental conditions (see
-\textit{Random conditions}, Fig.~\dynamicsRandom). In the feature-rich
-condition, the words' colors and locations varied randomly, and in the reduced
-condition, words were always presented in black, at the center of the display.
-Aggregating across all lists for each participant, we found no difference in
-recall accuracy (i.e., the proportions of words successfully recalled) for
-feature-rich versus reduced lists ($t(126) = -0.290,~p = 0.772$, Cohen's~$d =
--0.051$, bootstrap-estimated 95\% confidence interval (CI)~$= [-2.387,~
-1.768]$). However, participants in the feature-rich condition clustered their
-recalls substantially more along every dimension we examined (temporal
-clustering: $t(126) = 10.632,~p < 0.001,~d = 1.882,~\mathrm{CI} = [7.786,~
-14.386]$; semantic category clustering: $t(126) = 10.148,~p < 0.001,~d =
-1.796,~ \mathrm{CI} = [7.324,~13.778]$; size clustering: $t(126) = 12.033,~p <
-0.001,~ d = 2.129,~\mathrm{CI} = [9.030,~15.918]$; word length clustering:
-$t(126) = 10.720,~p < 0.001,~d = 1.897,~\mathrm{CI} = [7.442,~15.174]$; first
-letter clustering: $t(126) = 6.679,~p < 0.001,~d = 1.182,~\mathrm{CI} =
-[4.490,~ 9.611]$; see \textit{Permutation-corrected feature clustering scores}
-for more information about how we quantified each participant's clustering
-tendencies.) Taken together, these comparisons suggest that adding new features
-changes how participants organize their memories of studied words, even when
-those new features are independent of the words themselves and vary randomly
-across words. We found no evidence that those additional uninformative features
-were distracting (in terms of their impact on memory performance), but they did
+\textit{Random conditions}, Fig.~\dynamicsRandom,
+Tab.~\ref{tab:feature-rich-vs-reduced}). In the feature-rich condition, the
+words' colors and locations varied randomly, and in the reduced condition,
+words were always presented in black, at the center of the display. Aggregating
+across all lists for each participant, we found no difference in recall
+accuracy (i.e., the proportions of words successfully recalled) for
+feature-rich versus reduced lists. However, participants in the feature-rich
+condition clustered their recalls substantially more along every dimension we
+examined (see \textit{Permutation-corrected feature clustering scores} for more
+information about how we quantified each participant's clustering tendencies.)
+Taken together, these comparisons suggest that adding new features changes how
+participants organize their memories of studied words, even when those new
+features are independent of the words themselves and vary randomly across
+words. We found no evidence that those additional uninformative features were
+distracting (in terms of their impact on memory performance), but they did
 affect participants' recall dynamics (measured via their clustering scores).
@@
 
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/early_vs_late_fr.tex}
+
+\caption{\textbf{Comparing memory for early versus late lists in the
+feature-rich condition.} The $t$-tests reported in the table were carried out
+across-participants. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:early-vs-late-feature-rich} 
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/early_vs_late_r.tex}
+
+\caption{\textbf{Comparing memory for early versus late lists in the
+reduced condition.} The $t$-tests reported in the table were carried out
+across-participants. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:early-vs-late-reduced}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/early_vs_late_early.tex}
+
+\caption{\textbf{Comparing memory for early versus late lists in the
+reduced (early) condition.} The $t$-tests reported in the table were carried out
+across-participants. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:early-vs-late-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/early_vs_late_late.tex}
+
+\caption{\textbf{Comparing memory for early versus late lists in the
+reduced (late) condition.} The $t$-tests reported in the table were carried out
+across-participants. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:early-vs-late-late}
+\end{table}
+
 We next asked whether adding these incidental visual features to later lists
@@
 help interpret our results, we compared participants' memories on early versus
-late lists in the above feature-rich and reduced conditions. Participants in
-both conditions remembered more words on early versus late lists (feature-rich:
-$t(66) = 4.553,~p < 0.001,~d = 0.233,~\mathrm{CI} = [2.427,~7.262]$; reduced:
-$t(60) = 2.434,~p = 0.018,~d = 0.134,~\mathrm{CI} = [0.493,~4.910]$).
-Participants in the feature-rich (but not reduced) conditions exhibited more
-temporal clustering on early versus late lists (feature-rich: $t(66) = 2.268,~p
-= 0.027,~d = 0.181,~\mathrm{CI} = [0.437,~4.425]$; reduced: $t(60) = 0.986,~p =
-0.328,~d = 0.061,~\mathrm{CI} = [-0.897,~3.348]$). And participants in both
-conditions tended to exhibit more semantic clustering on early versus late
-lists (feature-rich, category: $t(66) = 3.684,~p < 0.001,~d = 0.220,~
-\mathrm{CI} = [1.733,~5.732]$; feature-rich, size: $t(66) = 1.629,~p = 0.108,~d
-= 0.100,~\mathrm{CI} = [-0.207,~3.905] $; reduced, category: $t(60) = 2.755,~p
-= 0.008,~d = 0.177,~\mathrm{CI} = [0.761,~5.189]$; reduced, size: $t(60) =
-3.081,~p = 0.003,~d = 0.201,~\mathrm{CI} = [1.210,~5.326]$). Participants in
-the reduced (but not feature-rich) conditions tended to exhibit more
-lexicographic clustering on early versus late lists (feature-rich, word length:
-$t(66) = -0.100,~p = 0.921,~d = -0.010,~\mathrm{CI} = [-2.217,~1.899]$; feature
-rich, first letter: $t(66) = 0.412,~p = 0.681,~d = 0.045,~\mathrm{CI} =
-[-1.645,~2.461]$; reduced, word length: $t(60) = 3.762,~p < 0.001,~d = 0.261,~
-\mathrm{CI} = [1.604,~6.821]$; reduced, first letter: $t(60) = 1.721,~p =
-0.090,~d = 0.175,~\mathrm{CI} = [-0.138,~4.098]$). Taken together, these
+late lists in the above feature-rich
+(Tab.~\ref{tab:early-vs-late-feature-rich}) and reduced
+(Tab.~\ref{tab:early-vs-late-reduced}) conditions. Participants in both
+conditions remembered more words on early versus late lists. Participants in
+the feature-rich (but not reduced) conditions exhibited more temporal
+clustering on early versus late lists. And participants in both conditions
+tended to exhibit more semantic clustering on early versus late lists.
+Participants in the reduced (but not feature-rich) conditions tended to exhibit
+more lexicographic clustering on early versus late lists. Taken together, these
 comparisons suggest that even when the presence or absence of incidental visual
@@
 condition and diminished in the reduced (late) condition. However, to our
-surprise, participants in \textit{neither} condition exhibited reliable early-versus-late
-differences in accuracy (reduced (early): $t(41) = 1.499, ~p =
-0.141,~d = 0.098,~\mathrm{CI} = [-0.345,~3.579]$; reduced (late): $t(40) =
-1.462,~p = 0.152,~d = 0.121,~\mathrm{CI} = [-0.376,~2.993]$), temporal
-clustering (reduced (early): $t(41) = 0.857,~p = 0.396,~d = 0.068,~\mathrm{CI}
-= [-1.012,~2.896]$; reduced (late): $t(40) = 1.244,~p = 0.221,~d = 0.128,~
-\mathrm{CI} = [-0.894,~3.088]$), nor feature-based clustering (reduced (early),
-category: $t(41) = 0.707,~p = 0.484,~d = 0.068,~\mathrm{CI} = [-1.314,~2.830]$;
-reduced (early), size: $t(41) = 0.803,~p = 0.427,~d = 0.079,~\mathrm{CI} =
-[-1.142,~2.953]$; reduced (early), length: $t(41) = 0.461,~p = 0.648,~d =
-0.060,~\mathrm{CI} = [-1.545,~2.462]$; reduced (early), first letter: $t(41) =
-0.781,~p = 0.439,~d = 0.101,~\mathrm{CI} = [-1.039,~2.881]$; reduced (late),
-category: $t(40) = 0.101,~p = 0.920,~d = 0.009,~\mathrm{CI} = [-1.776,~2.307]$;
-reduced (late), size: $t(40) = 0.555,~p = 0.582,~d = 0.058,~\mathrm{CI} =
-[-1.444,~2.274]$; reduced (late), length: $t(40) = 1.482,~p = 0.146,~d =
-0.126,~\mathrm{CI} = [-0.444,~3.743]$; reduced (late), first letter: $t(40) =
--0.143,~p = 0.887,~d = -0.017,~\mathrm{CI} = [-2.204,~1.830] $). We
+surprise, participants in \textit{neither} condition exhibited reliable
+early-versus-late differences in accuracy, temporal clustering, nor
+feature-based clustering
+(Tabs.~\ref{tab:early-vs-late-early},~\ref{tab:early-vs-late-late}). We
 hypothesized that adding or removing the variability in the visual features was
 acting as a sort of ``event boundary'' between early and late
-lists~\citep[e.g.,][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we
-(and others) have found that memories formed just after event boundaries can be
+lists~\citep[e.g.,][]{ClewEtal19,RadvZack17,RadvCope06}. In prior work, we (and
+others) have found that memories formed just after event boundaries can be
 enhanced~\citep[e.g., due to less contextual interference between pre- and
 post-boundary items;][]{MannEtal16, PettEtal16, GoldEtal17, FlorEtal17}.
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/feature_rich_vs_reduced_early.tex}
+
+\caption{\textbf{Comparing memory in the feature-rich versus reduced (early)
+conditions (all lists).} The $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all lists from each
+participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:feature-rich-vs-reduced-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/reduced_vs_reduced_early.tex}
+
+\caption{\textbf{Comparing memory in the reduced versus reduced (early)
+conditions (all lists).} The $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all lists from each
+participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:reduced-vs-reduced-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/feature_rich_vs_reduced_late.tex}
+
+\caption{\textbf{Comparing memory in the feature-rich versus reduced (late)
+conditions (all lists).} The $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all lists from each
+participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:feature-rich-vs-reduced-late}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/reduced_vs_reduced_late.tex}
+
+\caption{\textbf{Comparing memory in the reduced versus reduced (late)
+conditions (all lists).} The $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all lists from each
+participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:reduced-vs-reduced-late}
+\end{table}
 
@@
 served to enhance recall performance relative to conditions where all lists had
-the same blends of features (accuracy for feature-rich vs.~reduced (early):
-$t(107) = -2.230,~p = 0.028,~d = -0.439,~\mathrm{CI} = [-4.252,~-0.229]$;
-reduced vs.~reduced (early): $t(101) = -2.045,~p = 0.043,~d = -0.410,~
-\mathrm{CI} = [-3.826,~0.112]$; also see Fig.~\accuracyByList A). However,
-\textit{subtracting} irrelevant visual features on later lists that
-\textit{had} been present on early lists (as in the reduced (late) condition)
-did not appear to impact recall performance (accuracy for feature-rich vs.~
-reduced (late): $t(106) = -0.638,~p = 0.525,~d = -0.126,~\mathrm{CI} = [-2.720,~
-1.362]$; reduced vs.~reduced (late): $t(100) = -0.407,~p = 0.685,~d =
--0.082,~\mathrm{CI} = [-2.477,~1.626]$). These comparisons suggest that recall
-accuracy has a directional component: accuracy is affected differently by
-removing features that had initially been present versus adding features
-that had initially been absent. In contrast, we found that
-participants exhibited more temporal and feature-based clustering when we added
-incidental visual features to \textit{any} lists (comparisons of clustering on
-feature-rich vs.~reduced lists are reported above; temporal clustering in
-reduced vs.~reduced (early) and reduced vs.~reduced (late) conditions:
-$t$s $\leq -9.885$,~$p$s $< 0.001$; feature-based clustering in reduced versus
-reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
--4.555$,~$p$s $< 0.001$). Temporal and feature-based clustering were not
-reliably different in the feature-rich, reduced (early), and reduced (late)
-conditions (temporal clustering in feature-rich vs.~reduced (early) and
-feature-rich vs.~reduced (late) conditions: $t$s $\geq -1.379$,~$p$s $\geq
-0.171$; feature-based clustering in feature-rich vs.~reduced (early) and
-feature-rich vs.~reduced (late) conditions: $|t|$s $\leq 1.441$,~$p$s $\geq
-0.153$).
+the same blends of features
+(Tabs.~\ref{tab:feature-rich-vs-reduced-early},~\ref{tab:reduced-vs-reduced-early};
+also see Fig.~\accuracyByList A). However, \textit{subtracting} irrelevant
+visual features on later lists that \textit{had} been present on early lists
+(as in the reduced (late) condition) did not appear to impact recall
+performance
+(Tabs.~\ref{tab:feature-rich-vs-reduced-late},~\ref{tab:reduced-vs-reduced-late})
+These comparisons suggest that recall accuracy has a directional component:
+accuracy is affected differently by removing features that had initially been
+present versus adding features that had initially been absent. In contrast, we
+found that participants exhibited more temporal and feature-based clustering
+when we added incidental visual features to \textit{any} lists (feature-rich
+versus reduced: Tab.~\ref{tab:feature-rich-vs-reduced}; reduced versus reduced
+(early): Tab.~\ref{tab:reduced-vs-reduced-early}; reduced versus reduced
+(late): Tab.~\ref{tab:reduced-vs-reduced-late}). Temporal and feature-based
+clustering were not reliably different in the feature-rich versus reduced
+(early) or reduced (late) conditions
+(Tabs.~\ref{tab:feature-rich-vs-reduced-early},~\ref{tab:feature-rich-vs-reduced-late}).
 
@@
 
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/order_manip_vs_feature_rich_accuracy_early.tex}
+
+\caption{\textbf{Comparing accuracy on early lists in the order manipulation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:order-manip-vs-feature-rich-accuracy-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/order_manip_vs_feature_rich_temporal_early.tex}
+
+\caption{\textbf{Comparing temporal clustering on early lists in the order manipulation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:order-manip-vs-feature-rich-temporal-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/semantic_vs_feature_rich_feature_early.tex}
+
+\caption{\textbf{Comparing feature-based clustering on early lists in the semantic order manipulation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:semantic-vs-feature-rich-feature-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/lexicographic_vs_feature_rich_feature_early.tex}
+
+\caption{\textbf{Comparing feature-based clustering on early lists in the lexicographic order manipulation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:lexicographic-vs-feature-rich-feature-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/visual_vs_feature_rich_feature_early.tex}
+
+\caption{\textbf{Comparing feature-based clustering on early lists in the visual order manipulation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:visual-vs-feature-rich-feature-early}
+\end{table}
+
 Across each of six order manipulation conditions, we sorted early lists by one
@@
 \textit{Order manipulation conditions}; features: category, size, length, first
-letter, color, and location). Participants in the category-ordered condition
-showed an increase in memory performance on early lists (accuracy relative to
-early feature-rich lists; $t(95) = 3.034,~p = 0.003,~d = 0.667,~\mathrm{CI} =
-[1.048,~5.113]$). Participants in the color-ordered condition also showed a
-trending increase in memory performance on early lists (again, relative to
-early feature-rich lists: $t(96) = 1.850,~p = 0.067,~d = 0.402,~\mathrm{CI} =
-[-0.010,~3.712]$; Fig.~\ref{fig:barplots}A). Participants' performances on
-early lists in all of the other order manipulation conditions were
-indistinguishable from performance on the early feature-rich lists ($|t|$s
-$\leq 1.013,~p$s $\geq 0.314$). Participants in both of the semantically
-ordered conditions exhibited stronger temporal clustering on early lists
-(vs.~early feature-rich lists; category: $t(95) = 8.813,~p < 0.001,~d =
-1.936,~\mathrm{CI} = [6.793,~11.751] $; size: $t(95) = 2.630,~p = 0.010,~d =
-0.578,~\mathrm{CI} = [0.831,~4.866]$; Fig.~\ref{fig:barplots}B). Participants
-in the length-ordered condition tended to exhibit \textit{less} temporal
-clustering on early lists relative to early feature-rich lists ($t(95) =
--1.547,~p = 0.125,~d = -0.340,~\mathrm{CI} = [-3.693,~0.341]$), whereas
-participants in the first letter-ordered condition exhibited stronger temporal
-clustering on early lists ($t(95) = 2.858,~p = 0.005,~d = 0.628,~\mathrm{CI} =
-[1.031,~4.886]$). Participants in the visually ordered conditions exhibited
-more similar performance (accuracy) on early lists relative to early feature
-rich lists (we found a trending enhancement for participants in the
-color-ordered condition: $t(96) = 1.850,~p = 0.067,~d = 0.402,~\mathrm{CI} =
-[-0.010,~3.712]$; location: $t(95) = 0.043,~p = 0.966,~d = 0.010,~\mathrm{CI} =
-[-1.598,~1.729]$). Participants in the visually ordered conditions also showed
-similar temporal clustering on early lists relative to early feature-rich
-lists (color: $t(96) = -1.339,~p = 0.184,~d = -0.291,~\mathrm{CI} = [-3.238,~
-0.394]$, we found a trending increase for participants in the location-ordered
-condition: $t(95) = 1.705,~p = 0.092,~d = 0.374,~\mathrm{CI} = [-0.155,~
-3.521]$). We also compared feature-based clustering on early lists across the
-order manipulation and feature-rich conditions. Since these results were
-similar across both semantic conditions (category and size), both lexicographic
-conditions (length and first letter), and both visual conditions (color and
-location), here we aggregate data from conditions that manipulated each of
-these three feature groupings in our comparisons, to simplify the presentation.
-On early lists, participants in the semantically ordered conditions exhibited
-stronger semantic clustering relative to participants in the feature-rich
-condition (category: $t(125) = 2.722,~p = 0.007,~d = 0.484,~\mathrm{CI} =
-[0.827,~4.932]$; size: $t(125) = 3.866,~p < 0.001,~d = 0.687,~\mathrm{CI} =
-[2.020,~5.983]$), but showed no reliable differences in lexicographic (length:
-$t(125) = 0.521,~p = 0.603,~d = 0.093,~\mathrm{CI} = [-1.311,~2.333]$; first
-letter: $t(125) = -0.842,~p = 0.401,~d = -0.150,~\mathrm{CI} = [-2.825,~
-1.095]$) or visual (color: $t(125) = -0.650,~p = 0.517,~d = -0.116,~\mathrm{CI}
-= [-2.680,~1.249]$; location: $t(125) = -0.251,~p = 0.802,~d = -0.045,
-\mathrm{CI} = [-2.257,~1.524]$) clustering. Similarly, participants in the
-lexicographically ordered conditions exhibited stronger (relative to feature
-rich participants) lexicographic clustering (length: $t(125) = 3.682,~p <
-0.001,~d = 0.655,~\mathrm{CI} = [1.890,~5.569]$; first letter: $t(125) = 5.134,
-p < 0.001,~d = 0.912,~\mathrm{CI} = [3.251,~7.258]$) on early lists, but showed
-no reliable differences in semantic (category: $t(125) = -1.040,~p = 0.301,~d =
--0.185,~\mathrm{CI} = [-3.095,~1.092]$; size: $t(125) = 0.006,~p = 0.995,~d =
-0.001,~\mathrm{CI} = [-1.933,~1.952]$) or visual (color: $t(125) = 0.092,~p =
-0.927,~d = 0.016,~\mathrm{CI} = [-1.834,~1.867]$; location: $t(125) = 0.407, p
-= 0.685,~d = 0.072,~\mathrm{CI} = [-1.655,~2.463]$) clustering. And
-participants in the visually ordered conditions exhibited stronger visual
-clustering (again, relative to feature-rich participants, and on early lists;
-color: $t(126) = 2.022,~p = 0.045,~d = 0.358,~\mathrm{CI} = [0.056,~3.965]$;
-location: $t(126) = 4.390,~p < 0.001,~d = 0.777,~\mathrm{CI} = [2.730,~6.199]
-$), but showed no reliable differences in semantic (category: $t(126) = 0.012,
-p = 0.991,~d = 0.002,~\mathrm{CI} = [-1.988,~1.871]$; size: $t(126) = -0.104, p
-= 0.917,~d = -0.018,~\mathrm{CI} = [-2.166,~1.847] $) or lexicographic (length:
-$t(126) = 0.592,~p = 0.555,~d = 0.105,~\mathrm{CI} = [-1.361,~2.420]$; first
-letter: $t(126) = 0.040,~p = 0.968,~d = 0.007,~\mathrm{CI} = [-1.791,~1.863]$)
-clustering. Taken together, these order manipulation results suggest several
-broad patterns (Figs.~\ref{fig:recall-dynamics}A,~\ref{fig:fingerprints}).
-First, most of the order manipulations we carried out did \textit{not} reliably
-affect overall recall performance. Second, most of the order manipulations
-increased participants' tendencies to temporally cluster their recalls. Third,
-all of the order manipulations enhanced participants' clustering of each
-condition's target feature (i.e., semantic manipulations enhanced semantic
-clustering, lexicographic manipulations enhanced lexicographic clustering, and
-visual manipulations enhanced visual clustering; Fig.~\ref{fig:barplots}C)
-while leaving clustering along other feature dimensions roughly unchanged
-(i.e., semantic manipulations did not affect lexicographic or visual
-clustering, and so on). Although it is not possible to fully separate feature-based
-versus temporal clustering when considering sorted lists, we used a
-permutation-based procedure to identify the degree of feature clustering over
-and above what could be accounted for by temporal clustering alone (see
-\textit{Factoring out the effects of temporal clustering}). When we carried out
-this analysis (Fig.~\ref{fig:barplots}D), we found that participants exhibited
-more semantic clustering on semantically sorted lists than on randomly ordered
-lists, but the effects of the other order manipulations could not reliably be
-separated from temporal clustering alone (reliable comparisons are reported in
-the figure).
+letter, color, and location). When we compared participants' memories for early
+lists in each of these conditions to their memories for early lists in the
+feature-rich condition
+(Tab.~\ref{tab:order-manip-vs-feature-rich-accuracy-early}), we found that
+participants in the category-ordered condition remembered more words than
+participants in the feature-rich condition. Participants in the Participants in
+the color-ordered condition also showed a trending increase in memory
+performance on early lists. Participants' performances on early lists in all of
+the other order manipulation conditions were indistinguishable from performance
+on the early feature-rich lists. We also compared participants' temporal
+clustering on early lists in each of these conditions to their temporal
+clustering on early lists in the feature-rich condition
+(Tab.~\ref{tab:order-manip-vs-feature-rich-temporal-early}). Participants in
+both of the semantically ordered conditions exhibited stronger temporal
+clustering on early lists (vs.~early feature-rich lists). Participants in the
+length-ordered condition tended to exhibit \textit{less} temporal clustering on
+early lists relative to early feature-rich lists, whereas participants in the
+first letter-ordered condition exhibited stronger temporal clustering on early
+lists. Participants in the visually ordered conditions showed similar temporal
+clustering on early lists relative to early feature-rich lists. We also also
+compared feature-based clustering on early lists across the order manipulation
+and feature-rich conditions. Since these results were similar across both
+semantic conditions (category and size;
+Tab.~\ref{tab:semantic-vs-feature-rich-feature-early}), both lexicographic
+conditions (length and first letter;
+Tab.~\ref{tab:lexicographic-vs-feature-rich-feature-early}), and both visual
+conditions (color and location;
+Tab.~\ref{tab:visual-vs-feature-rich-feature-early}), here we aggregate data
+from conditions that manipulated each of these three feature groupings in our
+comparisons, to simplify the presentation. On early lists, participants in the
+semantically ordered conditions exhibited stronger semantic clustering relative
+to participants in the feature-rich condition, but showed no reliable
+differences in lexicographic or visual clustering. Similarly, participants in
+the lexicographically ordered conditions exhibited stronger (relative to
+feature rich participants) lexicographic clustering on early lists, but showed
+no reliable differences in semantic or visual clustering. And participants in
+the visually ordered conditions exhibited stronger visual clustering (again,
+relative to feature-rich participants, and on early lists), but showed no
+reliable differences in semantic or lexicographic clustering. Taken together,
+these order manipulation results suggest several broad patterns
+(Figs.~\ref{fig:recall-dynamics}A,~\ref{fig:fingerprints}). First, most of the
+order manipulations we carried out did \textit{not} reliably affect overall
+recall performance. Second, most of the order manipulations increased
+participants' tendencies to temporally cluster their recalls. Third, all of the
+order manipulations enhanced participants' clustering of each condition's
+target feature (i.e., semantic manipulations enhanced semantic clustering,
+lexicographic manipulations enhanced lexicographic clustering, and visual
+manipulations enhanced visual clustering; Fig.~\ref{fig:barplots}C) while
+leaving clustering along other feature dimensions roughly unchanged (i.e.,
+semantic manipulations did not affect lexicographic or visual clustering, and
+so on). Although it is not possible to fully separate feature-based versus
+temporal clustering when considering sorted lists, we used a permutation-based
+procedure to identify the degree of feature clustering over and above what
+could be accounted for by temporal clustering alone (see \textit{Factoring out
+the effects of temporal clustering}). When we carried out this analysis
+(Fig.~\ref{fig:barplots}D), we found that participants exhibited more semantic
+clustering on semantically sorted lists than on randomly ordered lists, but the
+effects of the other order manipulations could not reliably be separated from
+temporal clustering alone (reliable comparisons are reported in the figure).
 
@@
 
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/order_accuracy_comparisons_early.tex}
+
+\caption{\textbf{Comparing accuracy on early lists in different order manipulation
+conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:order-accuracy-comparisons-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/order_temporal_comparisons_early.tex}
+
+\caption{\textbf{Comparing temporal clustering on early lists in different order manipulation
+conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:order-temporal-comparisons-early}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/feature_comparisons_early.tex}
+
+\caption{\textbf{Comparing feature-based clustering on early lists in different order manipulation
+conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all early lists
+from each participant. The feature used to compute clustering is shown before the colon in each row, and the
+conditions being compared are shown after the colon.  Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:feature-comparisons-early}
+\end{table}
+
 The above comparisons between memory performance on early lists in the order
@@
 along each feature dimension influenced memory relative to sorting lists along
-the other feature dimensions. Participants trended towards remembering early
-lists that were sorted semantically better than those sorted lexicographically
-($t(118) = 1.936,~p = 0.055,~d = 0.353,~\mathrm{CI} = [0.057,~3.916]$).
-Participants also remembered visually sorted lists better than
-lexicographically sorted lists ($t(119) = 2.145,~p = 0.034,~d = 0.390,~
-\mathrm{CI} = [0.208,~4.254]$). However, participants showed no reliable
-differences in recall for semantically versus visually sorted lists ($t(119) =
-0.113,~p = 0.910,~d = 0.021,~\mathrm{CI} = [-1.987,~2.097]$). Participants
-temporally clustered semantically sorted lists more strongly than lists sorted either
-lexicographically ($t(118) = 5.620,~p < 0.001,~d = 1.026,~\mathrm{CI} = [3.486,~
-8.010]$) or visually ($t(119) = 6.613,~p < 0.001,~d = 1.202,~\mathrm{CI} =
-[4.481,~9.464]$), but did not show reliable differences in
-temporal clustering on lexicographically versus visually sorted lists ($t(119)
-= 0.589,~p = 0.557,~d = 0.107,~\mathrm{CI} = [-1.336,~2.539]$). Participants
-also showed reliably more semantic clustering on semantically sorted lists than
-lexicographically (category: $t(118) = 3.667,~p < 0.001,~d = 0.670,~\mathrm{CI}
-= [1.822,~5.942]$; size: $t(118) = 3.972,~p < 0.001$) or visually (category:
-$t(119) = 2.702,~p = 0.008$; size: $t(118) = 4.043,~p < 0.001,~d = 0.738,~
-\mathrm{CI} = [2.145,~6.296]$) sorted lists; more lexicographic clustering on
-lexicographically sorted lists than semantically (length: $t(118) = 3.390,~p <
-0.001,~d = 0.619,~\mathrm{CI} = [1.499,~5.661]$; first letter: $t(118) = 5.705,~
-p < 0.001,~d = 1.042,~\mathrm{CI} = [3.841,~7.790]$) or visually (length:
-$t(119) = 3.399,~p < 0.001,~d = 0.618,~\mathrm{CI} = [1.500,~5.527]$; first
-letter: $t(119) = 4.859,~p < 0.001,~d = 0.883,~\mathrm{CI} = [2.860,~6.849]$)
-sorted lists; and more visual clustering on visually sorted lists than
-semantically (color: $t(119) = 2.673,~p = 0.009,~d = 0.486,~\mathrm{CI} =
-[0.848,~4.567]$; location: $t(119) = 4.499,~p < 0.001,~d = 0.818,~\mathrm{CI} =
-[2.721,~6.399]$) or lexicographically (color: $t(119) = 1.988,~p = 0.049,~d =
-0.361,~\mathrm{CI} = [0.102,~3.894]$; location: $t(119) = 3.966,~p < 0.001,~d =
-0.721,~\mathrm{CI} = [2.099,~5.862]$) sorted lists. In summary, sorting lists
-by different features appeared to have slightly different effects on overall
+the other feature dimensions (accuracy:
+Tab.~\ref{tab:order-manip-vs-feature-rich-accuracy-early}; temporal clustering:
+Tab.~\ref{tab:order-manip-vs-feature-rich-temporal-early}; feature-based
+clustering: Tab~\ref{tab:feature-comparisons-early}). Participants trended
+towards remembering early lists that were sorted semantically better than those
+sorted lexicographically. Participants also remembered visually sorted lists
+better than lexicographically sorted lists. However, participants showed no
+reliable differences in recall for semantically versus visually sorted lists.
+Participants temporally clustered semantically sorted lists more strongly than
+lists sorted either lexicographically or visually, but did not show reliable
+differences in temporal clustering on lexicographically versus visually sorted
+lists. Participants also showed reliably more semantic clustering on
+semantically sorted lists than lexicographically or visually sorted lists; more
+lexicographic clustering on lexicographically sorted lists than semantically or
+visually sorted lists; and more visual clustering on visually sorted lists than
+semantically or lexicographically sorted lists. In summary, sorting lists by
+different features appeared to have slightly different effects on overall
 memory performance and temporal clustering. Participants also tended to cluster
@@
 
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/carryover_accuracy.tex}
+
+\caption{\textbf{Comparing accuracy on late lists in order-manipulatation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all late
+lists from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:accuracy-carryover}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/carryover_temporal_clustering.tex}
+
+\caption{\textbf{Comparing temporal clustering on late lists in order-manipulatation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all late
+lists from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:temporal-carryover}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/semantic_carryover_feature_clustering.tex}
+
+\caption{\textbf{Comparing feature-based clustering on late lists in semantic order-manipulatation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all late
+lists from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:semantic-feature-carryover}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/lexicographic_carryover_feature_clustering.tex}
+
+\caption{\textbf{Comparing feature-based clustering on late lists in
+lexicographic order-manipulatation versus feature-rich conditions.} The
+$t$-tests reported in the table were carried out across-participants, and
+reflect data aggregated across all late lists from each participant.
+Abbreviations used in this table are defined in Table~\abbreviations.}
+
+\label{tab:lexicographic-feature-carryover}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/visual_carryover_feature_clustering.tex}
+
+\caption{\textbf{Comparing feature-based clustering on late lists in visual order-manipulatation
+versus feature-rich conditions.} The $t$-tests reported in the table were
+carried out across-participants, and reflect data aggregated across all late
+lists from each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:visual-feature-carryover}
+\end{table}
+
 At the group level, there seemed to be almost no lingering impact of sorting
 early lists on memory for later lists. To simplify the presentation, we report
-these null results in aggregate across the three feature groupings. Relative to
-memory performance on late feature-rich lists, participants' memory performance
-in all six order manipulation conditions showed no reliable differences
-(semantic: $t(125) = 0.487,~p = 0.627,~d = 0.087,~\mathrm{CI} = [-1.661,~
-2.323]$; lexicographic: $t(125) = 0.878,~p = 0.382,~d = 0.156,~\mathrm{CI} =
-[-1.226,~3.044]$; visual: $t(126) = 1.437,~p = 0.153,~d = 0.254,~\mathrm{CI} =
-[-0.447,~3.519]$). Nor did we observe any reliable differences in temporal
-clustering on late lists (relative to late feature-rich lists; semantic:
-$t(125) = 0.157,~p = 0.875,~d = 0.028,~\mathrm{CI} = [-1.859,~1.974]$;
-lexicographic: $t(125) = 0.998,~p = 0.320,~d = 0.177,~\mathrm{CI} = [-0.902,~
-2.920]$; visual: $t(126) = 0.548,~p = 0.585,~d = 0.097,~\mathrm{CI} = [-1.450,~
-2.365]$). Aside from a slightly increased tendency for participants to cluster
-words by their length on late visual order manipulation lists (more than late
-feature-rich lists; $t(126) = 2.005,~p = 0.047,~d = 0.355,~\mathrm{CI} =
-[0.211,~3.722]$), we observed no reliable differences in any type of feature
-clustering on late order manipulation condition lists versus late feature-rich
-lists ($|t|$s $\leq 1.124,~p$s $\geq 0.263$).
+these null results in aggregate across the three feature groupings (accuracy:
+Tab.~\ref{tab:accuracy-carryover}; temporal clustering:
+Tab.~\ref{tab:temporal-carryover}; feature-based clustering:
+Tabs.~\ref{tab:semantic-feature-carryover},~\ref{tab:lexicographic-feature-carryover},
+and~\ref{tab:visual-feature-carryover}). Relative to memory performance on late
+feature-rich lists, participants' memory performance in all six order
+manipulation conditions showed no reliable differences. Nor did we observe any
+reliable differences in temporal clustering on late lists (relative to late
+feature-rich lists). Aside from a slightly increased tendency for participants
+to cluster words by their length on late visual order manipulation lists (more
+than late feature-rich lists), we observed no reliable differences in any type
+of feature clustering on late order manipulation condition lists versus late
+feature-rich lists.
 
@@
 
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/rank_table.tex}
+
+\caption{\textbf{Ranked clustering scores versus ``chance'' for participants in
+the feature-rich condition.} For each participant, we sorted their clustering
+scores in descending order (for each of the six feature dimensions, along with
+a seventh dimension to capture temporal clustering). The $t$-tests reported in
+the table (for the clustering scores at each ``rank'') were carried out
+across-participants, and reflect data aggregated across all lists from each
+participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
+\label{tab:ranked-clustering}
+\end{table}
+
 We note that the multivariate aspect of the adaptive condition (i.e., sorting
@@
 at least three dimensions, including temporal clustering (which was often
-ranked highest); Rank 1: $t(66) = 12.751,~p < 0.001,~d = 0.162,~\mathrm{CI} =
-[8.702,~20.013]$; Rank 2: $t(66) = 8.196,~p < 0.001,~d = 0.162,~\mathrm{CI} =
-[4.794,~12.978]$; Rank 3: $t(66) = 3.243,~p = 0.002,~d = 0.162,~\mathrm{CI} =
-[1.028,~7.051]$; Rank 4: $t(66) = -3.112,~p = 0.003,~d = 0.162,~\mathrm{CI} =
-[-5.282,~-1.920]$; Rank 5: $t(66) = -7.154,~p < 0.001,~d = 0.162,~\mathrm{CI} =
-[-12.649,~-5.568]$; Rank 6: $t(66) = -12.608,~p < 0.001,~d = 0.162,~\mathrm{CI}
-= [-22.114,~-9.347]$; Rank 7: $t(66) = -18.397,~p < 0.001,~d = 0.162,
-\mathrm{CI} = [-27.238,~-14.073]$.
+ranked highest; Tab.~\ref{tab:ranked-clustering}).
 

### 2023-11-20 code/analyze.py

@@ -663,6 +663,7 @@ def ttest(x, y, x_col=None, y_col=None, x_lists=None, y_lists=None, independent_
         p_string = f'p = {p:.3f}'
 
     print(f't({df}) = {t:.3f},~{p_string},~d = {d:.3f},~' + '\mathrm{CI}' + f' = [{low_conf:.3f},~{high_conf:.3f}]')
+    return t, df, d, p, low_conf, high_conf
     
 
 def stack_fried_eggs(*args):

### 2023-11-21 paper/main.tex
highlight significant rows in tables

@@
 \usepackage{booktabs}
+\usepackage[table]{xcolor}
 
@@
 
-We next asked whether adding these incidental visual features to later lists
-(after the participants had already studied impoverished lists), or removing
-the visual features from later lists (after the participants had already
-studied visually diverse lists) might affect memory performance. In other
-words, we sought to test for potential effects of changing the ``richness'' of
-participants' experiences over time. All participants studied and recalled a
-total of 16 lists; we defined \textit{early} lists as the first eight lists and
-\textit{late} lists as the last eight lists each participant encountered. To
-help interpret our results, we compared participants' memories on early versus
-late lists in the above feature-rich
-(Tab.~\ref{tab:early-vs-late-feature-rich}) and reduced
+Beyond affecting participants' memories for individual lists, we wondered how
+studying early lists (with versus without incidental visual features) might
+affect how participants remembered later lists (again, with versus without
+incidental visual features). In other words, we sought to test for potential
+effects of changing the ``richness'' of participants' experiences over time.
+All participants studied and recalled a total of 16 lists; we defined
+\textit{early} lists as the first eight lists and \textit{late} lists as the
+last eight lists each participant encountered. To help interpret our results,
+we compared participants' memories on early versus late lists in the above
+feature-rich (Tab.~\ref{tab:early-vs-late-feature-rich}) and reduced
 (Tab.~\ref{tab:early-vs-late-reduced}) conditions. Participants in both
 conditions remembered more words on early versus late lists. Participants in
-the feature-rich (but not reduced) conditions exhibited more temporal
+the feature-rich (but not reduced) condition exhibited more temporal
 clustering on early versus late lists. And participants in both conditions
@@
 images~\citep{GersEtal13, MannEtal16, ChiuEtal21}, background
-sounds~\citep{BeamJone10, SahaSmit14}, secondary tasks~\citep{MasiSaha14,
+sounds~\citep{BeamJone98, SahaSmit14}, secondary tasks~\citep{MasiSaha14,
 PolyEtal09, OberLewa08}, and more can all impact how participants remember, and

### 2023-11-21 paper/main.tex
another round of JRM edits

@@
 words on the to-be-presented list. Second, we compute the distances between the
-chosen word's feature(s) value(s) and the corresponding feature(s) value(s) of
-all yet-to-be-presented words. Third, we convert these distances (between the
+chosen word's feature value(s) and the corresponding feature value(s) of all
+yet-to-be-presented words. Third, we convert these distances (between the
 previously presented word's feature values, $a$, and each of the $W$
@@
 \mathrm{similarity}_{\mathrm{normalized}}(a, b_i) = \frac{\mathrm{similarity}(a,
-b)}{\sum_{j=1}^W \mathrm{similarity}(a, b_j)},
+b_i)}{\sum_{j=1}^W \mathrm{similarity}(a, b_j)},
 \end{equation}
@@
 fingerprints (for all lists and experimental conditions) to create a large
-matrix with number-of-lists (16) by number-of-conditions (10, including the
-adaptive condition) rows and seven columns (one for each feature clustering
+matrix with number-of-lists (16) $\times$ number-of-conditions (10, excluding
+the adaptive condition) rows and seven columns (one for each feature clustering
 score, plus an additional temporal clustering score column). We used principal
@@
 category and size) are correlated, then the clustering scores for those two
-dimensions will also be correlated. When lists are sorted along a given feature
-dimension, the sorted feature values will also tend to be correlated with the
-serial positions of the words in the list. This means that the temporal
-clustering score will \textit{also} tend to be correlated with the clustering
-scores for the sorted feature dimension. These correlations mean that it can be
-difficult to specifically identify when participants are using one feature
-versus another (or a manipulated feature versus temporal information) to
-organize or search their memories.
+dimensions will also be correlated (Figs.~\clusterCorrs,~\featureCorrs). When
+lists are sorted along a given feature dimension, the sorted feature values
+will also tend to be correlated with the serial positions of the words in the
+list. This means that the temporal clustering score will \textit{also} tend to
+be correlated with the clustering scores for the sorted feature dimension.
+These correlations mean that it can be difficult to specifically identify when
+participants are using one feature versus another (or a manipulated feature
+versus temporal information) to organize or search their memories.
 
@@
 occurring between any successive pair of words whose distances along the given
-feature dimension were greater than one standard deviation above the mean for
+feature dimension were greater than two standard deviations above the mean for
 that list. Note that, because event boundaries are defined for each feature
@@
 DrewMurd80, SochEtal09} or impair recall~\citep[e.g., by distracting or
-confusing participants with irrelevant information][]{MarsEtal15, MarsEtal12,
+confusing participants with irrelevant information;][]{MarsEtal15, MarsEtal12,
 Lang05, ReinEtal92}. Second, we manipulated the orders in which words were
@@
 0.162,~\mathrm{CI} = [3.086,~6.895]$). That within-participant fingerprint
-similarity was greater than between-participant fingerprint similarity suggests
-that participants' memory fingerprints are relatively stable across lists, and
-that each participant's fingerprint is unique to them.
+similarity (across lists) was greater than across-participant fingerprint
+similarity suggests that participants' memory fingerprints are relatively
+stable across lists, and that each participant's fingerprint is unique to them.
 
@@
 participants in the category-ordered condition remembered more words than
-participants in the feature-rich condition. Participants in the Participants in
-the color-ordered condition also showed a trending increase in memory
-performance on early lists. Participants' performances on early lists in all of
-the other order manipulation conditions were indistinguishable from performance
-on the early feature-rich lists. We also compared participants' temporal
-clustering on early lists in each of these conditions to their temporal
-clustering on early lists in the feature-rich condition
+participants in the feature-rich condition. Participants' performances on early
+lists in all of the other order manipulation conditions were indistinguishable
+from performance on the early feature-rich lists. We also compared
+participants' temporal clustering on early lists in each of these conditions to
+their temporal clustering on early lists in the feature-rich condition
 (Tab.~\ref{tab:order-manip-vs-feature-rich-temporal-early}). Participants in
-both of the semantically ordered conditions exhibited stronger temporal
-clustering on early lists (vs.~early feature-rich lists). Participants in the
-length-ordered condition tended to exhibit \textit{less} temporal clustering on
-early lists relative to early feature-rich lists, whereas participants in the
-first letter-ordered condition exhibited stronger temporal clustering on early
-lists. Participants in the visually ordered conditions showed similar temporal
-clustering on early lists relative to early feature-rich lists. We also also
-compared feature-based clustering on early lists across the order manipulation
-and feature-rich conditions. Since these results were similar across both
-semantic conditions (category and size;
+both of the semantically ordered conditions and the first letter-ordered
+condition all exhibited stronger temporal clustering on early lists (vs.~early
+feature-rich lists). Participants in the other order manipulation conditions
+all showed similar temporal clustering on early lists relative to early
+feature-rich lists. We also also compared feature-based clustering on early
+lists across the order manipulation and feature-rich conditions. Since these
+results were similar across both semantic conditions (category and size;
 Tab.~\ref{tab:semantic-vs-feature-rich-feature-early}), both lexicographic
@@
 the other feature dimensions (accuracy:
-Tab.~\ref{tab:order-manip-vs-feature-rich-accuracy-early}; temporal clustering:
-Tab.~\ref{tab:order-manip-vs-feature-rich-temporal-early}; feature-based
+Tab.~\ref{tab:order-accuracy-comparisons-early}; temporal clustering:
+Tab.~\ref{tab:order-temporal-comparisons-early}; feature-based
 clustering: Tab~\ref{tab:feature-comparisons-early}). Participants trended
-towards remembering early lists that were sorted semantically better than those
-sorted lexicographically. Participants also remembered visually sorted lists
+towards remembering early lists that were sorted visually
 better than lexicographically sorted lists. However, participants showed no
-reliable differences in recall for semantically versus visually sorted lists.
+reliable differences in recall for semantically versus lexicographically or visually sorted lists.
 Participants temporally clustered semantically sorted lists more strongly than
@@
 individual conditions: all $r$s $\geq 0.298$, all $p$s $\leq 0.110$; across
-conditions: $r(4) = 0.064,~p = 0.903,~\mathrm{CI} = [-0.972,$). Taken together,
+conditions: $r(4) = 0.064,~p = 0.903,~\mathrm{CI} = [-0.972,~0.965$). Taken together,
 the results displayed in Figure~\ref{fig:clustering-scatterplots} show that
@@
 weaker feature clustering on late lists
-(Fig.~\ref{fig:fingerprint-trajectories}) on \textit{average}, the associations
-between early and late lists for individual participants suggests that some
-influence of early order manipulations may linger on late lists. We found that
-participants who exhibited larger carryover in feature clustering (i.e.,
-continued to show strong feature clustering on late lists) for the semantic
-order manipulations (but not other manipulations) also tended to show a smaller
-decrease in recall on early versus late lists
+(Figs.~\ref{fig:fingerprints},~\ref{fig:barplots}) on \textit{average}, the
+associations between early and late lists for individual participants suggests
+that some influence of early order manipulations may linger on late lists. We
+found that participants who exhibited larger carryover in feature clustering
+(i.e., continued to show strong feature clustering on late lists) for the
+semantic order manipulations (but not other manipulations) also tended to show
+a smaller decrease in recall on early versus late lists
 (Fig.~\ref{fig:clustering-carryover}B; overall: $r(179) = 0.307,~p < 0.001,
@@
 Because the presentation orders on stabilize and destabilize lists are adjusted
-to best match each participant's (potentially unique) memory fingerprint, the
-adaptive condition removes uncertainty about whether participants' assigned
-conditions might just ``happen'' to match their preferred ways of organizing
-their memories.
+to best match each participant's personal memory fingerprint, the adaptive
+condition removes uncertainty about whether participants' assigned conditions
+might just ``happen'' to match their preferred ways of organizing their
+memories.
 
@@
 lists simultaneously along multiple feature dimensions) provides an important
-contrast with the order order manipulation conditions, where we sort lists
+contrast with the other order manipulation conditions, where we sort lists
 along only a single feature dimension in each condition. We found that
@@
 at least three dimensions, including temporal clustering (which was often
-ranked highest; Tab.~\ref{tab:ranked-clustering}).
+ranked highest; Tab.~\ref{tab:ranked-clustering}). That the clustering scores
+ranked in the top three dimensions were reliably greater than chance suggests
+that participants organize their memories along at least three feature
+dimensions, even when the words are studied in a random order.
 
@@
 \textit{increases} in richness (e.g., relative to a recency-weighted baseline)
-may be more important than the overall complexity or numbers of features.
+may be more important than the absolute complexity or numbers of meaningful
+features.
 
@@
 different ``strategies'' or ``pathways'' for searching for the target
-information. For example, modern accounts of free recall typically posit that
+information. For example, modern accounts of free recall often posit that
 memory search typically begins by matching the current state of mental context
@@
 
-In real-world experience, each moment's unique blend of contextual features
+In real-world experiences, each moment's unique blend of contextual features
 (where we are, who we are with, what else we are thinking of at the time, what
@@
 sensory properties (i.e., feature related to the item's appearance, sound,
-smell, etc.), emotional properties (i.e., features related to how meaningful the
-item is, whether the item evokes positive or negative feelings, etc.),
+smell, etc.), emotional properties (i.e., features related to how meaningful
+the item is, whether the item evokes positive or negative feelings, etc.),
 utility-related properties (e.g., features that describe how an item might be

### 2023-11-21 paper/main.tex
recompiled changes doc

@@
-%! suppress = LineBreak
 \documentclass[11pt]{article}
@@
 
-\caption{\textbf{Comparing accuracy on late lists in order-manipulatation
+\caption{\textbf{Comparing accuracy on late lists in order-manipulation
 versus feature-rich conditions.} The $t$-tests reported in the table were
@@
 
-\caption{\textbf{Comparing temporal clustering on late lists in order-manipulatation
+\caption{\textbf{Comparing temporal clustering on late lists in order-manipulation
 versus feature-rich conditions.} The $t$-tests reported in the table were
@@
 
-\caption{\textbf{Comparing feature-based clustering on late lists in semantic order-manipulatation
+\caption{\textbf{Comparing feature-based clustering on late lists in semantic order-manipulation
 versus feature-rich conditions.} The $t$-tests reported in the table were
@@
 \caption{\textbf{Comparing feature-based clustering on late lists in
-lexicographic order-manipulatation versus feature-rich conditions.} The
+lexicographic order-manipulation versus feature-rich conditions.} The
 $t$-tests reported in the table were carried out across-participants, and
@@
 
-\caption{\textbf{Comparing feature-based clustering on late lists in visual order-manipulatation
+\caption{\textbf{Comparing feature-based clustering on late lists in visual order-manipulation
 versus feature-rich conditions.} The $t$-tests reported in the table were

### 2023-11-21 paper/supplement.tex
moved most stats to tables
updated abbreviations table, removed page number from first page of supplement, re-ran stats notebook for consistency

@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 \renewcommand{\bibnumfmt}[1]{[S#1]}
@@
 \maketitle
+\thispagestyle{empty}
+
+\begin{table}[p]
+    \centering
+    \input{tables/abbreviations.tex}
+
+    \caption{\textbf{List of abbreviations.} Used in tables in the main text.}
+
+    \label{tab:abbreviations}
+\end{table}
 
@@
 panels.} The conditional probabilities of recalling each word are plotted as a
-function of the relative position (lag) to the word recalled just-prior.
+function of the relative position (Lag) to the word recalled just-prior.
 \textbf{Right panels.} The overall probabilities of recalling each word are
@@
 
-    \caption{\textbf{Recall dynamics in feature-rich free recall (adaptive condition).} \textbf{Left panel.} The probabilities of
-    initiating recall with each word are plotted as a function of presentation
-    position. \textbf{Middle panel.} The conditional probabilities of recalling
-    each word are plotted as a function of the relative position (lag) to the word
+    \caption{\textbf{Recall dynamics in feature-rich free recall (adaptive
+    condition).} \textbf{Left panel.} The probabilities of initiating recall
+    with each word are plotted as a function of presentation position.
+    \textbf{Middle panel.} The conditional probabilities of recalling each word
+    are plotted as a function of the relative position (Lag) to the word
     recalled just-prior. \textbf{Right panel.} The overall probabilities of
     recalling each word are plotted as a function of presentation position.
-    \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\% confidence
-    intervals (calculated across participants). Word-sorting policy (batch) is denoted by color.}
+    \textbf{All panels.} Error ribbons denote bootstrap-estimated 95\%
+    confidence intervals (calculated across participants). Word-sorting policy
+    (batch) is denoted by color.}
 
@@
 studied lists for the random conditions (\textbf{A.}), order manipulation
-conditions (\textbf{B.}), and word-sorting policy (batch) in the adaptive condition (\textbf{C.}). The conditions (or batches)
-are denoted by color. Note that words in the first four lists of the ``adaptive''
-condition were ordered randomly to compute a baseline fingerprint for each
-participant prior to initiating the adaptive ordering procedure. \textbf{All
-panels.} Error ribbons denote bootstrap-estimated 95\% confidence
-intervals (calculated across participants).} 
-\label{fig:accuracy-by-list}
+conditions (\textbf{B.}), and word-sorting policy (batch) in the adaptive
+condition (\textbf{C.}). The conditions (or batches) are denoted by color. Note
+that words in the first four lists of the ``adaptive'' condition were ordered
+randomly to compute a baseline fingerprint for each participant prior to
+initiating the adaptive ordering procedure. \textbf{All panels.} Error ribbons
+denote bootstrap-estimated 95\% confidence intervals (calculated across
+participants).} \label{fig:accuracy-by-list}
 
@@
 
-    \caption{\textbf{Correlations between feature clustering scores (order manipulation conditions).}  Each column reflects one experimental condition.  The matrices in the top and middle
-    rows display across-participant correlations between clustering scores for each feature dimension (top: order manipulation lists; middle: randomly ordered lists).  The matrices
-    in the bottom row display the differences between the top and middle rows.}
+    \caption{\textbf{Correlations between feature clustering scores (order
+    manipulation conditions).} Each column reflects one experimental condition.
+    The matrices in the top and middle rows display across-participant
+    correlations between clustering scores for each feature dimension (top:
+    early lists; middle: late lists). The matrices in the bottom row display
+    the differences between the top and middle rows.}
+
         \label{fig:clustering-correlations}

### 2023-11-21 paper/main.tex

@@
 between two words as the Euclidean distances between their $(r, g, b)$ color
-vectors and $(x, y)$ location vectors (specified as percentages of screen width
-and height), respectively. Therefore, the color and location distance measures
-always take on non-negative real values (upper-bounded at 439.94 for color, or
-124.52 for location, reflecting the distances between the corresponding
-maximally different vectors).
+vectors and $(x, y)$ location vectors (specified as a percentage of the
+viewable display's width), respectively. Therefore, the color and location
+distance measures always take on non-negative real values (upper-bounded at
+439.94 for color, or 124.52 for location, reflecting the distances between the
+corresponding maximally different vectors).
 
@@
 a given set of recalled items (whose presentation positions are given by $x_1,
-x_2, x_3, …, x_N$), we circularly shifted the presentation positions by a
-randomly chosen amount (between 1 and the list length) to obtain a new set of
+x_2, x_3, …, x_L$), we circularly shifted the presentation positions by a
+randomly chosen amount (between 1 and the list length, $L$) to obtain a new set of
 items at the (now altered) positions of the original recalls. Since the new set
@@
 clustering score within the distributions of circularly shifted feature
-clustering scores, across $N = 500$ repetitions of this procedure. This new
+clustering scores, across $n = 500$ repetitions of this procedure. This new
 temporally corrected score provides an estimate of the observed degree of
@@
 All of the data analyzed in this manuscript, along with all of the code for
-carrying out the analyses, may be found at https://github.com/ContextLab/FRFR-analyses.
-Code for running the non-adaptive
+carrying out the analyses, may be found at
+https://github.com/ContextLab/FRFR-analyses. Code for running the non-adaptive
 experimental conditions may be found at
@@
 \caption{\textbf{Comparing memory in the feature-rich versus reduced conditions
-(all lists).} The $t$-tests reported in the table were carried out
+(all lists).} The independent samples $t$-tests reported in the table were carried out
 across-participants, and reflect data aggregated across all lists from each
@@
 \caption{\textbf{Comparing memory for early versus late lists in the
-feature-rich condition.} The $t$-tests reported in the table were carried out
-across-participants. Abbreviations used in this table are defined in
+feature-rich condition.} The paired $t$-tests reported in the table were
+carried out within-participant. Abbreviations used in this table are defined in
 Table~\abbreviations.}
@@
 
-\caption{\textbf{Comparing memory for early versus late lists in the
-reduced condition.} The $t$-tests reported in the table were carried out
-across-participants. Abbreviations used in this table are defined in
+\caption{\textbf{Comparing memory for early versus late lists in the reduced
+condition.} The paired $t$-tests reported in the table were carried out
+within-participant. Abbreviations used in this table are defined in
 Table~\abbreviations.}
@@
 
-\caption{\textbf{Comparing memory for early versus late lists in the
-reduced (early) condition.} The $t$-tests reported in the table were carried out
-across-participants. Abbreviations used in this table are defined in
+\caption{\textbf{Comparing memory for early versus late lists in the reduced
+(early) condition.} The paired $t$-tests reported in the table were carried out
+within-participant. Abbreviations used in this table are defined in
 Table~\abbreviations.}
@@
 
-\caption{\textbf{Comparing memory for early versus late lists in the
-reduced (late) condition.} The $t$-tests reported in the table were carried out
-across-participants. Abbreviations used in this table are defined in
+\caption{\textbf{Comparing memory for early versus late lists in the reduced
+(late) condition.} The paired $t$-tests reported in the table were carried out
+within-participant. Abbreviations used in this table are defined in
 Table~\abbreviations.}
@@
 \caption{\textbf{Comparing memory in the feature-rich versus reduced (early)
-conditions (all lists).} The $t$-tests reported in the table were carried out
-across-participants, and reflect data aggregated across all lists from each
-participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+conditions (all lists).} The independent samples $t$-tests reported in the
+table were carried out across-participants, and reflect data aggregated across
+all lists from each participant. Abbreviations used in this table are defined
+in Table~\abbreviations.}
 
@@
 \caption{\textbf{Comparing memory in the reduced versus reduced (early)
-conditions (all lists).} The $t$-tests reported in the table were carried out
-across-participants, and reflect data aggregated across all lists from each
-participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+conditions (all lists).} The independent samples $t$-tests reported in the
+table were carried out across-participants, and reflect data aggregated across
+all lists from each participant. Abbreviations used in this table are defined
+in Table~\abbreviations.}
 
@@
 \caption{\textbf{Comparing memory in the feature-rich versus reduced (late)
-conditions (all lists).} The $t$-tests reported in the table were carried out
-across-participants, and reflect data aggregated across all lists from each
-participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+conditions (all lists).} The independent samples $t$-tests reported in the
+table were carried out across-participants, and reflect data aggregated across
+all lists from each participant. Abbreviations used in this table are defined
+in Table~\abbreviations.}
 
@@
 \caption{\textbf{Comparing memory in the reduced versus reduced (late)
-conditions (all lists).} The $t$-tests reported in the table were carried out
-across-participants, and reflect data aggregated across all lists from each
-participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+conditions (all lists).} The independent samples $t$-tests reported in the
+table were carried out across-participants, and reflect data aggregated across
+all lists from each participant. Abbreviations used in this table are defined
+in Table~\abbreviations.}
 
@@
 \caption{\textbf{Comparing accuracy on early lists in the order manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+versus feature-rich conditions.} The independent samples $t$-tests reported in
+the table were carried out across-participants, and reflect data aggregated
+across all early lists from each participant. Abbreviations used in this table
+are defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing temporal clustering on early lists in the order manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing temporal clustering on early lists in the order
+manipulation versus feature-rich conditions.} The independent samples $t$-tests
+reported in the table were carried out across-participants, and reflect data
+aggregated across all early lists from each participant. Abbreviations used in
+this table are defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing feature-based clustering on early lists in the semantic order manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing feature-based clustering on early lists in the
+semantic order manipulation versus feature-rich conditions.} The independent
+samples $t$-tests reported in the table were carried out across-participants,
+and reflect data aggregated across all early lists from each participant.
+Abbreviations used in this table are defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing feature-based clustering on early lists in the lexicographic order manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
+\caption{\textbf{Comparing feature-based clustering on early lists in the
+lexicographic order manipulation versus feature-rich conditions.} The
+independent samples $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all early lists from
+each participant. Abbreviations used in this table are defined in
 Table~\abbreviations.}
@@
 
-\caption{\textbf{Comparing feature-based clustering on early lists in the visual order manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing feature-based clustering on early lists in the
+visual order manipulation versus feature-rich conditions.} The independent
+samples $t$-tests reported in the table were carried out across-participants,
+and reflect data aggregated across all early lists from each participant.
+Abbreviations used in this table are defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing accuracy on early lists in different order manipulation
-conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing accuracy on early lists in different order
+manipulation conditions.} The independent samples $t$-tests reported in the
+table were carried out across-participants, and reflect data aggregated across
+all early lists from each participant. Abbreviations used in this table are
+defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing temporal clustering on early lists in different order manipulation
-conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing temporal clustering on early lists in different
+order manipulation conditions.} The independent samples $t$-tests reported in
+the table were carried out across-participants, and reflect data aggregated
+across all early lists from each participant. Abbreviations used in this table
+are defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing feature-based clustering on early lists in different order manipulation
-conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all early lists
-from each participant. The feature used to compute clustering is shown before the colon in each row, and the
-conditions being compared are shown after the colon.  Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing feature-based clustering on early lists in different
+order manipulation conditions.} The independent samples $t$-tests reported in
+the table were carried out across-participants, and reflect data aggregated
+across all early lists from each participant. The feature used to compute
+clustering is shown before the colon in each row, and the conditions being
+compared are shown after the colon. Abbreviations used in this table are
+defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing accuracy on late lists in order-manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all late
-lists from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing accuracy on late lists in order-manipulation versus
+feature-rich conditions.} The independent samples $t$-tests reported in the
+table were carried out across-participants, and reflect data aggregated across
+all late lists from each participant. Abbreviations used in this table are
+defined in Table~\abbreviations.}
 
@@
 
-\caption{\textbf{Comparing temporal clustering on late lists in order-manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all late
-lists from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
-
-\label{tab:temporal-carryover}
-\end{table}
-
-\begin{table}[tp]
-\centering
-\tiny
-\input{tables/semantic_carryover_feature_clustering.tex}
-
-\caption{\textbf{Comparing feature-based clustering on late lists in semantic order-manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all late
-lists from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
-
-\label{tab:semantic-feature-carryover}
-\end{table}
-
-\begin{table}[tp]
-\centering
-\tiny
-\input{tables/lexicographic_carryover_feature_clustering.tex}
-
-\caption{\textbf{Comparing feature-based clustering on late lists in
-lexicographic order-manipulation versus feature-rich conditions.} The
+\caption{\textbf{Comparing temporal clustering on late lists in
+order-manipulation versus feature-rich conditions.} The independent samples
 $t$-tests reported in the table were carried out across-participants, and
@@
 
+\label{tab:temporal-carryover}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/semantic_carryover_feature_clustering.tex}
+
+\caption{\textbf{Comparing feature-based clustering on late lists in semantic
+order-manipulation versus feature-rich conditions.} The independent samples
+$t$-tests reported in the table were carried out across-participants, and
+reflect data aggregated across all late lists from each participant.
+Abbreviations used in this table are defined in Table~\abbreviations.}
+
+\label{tab:semantic-feature-carryover}
+\end{table}
+
+\begin{table}[tp]
+\centering
+\tiny
+\input{tables/lexicographic_carryover_feature_clustering.tex}
+
+\caption{\textbf{Comparing feature-based clustering on late lists in
+lexicographic order-manipulation versus feature-rich conditions.} The
+independent samples $t$-tests reported in the table were carried out
+across-participants, and reflect data aggregated across all late lists from
+each participant. Abbreviations used in this table are defined in
+Table~\abbreviations.}
+
 \label{tab:lexicographic-feature-carryover}
@@
 
-\caption{\textbf{Comparing feature-based clustering on late lists in visual order-manipulation
-versus feature-rich conditions.} The $t$-tests reported in the table were
-carried out across-participants, and reflect data aggregated across all late
-lists from each participant. Abbreviations used in this table are defined in
-Table~\abbreviations.}
+\caption{\textbf{Comparing feature-based clustering on late lists in visual
+order-manipulation versus feature-rich conditions.} The independent samples
+$t$-tests reported in the table were carried out across-participants, and
+reflect data aggregated across all late lists from each participant.
+Abbreviations used in this table are defined in Table~\abbreviations.}
 
@@
 (Fig.~\ref{fig:clustering-scatterplots}G, H). Participants who showed stronger
-feature clustering on early lists showed stronger temporal
-clustering on later lists (overall: $r(179) = 0.464,~p < 0.001,~\mathrm{CI} =
-[0.321, 0.582]$; for individual conditions: all $r$s $\geq 0.377$, all $p$s
-$\leq 0.040$; across conditions: $ r(4) = 0.451,~p = 0.369,~\mathrm{CI} =
-[-0.986, 0.998]$). And participants who showed stronger temporal clustering on
-early lists trended towards showing stronger feature clustering on later lists
-(overall: $r(179) = 0.266,~p < 0.001,~\mathrm{CI} = [0.129, 0.396]$; for
-individual conditions: all $r$s $\geq 0.298$, all $p$s $\leq 0.110$; across
-conditions: $r(4) = 0.064,~p = 0.903,~\mathrm{CI} = [-0.972,~0.965$). Taken together,
-the results displayed in Figure~\ref{fig:clustering-scatterplots} show that
-participants who were more sensitive to the order manipulations (i.e.,
-participants who showed stronger feature clustering for their condition's
-feature on early lists) remembered more words and showed stronger temporal
-clustering. These associations also appeared to carry over across lists, even
-when the items on later lists were presented in a random order.
+feature clustering on early lists showed stronger temporal clustering on later
+lists (overall: $r(179) = 0.464,~p < 0.001,~\mathrm{CI} = [0.321, 0.582]$; for
+individual conditions: all $r$s $\geq 0.377$, all $p$s $\leq 0.040$; across
+conditions: $ r(4) = 0.451,~p = 0.369,~\mathrm{CI} = [-0.986, 0.998]$). And
+participants who showed stronger temporal clustering on early lists trended
+towards showing stronger feature clustering on later lists (overall: $r(179) =
+0.266,~p < 0.001,~\mathrm{CI} = [0.129, 0.396]$; for individual conditions: all
+$r$s $\geq 0.298$, all $p$s $\leq 0.110$; across conditions: $r(4) = 0.064,~p =
+0.903,~\mathrm{CI} = [-0.972,~0.965$). Taken together, the results displayed in
+Figure~\ref{fig:clustering-scatterplots} show that participants who were more
+sensitive to the order manipulations (i.e., participants who showed stronger
+feature clustering for their condition's feature on early lists) remembered
+more words and showed stronger temporal clustering. These associations also
+appeared to carry over to later lists, even when the items on those later lists
+were presented in a random order.
 
@@
 0.052,~p = 0.787,~\mathrm{CI} = [-0.307,~0.360]$; across conditions: $r(4) =
-0.635,~p = 0.176,~\mathrm{CI} = [-0.924,~0.981]$. Participants who exhibited
+0.635,~p = 0.176,~\mathrm{CI} = [-0.924,~0.981]$). Participants who exhibited
 larger carryover in feature clustering also tended to show stronger temporal

### 2023-11-22 paper/main.tex

@@
 
-\title{\Large Feature and order manipulations in a free recall task affect memory for current and future lists}
-
-\author{Jeremy R. Manning\textsuperscript{1, *}, Emily C.
-Whitaker\textsuperscript{1}, Paxton C. Fitzpatrick\textsuperscript{1},
-\\Madeline R. Lee\textsuperscript{1}, Allison M. Frantz\textsuperscript{1},
-Bryan J. Bollinger\textsuperscript{1},\\Darya Romanova\textsuperscript{1},
-Campbell E. Field\textsuperscript{1}, and Andrew C. Heusser\textsuperscript{1,
-2}\\\textsuperscript{1}Dartmouth College\\\textsuperscript{2}Akili
-Interactive Labs\\\textsuperscript{*}Corresponding author:
-jeremy.r.manning@dartmouth.edu}
+\title{\large Feature and order manipulations in a free recall task affect memory for current and future lists}
+
+\author{\small{Jeremy R. Manning\textsuperscript{1, *}}, \small{Emily C.
+Whitaker\textsuperscript{1}}, \small{Paxton C. Fitzpatrick\textsuperscript{1}},
+\\\small{Madeline R. Lee\textsuperscript{1}}, \small{Allison M. Frantz\textsuperscript{1}},
+\small{Bryan J. Bollinger\textsuperscript{1}},\\\small{Darya Romanova\textsuperscript{1}},
+\small{Campbell E. Field\textsuperscript{1}}, and \small{Andrew C. Heusser\textsuperscript{1,
+2}}\\\small{\textsuperscript{1}Dartmouth College}\\\small{\textsuperscript{2}Akili
+Interactive Labs}\\\small{\textsuperscript{*}Corresponding author:
+jeremy.r.manning@dartmouth.edu}}
 
@@
 \begin{document}
+\begin{titlepage}
+\singlespacing
+\footnotesize
 \maketitle
@@
 \end{abstract}
+
+\paragraph*{Author contributions.}
+
+Conceptualization: JRM and ACH. Data curation: JRM, PCF, ACH. Formal analysis:
+JRM, PCF, and ACH. Funding acquisition: JRM. Investigation: ECW, PCF, MRL, AMF,
+BJB, DR, CEF, and ACH. Methodology: JRM and ACH. Project administration: ECW
+and PCF. Resources: JRM. Software: JRM, PCF, CEF, and ACH. Supervision: JRM and
+ACH. Validation: JRM, PCF, and ACH. Writing (original draft): JRM. Writing
+(review and editing): ECW, PCF, MRL, AMF, BJB, DR, CEF, and ACH.
+
+\paragraph*{Author note.}
+
+All of the data analyzed in this manuscript, along with all of the code for
+carrying out the analyses may be found at
+https://github.com/ContextLab/FRFR-analyses. Note that this study was not
+preregistered.  Some of the ideas and data presented in this manuscript
+were also presented at the Annual Meeting of the Society for Neuroscience (2017)
+and the Context and Episodic Memory Symposium (2017).
+
+\paragraph*{Acknowledgements.}
+
+We acknowledge useful discussions, assistance in setting up an earlier
+(unpublished) version of this study, and assistance with some of the data
+collection efforts from Rachel Chacko, Joseph Finkelstein, Sheherzad Mohydin,
+Lucy Owen, Gal Perlman, Jake Rost, Jessica Tin, Marisol Tracy,
+Peter Tran, and Kirsten Ziman. Our work was supported in part by NSF CAREER Award Number
+2145172 to JRM. The content is solely the responsibility of the authors and
+does not necessarily represent the official views of our supporting
+organizations. The funders had no role in study design, data collection and
+analysis, decision to publish, or preparation of the manuscript.
+
+\end{titlepage}
+
+\normalsize
 
@@
 us to estimate which features participants are considering or leveraging in
-organizing their memories.
+organizing their memories.  Our study was approved by the Dartmouth College
+Committee for the Protection of Human Subjects (CPHS) under a protocol titled
+``\textit{Efficient Learning}.''
 
@@
 
-\subsection*{Author contributions}
-
-Conceptualization: JRM and ACH. Data curation: JRM, PCF, ACH. Formal analysis:
-JRM, PCF, and ACH. Funding acquisition: JRM. Investigation: ECW, PCF, MRL, AMF,
-BJB, DR, CEF, and ACH. Methodology: JRM and ACH. Project administration: ECW
-and PCF. Resources: JRM. Software: JRM, PCF, CEF, and ACH. Supervision: JRM and
-ACH. Validation: JRM, PCF, and ACH. Writing (original draft): JRM. Writing
-(review and editing): ECW, PCF, MRL, AMF, BJB, DR, CEF, and ACH.
-
-\subsection*{Author note}
-
-All of the data analyzed in this manuscript, along with all of the code for
-carrying out the analyses may be found at
-https://github.com/ContextLab/FRFR-analyses. Code for running the non-adaptive
-experimental conditions may be found at
-https://github.com/Con\-text\-Lab/efficient-learning-code. Code for running the
-adaptive experimental condition may be found at
-https://github.com/ContextLab/adaptiveFR. We have also released an associated
-Python toolbox for analyzing free recall data, which may be found at
-https://cdl-quail.read\-the\-docs.io/\-en/\-latest.  Note that this study was not
-preregistered.  Some of the ideas and data presented in this manuscript
-were also presented at the Annual Meeting of the Society for Neuroscience (2017)
-and the Context and Episodic Memory Symposium (2017).
-
-\subsection*{Acknowledgements}
-
-We acknowledge useful discussions, assistance in setting up an earlier
-(unpublished) version of this study, and assistance with some of the data
-collection efforts from Rachel Chacko, Joseph Finkelstein, Sheherzad Mohydin,
-Lucy Owen, Gal Perlman, Jake Rost, Jessica Tin, Marisol Tracy,
-Peter Tran, and Kirsten Ziman. Our work was supported in part by NSF CAREER Award Number
-2145172 to JRM. The content is solely the responsibility of the authors and
-does not necessarily represent the official views of our supporting
-organizations. The funders had no role in study design, data collection and
-analysis, decision to publish, or preparation of the manuscript.
-
 \bibliographystyle{apa}
